### 2018-09-29 mprop/mprop.tex
References

\documentclass{mprop}
\usepackage{graphicx}
% other potentially useful packages
%\uspackage{amssymb,amsmath}
%\usepackage{url}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}
\begin{document}

\title{Variational Inference for Inverse Reinforcement Learning with Gaussian
  Processes}
\author{Paulius Dilkas}
\date{Date of submission placed here} % TODO
\maketitle
\tableofcontents
\newpage

\section{Introduction}\label{intro}
briefly explain the context of the project problem
\subsection{A subsection}
Please note your proposal need not follow the included section headings - this is only a suggested structure. Also add subsections etc as required
\section{Statement of Problem}
clearly state the problem to be addressed in your forthcoming project. Explain why it would be worthwhile to solve this problem.
\section{Background Survey}
present an overview of relevant previous work including articles, books, and existing software products. Critically evaluate the strengths and weaknesses of the previous work.
\section{Proposed Approach}
state how you propose to solve the software development problem. Show that your proposed approach is feasible, but identify any risks.
\section{Work Plan}
show how you plan to organize your work, identifying intermediate deliverables and dates.

\section{Notes on papers (to be removed)}

\subsection{Miscellaneous}

(Directed) similarity between MDPs using restricted Boltzmann machines
\cite{9401f4eeb9a64c77afb3d087261d1080}

Chapter 6 on distance measures \cite{mccune2002analysis}

The PhD thesis behind maximum causal entropy \cite{Ziebart:2010:MPA:2049078}

\subsection{Gaussian Processes}

Your recommended book \cite{DBLP:books/lib/RasmussenW06}

Simple introduction to GPs for time-series modelling \cite{Roberts2013GaussianPF}

Spectral kernels \cite{pmlr-v28-wilson13}

GPs over graphs instead of vectors (haven't actually read)
\cite{DBLP:journals/corr/abs-1803-05776}

Another introduction from physics (skimmed through) \cite{introduction_to_gps}

Learning a GP from very little data \cite{DBLP:conf/nips/PlattBSWZ01}

One GP for multiple correlated output variables \cite{DBLP:journals/jcphy/BilionisZKL13}

Kernels for categorical and count data \cite{savitsky2011variable}

\subsection{Interpretability}

Learning latent factors \cite{DBLP:conf/nips/LiSE17}

The behaviour of Reddit users \cite{DBLP:conf/atal/DasL14}

\subsection{Inverse Reinforcement Learning}

One of the first papers on the topic \cite{DBLP:conf/icml/NgR00}

Follow-up on the previous paper \cite{DBLP:conf/icml/PieterN04}

Bayesian setting \cite{DBLP:conf/ijcai/RamachandranA07}

Learning optimal composite features \cite{DBLP:conf/ijcai/ChoiK13}

The main paper \cite{DBLP:conf/nips/LevinePK11}

A different take on IRL with GPs \cite{DBLP:journals/corr/abs-1208-2112}

The paper that introduced maximum entropy into IRL \cite{ziebart2008maximum}

IRL for large state spaces (haven't read) \cite{DBLP:journals/jmlr/BoulariasKP11}

Multiple reward functions \cite{DBLP:conf/nips/ChoiK12}

\subsubsection{Multiple Strategies}

EM clustering \cite{DBLP:conf/icml/BabesMLS11}

Structured priors \cite{DBLP:conf/ewrl/DimitrakakisR11}

There are more, but I haven't gotten to them yet.

\subsection{Variational Inference}

Chapter 10 on approximate inference \cite{DBLP:books/lib/Bishop07}

Part IV on probabilities and inference \cite{MacKay:2002:ITI:971143}

A recent review \cite{blei2017variational}

Normalizing flows \cite{DBLP:conf/icml/RezendeM15}

Linear VI for GPs \cite{DBLP:conf/nips/ChengB17}

Stochastic VI \cite{DBLP:journals/jmlr/HoffmanBWP13}

Structured stochastic VI (haven't read) \cite{DBLP:conf/aistats/HoffmanB15}

Another review of recent advances \cite{DBLP:journals/corr/abs-1711-05597}

Sparse VI for GP \cite{DBLP:journals/jmlr/HensmanDS17}

Sparse GPs \cite{DBLP:journals/jmlr/CandelaR05}

IRL via deep GP \cite{DBLP:conf/uai/JinDAS17}

\bibliographystyle{plain}
\bibliography{mprop}
\end{document}

### 2018-10-03 mprop/mprop.tex
Describing the problem

@@
 \usepackage{graphicx}
-% other potentially useful packages
-%\uspackage{amssymb,amsmath}
-%\usepackage{url}
+\usepackage{amssymb,amsmath,amsthm}
+\usepackage{url}
 %\usepackage{fancyvrb}
 %\usepackage[final]{pdfpages}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\DeclareMathOperator*{\argmax}{arg\,max}
+
 \begin{document}
-
 \title{Variational Inference for Inverse Reinforcement Learning with Gaussian
@@
 
-\section{Introduction}\label{intro}
-briefly explain the context of the project problem
-\subsection{A subsection}
-Please note your proposal need not follow the included section headings - this is only a suggested structure. Also add subsections etc as required
-\section{Statement of Problem}
-clearly state the problem to be addressed in your forthcoming project. Explain why it would be worthwhile to solve this problem.
-\section{Background Survey}
-present an overview of relevant previous work including articles, books, and existing software products. Critically evaluate the strengths and weaknesses of the previous work.
+\newcommand{\Kuu}{\mathbf{K}_{\mathbf{u},\mathbf{u}}}
+\newcommand{\Krr}{\mathbf{K}_{\mathbf{r},\mathbf{r}}}
+\newcommand{\Kru}{\mathbf{K}_{\mathbf{r},\mathbf{u}}}
+\newcommand{\DKL}{D_{\mathrm{KL}}}
+\newcommand{\pfull}{p(\mathcal{D}, \Theta, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
+\newcommand{\posterior}{p(\Theta, \mathbf{u}, \mathbf{r} | \mathcal{D}, \mathbf{X_u})}
+\newcommand{\approximation}{q(\Theta, \mathbf{u}, \mathbf{r}; \Lambda)}
+
+\section{Introduction}
+%briefly explain the context of the project problem
+
+Inverse reinforcement learning (IRL)---a problem proposed by Russell in 1998
+\cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a Markov
+decision process that best explains a set of given demonstrations. IRL is
+important because reward functions can be hard to define manually
+\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
+are not entirely specific to a given environment, allowing one to reuse the same
+reward structure in previously unseen environments
+\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
+Moreover, IRL has seen a wide array of applications in autonomous vehicle
+control
+\cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16}
+and learning to predict another agent's behaviour
+\cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
+Most approaches in the literature (see Section \ref{literature}) make a
+convenient yet unjustified assumption that the reward function can be expressed
+as a linear combination of features. One proven way to abandon this assumption
+is by representing the reward function as a Gaussian process
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
+
+\section{Statement of the Problem}
+%clearly state the problem to be addressed in your forthcoming project. Explain
+%why it would be worthwhile to solve this problem.
+
+\begin{definition}
+  A \emph{Markov decision process} (MDP) is a set $\mathcal{M} = \{ \mathcal{S},
+  \mathcal{A}, \mathcal{T}, \gamma, r \}$, where $\mathcal{S}$ and
+  $\mathcal{A}$ are sets of states and actions, respectively; $\mathcal{T} :
+  \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is a function
+  defined so that $\mathcal{T}(s, a, s')$ is the probability of moving to state $s'$
+  after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the discount
+  factor (with higher $\gamma$ values, it makes little difference whether a
+  reward is received now or later, while with lower $\gamma$ values the future
+  becomes gradually less and less important); and $r : \mathcal{S} \to
+  \mathbb{R}$ is the reward function.
+\end{definition}
+
+In \emph{inverse reinforcement learning}, one is presented with an MDP without a
+reward function $\mathcal{M} \setminus \{ r \}$ and a set of expert
+demonstrations $\mathcal{D} = \{ \zeta_i \}_{i=1}^N$, where each demonstration
+$\zeta_i = \{ (s_{i,0}, a_{i,0}), \dots, (s_{i,T}, a_{i,T}) \}$ is a multiset of
+state-action pairs representing the actions taken by the expert during a
+particular recorded session. Each state is also characterised by a number of
+features. The goal of IRL is then to find $r$ such that the optimal policy under
+$r$
+\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \sum_{t=0}^\infty \gamma^t r(s_t) | \pi
+  \right] \]
+matches the actions in $\mathcal{D}$.
+
+The likelihood of the data can be written down as
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
+\begin{equation} \label{pDr}
+  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q(s_{i,t}, a_{i,t}; r) - V(s_{i,t}; r) \right),
+\end{equation}
+where
+\[ Q(s_{i,t}, a_{i,t}; r) = r(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s_{i,t}, a_{i,t}, s')V(s'; r), \]
+and $V(s; r)$ can be obtained by repeatedly applying the equation
+\cite{supplementary_material}
+\[ V(s; r) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
+      \mathcal{S}} \mathcal{T}(s, a, s')V(s'; r) \right). \]
+However, a reward function learned by maximising this likelihood is not
+transferable to new situations
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
+reward structure in a way that would allow reward predictions for previously
+unseen states.
+
+One way to model rewards without assumptions of linearity is with a
+\emph{Gaussian process} (GP). A GP is a collection of random variables, any
+finite combination of which has a joint Gaussian distribution
+\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
+k_{\Theta})$ to say that $r$ is a GP with mean $0$ and covariance function
+$k_{\Theta}$, which uses a set of hyperparameters $\Theta$. Covariance functions
+take two state feature vectors as input and quantify how similar the two states
+are, in a sense that we would expect them to have similar rewards.
+
+% TODO: import info from the MSci paper
+As training a GP with $n$ data points has a time complexity of
+$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
+methods have been suggested, many of which select a subset of data called
+\emph{inducing points} and focus most of the training effort on them
+\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
+features at inducing states, $\mathbf{u}$ the rewards at those states, and
+$\mathbf{r}$ a vector with $r(\mathcal{S})$ as elements. Then the full joint
+probability distribution can be factorised as
+\begin{equation} \label{full}
+  \pfull = p(\mathbf{X_u}) \times p(\Theta | \mathbf{X_u}) \times p(\mathbf{u}
+  | \Theta, \mathbf{X_u}) \times p(\mathbf{r} | \Theta, \mathbf{X_u},
+  \mathbf{u}) \times p(\mathcal{D} | r).
+\end{equation}
+Here $p(\mathbf{X_u})$ and $p(\Theta | \mathbf{X_u})$ are freely chosen priors,
+\[
+  \begin{split}
+    p(\mathbf{u} | \Theta, \mathbf{X_u}) &= \mathcal{N}(\mathbf{0},
+    \Kuu) \\
+    &= \frac{1}{(2\pi)^{n/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2}
+      \mathbf{u}^T\Kuu^{-1}\mathbf{u} \right) \\
+    &= \exp\left(-\frac{1}{2}\mathbf{u}^T\Kuu^{-1}\mathbf{u} -
+      \frac{1}{2}\log|\Kuu| - \frac{n}{2}\log 2\pi\right)
+  \end{split}
+\]
+is the GP prior \cite{DBLP:books/lib/RasmussenW06}, the GP posterior is a
+multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11}
+\begin{equation} \label{eq:r}
+  p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) =
+  \mathcal{N}(\Kru^T\Kuu^{-1}\mathbf{u}, \Krr - \Kru^T\Kuu^{-1}\Kru),
+\end{equation}
+and $p(\mathcal{D} | r)$ is as in Equation \ref{pDr}. The matrices such as
+$\Kru$ are called \emph{covariance matrices} and are defined as
+$[\Kru]_{i,j} = k_\Theta(\mathbf{x}_{\mathbf{r},i}, \mathbf{x}_{\mathbf{u},j})$,
+where $\mathbf{x}_{\mathbf{r},i}$ and $\mathbf{x}_{\mathbf{u},j}$ denote feature
+vectors for the $i$th state in $\mathcal{S}$ and the $j$th state in
+$\mathbf{X_u}$, respectively.
+
+Given this model, data $\mathcal{D}$, and inducing feature matrix
+$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
+$\Theta$, inducing rewards $\mathbf{u}$, and the reward function $r$. While the
+previous paper that considered this IRL model computed maximum likelihood estimates
+for $\Theta$ and $\mathbf{u}$, and made an assumption that $\mathbf{r}$ in
+Equation \ref{eq:r} has $0$ variance \cite{DBLP:conf/nips/LevinePK11}, we aim to
+avoid this assumption and use variational inference to approximate the full
+posterior distribution $\posterior$. \emph{Variational inference} (VI) is an
+approximation technique for probability densities \cite{blei2017variational}.
+Let $\approximation$ be our approximating family of probability distributions
+for $\posterior$ with its own set of hyperparameters $\Lambda$. Then it is up to
+VI algorithms to optimise $\Lambda$ in order to minimise the
+\emph{Kullback-Leibler} (KL) divergence between the original probability
+distribution and our approximation.  KL divergence (asymmetrically) measures how
+different the two distributions are and in this case can be defined as follows
+\cite{blei2017variational}:
+\[ \DKL(\approximation || \posterior) = \mathbb{E}_q[\log\approximation] -
+  \mathbb{E}_q[\log \posterior]. \]
+Since KL divergence is typically hard to compute, instead of minimising it, VI
+typically tries to maximise the \emph{evidence lower bound} (ELBO) defined as
+\[
+  \begin{split}
+    \mathcal{L}(\Lambda) &= \mathbb{E}_q[\log \pfull] -
+    \mathbb{E}_q[\log\approximation] \\
+    &= \iiint \approximation \log
+    \frac{\pfull}{\approximation}\,d\Theta\,d\mathbf{u}\,d\mathbf{r}.
+  \end{split}
+\]
+% TODO: cite something
+
+By considering full probability distributions instead of point estimates,---as
+long as the approximations are able to capture important features of the
+posterior---our predictions are likely to be more accurate and rely on fewer
+assumptions. Moreover, we hope to make use of various recent advancements in VI
+for both time complexity and approximation distribution fit (see Section
+\ref{literature}), making the resulting algorithm competitive both in terms of
+speed and model fit.
+
+\section{Literature Survey} \label{literature}
+
+present an overview of relevant previous work including articles, books, and
+existing software products. Critically evaluate the strengths and weaknesses of
+the previous work.
+
 \section{Proposed Approach}
-state how you propose to solve the software development problem. Show that your proposed approach is feasible, but identify any risks.
+
+%state how you propose to solve the software development problem. Show that your
+%proposed approach is feasible, but identify any risks.
+
+We begin by rewriting the posterior by using the chain rule and Bayes' theorem,
+trying to extract parts of the distribution we know how to compute, namely those
+in Equation \ref{full}:
+\[
+  \begin{split}
+    \posterior &= p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
+    \Theta, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \Theta,
+    \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
+    &\propto p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} | \Theta,
+    \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times p(\mathbf{r} |
+    \Theta, \mathbf{X_u}, \mathbf{u}) \\
+    &\propto p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
+    \Theta, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \Theta,
+    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta,
+    \mathbf{X_u}, \mathbf{u}) \\
+    &\propto p(\mathcal{D} | \Theta, \mathbf{X_u}) \times p(\Theta |
+    \mathbf{X_u}) \times p(\mathcal{D} | \Theta, \mathbf{X_u}, \mathbf{u})
+    \times p(\mathbf{u} | \Theta, \mathbf{X_u}) \times p(\mathcal{D} | r) \times
+    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u})
+  \end{split}
+\]
+Note that now there are only two unknown probability distributions:
+$p(\mathcal{D} | \Theta, \mathbf{X_u})$ and $p(\mathcal{D} | \Theta,
+\mathbf{X_u}, \mathbf{u})$. They can be computed as follows:
+\begin{align*}
+  p(\mathcal{D} | \Theta, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
+  p(\mathcal{D} | \Theta, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \Theta, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
+\end{align*}
+
 \section{Work Plan}
-show how you plan to organize your work, identifying intermediate deliverables and dates.
+
+show how you plan to organize your work, identifying intermediate deliverables
+and dates.
 
@@
 \subsection{Gaussian Processes}
-
-Your recommended book \cite{DBLP:books/lib/RasmussenW06}
 
@@
 
+Scalability/Approximations thesis \cite{kth}
+
 \subsection{Interpretability}
@@
 
-Follow-up on the previous paper \cite{DBLP:conf/icml/PieterN04}
-
 Bayesian setting \cite{DBLP:conf/ijcai/RamachandranA07}
@@
 
-The main paper \cite{DBLP:conf/nips/LevinePK11}
-
 A different take on IRL with GPs \cite{DBLP:journals/corr/abs-1208-2112}
-
-The paper that introduced maximum entropy into IRL \cite{ziebart2008maximum}
 
@@
 Multiple reward functions \cite{DBLP:conf/nips/ChoiK12}
+
+A recent survey \cite{DBLP:journals/corr/abs-1806-06877}
+
+Some not-very-successful method \cite{DBLP:conf/uai/NeuS07}
 
@@
 
-Chapter 10 on approximate inference \cite{DBLP:books/lib/Bishop07}
-
 Part IV on probabilities and inference \cite{MacKay:2002:ITI:971143}
-
-A recent review \cite{blei2017variational}
 
@@
 
-Sparse GPs \cite{DBLP:journals/jmlr/CandelaR05}
+Sparse GPs \cite{DBLP:journals/jmlr/CandelaR05}.
 
-IRL via deep GP \cite{DBLP:conf/uai/JinDAS17}
+Tighter ELBOs are not necessarily better \cite{DBLP:conf/icml/RainforthKLMIWT18}.
+
+\cite{DBLP:books/lib/Bishop07}
 

### 2018-10-23 mprop/mprop.tex
I noticed I haven't commited in a while...

@@
 \usepackage{url}
+\usepackage{bm}
+\usepackage{bbm}
+\usepackage{booktabs}
 %\usepackage{fancyvrb}
@@
 
+\newtheorem{lemma}{Lemma}
 \theoremstyle{definition}
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator{\Var}{Var}
+\DeclareMathOperator{\diag}{diag}
 
@@
 
+% TODO: table of notation?
 \newcommand{\Kuu}{\mathbf{K}_{\mathbf{u},\mathbf{u}}}
@@
 \newcommand{\pfull}{p(\mathcal{D}, \Theta, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
+\newcommand{\pfullS}{p(\mathcal{D}, \Theta_s, \mathbf{X_u}, \mathbf{u}_s, \mathbf{r}_s)}
 \newcommand{\posterior}{p(\Theta, \mathbf{u}, \mathbf{r} | \mathcal{D}, \mathbf{X_u})}
-\newcommand{\approximation}{q(\Theta, \mathbf{u}, \mathbf{r}; \Lambda)}
+\newcommand{\approximation}{q_{\bm\nu}(\Theta, \mathbf{u}, \mathbf{r})}
+\newcommand{\approximationS}{q_{\bm\nu}(\Theta_s, \mathbf{u}_s, \mathbf{r}_s)}
+\newcommand{\Eq}{\mathbb{E}_{(\Theta, \mathbf{u}, \mathbf{r}) \sim \approximation}}
 
@@
 Here $p(\mathbf{X_u})$ and $p(\Theta | \mathbf{X_u})$ are freely chosen priors,
-\[
-  \begin{split}
-    p(\mathbf{u} | \Theta, \mathbf{X_u}) &= \mathcal{N}(\mathbf{0},
-    \Kuu) \\
-    &= \frac{1}{(2\pi)^{n/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2}
-      \mathbf{u}^T\Kuu^{-1}\mathbf{u} \right) \\
-    &= \exp\left(-\frac{1}{2}\mathbf{u}^T\Kuu^{-1}\mathbf{u} -
-      \frac{1}{2}\log|\Kuu| - \frac{n}{2}\log 2\pi\right)
-  \end{split}
-\]
+\begin{align}
+  p(\mathbf{u} | \Theta, \mathbf{X_u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \nonumber \\
+                                       &= \frac{1}{(2\pi)^{n/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^T\Kuu^{-1}\mathbf{u} \right) \nonumber \\
+                                       &= \exp\left(-\frac{1}{2}\mathbf{u}^T\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right) \label{eq:normal}
+  \end{align} % TODO: introduce m here, deleting later introduction
 is the GP prior \cite{DBLP:books/lib/RasmussenW06}, the GP posterior is a
@@
   p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) =
-  \mathcal{N}(\Kru^T\Kuu^{-1}\mathbf{u}, \Krr - \Kru^T\Kuu^{-1}\Kru),
+  \mathcal{N}(\mathbf{r}; \Kru^T\Kuu^{-1}\mathbf{u}, \Krr - \Kru^T\Kuu^{-1}\Kru),
 \end{equation}
-and $p(\mathcal{D} | r)$ is as in Equation \ref{pDr}. The matrices such as
+and $p(\mathcal{D} | r)$ is as in \eqref{pDr}. The matrices such as
 $\Kru$ are called \emph{covariance matrices} and are defined as
@@
 vectors for the $i$th state in $\mathcal{S}$ and the $j$th state in
-$\mathbf{X_u}$, respectively.
+$\mathbf{X_u}$, respectively \cite{DBLP:conf/uai/JinDAS17}.
 
@@
 for $\Theta$ and $\mathbf{u}$, and made an assumption that $\mathbf{r}$ in
-Equation \ref{eq:r} has $0$ variance \cite{DBLP:conf/nips/LevinePK11}, we aim to
+\eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11}, we aim to
 avoid this assumption and use variational inference to approximate the full
@@
 Let $\approximation$ be our approximating family of probability distributions
-for $\posterior$ with its own set of hyperparameters $\Lambda$. Then it is up to
-VI algorithms to optimise $\Lambda$ in order to minimise the
+for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up
+to VI algorithms to optimise $\bm\nu$ in order to minimise the
 \emph{Kullback-Leibler} (KL) divergence between the original probability
@@
 \cite{blei2017variational}:
-\[ \DKL(\approximation || \posterior) = \mathbb{E}_q[\log\approximation] -
-  \mathbb{E}_q[\log \posterior]. \]
+\[ \DKL(\approximation || \posterior) = \Eq \left[ \log
+  \frac{\approximation}{\posterior} \right]. \]
 Since KL divergence is typically hard to compute, instead of minimising it, VI
 typically tries to maximise the \emph{evidence lower bound} (ELBO) defined as
+\cite{DBLP:books/lib/Bishop07,blei2017variational}
 \[
   \begin{split}
-    \mathcal{L}(\Lambda) &= \mathbb{E}_q[\log \pfull] -
-    \mathbb{E}_q[\log\approximation] \\
+    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
+    \right] \\
     &= \iiint \approximation \log
@@
 \]
-% TODO: cite something
 
@@
 \section{Literature Survey} \label{literature}
+%present an overview of relevant previous work including articles, books, and
+%existing software products. Critically evaluate the strengths and weaknesses of
+%the previous work.
 
-present an overview of relevant previous work including articles, books, and
-existing software products. Critically evaluate the strengths and weaknesses of
-the previous work.
+Variational inference has seen a recent increase in interest among academics,
+with different approaches focusing on different goals: better time complexity,
+handling a wider variety of models, making approximations more accurate, and
+using more complex function approximation techniques (such as neural networks)
+to infer local latent variable values without having to calculate them
+individually for each data point \cite{DBLP:journals/corr/abs-1711-05597}. As
+our IRL model is based on a GP, we will begin by reviewing some of the VI
+approaches applied specifically to GP regression. Based on a recent review of
+scalable GPs \cite{DBLP:journals/corr/abs-1711-05597}, we will concentrate on
+\emph{stochastic variational sparse approximations}, as they have achieved
+modelling accuracy close to that of the full GP with no approximations, while
+providing a time complexity of $\mathcal{O}(m^3)$, where $m$ is the number of
+inducing points. Below we provide a short overview of various assumptions that
+have been used in approximating \emph{sparse} GPs (i.e., GPs that use inducing
+points), following on a paper by Qui\~{n}onero-Candela and Rasmussen
+\cite{DBLP:journals/jmlr/CandelaR05}.
 
-\section{Proposed Approach}
+\begin{description}
+  \item [Subset of data ($\mathcal{O}(m^3)$)] is a baseline method of simply
+    using a subset of data points.
+  \item [Subset of regressors ($\mathcal{O}(nm^2)$)
+    \cite{silverman1985some,DBLP:conf/nips/SmolaB00,DBLP:conf/nips/WahbaLGXKK98}]
+    is a degenerate approximation that uses a weight for each inducing point. A
+    GP is called \emph{degenerate} if the covariance function has a finite
+    number of non-zero eigenvalues, restricting the prior distribution to only a
+    finite number of linearly independent functions
+    \cite{DBLP:journals/jmlr/CandelaR05}.
+  \item [Deterministic training conditional ($\mathcal{O}(nm^2)$)
+    \cite{DBLP:conf/aistats/SeegerWL03}] approximation imposes a zero-variance
+    normal distribution for $\mathbf{r} | \mathbf{u}$, resulting in the same mean
+    but different variance predictions compared to the subset of regressors.
+  \item [Fully independent training conditional ($\mathcal{O}(nm^2)$)
+    \cite{DBLP:conf/nips/SnelsonG05}] has the assumption that the GP values are
+    independent of each other when conditioned on the inducing values:
+    \[ q(\mathbf{r} | \mathbf{u}) = \prod_{i=1}^n p(r_i | \mathbf{u}). \]
+  \item [Partially independent training conditional ($\mathcal{O}(nm^2)$)
+    \cite{DBLP:journals/neco/Tresp00,DBLP:conf/nips/SchwaighoferT02}]
+    approximates the same distribution as the fully independent training
+    conditional, but considers a block diagonal rather than a diagonal
+    covariance matrix.
+  \item [Transduction] tailors the predictive distribution to
+    specific test inputs \cite{DBLP:journals/jmlr/CandelaR05}. As we are not too
+    concerned about a specific set of test inputs in the IRL setting,
+    transduction is of limited interest to us.
+  \item [Augmentation \cite{rasmussen2002reduced}] aims to improve predictive
+    accuracy by adding each test input to the inducing points.
+  \item [Nystr\"{o}m approximation ($\mathcal{O}(nm^2)$)
+    \cite{DBLP:conf/nips/WilliamsS00}] approximates the prior covariance of
+    $\mathbf{r}$, but can lead to negative predictive variances.
+  \item [Relevance vector machine ($\mathcal{O}(m^3)$)
+    \cite{DBLP:journals/jmlr/Tipping01}] is a degenerative approximation
+    supporting a limited range of covariance functions.
+\end{description}
+
+\begin{table}
+  \centering % TODO: recheck and explain
+  \begin{tabular}{llllll} % Lower bound? Model? % TODO: comment about 'modelled' no longer needed
+    \toprule
+    Authors, year & Inducing points & Hyperparameters & Complexity \\
+    \midrule
+    Titsias, 2009 \cite{DBLP:journals/jmlr/Titsias09} & variational & variational & $\mathcal{O}(nm^2)$ \\
+    Hensman et al., 2013 \cite{DBLP:conf/uai/HensmanFL13} & fixed & variational & $\mathcal{O}(m^3)$ \\
+    Gal et al., 2014 \cite{DBLP:conf/nips/GalWR14} & variational & variational & $\mathcal{O}(nm^2)$ \\
+    Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
+    Cheng and Boots, 2017 \cite{DBLP:conf/nips/ChengB17} & variational & variational & $\mathcal{O}(nm_\alpha + nm_\beta^2)$ \\
+    Hensman et al., 2017 \cite{DBLP:journals/jmlr/HensmanDS17} & fixed & variational & $\mathcal{O}(nm)$ \\
+    Peng et al., 2017 \cite{DBLP:conf/icml/PengZZQ17} & variational & variational & $\mathcal{O}(m^3)$ \\
+    \bottomrule
+  \end{tabular}
+  \caption{Summary of relevant VI approximations to GPs. For both inducing
+    points and hyperparameters, 'fixed' means 'chosen before the algorithm
+    starts', 'modelled' means 'represented by a probability distribution', and
+    'variational' means 'included amongst the variational parameters'.
+    Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of bases used
+    to represent the GP's mean and covariance, respectively.}
+  \label{table:approximations}
+\end{table}
+
+Table \ref{table:approximations} further summarises some of the recent and/or
+influential GP approximation approaches that might be suitable for our GP IRL
+model. In order to derive a reliable ELBO, the inducing points should be either
+fixed or modelled, while hyperparameters cannot be variational (the reason will
+be explained in Section \ref{sec:proposed_approach}).
+
+\begin{itemize}
+\item We would also like to avoid having the hyperparameters fixed, as learning
+  them efficiently would introduce a separate problem. As the approach by Hoang
+  et al. does not seem to be easily extendable to modelled hyperparameters, it
+  is unsuitable to our needs.
+\item (TODO: this will have comments about other papers as well and will be
+  restructured into a paragraph) The variational Fourier features (VFF)
+  algorithm by Hensman et al. \cite{DBLP:journals/jmlr/HensmanDS17} is only
+  defined for Mat\'ern kernels, which is likely to be too restrictive for our
+  situation (the original paper on using GPs for IRL
+  \cite{DBLP:conf/nips/LevinePK11} used the automatic relevance detection kernel
+  that has weights controlling how important each feature is). While extending
+  VFF to support a flexible class of kernels defined by Wilson and Adams
+  \cite{pmlr-v28-wilson13} is an interesting and promising avenue of work, it is
+  likely to be beyond the scope of this project.
+\item The derivation of the ELBO in the work of Peng et al.
+  \cite{DBLP:conf/icml/PengZZQ17} relies primarily on the fact that the evidence
+  for a GP is a Gaussian, while in our case the evidence can be defined in
+  several ways depending on the model, but must always involve $\mathcal{D}$,
+  making the main idea of the paper inapplicable to IRL.
+\end{itemize}
+
+Since we expect $\posterior$ to be highly irregular, we would like our
+approximation to be capable of representing more than just a unimodal Gaussian.
+The primary way to represent complex posteriors in VI is by using
+\emph{normalising flows}, i.e., a collection of invertible
+functions---parametrised by additional variational parameters---that are applied
+to latent variables \cite{DBLP:conf/icml/RezendeM15}. Unfortunately, this
+parametrisation also means that the gradient of the joint probability
+distribution w.r.t. variational parameters $\nabla_{\bm\nu}\pfull$ is no
+longer zero, making an analytic expression for ELBO impossible using the usual
+methods. % TODO: not full (I think) + discuss the possibility of making it
+         % variational
+
+\section{Proposed Approach} \label{sec:proposed_approach}
 
@@
 
-We begin by rewriting the posterior by using the chain rule and Bayes' theorem,
-trying to extract parts of the distribution we know how to compute, namely those
-in Equation \ref{full}:
+In order to properly investigate the difference between variational inference
+and maximum likelihood estimation for the model, we keep other parts of the
+model the same. Namely, we set the covariance function to a version of the
+automatic relevance detection kernel
+\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
+\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \zeta\exp\left(
+    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^T\bm\Lambda(\mathbf{x}_i -
+    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
+where $\zeta$ is the overall ``scale'' factor for how similar or distant the
+states are,
+\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) = \diag(\bm\lambda) \]
+is a diagonal matrix that determines how relevant each feature is (where $d$
+denotes the number of features), $\mathbbm{1}$ is defined as
+\[ \mathbbm{1}[b] = \begin{cases}
+    1 & \mbox{if $b$ is true} \\
+    0 & \mbox{otherwise},
+  \end{cases} \]
+and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
+the value makes little difference to the performance of the algorithm
+\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\Theta = \{ \zeta,
+\bm\lambda \}$. Similarly, we keep the expression for the prior of $\Theta$:
+\[ p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
+    \sum_{i=1}^d \log(\lambda_i + 1) \right). \]
+
+Next, we can rewrite the posterior by using the chain rule and Bayes' theorem in
+order to get a better sense of what we are trying to approximate:
 \[
@@
 $p(\mathcal{D} | \Theta, \mathbf{X_u})$ and $p(\mathcal{D} | \Theta,
-\mathbf{X_u}, \mathbf{u})$. They can be computed as follows:
+\mathbf{X_u}, \mathbf{u})$, which can be computed as follows:
 \begin{align*}
@@
 \end{align*}
+This suggests the following form for the approximation:
+\[ \approximation = q(\Theta) \times q(\mathbf{u} | \Theta) \times q(\mathbf{r}
+  | \Theta, \mathbf{u}). \]
+
+%where $q(\Theta)$ can be of the same form as $p(\Theta | \mathbf{X_u})$,
+%$q(\mathbf{u} | \Theta) = \mathcal{N}(\mathbf{0}, \bm\Sigma_{\mathbf{u}})$, and
+%$q(\mathbf{r} | \Theta, \mathbf{u}) = \mathcal{N}(\bm\mu_{\mathbf{r}},
+%\bm\Sigma_{\mathbf{r}})$.
+
+\subsection{The Structure of the Approximating Distribution}
+
+At this point we are forced to make assumptions about the approximate posterior
+in order to arrive at an implementable solution. Ways to relax the assumptions
+may be investigated towards the end of the project, if time permits.
+
+First, as is common in the literature for applying VI to GPs
+\cite{DBLP:conf/nips/ChengB17,DBLP:conf/uai/HensmanFL13,DBLP:conf/icml/HoangHL15,DBLP:journals/jmlr/Titsias09},
+we simply set
+\[ q(\mathbf{r} | \Theta, \mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u},
+  \mathbf{u}). \]
+We can make a similarly justified choice for $q(\mathbf{u} | \Theta)$:
+\[ q(\mathbf{u} | \Theta) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) \]
+for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and $\mathbf{S}
+\in \mathbb{R}^{m \times m}$
+\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
+
+In order to have a reasonable way of calculating/approximating the ELBO (see
+Section \ref{vi_algs}), we need to have an approximating distribution for
+$\Theta$, but unfortunately all the papers in Table \ref{table:approximations}
+either fix it or treat it as a variational parameter. Hence we make our first
+assumption without justification from previous literature:
+\[ q(\Theta) = q(\zeta) \times q(\bm\lambda). \]
+As it is reasonable for $\bm\lambda$ to take any value in $\mathbb{R}^d$, we can
+approximate it simply as
+\[ q(\bm\lambda) = \mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma) \]
+for some variational parameters $\bm\mu \in \mathbb{R}^d$ and $\bm\Sigma \in
+\mathbb{R}^{d \times d}$. In contrast, as the covariance function must produce
+non-negative values, we want to restrict $\zeta$ to positive real numbers. At
+the same time, the distribution for $\zeta$ should have a flexible mean (i.e.,
+not be tied to zero, like in the exponential distribution) and tails that
+converge to zero as the value of the random variable moves away from the mean.
+We might want to support some right skew, but the distribution should be close
+to symmetric with at least some parameter values. This limits our choice of
+distributions quite significantly, and we decide to go with the gamma
+distribution as it is fairly flexible and commonly used
+\cite{hogg2018introduction}. We then define the probability density function of
+the gamma distribution as
+\[ q(\zeta) = \Gamma(\zeta; \alpha, \beta) =
+  \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta}, \]
+where $\alpha > 0$ and $\beta > 0$ are two parameters of the distribution, and
+$\Gamma(\cdot)$ is the gamma function. This gives us our list of variational
+parameters: $\mathbf{m}, \mathbf{S}, \bm\mu, \bm\Sigma, \alpha, \beta$.
+
+\subsection{Evidence Lower Bound}
+
+In this section we derive and simplify the ELBO for this (now fully specified)
+model. We begin with a lemma that will be useful during the derivation.
+
+\begin{lemma}
+  Let $\mathbf{x} \sim \mathcal{N}(\bm\mu, \bm\Sigma)$ be a random Gaussian
+  vector with $n$ elements, and let $f$ be its probability density function.
+  Then
+  \[ \mathbb{E}[\log f(\mathbf{x})] = -\frac{1}{2}\log |\bm\Sigma| + c, \]
+  where $c \in \mathbb{R}$ is a constant w.r.t. $\bm\mu$ and $\bm\Sigma$.
+\end{lemma}
+
+\begin{proof} % TODO: cite the matrix cookbook
+  Similarly to \eqref{eq:normal}, we can write
+  \[ \mathbb{E}[\log f(\mathbf{x})] = \mathbb{E}\left[ -\frac{1}{2}(\mathbf{x} -
+      \bm\mu)^T\bm\Sigma^{-1}(\mathbf{x} - \bm\mu) - \frac{1}{2}\log |\bm\Sigma|
+      - \frac{n}{2}\log 2\pi \right]. \]
+  Note that the term
+  \[ \mathbb{E}\left[ -\frac{n}{2}\log 2\pi \right] = -\frac{n}{2}\log 2\pi \]
+  can be absorbed into $c$. Now
+  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^T\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
+    \Tr(\bm\Sigma^{-1}\mathbf{M}) + \mathbf{c}^T\bm\Sigma^{-1}\mathbf{c}, \]
+  where $\mathbf{M} = \Var[\mathbf{x} - \bm\mu]$ and $\mathbf{c} =
+  \mathbb{E}[\mathbf{x} - \bm\mu]$ \cite{petersen2008matrix}. But
+  \[ \mathbf{c} = \mathbb{E}[\mathbf{x} - \bm\mu] = \mathbb{E}[\mathbf{x}] -
+    \bm\mu = \bm\mu - \bm\mu = \mathbf{0}, \]
+  so
+  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^T\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
+    \Tr(\bm\Sigma^{-1}\mathbf{M}) = \Tr(\mathbf{I}), \]
+  since
+  \[ \mathbf{M} = \Var[\mathbf{x} - \bm\mu] = \Var[\mathbf{x}] = \bm\Sigma. \]
+  Thus, since we have that
+  \[ \mathbb{E}\left[ -\frac{1}{2}\log |\bm\Sigma| \right] = -\frac{1}{2}\log
+    |\bm\Sigma|, \]
+  we get that
+  \[ \mathbb{E}[\log f(\mathbf{x})] = -\frac{1}{2}\log |\bm\Sigma| + c \]
+  for some $c \in \mathbb{R}$ that depends on $n$, but not on $\bm\mu$ or
+  $\bm\Sigma$.
+\end{proof}
+
+\subsection{Variational Inference Algorithms} \label{vi_algs}
+
+% TODO: will need to make this equation more precise, once I know what
+% approximation I want to use
+The typical way to optimise a quantity (the ELBO, in this case) involves
+computing its gradient \cite{blei2017variational}. Due to terms involving
+$\mathcal{D}$ that are computed by solving an MDP and do not resemble a typical
+probability distribution, we turn our attention to the \emph{black box
+  variational inference} \cite{DBLP:conf/aistats/RanganathGB14} paper that
+suggests a convenient trick:
+\[ \nabla_{\bm\nu} \mathcal{L} = \Eq[\nabla_{\bm\nu} \log \approximation
+  (\log \pfull - \log \approximation)]. \]
+With this trick in mind, we only need to evaluate $\log \pfull$ and take the
+gradient of $\log \approximation$. Following the same paper,
+$\nabla_{\bm\nu} \mathcal{L}$ then has an unbiased estimate
+\[ \nabla_{\bm\nu} \mathcal{L} \approx \frac{1}{S} \sum_{s=1}^S
+  \nabla_{\bm\nu} \log \approximationS (\log \pfullS - \log
+  \approximationS) \]
+computed by drawing $S$ Monte Carlo samples $(\Theta_s, \mathbf{u}_s,
+\mathbf{r}_s) \sim \approximation$.
 
@@
 Simple introduction to GPs for time-series modelling \cite{Roberts2013GaussianPF}
-
-Spectral kernels \cite{pmlr-v28-wilson13}
 
@@
 
-Normalizing flows \cite{DBLP:conf/icml/RezendeM15}
+Stochastic VI \cite{DBLP:journals/jmlr/HoffmanBWP13}
+
+Structured stochastic VI \cite{DBLP:conf/aistats/HoffmanB15}
+
+Another review of recent advances \cite{DBLP:journals/corr/abs-1711-05597}
+
+Tighter ELBOs are not necessarily better \cite{DBLP:conf/icml/RainforthKLMIWT18}
+
+For details on Lebesgue's dominated convergence theorem \cite{cinlar_probability}
+
+Still looking for the relevant version of the theorem \cite{knill2017probability}
+
+Approximation as a multivariate Gaussian (haven't read) \cite{DBLP:journals/sac/TanN18}
+
+How black box VI can be even more black box \cite{li2016wild}
+
+Evaluating VI \cite{DBLP:conf/icml/YaoVSG18}
+
+\subsubsection{for GPs}
 
@@
 
-Stochastic VI \cite{DBLP:journals/jmlr/HoffmanBWP13}
-
-Structured stochastic VI (haven't read) \cite{DBLP:conf/aistats/HoffmanB15}
-
-Another review of recent advances \cite{DBLP:journals/corr/abs-1711-05597}
-
 Sparse VI for GP \cite{DBLP:journals/jmlr/HensmanDS17}
 
-Sparse GPs \cite{DBLP:journals/jmlr/CandelaR05}.
+Stochastic VI for sparse spectral GPs (no inducing points, so not that relevant) \cite{DBLP:conf/aaai/HoangHL17}
 
-Tighter ELBOs are not necessarily better \cite{DBLP:conf/icml/RainforthKLMIWT18}.
+Sparse GPs 2 \cite{DBLP:conf/nips/GalWR14}
 
-\cite{DBLP:books/lib/Bishop07}
+The 3 papers I need to focus on:
+
+SVI for sparse GPs \cite{DBLP:conf/uai/HensmanFL13}
+
+distributed 1 \cite{DBLP:conf/icml/HoangHL15}
+
+distributed 2 (haven't read) \cite{DBLP:conf/icml/PengZZQ17}
+
+The paper at the core of the 3 approaches (approximates posterior) \cite{DBLP:journals/jmlr/Titsias09}
+
+Generalized version? \cite{DBLP:conf/icml/ShethWK15}
 

### 2018-10-26 mprop/mprop.tex
Small changes

@@
 
-% TODO: table of notation?
 \newcommand{\Kuu}{\mathbf{K}_{\mathbf{u},\mathbf{u}}}
@@
 \begin{equation} \label{pDr}
-  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q(s_{i,t}, a_{i,t}; r) - V(s_{i,t}; r) \right),
+  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right),
 \end{equation}
 where
-\[ Q(s_{i,t}, a_{i,t}; r) = r(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}}
-  \mathcal{T}(s_{i,t}, a_{i,t}, s')V(s'; r), \]
-and $V(s; r)$ can be obtained by repeatedly applying the equation
+\begin{equation}
+  Q_r(s_{i,t}, a_{i,t}) = r(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s_{i,t}, a_{i,t}, s')V_r(s'),
+\end{equation}
+and $V_r(s)$ can be obtained by repeatedly applying the equation
 \cite{supplementary_material}
-\[ V(s; r) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
-      \mathcal{S}} \mathcal{T}(s, a, s')V(s'; r) \right). \]
+\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
+      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
 However, a reward function learned by maximising this likelihood is not
@@
 
-% TODO: import info from the MSci paper
+% TODO: positive definite
 As training a GP with $n$ data points has a time complexity of
@@
   p(\mathbf{u} | \Theta, \mathbf{X_u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \nonumber \\
-                                       &= \frac{1}{(2\pi)^{n/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^T\Kuu^{-1}\mathbf{u} \right) \nonumber \\
-                                       &= \exp\left(-\frac{1}{2}\mathbf{u}^T\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right) \label{eq:normal}
-  \end{align} % TODO: introduce m here, deleting later introduction
-is the GP prior \cite{DBLP:books/lib/RasmussenW06}, the GP posterior is a
+                                       &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \nonumber \\
+                                       &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right) \label{eq:normal}
+\end{align}
+is the GP prior \cite{DBLP:books/lib/RasmussenW06}, where $m \in \mathbb{N}$ is
+the number of inducing points. The GP posterior is a
 multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11}
@@
   p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) =
-  \mathcal{N}(\mathbf{r}; \Kru^T\Kuu^{-1}\mathbf{u}, \Krr - \Kru^T\Kuu^{-1}\Kru),
+  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
 \end{equation}
@@
 \cite{DBLP:books/lib/Bishop07,blei2017variational}
-\[
+\begin{equation} \label{eq:elbo}
   \begin{split}
@@
   \end{split}
-\]
+\end{equation}
 
@@
 modelling accuracy close to that of the full GP with no approximations, while
-providing a time complexity of $\mathcal{O}(m^3)$, where $m$ is the number of
-inducing points. Below we provide a short overview of various assumptions that
-have been used in approximating \emph{sparse} GPs (i.e., GPs that use inducing
-points), following on a paper by Qui\~{n}onero-Candela and Rasmussen
-\cite{DBLP:journals/jmlr/CandelaR05}.
+providing a time complexity of $\mathcal{O}(m^3)$. Below we provide a short
+overview of various assumptions that have been used in approximating
+\emph{sparse} GPs (i.e., GPs that use inducing points), following on a paper by
+Qui\~{n}onero-Candela and Rasmussen \cite{DBLP:journals/jmlr/CandelaR05}.
 
@@
 \begin{table}
-  \centering % TODO: recheck and explain
-  \begin{tabular}{llllll} % Lower bound? Model? % TODO: comment about 'modelled' no longer needed
+  \centering % TODO: recheck
+  \begin{tabular}{llllll}
     \toprule
@@
     points and hyperparameters, 'fixed' means 'chosen before the algorithm
-    starts', 'modelled' means 'represented by a probability distribution', and
-    'variational' means 'included amongst the variational parameters'.
-    Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of bases used
-    to represent the GP's mean and covariance, respectively.}
+    starts' and 'variational' means 'included amongst the variational
+    parameters'. Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of
+    bases used to represent the GP's mean and covariance, respectively.}
   \label{table:approximations}
@@
 
+% TODO: expand on this
 Table \ref{table:approximations} further summarises some of the recent and/or
@@
 model. In order to derive a reliable ELBO, the inducing points should be either
-fixed or modelled, while hyperparameters cannot be variational (the reason will
-be explained in Section \ref{sec:proposed_approach}).
+fixed or modelled by a probability distribution (none of the approaches do
+that), while hyperparameters cannot be variational (the reason will be explained
+in Section \ref{sec:proposed_approach}).
 
@@
 Since we expect $\posterior$ to be highly irregular, we would like our
-approximation to be capable of representing more than just a unimodal Gaussian.
-The primary way to represent complex posteriors in VI is by using
+approximation to be capable of representing a wide range of possible probability
+distributions. The primary way to represent complex posteriors in VI is by using
 \emph{normalising flows}, i.e., a collection of invertible
@@
 longer zero, making an analytic expression for ELBO impossible using the usual
-methods. % TODO: not full (I think) + discuss the possibility of making it
-         % variational
+methods. While it may be possible to have an approximating distribution for the
+flow parameters, it is uncertain how such an algorithm would behave, as it
+would have to perform optimisation in a space with significantly more dimensions.
 
@@
 \[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \zeta\exp\left(
-    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^T\bm\Lambda(\mathbf{x}_i -
+    -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
     \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
@@
 \bm\lambda \}$. Similarly, we keep the expression for the prior of $\Theta$:
-\[ p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
-    \sum_{i=1}^d \log(\lambda_i + 1) \right). \]
+\begin{equation} \label{eq:theta_prior}
+  p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
+    \sum_{i=1}^d \log(\lambda_i + 1) \right).
+\end{equation}
 
@@
 This suggests the following form for the approximation:
-\[ \approximation = q(\Theta) \times q(\mathbf{u} | \Theta) \times q(\mathbf{r}
-  | \Theta, \mathbf{u}). \]
+\begin{equation} \label{eq:approximation}
+  \approximation = q(\Theta) \times q(\mathbf{u} | \Theta) \times q(\mathbf{r}
+  | \Theta, \mathbf{u}).
+\end{equation}
 
@@
 we simply set
-\[ q(\mathbf{r} | \Theta, \mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u},
-  \mathbf{u}). \]
+\begin{equation}
+  q(\mathbf{r} | \Theta, \mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u},
+  \mathbf{u}).
+\end{equation}
 We can make a similarly justified choice for $q(\mathbf{u} | \Theta)$:
-\[ q(\mathbf{u} | \Theta) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) \]
+\begin{equation}
+  q(\mathbf{u} | \Theta) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
+\end{equation}
 for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and $\mathbf{S}
@@
 assumption without justification from previous literature:
-\[ q(\Theta) = q(\zeta) \times q(\bm\lambda). \]
+\begin{equation} \label{eq:theta}
+  q(\Theta) = q(\zeta) \times q(\bm\lambda).
+\end{equation}
 As it is reasonable for $\bm\lambda$ to take any value in $\mathbb{R}^d$, we can
 approximate it simply as
-\[ q(\bm\lambda) = \mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma) \]
+\begin{equation} \label{eq:lambda}
+  q(\bm\lambda) = \mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma)
+\end{equation}
 for some variational parameters $\bm\mu \in \mathbb{R}^d$ and $\bm\Sigma \in
@@
 the gamma distribution as
-\[ q(\zeta) = \Gamma(\zeta; \alpha, \beta) =
-  \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta}, \]
+\begin{equation} \label{eq:zeta}
+  q(\zeta) = \Gamma(\zeta; \alpha, \beta) =
+  \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta},
+\end{equation}
 where $\alpha > 0$ and $\beta > 0$ are two parameters of the distribution, and
-$\Gamma(\cdot)$ is the gamma function. This gives us our list of variational
-parameters: $\mathbf{m}, \mathbf{S}, \bm\mu, \bm\Sigma, \alpha, \beta$.
+$\Gamma(\cdot)$ is the gamma function. This gives us our vector of variational
+parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \bm\mu, \bm\Sigma, \alpha,
+\beta)^\intercal$.
 
@@
 
-\begin{lemma}
+% TODO: this is actually the entropy of the multivariate Gaussian
+\begin{lemma} \label{expectation_lemma}
   Let $\mathbf{x} \sim \mathcal{N}(\bm\mu, \bm\Sigma)$ be a random Gaussian
@@
 
-\begin{proof} % TODO: cite the matrix cookbook
+\begin{proof}
   Similarly to \eqref{eq:normal}, we can write
   \[ \mathbb{E}[\log f(\mathbf{x})] = \mathbb{E}\left[ -\frac{1}{2}(\mathbf{x} -
-      \bm\mu)^T\bm\Sigma^{-1}(\mathbf{x} - \bm\mu) - \frac{1}{2}\log |\bm\Sigma|
+      \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{x} - \bm\mu) - \frac{1}{2}\log |\bm\Sigma|
       - \frac{n}{2}\log 2\pi \right]. \]
@@
   can be absorbed into $c$. Now
-  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^T\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
-    \Tr(\bm\Sigma^{-1}\mathbf{M}) + \mathbf{c}^T\bm\Sigma^{-1}\mathbf{c}, \]
+  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
+    \Tr(\bm\Sigma^{-1}\mathbf{M}) + \mathbf{c}^\intercal\bm\Sigma^{-1}\mathbf{c}, \]
   where $\mathbf{M} = \Var[\mathbf{x} - \bm\mu]$ and $\mathbf{c} =
@@
   so
-  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^T\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
+  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
     \Tr(\bm\Sigma^{-1}\mathbf{M}) = \Tr(\mathbf{I}), \]
@@
 \end{proof}
+
+In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
+write\footnote{At this point, we will drop the subscript denoting which
+  variables the expectation is taken over. Also note that during the derivation
+  equality is taken to mean ``equality up to an additive constant''.}
+\[ \mathcal{L}(\bm\nu) = \mathbb{E}[\log\pfull] -
+  \mathbb{E}[\log\approximation]. \]
+By plugging in \eqref{full} and \eqref{eq:approximation}, we get
+\[
+  \begin{split}
+    \mathcal{L}(\bm\nu) &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\Theta |
+    \mathbf{X_u}) + \log p(\mathbf{u} | \Theta, \mathbf{X_u}) + \log
+    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) + \log p(\mathcal{D} | r)] \\
+    &- \mathbb{E}[\log q(\Theta) + \log q(\mathbf{u}) + \log q(\mathbf{r} |
+    \Theta, \mathbf{u})].
+  \end{split}
+\]
+Note that $\mathbb{E}[\log p(\mathbf{X_u})]$ is just a constant, so we can
+simply drop it from the expression. Furthermore, since $q(\mathbf{r} | \Theta,
+\mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u})$, they cancel
+each other out. Then we can substitute various terms with their definitions to
+get
+\begin{alignat*}{3}
+  \mathcal{L}(\bm\nu) &= \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] -
+    \sum_{i=1}^d \log(\lambda_i + 1)\right] &&+ \mathbb{E}[\log
+  \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)] \\
+  &+ \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) -
+    V_r(s_{i,t}) \right] &&-
+  \mathbb{E}\left[\log\left(\frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha -
+        1}e^{-\beta\zeta}\right)\right] \\
+  &- \mathbb{E}[\log\mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma)] &&-
+  \mathbb{E}[\log\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})].
+\end{alignat*}
+Next, we can plug in the definitions of $\mathcal{N}(\mathbf{u}; \mathbf{0},
+\Kuu)$ and $Q_r$, distribute the log across the gamma distribution, and apply
+Lemma \ref{expectation_lemma} to the last two terms\footnote{We cannot
+  apply the lemma to $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0},
+  \Kuu)]$ because $\Kuu$ depends on $\Theta$.} to get
+\begin{align*}
+  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] -
+    \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
+  &+ \mathbb{E}\left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
+    \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right] \\
+  &+ \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - V_r(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}}
+    \mathcal{T}(s_{i,t}, a_{i,t}, s')V_r(s') \right] \\
+  &- \mathbb{E}[\alpha\log\beta - \log\Gamma(\alpha) + (\alpha - 1)\log\zeta - \beta\zeta].
+\end{align*}
+Now we can remove $\mathbb{E}\left[-\frac{m}{2}\log2\pi\right]$ since it is
+constant w.r.t. both the variational parameters and the variables the
+expectation is over, and move constants (or variational parameters) independent
+of the approximated variables outside of the expectations. Also note that
+$\mathbb{E}[\zeta] = \alpha/\beta$ and
+$\mathbb{E}[\log \zeta] = \psi(\alpha) - \log \beta$, where $\psi$ is the
+digamma function defined as $\psi(x) =
+\frac{d}{dx}\log\Gamma(x)$\cite{DBLP:books/lib/Bishop07}. This allows us to
+simplify the expression to the following:
+\begin{align*}
+  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2} \mathbb{E}[\Tr[\Kuu^{-2}]] -
+    \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] \\
+  &- \frac{1}{2}\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \alpha - \log\beta + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha) \\
+  &+ \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] - \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')]. \\
+\end{align*}
+
+\subsection{Derivatives}
 

### 2018-10-27 mprop/mprop.tex
Lots of derivatives

@@
 \newcommand{\Kru}{\mathbf{K}_{\mathbf{r},\mathbf{u}}}
+
 \newcommand{\DKL}{D_{\mathrm{KL}}}
+
 \newcommand{\pfull}{p(\mathcal{D}, \Theta, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
@@
 \newcommand{\approximationS}{q_{\bm\nu}(\Theta_s, \mathbf{u}_s, \mathbf{r}_s)}
+
 \newcommand{\Eq}{\mathbb{E}_{(\Theta, \mathbf{u}, \mathbf{r}) \sim \approximation}}
+
+\newcommand{\dm}{\frac{\partial}{\partial\mathbf{m}}}
+\newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
+\newcommand{\da}{\frac{\partial}{\partial\alpha}}
+\newcommand{\db}{\frac{\partial}{\partial\beta}}
 
@@
 %briefly explain the context of the project problem
+% TODO: introduce a data set
+% TODO: example application for motivation (why would the world be better)
+% TODO: (possibly) a figure with my notation
+% TODO: explain the MDP Q/V stuff
+% TODO: introduce research questions
 
@@
 \end{equation}
-Here $p(\mathbf{X_u})$ and $p(\Theta | \mathbf{X_u})$ are freely chosen priors,
+Here $p(\mathbf{X_u})$ and $p(\Theta | \mathbf{X_u})$ are customisable priors,
 \begin{align}
@@
 
+% TODO: where does the lower bound come from?
 Given this model, data $\mathcal{D}$, and inducing feature matrix
@@
 %the previous work.
+
+% TODO: review IRL papers, their weaknesses
 
@@
 At this point we are forced to make assumptions about the approximate posterior
-in order to arrive at an implementable solution. Ways to relax the assumptions
-may be investigated towards the end of the project, if time permits.
+in order to arrive at an implementable solution. Time permitting, ways to relax
+the assumptions may be investigated towards the end of the project.
 
@@
 \end{equation}
-for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and $\mathbf{S}
-\in \mathbb{R}^{m \times m}$
+for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and a positive
+semi-definite matrix $\mathbf{S} \in \mathbb{R}^{m \times m}$.
 \cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
 
+% TODO: uncorrelated gammas instead of a multivariate Gaussian? review several options
 In order to have a reasonable way of calculating/approximating the ELBO (see
@@
 \subsection{Evidence Lower Bound}
+% TODO: remove/add niu from/to q everywhere
 
 In this section we derive and simplify the ELBO for this (now fully specified)
-model. We begin with a lemma that will be useful during the derivation.
-
-% TODO: this is actually the entropy of the multivariate Gaussian
-\begin{lemma} \label{expectation_lemma}
-  Let $\mathbf{x} \sim \mathcal{N}(\bm\mu, \bm\Sigma)$ be a random Gaussian
-  vector with $n$ elements, and let $f$ be its probability density function.
-  Then
-  \[ \mathbb{E}[\log f(\mathbf{x})] = -\frac{1}{2}\log |\bm\Sigma| + c, \]
-  where $c \in \mathbb{R}$ is a constant w.r.t. $\bm\mu$ and $\bm\Sigma$.
-\end{lemma}
-
-\begin{proof}
-  Similarly to \eqref{eq:normal}, we can write
-  \[ \mathbb{E}[\log f(\mathbf{x})] = \mathbb{E}\left[ -\frac{1}{2}(\mathbf{x} -
-      \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{x} - \bm\mu) - \frac{1}{2}\log |\bm\Sigma|
-      - \frac{n}{2}\log 2\pi \right]. \]
-  Note that the term
-  \[ \mathbb{E}\left[ -\frac{n}{2}\log 2\pi \right] = -\frac{n}{2}\log 2\pi \]
-  can be absorbed into $c$. Now
-  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
-    \Tr(\bm\Sigma^{-1}\mathbf{M}) + \mathbf{c}^\intercal\bm\Sigma^{-1}\mathbf{c}, \]
-  where $\mathbf{M} = \Var[\mathbf{x} - \bm\mu]$ and $\mathbf{c} =
-  \mathbb{E}[\mathbf{x} - \bm\mu]$ \cite{petersen2008matrix}. But
-  \[ \mathbf{c} = \mathbb{E}[\mathbf{x} - \bm\mu] = \mathbb{E}[\mathbf{x}] -
-    \bm\mu = \bm\mu - \bm\mu = \mathbf{0}, \]
-  so
-  \[ \mathbb{E}[(\mathbf{x} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{x} - \bm\mu)] =
-    \Tr(\bm\Sigma^{-1}\mathbf{M}) = \Tr(\mathbf{I}), \]
-  since
-  \[ \mathbf{M} = \Var[\mathbf{x} - \bm\mu] = \Var[\mathbf{x}] = \bm\Sigma. \]
-  Thus, since we have that
-  \[ \mathbb{E}\left[ -\frac{1}{2}\log |\bm\Sigma| \right] = -\frac{1}{2}\log
-    |\bm\Sigma|, \]
-  we get that
-  \[ \mathbb{E}[\log f(\mathbf{x})] = -\frac{1}{2}\log |\bm\Sigma| + c \]
-  for some $c \in \mathbb{R}$ that depends on $n$, but not on $\bm\mu$ or
-  $\bm\Sigma$.
-\end{proof}
-
-In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
+model. In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
 write\footnote{At this point, we will drop the subscript denoting which
-  variables the expectation is taken over. Also note that during the derivation
+  variables the expectation is taken over. Also note that throughout the derivation
   equality is taken to mean ``equality up to an additive constant''.}
@@
 \end{alignat*}
-Next, we can plug in the definitions of $\mathcal{N}(\mathbf{u}; \mathbf{0},
-\Kuu)$ and $Q_r$, distribute the log across the gamma distribution, and apply
-Lemma \ref{expectation_lemma} to the last two terms\footnote{We cannot
-  apply the lemma to $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0},
-  \Kuu)]$ because $\Kuu$ depends on $\Theta$.} to get
+For the last two terms we can use the fact that, e.g., $-\mathbb{E}[\log
+\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})] = \frac{1}{2}\log|\mathbf{S}|$
+up to an additive constant that depends on the number of dimensions of the
+distribution \cite{DBLP:journals/tit/AhmedG89}. Note that we cannot use this
+fact for $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)]$  because
+$\Kuu$ depends on $\Theta$. We also plug in the definitions of
+$\mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)$ and $Q_r$, and get:
 \begin{align*}
@@
 $\mathbb{E}[\log \zeta] = \psi(\alpha) - \log \beta$, where $\psi$ is the
-digamma function defined as $\psi(x) =
-\frac{d}{dx}\log\Gamma(x)$\cite{DBLP:books/lib/Bishop07}. This allows us to
+digamma function defined as $\psi(x) = \frac{d}{dx}\log\Gamma(x)$
+\cite{DBLP:books/lib/Bishop07}. This allows us to
 simplify the expression to the following:
@@
 \subsection{Derivatives}
+% TODO: integrals always mean "on all range"
+% TODO: some kind of intro
+
+\subsubsection{$\partial/\partial\mathbf{m}$}
+
+We begin by removing terms independent of $\mathbf{m}$:
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
+    &-\frac{1}{2}\dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] \\
+    &+ \dm\left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
+      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
+      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')]\right],
+  \end{split}
+\]
+where
+\begin{equation} \label{eq:1}
+\begin{split} % TODO: dominated convergence theorem
+    \dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
+    \dm\iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\mathcal{N}(\mathbf{u};
+    \mathbf{m},
+    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
+    &= \iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{m}} q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda.
+  \end{split}
+\end{equation}
+$\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{m}}$ can be calculated as
+follows: % TODO: by the cookbook
+\[
+  \begin{split}
+    \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial\mathbf{m}} &=
+    \mathcal{N}(\mathbf{u}; \mathbf{m},
+    \mathbf{S})\dm\left[-\frac{1}{2}(\mathbf{u} -
+      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right] \\
+    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
+    \mathbf{S})\left(-\frac{1}{2}\right)(\mathbf{S}^{-1} +
+    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\dm[\mathbf{u} -
+    \mathbf{m}] \\
+    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
+    \mathbf{S})\frac{1}{2}(\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u}
+    - \mathbf{m}).
+  \end{split}
+\]
+Plugging it back to \eqref{eq:1} gives:
+\[
+  \begin{split}
+    \dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
+    \iiint(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})\frac{1}{2}(\mathbf{S}^{-1} +
+    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\mathcal{N}(\mathbf{u};
+    \mathbf{m},
+    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
+    &=
+    \frac{1}{2}\mathbb{E}[(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})(\mathbf{S}^{-1}
+    + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})]
+  \end{split}
+\]
+
+\subsubsection{$\partial/\partial\mathbf{S}$}
+
+Similarly to the previous section,
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
+    \frac{1}{2}\dS[\log|\mathbf{S}|]
+    -\frac{1}{2}\dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] \\
+    &+ \dS\left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
+      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
+      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')]\right],
+  \end{split}
+\]
+where % TODO: cite cookbook
+\[ \dS[\log|\mathbf{S}|] =
+  \frac{1}{|\mathbf{S}|}\frac{\partial|\mathbf{S}|}{\partial \mathbf{S}} =
+  \frac{1}{|\mathbf{S}|}|\mathbf{S}|\mathbf{S}^{-\intercal} =
+  \mathbf{S}^{-\intercal} \]
+and
+\[
+  \begin{split}
+    \dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
+    \dS\iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\mathcal{N}(\mathbf{u};
+    \mathbf{m},
+    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
+    &=
+    \iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\frac{\partial \mathcal{N}(\mathbf{u};
+      \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}}q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda.
+  \end{split}
+\]
+$\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}}$ can be simplified to
+\[
+  \begin{split}
+    \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}} &=
+    \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\exp \left( -\frac{1}{2}
+        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
+        \mathbf{m})\right)\right] \\
+    &= \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\right]\exp \left( -\frac{1}{2}
+        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
+        \mathbf{m})\right) \\
+      &+ \frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\dS\left[\exp\left( -\frac{1}{2}
+        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
+        \mathbf{m})\right)\right] \\
+    &=
+    \frac{1}{(2\pi)^{m/2}}\dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right]\exp
+    \left( -\frac{1}{2} (\mathbf{u} -
+      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right) \\
+    &- \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})\dS[(\mathbf{u}
+    - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})], \\
+  \end{split}
+\]
+where
+\[ \dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right] =
+  -\frac{1}{2}|\mathbf{S}|^{-3/2}\frac{\partial |\mathbf{S}|}{\partial \mathbf{S}} =
+  -\frac{1}{2}|\mathbf{S}|^{-3/2}|\mathbf{S}|\mathbf{S}^{-\intercal} =
+  -\frac{1}{2|\mathbf{S}|^{1/2}}\mathbf{S}^{-\intercal} \]
+and
+\[ \dS[(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
+  \mathbf{m})] = -\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
+  \mathbf{m})^\intercal\mathbf{S}^{-\intercal} \]
+according to \textit{The Matrix Cookbook} \cite{petersen2008matrix}. Plugging
+the two expression back in gives
+\[ \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial
+    \mathbf{S}} = -\frac{1}{2}\mathbf{S}^{-\intercal}\mathcal{N}(\mathbf{u};
+  \mathbf{m}, \mathbf{S}) + \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m},
+  \mathbf{S})\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
+  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}. \]
+Thus
+\[ \dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
+  \frac{1}{2}\mathbb{E}[(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})(\mathbf{S}^{-\intercal}(\mathbf{u}
+  - \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
+  \mathbf{S}^{-\intercal})]. \]
+
+\subsubsection{$\partial/\partial\alpha$}
+
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\alpha} = &-
+    \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-2}]] -
+    \frac{1}{2}\da\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] -
+    \frac{1}{2}\da\mathbb{E}[\log|\Kuu|] \\
+    &+ \da [\alpha + \log\Gamma(\alpha) + (1 -
+      \alpha)\psi(\alpha)] \\
+    &+ \da \left[ \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
+      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
+      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \right].
+  \end{split}
+\]
+The first three terms can all be treated in the same way, since all three
+expectations contain expressions without $\alpha$, but $\alpha$ appears in
+$q(\zeta)$. Thus we can work with an abstract function as follows:
+\[
+  \begin{split}
+    \da\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \da \iiint f(k_\Theta, \mathbf{u})
+    \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} e^{-\beta\zeta}
+    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
+    &= \iiint f(k_\Theta, \mathbf{u})
+    \da \left[ \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} \right]
+      e^{-\beta\zeta}
+    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u}.
+  \end{split}
+\]
+Then
+\[
+  \begin{split}
+    \da \left[ \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} \right] &=
+    \frac{\da[\beta^\alpha\zeta^{\alpha - 1}]\Gamma(\alpha) -
+      \beta^\alpha\zeta^{\alpha - 1}\Gamma'(\alpha)}{(\Gamma(\alpha))^2} \\
+    &= \frac{\beta^\alpha\zeta^{\alpha - 1}\da[\alpha\log\beta + (\alpha -
+      1)\log\zeta]\Gamma(\alpha) - \beta^\alpha\zeta^{\alpha -
+        1}\Gamma'(\alpha)}{(\Gamma(\alpha))^2} \\
+    &= \frac{\beta^\alpha\zeta^{\alpha - 1}(\log\beta + \log\zeta)\Gamma(\alpha)
+      - \beta^\alpha\zeta^{\alpha - 1}\Gamma'(\alpha)}{(\Gamma(\alpha))^2} \\
+    &= \frac{\beta^\alpha\zeta^{\alpha - 1}}{\Gamma(\alpha)}\left(\log\beta +
+    \log\zeta - \frac{\Gamma'(\alpha)}{\Gamma(\alpha)}\right),
+  \end{split}
+\]
+which means that
+\[
+  \begin{split}
+    \da\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \iiint f(k_\Theta, \mathbf{u})
+    \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta} \left(
+      \log\beta + \log\zeta - \frac{\Gamma'(\alpha)}{\Gamma(\alpha)} \right)
+    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
+    &= \iiint f(k_\Theta, \mathbf{u}) \left(\log\beta + \log\zeta -
+      \frac{\Gamma'(\alpha)}{\Gamma(\alpha)} \right) q(\zeta) q(\bm\lambda)
+    q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
+    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{u}) \left(\log\beta + \log\zeta -
+        \frac{\Gamma'(\alpha)}{\Gamma(\alpha)} \right) \right] \\
+    &= \left( \log\beta - \frac{\Gamma'(\alpha)}{\Gamma(\alpha)}
+    \right)\mathbb{E}[f(k_\Theta, \mathbf{u})] + \mathbb{E}[f(k_\Theta,
+    \mathbf{u})\log\zeta].
+  \end{split}
+\]
+Next,
+\[ \da [\alpha + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha)] =
+  1 + \psi(\alpha) - \psi(\alpha) + (1 - \alpha)\psi'(\alpha) = 1 + (1 -
+  \alpha)\psi'(\alpha). \]
+
+\subsubsection{$\partial/\partial\beta$}
+
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\beta} = &-
+    \frac{1}{2}\db\mathbb{E}[\Tr[\Kuu^{-2}]] -
+    \frac{1}{2}\db\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] -
+    \frac{1}{2}\db\mathbb{E}[\log|\Kuu|] - \db[\log\beta] \\
+    &+ \db \left[ \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
+      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
+      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \right].
+  \end{split}
+\]
+Similarly to the previous section, we can deal with three cases at once:
+\[
+  \begin{split}
+    \db\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \db \iiint f(k_\Theta, \mathbf{u})
+    \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} e^{-\beta\zeta}
+    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
+    &= \iiint f(k_\Theta, \mathbf{u}) \frac{\zeta^{\alpha - 1}}{\Gamma(\alpha)}
+    \db [\beta^\alpha e^{-\beta\zeta}] q(\bm\lambda)
+    q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u}.
+  \end{split}
+\]
+Since
+\[ \db [\beta^\alpha e^{-\beta\zeta}] = \alpha\beta^{\alpha - 1}e^{-\beta\zeta}
+  - \beta^\alpha e^{-\beta\zeta}\beta = \beta^\alpha
+  e^{-\beta\zeta} \left( \frac{\alpha}{\beta} - \beta \right), \]
+we have that
+\[
+  \begin{split}
+    \db\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \iiint f(k_\Theta, \mathbf{u})
+    \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta} \left(
+      \frac{\alpha}{\beta} - \beta \right)  q(\bm\lambda)
+    q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
+    &= \iiint f(k_\Theta, \mathbf{u}) \left(\frac{\alpha}{\beta} - \beta \right)
+    q(\zeta)q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
+    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{u}) \left( \frac{\alpha}{\beta} -
+        \beta \right) \right] =  \left( \frac{\alpha}{\beta} - \beta \right)
+    \mathbb{E} [f(k_\Theta, \mathbf{u})].
+  \end{split}
+\]
 
@@
 
+% TODO: integrate this into the previous subsection
 % TODO: will need to make this equation more precise, once I know what

### 2018-10-28 mprop/mprop.tex
Just before I rewrite everything

@@
 \begin{align*}
-  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] -
-    \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
+  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
   &+ \mathbb{E}\left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
@@
 digamma function defined as $\psi(x) = \frac{d}{dx}\log\Gamma(x)$
-\cite{DBLP:books/lib/Bishop07}. This allows us to
-simplify the expression to the following:
+\cite{DBLP:books/lib/Bishop07}. Finally, we can simplify
+$\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by choosing to represent all visited states
+in $\mathbf{r} = (r_1, \dots, r_k)^\intercal$, and defining a new vector
+$\mathbf{t} = (t_1, \dots, t_k)^\intercal$, where $t_i$ is the number of times
+the state associated with the reward $r_i$ has been visited across all
+demonstrations. Then
+\begin{align*}
+\mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &=
+    \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] =
+    \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] =
+    \mathbf{t}^\intercal\mathbb{E}_{(\Theta, \mathbf{u}) \sim
+  q(\Theta)q(\mathbf{u})}[\Kru^\intercal\Kuu^{-1}\mathbf{u}] \\
+  &= \mathbf{t}^\intercal\mathbb{E}_{\Theta \sim q(\Theta)}[\Kru^\intercal\Kuu^{-1}\mathbf{m}] = \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}.
+\end{align*}
+This allows us to simplify $\mathcal{L}(\bm\nu)$ to the following:
 \begin{align*}
@@
   &- \frac{1}{2}\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \alpha - \log\beta + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha) \\
-  &+ \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] - \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')]. \\
+  &+ \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m} - \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')].
+\\
 \end{align*}
@@
     \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
-    &-\frac{1}{2}\dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] \\
-    &+ \dm\left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
-      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
-      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')]\right],
+    &-\frac{1}{2}\dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}]
+    + \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] \\
+    &- \sum_{i=1}^N \sum_{t=1}^T \dm\mathbb{E}[V_r(s_{i,t})] -
+      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
+      s')\dm\mathbb{E}[V_r(s')],
   \end{split}
@@
     -\frac{1}{2}\dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] \\
-    &+ \dS\left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
-      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
-      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')]\right],
+    &- \dS \left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] -
+      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
+      s')\mathbb{E}[V_r(s')] \right],
   \end{split}
@@
     \frac{1}{2}\da\mathbb{E}[\log|\Kuu|] \\
-    &+ \da [\alpha + \log\Gamma(\alpha) + (1 -
-      \alpha)\psi(\alpha)] \\
-    &+ \da \left[ \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
-      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
-      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \right].
+    &+ \mathbf{t}^\intercal\da\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m} +
+    \da [\alpha + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha)] \\
+      &- \da \left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] -
+        \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
+        s')\mathbb{E}[V_r(s')] \right].
   \end{split}
 \]
-The first three terms can all be treated in the same way, since all three
+The first four terms can all be treated in the same way, since all four
 expectations contain expressions without $\alpha$, but $\alpha$ appears in
@@
     \frac{1}{2}\db\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] -
-    \frac{1}{2}\db\mathbb{E}[\log|\Kuu|] - \db[\log\beta] \\
-    &+ \db \left[ \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[r(s_{i,t})] -
-      \mathbb{E}[V_r(s_{i,t})] + \gamma\sum_{s' \in \mathcal{S}}
+    \frac{1}{2}\db\mathbb{E}[\log|\Kuu|] +
+    \mathbf{t}^\intercal\db\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m} \\
+    &- \db[\log\beta] - \db \left[\sum_{i=1}^N \sum_{t=1}^T
+      \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s' \in \mathcal{S}}
       \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \right].
@@
 \]
-Similarly to the previous section, we can deal with three cases at once:
+Similarly to the previous section, we can deal with four cases at once:
 \[

### 2018-11-04 mprop/mprop.tex
Derivatives

@@
 \usepackage{graphicx}
-\usepackage{amssymb,amsmath,amsthm}
+\usepackage{amssymb,amsmath,amsthm,mathtools}
 \usepackage{url}
@@
 \newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
-\newcommand{\da}{\frac{\partial}{\partial\alpha}}
-\newcommand{\db}{\frac{\partial}{\partial\beta}}
+\newcommand{\da}{\frac{\partial}{\partial\alpha_j}}
+\newcommand{\db}{\frac{\partial}{\partial\beta_j}}
 
@@
 %proposed approach is feasible, but identify any risks.
+% TODO: Theta should be equal to lambda
 
@@
 \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
-\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \zeta\exp\left(
+\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
     -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
     \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
-where $\zeta$ is the overall ``scale'' factor for how similar or distant the
+where $\lambda_0$ is the overall ``scale'' factor for how similar or distant the
 states are,
-\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) = \diag(\bm\lambda) \]
+\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) \]
 is a diagonal matrix that determines how relevant each feature is (where $d$
@@
 the value makes little difference to the performance of the algorithm
-\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\Theta = \{ \zeta,
-\bm\lambda \}$. Similarly, we keep the expression for the prior of $\Theta$:
+\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\bm\lambda = (\lambda_0,
+\dots, \lambda_d)^\intercal$, and $\Theta = \{ \bm\lambda \}$. Similarly, we
+keep the expression for the prior of $\Theta$:
 \begin{equation} \label{eq:theta_prior}
@@
 
-% TODO: uncorrelated gammas instead of a multivariate Gaussian? review several options
 In order to have a reasonable way of calculating/approximating the ELBO (see
@@
 \begin{equation} \label{eq:theta}
-  q(\Theta) = q(\zeta) \times q(\bm\lambda).
+  q(\Theta) = \prod_{i=0}^dq(\lambda_i).
 \end{equation}
-As it is reasonable for $\bm\lambda$ to take any value in $\mathbb{R}^d$, we can
-approximate it simply as
-\begin{equation} \label{eq:lambda}
-  q(\bm\lambda) = \mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma)
+We want to restrict $\lambda_0$ to be positive so that $k_\Theta$ would produce
+non-negative values and not become trivial. Similarly, we want that $\lambda_i
+\ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a positive-definite matrix.
+Considering the possible distributions for all $d + 1$ variables, we would like
+the mean to be flexible (i.e., not tied to zero, like with the exponential
+distribution), and the tails to converge to zero as the value of the random
+variable moves away from the mean. We might want to support some right skew, but
+the distribution should be close to symmetric with at least some parameter
+values. This limits our choice of distributions quite significantly, and we
+decide to go with the gamma distribution as it is fairly flexible and commonly
+used \cite{hogg2018introduction}. We then define the probability density
+functions as
+\begin{equation} \label{eq:zeta}
+  q(\lambda_i) = \Gamma(\lambda_i; \alpha_i, \beta_i) =
+  \frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i - 1}e^{-\beta_i\lambda_i}, \quad i = 0, \dots, d,
 \end{equation}
-for some variational parameters $\bm\mu \in \mathbb{R}^d$ and $\bm\Sigma \in
-\mathbb{R}^{d \times d}$. In contrast, as the covariance function must produce
-non-negative values, we want to restrict $\zeta$ to positive real numbers. At
-the same time, the distribution for $\zeta$ should have a flexible mean (i.e.,
-not be tied to zero, like in the exponential distribution) and tails that
-converge to zero as the value of the random variable moves away from the mean.
-We might want to support some right skew, but the distribution should be close
-to symmetric with at least some parameter values. This limits our choice of
-distributions quite significantly, and we decide to go with the gamma
-distribution as it is fairly flexible and commonly used
-\cite{hogg2018introduction}. We then define the probability density function of
-the gamma distribution as
-\begin{equation} \label{eq:zeta}
-  q(\zeta) = \Gamma(\zeta; \alpha, \beta) =
-  \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta},
-\end{equation}
-where $\alpha > 0$ and $\beta > 0$ are two parameters of the distribution, and
-$\Gamma(\cdot)$ is the gamma function. This gives us our vector of variational
-parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \bm\mu, \bm\Sigma, \alpha,
-\beta)^\intercal$.
+where $\alpha_i > 0$ and $\beta_i > 0$ are parameters of the distribution, and
+$\Gamma(\cdot)$ is the gamma function. This gives us our
+vector of variational parameters $\bm\nu = (\mathbf{m}, \mathbf{S}, \alpha_0,
+\beta_0, \dots, \alpha_d, \beta_d)^\intercal$.
 
@@
     V_r(s_{i,t}) \right] &&-
-  \mathbb{E}\left[\log\left(\frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha -
-        1}e^{-\beta\zeta}\right)\right] \\
-  &- \mathbb{E}[\log\mathcal{N}(\bm\lambda; \bm\mu, \bm\Sigma)] &&-
+  \sum_{i=0}^d\mathbb{E}\left[\log\left(\frac{\beta_i^{\alpha_i}}{\Gamma(\alpha_i)}\lambda_i^{\alpha_i
+        - 1}e^{-\beta_i\lambda_i}\right)\right] \\
+  &-
   \mathbb{E}[\log\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})].
 \end{alignat*}
-For the last two terms we can use the fact that, e.g., $-\mathbb{E}[\log
+We can simplify the last term by noting that $-\mathbb{E}[\log
 \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})] = \frac{1}{2}\log|\mathbf{S}|$
 up to an additive constant that depends on the number of dimensions of the
-distribution \cite{DBLP:journals/tit/AhmedG89}. Note that we cannot use this
-fact for $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)]$  because
-$\Kuu$ depends on $\Theta$. We also plug in the definitions of
+distribution \cite{DBLP:journals/tit/AhmedG89}. Also note that we cannot use
+this fact for $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)]$
+because $\Kuu$ depends on $\Theta$. We also plug in the definitions of
 $\mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)$ and $Q_r$, and get:
 \begin{align*}
-  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
+  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
   &+ \mathbb{E}\left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
@@
     \mathcal{T}(s_{i,t}, a_{i,t}, s')V_r(s') \right] \\
-  &- \mathbb{E}[\alpha\log\beta - \log\Gamma(\alpha) + (\alpha - 1)\log\zeta - \beta\zeta].
+  &- \sum_{i=0}^d \mathbb{E}[\alpha_i\log\beta_i - \log\Gamma(\alpha_i) + (\alpha_i - 1)\log\lambda_i - \beta_i\lambda_i]
 \end{align*}
@@
 of the approximated variables outside of the expectations. Also note that
-$\mathbb{E}[\zeta] = \alpha/\beta$ and
-$\mathbb{E}[\log \zeta] = \psi(\alpha) - \log \beta$, where $\psi$ is the
+$\mathbb{E}[\lambda_i] = \alpha_i/\beta_i$ and
+$\mathbb{E}[\log \lambda_i] = \psi(\alpha_i) - \log\beta_i$, where $\psi$ is the
 digamma function defined as $\psi(x) = \frac{d}{dx}\log\Gamma(x)$
-\cite{DBLP:books/lib/Bishop07}. Finally, we can simplify
+\cite{DBLP:books/lib/Bishop07}. Moreover, we can simplify
 $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by choosing to represent all visited states
@@
 \end{align*}
+Finally, as $\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}$ is a function of
+$\mathbf{u}$ and $\Theta$, we can take the expectation of $\mathbf{u}$, leaving
+the expectation of $\Theta$:
+\[ \mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
+  \mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}] +
+  \mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] =
+  \Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] +
+  \mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}. \]
 This allows us to simplify $\mathcal{L}(\bm\nu)$ to the following:
 \begin{align*}
-  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\bm\Sigma| + \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2} \mathbb{E}[\Tr[\Kuu^{-2}]] -
-    \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] \\
-  &- \frac{1}{2}\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \alpha - \log\beta + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha) \\
+  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] - \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\
+  &- \frac{1}{2}\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m} - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \sum_{i=0}^d \alpha_i - \log\beta_i + \log\Gamma(\alpha_i) + (1 - \alpha_i)\psi(\alpha_i) \\
   &+ \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m} - \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')].
@@
 % TODO: some kind of intro
+% TODO: prove each application of the dominated convergence theorem
 
@@
     \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
-    &-\frac{1}{2}\dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}]
+    &- \frac{1}{2}\dm[\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}]
     + \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] \\
@@
       \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-      s')\dm\mathbb{E}[V_r(s')],
+      s')\dm\mathbb{E}[V_r(s')].
   \end{split}
 \]
-where
+Here
+\begin{align*}
+  \dm[\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}] &= (\mathbb{E}[\Kuu^{-1}] + \mathbb{E}[\Kuu^{-1}]^\intercal)\mathbf{m}, \\
+  \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] &= \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}],
+\end{align*}
+and
 \begin{equation} \label{eq:1}
-\begin{split} % TODO: dominated convergence theorem
-    \dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
-    \dm\iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\mathcal{N}(\mathbf{u};
-    \mathbf{m},
-    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
-    &= \iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{m}} q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda.
+  \begin{split}
+    \dm\mathbb{E}[V_r(s)] &= \dm\iiint V_r(s) p(\mathbf{r} | \Theta,
+    \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
+    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
+    &= \iiint V_r(s) p(\mathbf{r} | \Theta,
+    \mathbf{X_u}, \mathbf{u}) \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
+    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda,
   \end{split}
 \end{equation}
-$\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{m}}$ can be calculated as
-follows: % TODO: by the cookbook
+where
 \[
   \begin{split}
-    \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial\mathbf{m}} &=
+    \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) &=
     \mathcal{N}(\mathbf{u}; \mathbf{m},
@@
 \]
-Plugging it back to \eqref{eq:1} gives:
+Substituting it back into \eqref{eq:1} gives
 \[
   \begin{split}
-    \dm\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
-    \iiint(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})\frac{1}{2}(\mathbf{S}^{-1} +
-    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\mathcal{N}(\mathbf{u};
-    \mathbf{m},
-    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
-    &=
-    \frac{1}{2}\mathbb{E}[(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})(\mathbf{S}^{-1}
-    + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})]
+    \dm\mathbb{E}[V_r(s)] &= \frac{1}{2}\iiint V_r(s)
+    (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})
+    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u};
+    \mathbf{m}, \mathbf{S})
+    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
+    &= \frac{1}{2}\mathbb{E}[V_r(s) (\mathbf{S}^{-1} +
+    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})].
+\end{split}
+\]
+Hence
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
+    &-\frac{1}{2}(\mathbb{E}[\Kuu^{-1}] +
+    \mathbb{E}[\Kuu^{-1}]^\intercal)\mathbf{m} +
+    \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}] \\
+    &- \frac{1}{2}\sum_{i=1}^N\sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})
+    (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})] \\
+    &- \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')
+    \mathbb{E}[V_r(s') (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} -
+    \mathbf{m})].
 \end{split}
@@
     \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
-    \frac{1}{2}\dS[\log|\mathbf{S}|]
-    -\frac{1}{2}\dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] \\
-    &- \dS \left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] -
+    \frac{1}{2}\dS\log|\mathbf{S}|
+    -\frac{1}{2}\dS\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\
+    &- \sum_{i=1}^N \sum_{t=1}^T \dS\mathbb{E}[V_r(s_{i,t})] -
       \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-      s')\mathbb{E}[V_r(s')] \right],
+      s')\dS\mathbb{E}[V_r(s')],
   \end{split}
 \]
-where % TODO: cite cookbook
-\[ \dS[\log|\mathbf{S}|] =
-  \frac{1}{|\mathbf{S}|}\frac{\partial|\mathbf{S}|}{\partial \mathbf{S}} =
-  \frac{1}{|\mathbf{S}|}|\mathbf{S}|\mathbf{S}^{-\intercal} =
-  \mathbf{S}^{-\intercal} \]
+where
+\[ \dS\log|\mathbf{S}| = \mathbf{S}^{-\intercal}, \]
 and
+\[ \dS\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] = \mathbb{E}[\Kuu^{-1}]^\intercal \]
+by \emph{The Matrix Cookbook} \cite{petersen2008matrix}. Then
+\[ \dS\mathbb{E}[V_r(s)] = \iiint V_r(s) q(\mathbf{r}) \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
+  q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda, \]
+where
 \[
   \begin{split}
-    \dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] &=
-    \dS\iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\mathcal{N}(\mathbf{u};
-    \mathbf{m},
-    \mathbf{S})q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda \\
-    &=
-    \iiint\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}\frac{\partial \mathcal{N}(\mathbf{u};
-      \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}}q(\zeta)q(\bm\lambda)\,d\mathbf{u}\,d\zeta\,d\bm\lambda.
-  \end{split}
-\]
-$\frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}}$ can be simplified to
-\[
-  \begin{split}
-    \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial \mathbf{S}} &=
+    \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) &=
     \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\exp \left( -\frac{1}{2}
@@
     &- \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})\dS[(\mathbf{u}
-    - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})], \\
+    - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})]. \\
   \end{split}
 \]
-where
-\[ \dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right] =
+The two remaining derivatives can be taken with the help of \emph{The Matrix
+  Cookbook} \cite{petersen2008matrix}:
+\begin{align*}
+  \dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right] =
   -\frac{1}{2}|\mathbf{S}|^{-3/2}\frac{\partial |\mathbf{S}|}{\partial \mathbf{S}} =
-  -\frac{1}{2}|\mathbf{S}|^{-3/2}|\mathbf{S}|\mathbf{S}^{-\intercal} =
-  -\frac{1}{2|\mathbf{S}|^{1/2}}\mathbf{S}^{-\intercal} \]
-and
-\[ \dS[(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
-  \mathbf{m})] = -\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
-  \mathbf{m})^\intercal\mathbf{S}^{-\intercal} \]
-according to \textit{The Matrix Cookbook} \cite{petersen2008matrix}. Plugging
-the two expression back in gives
-\[ \frac{\partial \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})}{\partial
-    \mathbf{S}} = -\frac{1}{2}\mathbf{S}^{-\intercal}\mathcal{N}(\mathbf{u};
+  -\frac{1}{2}|\mathbf{S}|^{-3/2}|\mathbf{S}|\mathbf{S}^{-\intercal} &=
+  -\frac{1}{2|\mathbf{S}|^{1/2}}\mathbf{S}^{-\intercal}, \\
+  \dS[(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
+  \mathbf{m})] &= -\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
+  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}.
+\end{align*}
+Plugging them back in gives
+\[ \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) =
+  -\frac{1}{2}\mathbf{S}^{-\intercal}\mathcal{N}(\mathbf{u};
   \mathbf{m}, \mathbf{S}) + \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m},
   \mathbf{S})\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
-  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}. \]
-Thus
-\[ \dS\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
-  \frac{1}{2}\mathbb{E}[(\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u})(\mathbf{S}^{-\intercal}(\mathbf{u}
-  - \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
-  \mathbf{S}^{-\intercal})]. \]
-
-\subsubsection{$\partial/\partial\alpha$}
-
+  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}, \]
+and
 \[
   \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\alpha} = &-
-    \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-2}]] -
-    \frac{1}{2}\da\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] -
-    \frac{1}{2}\da\mathbb{E}[\log|\Kuu|] \\
-    &+ \mathbf{t}^\intercal\da\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m} +
-    \da [\alpha + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha)] \\
-      &- \da \left[\sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] -
-        \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-        s')\mathbb{E}[V_r(s')] \right].
+    \dS\mathbb{E}[V_r(s)] &= \frac{1}{2}\iiint V_r(s)
+    (\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
+    \mathbf{m})^\intercal\mathbf{S}^{-\intercal} - \mathbf{S}^{-\intercal})
+    q(\mathbf{r}) \mathcal{N}(\mathbf{u};
+    \mathbf{m}, \mathbf{S}) q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda
+    \\
+    &= \frac{1}{2}\mathbb{E}[V_r(s)(\mathbf{S}^{-\intercal}(\mathbf{u} -
+    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
+    \mathbf{S}^{-\intercal})].
   \end{split}
 \]
-The first four terms can all be treated in the same way, since all four
-expectations contain expressions without $\alpha$, but $\alpha$ appears in
-$q(\zeta)$. Thus we can work with an abstract function as follows:
+Therefore
 \[
   \begin{split}
-    \da\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \da \iiint f(k_\Theta, \mathbf{u})
-    \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} e^{-\beta\zeta}
-    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
-    &= \iiint f(k_\Theta, \mathbf{u})
-    \da \left[ \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} \right]
-      e^{-\beta\zeta}
-    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u}.
+    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
+    \frac{1}{2}\mathbf{S}^{-\intercal} -
+    \frac{1}{2}\mathbb{E}[\Kuu^{-1}]^\intercal - \frac{1}{2}
+    \sum_{i=1}^N\sum_{t=1}^T
+    \mathbb{E}[V_r(s_{i,t})(\mathbf{S}^{-\intercal}(\mathbf{u} -
+    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
+    \mathbf{S}^{-\intercal})] \\
+    &- \gamma\sum_{s' \in \mathcal{S}}\mathcal{T}(s_{i,t}, a_{i,t}, s')
+    \mathbb{E}[V_r(s')(\mathbf{S}^{-\intercal}(\mathbf{u} -
+    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
+    \mathbf{S}^{-\intercal})].
+  \end{split}
+\]
+
+\subsubsection{$\partial/\partial\alpha_j$}
+
+We begin in the usual way:
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\alpha_j} = &-
+    \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-2}]] - \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]] - \frac{1}{2}\da\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] - \frac{1}{2}\da\mathbb{E}[\log|\Kuu|] \\
+    &+ \da\mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] +
+    \da [\alpha_j + \log\Gamma(\alpha_j) + (1 - \alpha_j)\psi(\alpha_j)] \\
+      &- \sum_{i=1}^N \sum_{t=1}^T \da\mathbb{E}[V_r(s_{i,t})] -
+        \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
+        s')\da\mathbb{E}[V_r(s')].
+  \end{split}
+\]
+First,
+\[ \da [\alpha_j + \log\Gamma(\alpha_j) + (1 - \alpha_j)\psi(\alpha_j)] = 1 +
+  \psi(\alpha_j) - \psi(\alpha_j) + (1 - \alpha_j)\psi'(\alpha_j) = 1 + (1 -
+  \alpha_j)\psi'(\alpha_j) \]
+by the definition of $\psi$. The remaining terms can all be treated in the same
+way, as they all contain expectations of scalar functions that are independent
+of $\alpha_j$, and $\alpha_j$ only occurs in $\Gamma(\lambda_j; \alpha_j,
+\beta_j)$. Thus we can work with an abstract function as follows:
+\[
+  \begin{split}
+    \da\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \da \iiint f(k_\Theta, \mathbf{r})
+    q(\bm\lambda) q(\mathbf{r})
+    q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
+    &= \!\begin{multlined}[t]
+      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1}) \da
+      \left[ \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)} \lambda_j^{\alpha_j -
+          1} \right] e^{-\beta_j\lambda_j} \\
+      q(\lambda_{j+1}) \cdots q(\lambda_d) q(\mathbf{r})
+      q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}.
+    \end{multlined}
   \end{split}
@@
   \begin{split}
-    \da \left[ \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} \right] &=
-    \frac{\da[\beta^\alpha\zeta^{\alpha - 1}]\Gamma(\alpha) -
-      \beta^\alpha\zeta^{\alpha - 1}\Gamma'(\alpha)}{(\Gamma(\alpha))^2} \\
-    &= \frac{\beta^\alpha\zeta^{\alpha - 1}\da[\alpha\log\beta + (\alpha -
-      1)\log\zeta]\Gamma(\alpha) - \beta^\alpha\zeta^{\alpha -
-        1}\Gamma'(\alpha)}{(\Gamma(\alpha))^2} \\
-    &= \frac{\beta^\alpha\zeta^{\alpha - 1}(\log\beta + \log\zeta)\Gamma(\alpha)
-      - \beta^\alpha\zeta^{\alpha - 1}\Gamma'(\alpha)}{(\Gamma(\alpha))^2} \\
-    &= \frac{\beta^\alpha\zeta^{\alpha - 1}}{\Gamma(\alpha)}\left(\log\beta +
-    \log\zeta - \frac{\Gamma'(\alpha)}{\Gamma(\alpha)}\right),
+    \da \left[ \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)} \lambda_j^{\alpha_j - 1} \right] &=
+    \frac{\da[\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}]\Gamma(\alpha_j) -
+      \beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}\Gamma'(\alpha_j)}{(\Gamma(\alpha_j))^2} \\
+    &= \frac{\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}\da[\alpha_j\log\beta_j + (\alpha_j -
+      1)\log\lambda_j]\Gamma(\alpha_j) - \beta_j^{\alpha_j}\lambda_j^{\alpha_j -
+        1}\Gamma'(\alpha_j)}{(\Gamma(\alpha_j))^2} \\
+    &= \frac{\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}(\log\beta_j + \log\lambda_i)\Gamma(\alpha_j)
+      - \beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}\Gamma'(\alpha_j)}{(\Gamma(\alpha_j))^2} \\
+    &= \frac{\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}}{\Gamma(\alpha_j)}\left(\log\beta_j +
+    \log\lambda_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)}\right),
   \end{split}
@@
   \begin{split}
-    \da\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \iiint f(k_\Theta, \mathbf{u})
-    \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta} \left(
-      \log\beta + \log\zeta - \frac{\Gamma'(\alpha)}{\Gamma(\alpha)} \right)
-    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
-    &= \iiint f(k_\Theta, \mathbf{u}) \left(\log\beta + \log\zeta -
-      \frac{\Gamma'(\alpha)}{\Gamma(\alpha)} \right) q(\zeta) q(\bm\lambda)
-    q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
-    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{u}) \left(\log\beta + \log\zeta -
-        \frac{\Gamma'(\alpha)}{\Gamma(\alpha)} \right) \right] \\
-    &= \left( \log\beta - \frac{\Gamma'(\alpha)}{\Gamma(\alpha)}
-    \right)\mathbb{E}[f(k_\Theta, \mathbf{u})] + \mathbb{E}[f(k_\Theta,
-    \mathbf{u})\log\zeta].
+    \da\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \!\begin{multlined}[t]
+      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
+      \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)}\lambda_j^{\alpha_j -
+        1}e^{-\beta_j\lambda_j} \left(\log\beta_j + \log\lambda_j -
+        \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) \\
+      q(\lambda_{j+1}) \cdots q(\lambda_d)
+      q(\mathbf{r})q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}
+    \end{multlined} \\
+    &= \iiint f(k_\Theta, \mathbf{r}) \left(\log\beta_j + \log\lambda_j -
+      \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) q(\bm\lambda)
+    q(\mathbf{r})q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
+    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{r}) \left(\log\beta_j + \log\lambda_j -
+        \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) \right] \\
+    &= \left( \log\beta_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)}
+    \right)\mathbb{E}[f(k_\Theta, \mathbf{r})] + \mathbb{E}[f(k_\Theta,
+    \mathbf{r})\log\lambda_j].
   \end{split}
 \]
-Next,
-\[ \da [\alpha + \log\Gamma(\alpha) + (1 - \alpha)\psi(\alpha)] =
-  1 + \psi(\alpha) - \psi(\alpha) + (1 - \alpha)\psi'(\alpha) = 1 + (1 -
-  \alpha)\psi'(\alpha). \]
-
-\subsubsection{$\partial/\partial\beta$}
-
+With these results in mind, we can simplify the initial expression to
 \[
   \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\beta} = &-
-    \frac{1}{2}\db\mathbb{E}[\Tr[\Kuu^{-2}]] -
-    \frac{1}{2}\db\mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] -
-    \frac{1}{2}\db\mathbb{E}[\log|\Kuu|] +
-    \mathbf{t}^\intercal\db\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m} \\
-    &- \db[\log\beta] - \db \left[\sum_{i=1}^N \sum_{t=1}^T
-      \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s' \in \mathcal{S}}
-      \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \right].
+    \frac{\partial\mathcal{L}}{\partial\alpha_j} &= 1 + (1 -
+    \alpha_j)\psi'(\alpha_j) + \left( \log\beta_j -
+      \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right)
+    \left(\vphantom{\sum_{i=1}^N} \right. - \frac{1}{2}
+    \mathbb{E}[\Tr[\Kuu^{-2}]] - \frac{1}{2}\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]]
+    \\
+    &- \frac{1}{2}\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] -
+    \frac{1}{2}\mathbb{E}[\log|\Kuu|] +
+    \mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] \\
+    &- \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s' \in
+      \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \left.
+      \vphantom{\sum_{i=1}^N} \right) \\
+    & - \frac{1}{2}\mathbb{E}[\Tr[\Kuu^{-2}]\log\lambda_j] -
+    \frac{1}{2}\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]\log\lambda_j] -
+    \frac{1}{2}\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}\log\lambda_j]
+    \\
+    &- \frac{1}{2}\mathbb{E}[\log|\Kuu|\log\lambda_j] +
+    \mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}\log\lambda_j]
+    \\
+    &- \sum_{i=1}^N\sum_{t=1}^T\mathbb{E}[V_r(s_{i,t})\log\lambda_j] -
+    \gamma\sum_{s' \in \mathcal{S}}\mathcal{T}(s_{i,t}, a_{i,t},
+    s')\mathbb{E}[V_r(s')\log\lambda_j].
   \end{split}
 \]
-Similarly to the previous section, we can deal with four cases at once:
+
+\subsubsection{$\partial/\partial\beta_j$}
+Finally,
 \[
   \begin{split}
-    \db\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \db \iiint f(k_\Theta, \mathbf{u})
-    \frac{\beta^\alpha}{\Gamma(\alpha)} \zeta^{\alpha - 1} e^{-\beta\zeta}
-    q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
-    &= \iiint f(k_\Theta, \mathbf{u}) \frac{\zeta^{\alpha - 1}}{\Gamma(\alpha)}
-    \db [\beta^\alpha e^{-\beta\zeta}] q(\bm\lambda)
-    q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u}.
+    \frac{\partial\mathcal{L}}{\partial\beta_j} = &-
+    \frac{1}{2}\db\mathbb{E}[\Tr[\Kuu^{-2}]] -
+    \frac{1}{2}\db\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]] -
+    \frac{1}{2}\db\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] \\
+    &- \frac{1}{2}\db\mathbb{E}[\log|\Kuu|] +
+    \db\mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] -
+    \db[\log\beta_j] \\
+    &- \sum_{i=1}^N \sum_{t=1}^T \db \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s'
+      \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\db\mathbb{E}[V_r(s')].
+  \end{split}
+\]
+Similarly to the previous section, we can handle all derivatives of expectations
+in the same way:
+\[
+  \begin{split}
+    \db\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \db \iiint f(k_\Theta, \mathbf{r})
+    q(\bm\lambda) q(\mathbf{r}) q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
+    &= \!\begin{multlined}[t]
+      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
+      \frac{\lambda_j^{\alpha_j - 1}}{\Gamma(\alpha_j)} \db [\beta_j^{\alpha_j}
+      e^{-\beta_j\lambda_j}] \\
+      q(\lambda_{j+1}) \cdots q(\lambda_d) q(\mathbf{r})
+      q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}.
+    \end{multlined}
   \end{split}
@@
 Since
-\[ \db [\beta^\alpha e^{-\beta\zeta}] = \alpha\beta^{\alpha - 1}e^{-\beta\zeta}
-  - \beta^\alpha e^{-\beta\zeta}\beta = \beta^\alpha
-  e^{-\beta\zeta} \left( \frac{\alpha}{\beta} - \beta \right), \]
+\[ \db [\beta_j^{\alpha_j} e^{-\beta_j\lambda_j}] = \alpha_j\beta_j^{\alpha_j - 1}e^{-\beta_j\lambda_j}
+  - \beta_j^{\alpha_j} e^{-\beta_j\lambda_j}\lambda_j = \beta_j^{\alpha_j}
+  e^{-\beta_j\lambda_j} \left( \frac{\alpha_j}{\beta_j} - \lambda_j \right), \]
 we have that
@@
   \begin{split}
-    \db\mathbb{E}[f(k_\Theta, \mathbf{u})] &= \iiint f(k_\Theta, \mathbf{u})
-    \frac{\beta^\alpha}{\Gamma(\alpha)}\zeta^{\alpha - 1}e^{-\beta\zeta} \left(
-      \frac{\alpha}{\beta} - \beta \right)  q(\bm\lambda)
-    q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
-    &= \iiint f(k_\Theta, \mathbf{u}) \left(\frac{\alpha}{\beta} - \beta \right)
-    q(\zeta)q(\bm\lambda) q(\mathbf{u})\,d\zeta\,d\bm\lambda\,d\mathbf{u} \\
-    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{u}) \left( \frac{\alpha}{\beta} -
-        \beta \right) \right] =  \left( \frac{\alpha}{\beta} - \beta \right)
-    \mathbb{E} [f(k_\Theta, \mathbf{u})].
+    \db\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \!\begin{multlined}[t]
+      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
+      \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)}\lambda_j^{\alpha_j -
+        1}e^{-\beta_j\lambda_j} \left(\frac{\alpha_j}{\beta_j} - \lambda_j
+      \right) \\
+      q(\lambda_{j+1}) \cdots q(\lambda_d) q(\mathbf{r})
+      q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}
+    \end{multlined}
+    \\
+    &= \iiint f(k_\Theta, \mathbf{r}) \left(\frac{\alpha_j}{\beta_j} - \lambda_j
+    \right) q(\bm\lambda) q(\mathbf{r})
+    q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
+    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{r}) \left( \frac{\alpha_j}{\beta_j} -
+        \lambda_j \right) \right] =
+    \frac{\alpha_j}{\beta_j}\mathbb{E}[f(k_\Theta, \mathbf{r})] -
+    \mathbb{E}[f(k_\Theta, \mathbf{r})\lambda_j].
+  \end{split}
+\]
+This gives us the final expression of $\frac{\partial\mathcal{L}}{\partial\beta_j}$:
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\beta_j} = &- \frac{1}{\beta_j} -
+    \frac{1}{2}\mathbb{E} \left[ \Tr[\Kuu^{-2}] \left( \frac{\alpha_j}{\beta_j}
+        - \lambda_j \right) \right] - \frac{1}{2}\mathbb{E} \left[
+      \Tr[\Kuu^{-1}\mathbf{S}] \left(\frac{\alpha_j}{\beta_j} - \lambda_j
+      \right) \right] \\
+    &- \frac{1}{2}\mathbb{E} \left[ \mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}
+      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] -
+    \frac{1}{2}\mathbb{E} \left[ \log|\Kuu| \left(\frac{\alpha_j}{\beta_j} -
+        \lambda_j \right) \right] \\
+    &+ \mathbb{E} \left[ \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}
+      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] \\
+    &- \sum_{i=1}^N \sum_{t=1}^T \mathbb{E} \left[ V_r(s_{i,t})
+      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] -
+    \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}
+    \left[V_r(s') \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right].
   \end{split}

### 2018-11-04 mprop/mprop.tex
Set Theta = lambda, and explain the MDP representation a little better

@@
 
-\newcommand{\pfull}{p(\mathcal{D}, \Theta, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
-\newcommand{\pfullS}{p(\mathcal{D}, \Theta_s, \mathbf{X_u}, \mathbf{u}_s, \mathbf{r}_s)}
-\newcommand{\posterior}{p(\Theta, \mathbf{u}, \mathbf{r} | \mathcal{D}, \mathbf{X_u})}
-\newcommand{\approximation}{q_{\bm\nu}(\Theta, \mathbf{u}, \mathbf{r})}
-\newcommand{\approximationS}{q_{\bm\nu}(\Theta_s, \mathbf{u}_s, \mathbf{r}_s)}
-
-\newcommand{\Eq}{\mathbb{E}_{(\Theta, \mathbf{u}, \mathbf{r}) \sim \approximation}}
+\newcommand{\pfull}{p(\mathcal{D}, \bm\lambda, \mathbf{X_u}, \mathbf{u}, \mathbf{r})}
+\newcommand{\pfullS}{p(\mathcal{D}, \bm\lambda_s, \mathbf{X_u}, \mathbf{u}_s, \mathbf{r}_s)}
+\newcommand{\posterior}{p(\bm\lambda, \mathbf{u}, \mathbf{r} | \mathcal{D}, \mathbf{X_u})}
+\newcommand{\approximation}{q_{\bm\nu}(\bm\lambda, \mathbf{u}, \mathbf{r})}
+\newcommand{\approximationS}{q_{\bm\nu}(\bm\lambda_s, \mathbf{u}_s, \mathbf{r}_s)}
+
+\newcommand{\Eq}{\mathbb{E}_{(\bm\lambda, \mathbf{u}, \mathbf{r}) \sim \approximation}}
 
@@
 % TODO: example application for motivation (why would the world be better)
-% TODO: (possibly) a figure with my notation
-% TODO: explain the MDP Q/V stuff
 % TODO: introduce research questions
@@
 $r$
-\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \sum_{t=0}^\infty \gamma^t r(s_t) | \pi
-  \right] \]
+\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \left. \sum_{t=0}^\infty \gamma^t r(s_t)
+    \right| \pi \right] \]
 matches the actions in $\mathcal{D}$.
 
-The likelihood of the data can be written down as
+Following previous work on GP IRL
+\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
+entropy IRL model \cite{ziebart2008maximum}, under which we have that
+\[ P(a | s) \propto \exp(Q_r(s, a)), \]
+where
+\begin{equation}
+  Q_r(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s, a, s')V_r(s'),
+\end{equation}
+and $V_r(s)$ is a ``soft'' version of the Bellman backup operator, which can be
+obtained by repeatedly applying the following equation until convergence:
+\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
+\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
+      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
+The likelihood of the data can then be written down as
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
 \begin{equation} \label{pDr}
-  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right),
-\end{equation}
-where
-\begin{equation}
-  Q_r(s_{i,t}, a_{i,t}) = r(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}}
-  \mathcal{T}(s_{i,t}, a_{i,t}, s')V_r(s'),
-\end{equation}
-and $V_r(s)$ can be obtained by repeatedly applying the equation
-\cite{supplementary_material}
-\[ V_r(s) = \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in
-      \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
+  p(\mathcal{D} | r) = \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) = \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_r(s_{i,t}, a_{i,t}) - V_r(s_{i,t}) \right).
+\end{equation}
 However, a reward function learned by maximising this likelihood is not
@@
 \cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
-k_{\Theta})$ to say that $r$ is a GP with mean $0$ and covariance function
-$k_{\Theta}$, which uses a set of hyperparameters $\Theta$. Covariance functions
-take two state feature vectors as input and quantify how similar the two states
-are, in a sense that we would expect them to have similar rewards.
-
-% TODO: positive definite
+k_{\bm\lambda})$ to say that $r$ is a GP with mean $0$ and covariance function
+$k_{\bm\lambda}$, which uses a vector of hyperparameters $\bm\lambda$.
+Covariance functions take two state feature vectors as input and quantify how
+similar the two states are, in a sense that we would expect them to have similar
+rewards.
+
 As training a GP with $n$ data points has a time complexity of
@@
 \begin{equation} \label{full}
-  \pfull = p(\mathbf{X_u}) \times p(\Theta | \mathbf{X_u}) \times p(\mathbf{u}
-  | \Theta, \mathbf{X_u}) \times p(\mathbf{r} | \Theta, \mathbf{X_u},
+  \pfull = p(\mathbf{X_u}) \times p(\bm\lambda | \mathbf{X_u}) \times p(\mathbf{u}
+  | \bm\lambda, \mathbf{X_u}) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u},
   \mathbf{u}) \times p(\mathcal{D} | r).
 \end{equation}
-Here $p(\mathbf{X_u})$ and $p(\Theta | \mathbf{X_u})$ are customisable priors,
+Here $p(\mathbf{X_u})$ and $p(\bm\lambda | \mathbf{X_u})$ are customisable priors,
 \begin{align}
-  p(\mathbf{u} | \Theta, \mathbf{X_u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \nonumber \\
+  p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \nonumber \\
                                        &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \nonumber \\
@@
 \begin{equation} \label{eq:r}
-  p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) =
+  p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) =
   \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
@@
 $\Kru$ are called \emph{covariance matrices} and are defined as
-$[\Kru]_{i,j} = k_\Theta(\mathbf{x}_{\mathbf{r},i}, \mathbf{x}_{\mathbf{u},j})$,
-where $\mathbf{x}_{\mathbf{r},i}$ and $\mathbf{x}_{\mathbf{u},j}$ denote feature
-vectors for the $i$th state in $\mathcal{S}$ and the $j$th state in
-$\mathbf{X_u}$, respectively \cite{DBLP:conf/uai/JinDAS17}.
+$[\Kru]_{i,j} = k_{\bm\lambda}(\mathbf{x}_{\mathbf{r},i},
+\mathbf{x}_{\mathbf{u},j})$, where $\mathbf{x}_{\mathbf{r},i}$ and
+$\mathbf{x}_{\mathbf{u},j}$ denote feature vectors for the $i$th state in
+$\mathcal{S}$ and the $j$th state in $\mathbf{X_u}$, respectively
+\cite{DBLP:conf/uai/JinDAS17}.
 
@@
 $\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
-$\Theta$, inducing rewards $\mathbf{u}$, and the reward function $r$. While the
-previous paper that considered this IRL model computed maximum likelihood estimates
-for $\Theta$ and $\mathbf{u}$, and made an assumption that $\mathbf{r}$ in
-\eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11}, we aim to
-avoid this assumption and use variational inference to approximate the full
-posterior distribution $\posterior$. \emph{Variational inference} (VI) is an
-approximation technique for probability densities \cite{blei2017variational}.
+$\bm\lambda$, inducing rewards $\mathbf{u}$, and the reward function $r$. While
+the previous paper that considered this IRL model computed maximum likelihood
+estimates for $\bm\lambda$ and $\mathbf{u}$, and made an assumption that
+$\mathbf{r}$ in \eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11},
+we aim to avoid this assumption and use variational inference to approximate the
+full posterior distribution $\posterior$. \emph{Variational inference} (VI) is
+an approximation technique for probability densities \cite{blei2017variational}.
 Let $\approximation$ be our approximating family of probability distributions
-for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up
-to VI algorithms to optimise $\bm\nu$ in order to minimise the
+for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up to
+VI algorithms to optimise $\bm\nu$ in order to minimise the
 \emph{Kullback-Leibler} (KL) divergence between the original probability
@@
     &= \iiint \approximation \log
-    \frac{\pfull}{\approximation}\,d\Theta\,d\mathbf{u}\,d\mathbf{r}.
+    \frac{\pfull}{\approximation}\,d\bm\lambda\,d\mathbf{u}\,d\mathbf{r}.
   \end{split}
@@
 \begin{table}
-  \centering % TODO: recheck
+  \centering
   \begin{tabular}{llllll}
@@
 %proposed approach is feasible, but identify any risks.
-% TODO: Theta should be equal to lambda
 
@@
 \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
-\[ k_\Theta(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
+\[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
     -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
@@
 the value makes little difference to the performance of the algorithm
-\cite{DBLP:conf/nips/LevinePK11}). Thus we can set $\bm\lambda = (\lambda_0,
-\dots, \lambda_d)^\intercal$, and $\Theta = \{ \bm\lambda \}$. Similarly, we
-keep the expression for the prior of $\Theta$:
+\cite{DBLP:conf/nips/LevinePK11}). Our vector of hyperparameters for the
+covariance function is then $\bm\lambda = (\lambda_0, \dots,
+\lambda_d)^\intercal$. Similarly, we keep the expression for the prior of
+$\bm\lambda$:
 \begin{equation} \label{eq:theta_prior}
-  p(\Theta | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
+  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
     \sum_{i=1}^d \log(\lambda_i + 1) \right).
@@
   \begin{split}
-    \posterior &= p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
-    \Theta, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \Theta,
+    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u}
+    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \bm\lambda,
     \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
-    &\propto p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} | \Theta,
-    \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times p(\mathbf{r} |
-    \Theta, \mathbf{X_u}, \mathbf{u}) \\
-    &\propto p(\Theta | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
-    \Theta, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \Theta,
-    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta,
+    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
+    \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times
+    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
+    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
+    \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda,
+    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda,
     \mathbf{X_u}, \mathbf{u}) \\
-    &\propto p(\mathcal{D} | \Theta, \mathbf{X_u}) \times p(\Theta |
-    \mathbf{X_u}) \times p(\mathcal{D} | \Theta, \mathbf{X_u}, \mathbf{u})
-    \times p(\mathbf{u} | \Theta, \mathbf{X_u}) \times p(\mathcal{D} | r) \times
-    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u})
+    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) \times p(\bm\lambda |
+    \mathbf{X_u}) \times p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
+    \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \times p(\mathcal{D} | r)
+    \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
   \end{split}
@@
 Note that now there are only two unknown probability distributions:
-$p(\mathcal{D} | \Theta, \mathbf{X_u})$ and $p(\mathcal{D} | \Theta,
+$p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
 \mathbf{X_u}, \mathbf{u})$, which can be computed as follows:
 \begin{align*}
-  p(\mathcal{D} | \Theta, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
-  p(\mathcal{D} | \Theta, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \Theta, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
+  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
+  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
 \end{align*}
@@
 \begin{equation} \label{eq:approximation}
-  \approximation = q(\Theta) \times q(\mathbf{u} | \Theta) \times q(\mathbf{r}
-  | \Theta, \mathbf{u}).
-\end{equation}
-
-%where $q(\Theta)$ can be of the same form as $p(\Theta | \mathbf{X_u})$,
-%$q(\mathbf{u} | \Theta) = \mathcal{N}(\mathbf{0}, \bm\Sigma_{\mathbf{u}})$, and
-%$q(\mathbf{r} | \Theta, \mathbf{u}) = \mathcal{N}(\bm\mu_{\mathbf{r}},
-%\bm\Sigma_{\mathbf{r}})$.
+  \approximation = q(\bm\lambda) \times q(\mathbf{u} | \bm\lambda) \times q(\mathbf{r}
+  | \bm\lambda, \mathbf{u}).
+\end{equation}
 
@@
 \begin{equation}
-  q(\mathbf{r} | \Theta, \mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u},
+  q(\mathbf{r} | \bm\lambda, \mathbf{u}) = p(\mathbf{r} | \bm\lambda, \mathbf{X_u},
   \mathbf{u}).
 \end{equation}
-We can make a similarly justified choice for $q(\mathbf{u} | \Theta)$:
+We can make a similarly justified choice for $q(\mathbf{u} | \bm\lambda)$:
 \begin{equation}
-  q(\mathbf{u} | \Theta) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
+  q(\mathbf{u} | \bm\lambda) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
 \end{equation}
@@
 Section \ref{vi_algs}), we need to have an approximating distribution for
-$\Theta$, but unfortunately all the papers in Table \ref{table:approximations}
+$\bm\lambda$, but unfortunately all the papers in Table \ref{table:approximations}
 either fix it or treat it as a variational parameter. Hence we make our first
@@
 \begin{equation} \label{eq:theta}
-  q(\Theta) = \prod_{i=0}^dq(\lambda_i).
-\end{equation}
-We want to restrict $\lambda_0$ to be positive so that $k_\Theta$ would produce
-non-negative values and not become trivial. Similarly, we want that $\lambda_i
-\ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a positive-definite matrix.
-Considering the possible distributions for all $d + 1$ variables, we would like
-the mean to be flexible (i.e., not tied to zero, like with the exponential
-distribution), and the tails to converge to zero as the value of the random
-variable moves away from the mean. We might want to support some right skew, but
-the distribution should be close to symmetric with at least some parameter
-values. This limits our choice of distributions quite significantly, and we
-decide to go with the gamma distribution as it is fairly flexible and commonly
-used \cite{hogg2018introduction}. We then define the probability density
-functions as
+  q(\bm\lambda) = \prod_{i=0}^dq(\lambda_i).
+\end{equation}
+We want to restrict $\lambda_0$ to be positive so that $k_{\bm\lambda}$ would
+produce non-negative values and not become trivial. Similarly, we want that
+$\lambda_i \ge 0$ for $i = 1, \dots, d$ so that $\bm\Lambda$ is a
+positive-definite matrix. Considering the possible distributions for all $d + 1$
+variables, we would like the mean to be flexible (i.e., not tied to zero, like
+with the exponential distribution), and the tails to converge to zero as the
+value of the random variable moves away from the mean. We might want to support
+some right skew, but the distribution should be close to symmetric with at least
+some parameter values. This limits our choice of distributions quite
+significantly, and we decide to go with the gamma distribution as it is fairly
+flexible and commonly used \cite{hogg2018introduction}. We then define the
+probability density functions as
 \begin{equation} \label{eq:zeta}
@@
 \subsection{Evidence Lower Bound}
-% TODO: remove/add niu from/to q everywhere
 
@@
   \begin{split}
-    \mathcal{L}(\bm\nu) &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\Theta |
-    \mathbf{X_u}) + \log p(\mathbf{u} | \Theta, \mathbf{X_u}) + \log
-    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) + \log p(\mathcal{D} | r)] \\
-    &- \mathbb{E}[\log q(\Theta) + \log q(\mathbf{u}) + \log q(\mathbf{r} |
-    \Theta, \mathbf{u})].
+    \mathcal{L}(\bm\nu) &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\bm\lambda |
+    \mathbf{X_u}) + \log p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) + \log
+    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) + \log p(\mathcal{D} | r)] \\
+    &- \mathbb{E}[\log q(\bm\lambda) + \log q(\mathbf{u}) + \log q(\mathbf{r} |
+    \bm\lambda, \mathbf{u})].
   \end{split}
@@
 Note that $\mathbb{E}[\log p(\mathbf{X_u})]$ is just a constant, so we can
-simply drop it from the expression. Furthermore, since $q(\mathbf{r} | \Theta,
-\mathbf{u}) = p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u})$, they cancel
+simply drop it from the expression. Furthermore, since $q(\mathbf{r} | \bm\lambda,
+\mathbf{u}) = p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})$, they cancel
 each other out. Then we can substitute various terms with their definitions to
@@
 this fact for $\mathbb{E}[\log \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)]$
-because $\Kuu$ depends on $\Theta$. We also plug in the definitions of
+because $\Kuu$ depends on $\bm\lambda$. We also plug in the definitions of
 $\mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)$ and $Q_r$, and get:
@@
     \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] =
-    \mathbf{t}^\intercal\mathbb{E}_{(\Theta, \mathbf{u}) \sim
-  q(\Theta)q(\mathbf{u})}[\Kru^\intercal\Kuu^{-1}\mathbf{u}] \\
-  &= \mathbf{t}^\intercal\mathbb{E}_{\Theta \sim q(\Theta)}[\Kru^\intercal\Kuu^{-1}\mathbf{m}] = \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}.
+    \mathbf{t}^\intercal\mathbb{E}_{(\bm\lambda, \mathbf{u}) \sim
+  q(\bm\lambda)q(\mathbf{u})}[\Kru^\intercal\Kuu^{-1}\mathbf{u}] \\
+  &= \mathbf{t}^\intercal\mathbb{E}_{\bm\lambda \sim q(\bm\lambda)}[\Kru^\intercal\Kuu^{-1}\mathbf{m}] = \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}.
 \end{align*}
 Finally, as $\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}$ is a function of
-$\mathbf{u}$ and $\Theta$, we can take the expectation of $\mathbf{u}$, leaving
-the expectation of $\Theta$:
+$\mathbf{u}$ and $\bm\lambda$, we can take the expectation of $\mathbf{u}$, leaving
+the expectation of $\bm\lambda$:
 \[ \mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
@@
 % TODO: some kind of intro
-% TODO: prove each application of the dominated convergence theorem
 
@@
   \begin{split}
-    \dm\mathbb{E}[V_r(s)] &= \dm\iiint V_r(s) p(\mathbf{r} | \Theta,
+    \dm\mathbb{E}[V_r(s)] &= \dm\iiint V_r(s) p(\mathbf{r} | \bm\lambda,
     \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
     q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
-    &= \iiint V_r(s) p(\mathbf{r} | \Theta,
+    &= \iiint V_r(s) p(\mathbf{r} | \bm\lambda,
     \mathbf{X_u}, \mathbf{u}) \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
@@
     (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})
-    p(\mathbf{r} | \Theta, \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u};
+    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u};
     \mathbf{m}, \mathbf{S})
@@
   \begin{split}
-    \da\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \da \iiint f(k_\Theta, \mathbf{r})
-    q(\bm\lambda) q(\mathbf{r})
+    \da\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \da \iiint f(k_{\bm\lambda},
+    \mathbf{r}) q(\bm\lambda) q(\mathbf{r})
     q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
     &= \!\begin{multlined}[t]
-      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1}) \da
+      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1}) \da
       \left[ \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)} \lambda_j^{\alpha_j -
@@
   \begin{split}
-    \da\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \!\begin{multlined}[t]
-      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
+    \da\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \!\begin{multlined}[t]
+      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
       \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)}\lambda_j^{\alpha_j -
@@
     \end{multlined} \\
-    &= \iiint f(k_\Theta, \mathbf{r}) \left(\log\beta_j + \log\lambda_j -
+    &= \iiint f(k_{\bm\lambda}, \mathbf{r}) \left(\log\beta_j + \log\lambda_j -
       \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) q(\bm\lambda)
     q(\mathbf{r})q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
-    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{r}) \left(\log\beta_j + \log\lambda_j -
-        \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) \right] \\
+    &= \mathbb{E} \left[ f(k_{\bm\lambda}, \mathbf{r}) \left(\log\beta_j +
+        \log\lambda_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right)
+    \right] \\
     &= \left( \log\beta_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)}
-    \right)\mathbb{E}[f(k_\Theta, \mathbf{r})] + \mathbb{E}[f(k_\Theta,
-    \mathbf{r})\log\lambda_j].
+    \right)\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] +
+    \mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})\log\lambda_j].
   \end{split}
@@
   \begin{split}
-    \db\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \db \iiint f(k_\Theta, \mathbf{r})
-    q(\bm\lambda) q(\mathbf{r}) q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
+    \db\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \db \iiint f(k_{\bm\lambda},
+    \mathbf{r}) q(\bm\lambda) q(\mathbf{r})
+    q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
     &= \!\begin{multlined}[t]
-      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
+      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
       \frac{\lambda_j^{\alpha_j - 1}}{\Gamma(\alpha_j)} \db [\beta_j^{\alpha_j}
@@
   \begin{split}
-    \db\mathbb{E}[f(k_\Theta, \mathbf{r})] &= \!\begin{multlined}[t]
-      \iiint f(k_\Theta, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
+    \db\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \!\begin{multlined}[t]
+      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
       \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)}\lambda_j^{\alpha_j -
@@
     \\
-    &= \iiint f(k_\Theta, \mathbf{r}) \left(\frac{\alpha_j}{\beta_j} - \lambda_j
-    \right) q(\bm\lambda) q(\mathbf{r})
+    &= \iiint f(k_{\bm\lambda}, \mathbf{r}) \left(\frac{\alpha_j}{\beta_j} -
+      \lambda_j \right) q(\bm\lambda) q(\mathbf{r})
     q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
-    &= \mathbb{E} \left[ f(k_\Theta, \mathbf{r}) \left( \frac{\alpha_j}{\beta_j} -
-        \lambda_j \right) \right] =
-    \frac{\alpha_j}{\beta_j}\mathbb{E}[f(k_\Theta, \mathbf{r})] -
-    \mathbb{E}[f(k_\Theta, \mathbf{r})\lambda_j].
+    &= \mathbb{E} \left[ f(k_{\bm\lambda}, \mathbf{r})
+      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] =
+    \frac{\alpha_j}{\beta_j}\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] -
+    \mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})\lambda_j].
   \end{split}
@@
   \approximationS) \]
-computed by drawing $S$ Monte Carlo samples $(\Theta_s, \mathbf{u}_s,
+computed by drawing $S$ Monte Carlo samples $(\bm\lambda_s, \mathbf{u}_s,
 \mathbf{r}_s) \sim \approximation$.
@@
 and dates.
+
+1. prove all applications of the dominated convergence theorem
+
+2. develop the algorithm
+
+3. test in on some example problems, drawing some pretty graphs
+
+4. integrate a few extra improvements from various papers
 

### 2018-11-06 mprop/mprop.tex
Literature survey on IRL

@@
 
-% TODO: where does the lower bound come from?
 Given this model, data $\mathcal{D}$, and inducing feature matrix
 $\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
-$\bm\lambda$, inducing rewards $\mathbf{u}$, and the reward function $r$. While
-the previous paper that considered this IRL model computed maximum likelihood
-estimates for $\bm\lambda$ and $\mathbf{u}$, and made an assumption that
-$\mathbf{r}$ in \eqref{eq:r} has zero variance \cite{DBLP:conf/nips/LevinePK11},
-we aim to avoid this assumption and use variational inference to approximate the
-full posterior distribution $\posterior$. \emph{Variational inference} (VI) is
-an approximation technique for probability densities \cite{blei2017variational}.
-Let $\approximation$ be our approximating family of probability distributions
-for $\posterior$ with its own hyperparameter vector $\bm\nu$. Then it is up to
-VI algorithms to optimise $\bm\nu$ in order to minimise the
-\emph{Kullback-Leibler} (KL) divergence between the original probability
-distribution and our approximation.  KL divergence (asymmetrically) measures how
-different the two distributions are and in this case can be defined as follows
-\cite{blei2017variational}:
-\[ \DKL(\approximation || \posterior) = \Eq \left[ \log
-  \frac{\approximation}{\posterior} \right]. \]
-Since KL divergence is typically hard to compute, instead of minimising it, VI
-typically tries to maximise the \emph{evidence lower bound} (ELBO) defined as
+$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
+states $\mathbf{r}$. While the previous paper to consider this IRL model
+computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
+made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
+\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
+variational inference to approximate the full posterior distribution
+$\posterior$. \emph{Variational inference} (VI) is an approximation technique
+for probability densities \cite{blei2017variational}. Let $\approximation$ be
+our approximating family of probability distributions for $\posterior$ with its
+own hyperparameter vector $\bm\nu$. Then the job of VI algorithms is to optimise
+$\bm\nu$ in order to minimise the \emph{Kullback-Leibler} (KL) divergence
+between the original probability distribution and our approximation.  KL
+divergence (asymmetrically) measures how different the two distributions are,
+and in this case can be defined as \cite{blei2017variational}
+\[ \begin{split}
+    \DKL(\approximation || \posterior) &= \Eq [\log\approximation -
+    \log\posterior ] \\
+    &= \Eq[\log\approximation - \log\pfull] \\
+    &+ \Eq[\log p(\mathcal{D}, \mathbf{X_u})].
+  \end{split}
+\]
+The last term is both hard to compute and constant w.r.t. $\approximation$
+\cite{blei2017variational}, so we can remove it from our optimisation objective.
+The negation of what remains is often called the \emph{evidence lower bound}
+(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
+  be interpreted as Lebesgue integrals over the entire sample space.}
 \cite{DBLP:books/lib/Bishop07,blei2017variational}
@@
 
-By considering full probability distributions instead of point estimates,---as
+By considering full probability distributions instead of point estimates---as
 long as the approximations are able to capture important features of the
@@
 \ref{literature}), making the resulting algorithm competitive both in terms of
-speed and model fit.
+running time and model fit.
 
@@
 
-% TODO: review IRL papers, their weaknesses
+As mentioned in the introduction, most IRL algorithms assume that the reward
+function can be represented as a linear combination of features. This assumption
+originated in one of the earliest papers on the topic by Ng and Russell
+\cite{DBLP:conf/icml/NgR00}, which introduced several linear programming
+approaches to the problem. The authors also noticed that often
+multiple reward functions can explain the same behaviour, and suggested
+heuristics for reward functions that are ``far away'' from reward functions that
+do not fit the data.
+
+A few years later, Abbeel and Ng \cite{DBLP:conf/icml/PieterN04} developed an
+algorithm for the same formulation of the problem with a guarantee to converge
+quickly. Neu and Szepesv\'ari \cite{DBLP:conf/uai/NeuS07} identified a weakness
+in Abbeel and Ng's approach: the algorithm requires features to be
+``appropriately'' scaled, and optimal scaling may not be known. Instead, they
+suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
+supervised learning task for optimal \emph{policy} recovery (whereas IRL focuses
+on recovering the reward function).
+
+Ramachandran and Amir \cite{DBLP:conf/ijcai/RamachandranA07} were the first to
+formulate IRL in terms of Bayesian learning. While the model is easily
+interpretable and able to handle experts that make mistakes, the algorithm can
+only handle small state spaces and requires Monte Carlo Markov Chain (MCMC)
+sampling for inference.
+
+Ziebart et al. \cite{ziebart2008maximum} keep the linearity assumption, but
+introduce an influential idea: resolving the ambiguity when multiple reward
+functions explain the data by appealing to the maximum entropy principle.
+
+Choi and Kim \cite{DBLP:conf/ijcai/ChoiK13} extend the Bayesian model to learn
+good features as well as the reward function, trying to overcome the limitation
+of linearity. However, the approach is quite limiting: all features are assumed
+to have Boolean values, and the algorithm simply learns their conjunctions.
+
+Levine et al. \cite{DBLP:conf/nips/LevinePK11} are the first to suggest a way to
+learn nonlinear reward functions without harsh restrictions on the problem
+domain by using GPs. We base our work primarily on their paper, and the
+weaknesses we hope to address have already been covered in the previous section.
+A recent extension to their work by Jin et al. \cite{DBLP:conf/uai/JinDAS17}
+aims to harness the power of deep learning by using several layers of GPs,
+making the model less dependent on being provided good features. They also use
+VI, but with a few simplifying assumptions: deterministic training conditional
+for the reward vector, and fully independent training conditional for the latent
+state (see Section \ref{sec:vi_literature} for more details).
+
+%A different take on IRL with GPs \cite{DBLP:journals/corr/abs-1208-2112}
+%IRL for large state spaces (haven't read) \cite{DBLP:journals/jmlr/BoulariasKP11}
+%Multiple reward functions \cite{DBLP:conf/nips/ChoiK12}
+%A recent survey \cite{DBLP:journals/corr/abs-1806-06877}
+%EM clustering \cite{DBLP:conf/icml/BabesMLS11}
+%Structured priors \cite{DBLP:conf/ewrl/DimitrakakisR11}
+
+\subsection{Variational Inference} \label{sec:vi_literature}
 
@@
 \subsection{Derivatives}
-% TODO: integrals always mean "on all range"
 % TODO: some kind of intro
@@
 \section{Work Plan}
-
-show how you plan to organize your work, identifying intermediate deliverables
-and dates.
-
-1. prove all applications of the dominated convergence theorem
-
-2. develop the algorithm
-
-3. test in on some example problems, drawing some pretty graphs
-
-4. integrate a few extra improvements from various papers
+%show how you plan to organize your work, identifying intermediate deliverables
+%and dates.
+
+\begin{enumerate}
+\item Prove all applications of the dominated convergence theorem.
+\item Implement the algorithm as a more specialised version of Algorithm 2 in
+  the black box VI paper.
+\item <Evaluation>
+\item Experiment with various improvements mentioned in the black box VI paper,
+  other VI papers, especially the ones about GPs.
+\item Write/finalise the paper.
+\end{enumerate}
+
+\subsection{Evaluation}
+
+In order to evaluate our approach, we compare it to two related alternatives:
+the original GP IRL paper \cite{DBLP:conf/nips/LevinePK11} and the deep GP
+extension \cite{DBLP:conf/uai/JinDAS17}. To the best of our knowledge,... TODO: continue
+Against what? GPIRL and DGP-IRL.
+
+For what problems? GPIRL paper has a big toolkit, but with MATLAB.
+\begin{itemize}
+\item Object world. Fix one, vary the other: number of colours, number of
+  examples.
+\item Highway driving behaviour: both synthetic and human performance.
+\end{itemize}
+
+Using what metrics?
+\begin{itemize}
+\item Visualise the (true and learned) rewards as shades of grey.
+\item Also measure expected value difference in four situations:
+  \begin{itemize}
+  \item discrete features
+  \item discrete feature transfer
+  \item continuous features
+  \item continuous feature transfer
+  \end{itemize}
+\item Expected value difference over time rather than #examples (since VI tends
+  to be faster than MCMC).
+\end{itemize}
 

### 2018-11-16 mpaper/paper.tex
Making the proposal more readable and moving the derivatives out of the proposal

\documentclass{mpaper}

\begin{document}

\title{Variational Inference for Inverse Reinforcement Learning with Gaussian Processes}
\author{Paulius Dilkas}
\matricnum{2146879}

\maketitle

\begin{abstract}
According to Simon Peyton Jones, an abstract should address
four key questions. First, what is the problem that this
paper tackles? Second, why is this an interesting problem?
Third, what is the solution this paper proposes?
Finally, why is the proposed solution a good one?
\end{abstract}

\section{Introduction}

This paper outlines the standard template for an MSci submission.
In earlier years, MSci students at the School of Computing
Science\footnote{\url{http://www.dcs.gla.ac.uk}},
University of Glasgow, were expected to produce a full-length
dissertation. Now, the requirement is for MSci students to
write a paper of up to 14 pages in length, using the supplied
\texttt{mpaper} \LaTeX style file.

The precise structure of an MSci paper is not mandated, but it should
probably cover in detail the following aspects of the project.
\begin{enumerate}
\item General description of the problem, motivation, relevance
\item Background information, possibly including a literature survey
\item Description of approach taken to solve the problem, including
  high-level design and lower-level implementation details as appropriate
\item Evaluation, qualitative or quantitative as appropriate
\item Conclusion, including scope for future work
\end{enumerate}

\section{Background}

This \LaTeX template is based on the ACM \texttt{sig-alternate} class.
The layout is two-column text. Generally figures and tables only
extend to one column width, e.g.\ Table \ref{tab-eg},
but it is possible to make them
stretch over both columns using the \texttt{figure*} and
\texttt{table*} environments. For an example, see Figure \ref{fig-eg}.

\begin{table}
\begin{tabular}{l||c||p{2cm}}
\emph{Operating System} & \emph{Version} & \emph{Verdict} \\ \hline \hline
Ubuntu & 12.04 & Everyone's favourite Linux, unless you grew up with
RedHat \\ \hline
Slackware & xxx & Pseudo-hacker's Linux, how often do you recompile
your kernel? \\ \hline
Mac OS & 10.7 & For people with more money than sense \\ \hline
\end{tabular}
\caption{\label{tab-eg}Single column table of figures}
\end{table}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.3]{alice.pdf}
\end{center}
\caption{\label{fig-eg}An example figure stretching over two columns}
\end{figure*}

\section{The WizWoz System}

Again, Simon Peyton Jones has a lovely description of how to write a
paper on his
website\footnote{\url{http://research.microsoft.com/en-us/um/people/simonpj/papers/giving-a-talk/giving-a-talk.htm}}.
Personally, I put URLs in footnotes and \emph{bona fide} references
in the bibliography. How many references? Hard to say. Five is not enough, 50 is
pushing it.

\subsection{User Interface}

Blah blah blah
Blah blah blah
Blah blah blah
Blah blah blah

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Foo}

Blah blah blah
Blah blah blah
Blah blah blah

\subsubsection{Bar}
\textsf{Blah} \textit{blah} \textbf{blah}

\section{Evaluation}

Graphs are always good. I recommend getting to grips with Matlab, R or
gnuplot rather than exporting horribly Excel bitmapped graphs.

\begin{quotation}
 The Assyrian came down like the wolf on the fold,
 And his cohorts were gleaming in purple and gold;
 And the sheen of their spears was like stars on the sea,
 When the blue wave rolls nightly on deep Galilee.

 Like the leaves of the forest when Summer is green,
 That host with their banners at sunset were seen:
 Like the leaves of the forest when Autumn hath blown,
 That host on the morrow lay withered and strown.
\end{quotation}

\section{Conclusions}

The standard Lorem Ipsum passage, used since the 1500s

``Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.''
\vskip8pt \noindent
{\bf Acknowledgments.}
This is optional; it is a location for you to thank people

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}

### 2018-11-16 mprop/mprop.tex
Making the proposal more readable and moving the derivatives out of the proposal

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 \theoremstyle{definition}
@@
 
-\newcommand{\dm}{\frac{\partial}{\partial\mathbf{m}}}
-\newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
-\newcommand{\da}{\frac{\partial}{\partial\alpha_j}}
-\newcommand{\db}{\frac{\partial}{\partial\beta_j}}
-
 \section{Introduction}
 %briefly explain the context of the project problem
-% TODO: introduce a data set
-% TODO: example application for motivation (why would the world be better)
-% TODO: introduce research questions
 
-Inverse reinforcement learning (IRL)---a problem proposed by Russell in 1998
-\cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a Markov
-decision process that best explains a set of given demonstrations. IRL is
+%Imagine a scenario where the agent is driving on a three-lane highway. The
+%features of the environment include the agent's speed, lane number, and
+%various distances to different types of vehicles in front of and behind the
+%agent. The agent can regulate its speed and position. The main objective is, of
+%course, to not crash into other vehicles, but one could also add objectives such
+%as staying within the speed limit when close to a police vehicle.
+
+Inverse reinforcement learning (IRL)---a problem proposed by Russell in
+1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
+Markov decision process that best explains a set of given demonstrations. IRL is
 important because reward functions can be hard to define manually
@@
 Moreover, IRL has seen a wide array of applications in autonomous vehicle
-control
-\cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16}
-and learning to predict another agent's behaviour
+control \cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} and
+learning to predict another agent's behaviour
 \cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
@@
 as a linear combination of features. One proven way to abandon this assumption
-is by representing the reward function as a Gaussian process
+is by representing the reward function as a Gaussian process (GP)
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
+Our main goal is employ variational inference (VI) for recovering the reward
+function, which can prove useful in two ways:
+\begin{enumerate}
+\item As VI tends to be faster than Markov chain Monte Carlo
+  sampling \cite{blei2017variational}, we should be able to handle more data.
+\item Modelling full posterior distributions for various parameters can result
+  in more precise reward predictions, as the model simply holds more
+  information.
+\end{enumerate}
 
@@
 \cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
-variational inference to approximate the full posterior distribution
-$\posterior$. \emph{Variational inference} (VI) is an approximation technique
-for probability densities \cite{blei2017variational}. Let $\approximation$ be
-our approximating family of probability distributions for $\posterior$ with its
-own hyperparameter vector $\bm\nu$. Then the job of VI algorithms is to optimise
-$\bm\nu$ in order to minimise the \emph{Kullback-Leibler} (KL) divergence
-between the original probability distribution and our approximation.  KL
-divergence (asymmetrically) measures how different the two distributions are,
-and in this case can be defined as \cite{blei2017variational}
+VI to approximate the full posterior distribution $\posterior$.
+\emph{Variational inference} is an approximation technique for probability
+densities \cite{blei2017variational}. Let $\approximation$ be our approximating
+family of probability distributions for $\posterior$ with its own hyperparameter
+vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
+to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
+probability distribution and our approximation.  KL divergence (asymmetrically)
+measures how different the two distributions are, and in this case can be
+defined as \cite{blei2017variational}
 \[ \begin{split}
@@
 \end{equation}
+% TODO: are they actually Lebesgue? What about notation?
 
@@
 \ref{literature}), making the resulting algorithm competitive both in terms of
-running time and model fit.
+running time and model fit. Therefore, we raise the following research
+questions\footnote{For each question, we are primarily comparing the original GP
+IRL approach \cite{DBLP:conf/nips/LevinePK11} with our variational version.}:
+\begin{enumerate}
+\item Does VI allow us to make more accurate reward predictions when given the
+  same amount of training time?
+\item How does the reward prediction accuracy differ between VI and sampling
+  approaches when given the same amount of training data?
+\item How well does our VI model approximate the real posterior distribution?
+\item Is VI better at learning behaviour driven by multiple goals?
+\item Is VI more robust at handling mistakes in demonstrations?
+\end{enumerate}
 
@@
 
+Finally, instead of using GPs to model nonlinear reward functions, one can use a
+neural network (NN), as demonstrated by Wulfmeier et al.
+\cite{wulfmeier2015maximum}. Their approach benefits from constant time
+inference and the ability to learn complex features either from already-given
+features or even from raw data. The only disadvantage (as demonstrated in the
+paper) is that NNs take longer to learn compared to GPs.
+
 %A different take on IRL with GPs \cite{DBLP:journals/corr/abs-1208-2112}
@@
 \emph{stochastic variational sparse approximations}, as they have achieved
-modelling accuracy close to that of the full GP with no approximations, while
-providing a time complexity of $\mathcal{O}(m^3)$. Below we provide a short
-overview of various assumptions that have been used in approximating
+modelling accuracy close to that of the full GP with no approximations, with
+many methods providing a time complexity of $\mathcal{O}(m^3)$. Below we provide
+a short overview of various assumptions that have been used in approximating
 \emph{sparse} GPs (i.e., GPs that use inducing points), following on a paper by
@@
 
-% TODO: expand on this
 Table \ref{table:approximations} further summarises some of the recent and/or
@@
 fixed or modelled by a probability distribution (none of the approaches do
-that), while hyperparameters cannot be variational (the reason will be explained
-in Section \ref{sec:proposed_approach}).
+that), while hyperparameters cannot be variational (the reason will become clear
+in Section \ref{sec:proposed_approach}). We would also like to avoid having the
+hyperparameters fixed, as learning them efficiently would introduce a separate
+problem. As the approach by Hoang et al. does not seem to be easily extendable
+to modelled hyperparameters, it is unsuitable to our needs.
 
-\begin{itemize}
-\item We would also like to avoid having the hyperparameters fixed, as learning
-  them efficiently would introduce a separate problem. As the approach by Hoang
-  et al. does not seem to be easily extendable to modelled hyperparameters, it
-  is unsuitable to our needs.
-\item (TODO: this will have comments about other papers as well and will be
-  restructured into a paragraph) The variational Fourier features (VFF)
-  algorithm by Hensman et al. \cite{DBLP:journals/jmlr/HensmanDS17} is only
-  defined for Mat\'ern kernels, which is likely to be too restrictive for our
-  situation (the original paper on using GPs for IRL
-  \cite{DBLP:conf/nips/LevinePK11} used the automatic relevance detection kernel
-  that has weights controlling how important each feature is). While extending
-  VFF to support a flexible class of kernels defined by Wilson and Adams
-  \cite{pmlr-v28-wilson13} is an interesting and promising avenue of work, it is
-  likely to be beyond the scope of this project.
-\item The derivation of the ELBO in the work of Peng et al.
+The variational Fourier features (VFF) algorithm by Hensman et al.
+\cite{DBLP:journals/jmlr/HensmanDS17} is only defined for Mat\'ern kernels,
+which is likely to be too restrictive for our situation (the original paper on
+using GPs for IRL \cite{DBLP:conf/nips/LevinePK11} used the automatic relevance
+detection kernel that has weights controlling how important each feature is).
+While extending VFF to support a flexible class of kernels defined by Wilson and
+Adams \cite{pmlr-v28-wilson13} is an interesting and promising avenue of work,
+it is likely to be beyond the scope of this project.
+
+The derivation of the ELBO in the work of Peng et al.
 \cite{DBLP:conf/icml/PengZZQ17} relies primarily on the fact that the evidence
-  for a GP is a Gaussian, while in our case the evidence can be defined in
-  several ways depending on the model, but must always involve $\mathcal{D}$,
-  making the main idea of the paper inapplicable to IRL.
-\end{itemize}
+for a GP is a Gaussian, while in our case the evidence can be defined in several
+ways depending on the model, but must always involve $\mathcal{D}$, making the
+main idea of the paper inapplicable to IRL. Remaining papers can be rejected for
+similar reasons, as they are all highly dependent on the specific situation of
+approximating a GP regression. Thus we cannot simply apply a GP VI model as a
+whole, although there might be some smaller ideas that we can use (some of which
+will be mentioned in the next section).
 
-Since we expect $\posterior$ to be highly irregular, we would like our
+Finally, since we expect $\posterior$ to be highly irregular, we would like our
 approximation to be capable of representing a wide range of possible probability
-distributions. The primary way to represent complex posteriors in VI is by using
-\emph{normalising flows}, i.e., a collection of invertible
+distributions. The primary way of representing complex posteriors in VI is by
+using \emph{normalising flows}, i.e., a collection of invertible
 functions---parametrised by additional variational parameters---that are applied
@@
 parametrisation also means that the gradient of the joint probability
-distribution w.r.t. variational parameters $\nabla_{\bm\nu}\pfull$ is no
-longer zero, making an analytic expression for ELBO impossible using the usual
+distribution w.r.t. variational parameters $\nabla_{\bm\nu}\pfull$ is no longer
+zero, making an analytic expression for the ELBO impossible using the usual
 methods. While it may be possible to have an approximating distribution for the
-flow parameters, it is uncertain how such an algorithm would behave, as it
-would have to perform optimisation in a space with significantly more dimensions.
+flow parameters, it is uncertain how such an algorithm would behave, as it would
+have to perform optimisation in a space with significantly more dimensions.
 
@@
 
-\subsection{The Structure of the Approximating Distribution}
+\subsection{Structure of the Approximating Distribution}
 
@@
 
-\subsection{Derivatives}
-% TODO: some kind of intro
+\subsection{Variational Inference} \label{vi_algs}
 
-\subsubsection{$\partial/\partial\mathbf{m}$}
-
-We begin by removing terms independent of $\mathbf{m}$:
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
-    &- \frac{1}{2}\dm[\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}]
-    + \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] \\
-    &- \sum_{i=1}^N \sum_{t=1}^T \dm\mathbb{E}[V_r(s_{i,t})] -
-      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-      s')\dm\mathbb{E}[V_r(s')].
-  \end{split}
-\]
-Here
-\begin{align*}
-  \dm[\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}] &= (\mathbb{E}[\Kuu^{-1}] + \mathbb{E}[\Kuu^{-1}]^\intercal)\mathbf{m}, \\
-  \dm[\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}] &= \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}],
-\end{align*}
-and
-\begin{equation} \label{eq:1}
-  \begin{split}
-    \dm\mathbb{E}[V_r(s)] &= \dm\iiint V_r(s) p(\mathbf{r} | \bm\lambda,
-    \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
-    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
-    &= \iiint V_r(s) p(\mathbf{r} | \bm\lambda,
-    \mathbf{X_u}, \mathbf{u}) \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
-    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda,
-  \end{split}
-\end{equation}
-where
-\[
-  \begin{split}
-    \dm\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) &=
-    \mathcal{N}(\mathbf{u}; \mathbf{m},
-    \mathbf{S})\dm\left[-\frac{1}{2}(\mathbf{u} -
-      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right] \\
-    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
-    \mathbf{S})\left(-\frac{1}{2}\right)(\mathbf{S}^{-1} +
-    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})\dm[\mathbf{u} -
-    \mathbf{m}] \\
-    &= \mathcal{N}(\mathbf{u}; \mathbf{m},
-    \mathbf{S})\frac{1}{2}(\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u}
-    - \mathbf{m}).
-  \end{split}
-\]
-Substituting it back into \eqref{eq:1} gives
-\[
-  \begin{split}
-    \dm\mathbb{E}[V_r(s)] &= \frac{1}{2}\iiint V_r(s)
-    (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})
-    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \mathcal{N}(\mathbf{u};
-    \mathbf{m}, \mathbf{S})
-    q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda \\
-    &= \frac{1}{2}\mathbb{E}[V_r(s) (\mathbf{S}^{-1} +
-    \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})].
-\end{split}
-\]
-Hence
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\mathbf{m}} =
-    &-\frac{1}{2}(\mathbb{E}[\Kuu^{-1}] +
-    \mathbb{E}[\Kuu^{-1}]^\intercal)\mathbf{m} +
-    \mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}] \\
-    &- \frac{1}{2}\sum_{i=1}^N\sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})
-    (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} - \mathbf{m})] \\
-    &- \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')
-    \mathbb{E}[V_r(s') (\mathbf{S}^{-1} + \mathbf{S}^{-\intercal})(\mathbf{u} -
-    \mathbf{m})].
-\end{split}
-\]
-
-\subsubsection{$\partial/\partial\mathbf{S}$}
-
-Similarly to the previous section,
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
-    \frac{1}{2}\dS\log|\mathbf{S}|
-    -\frac{1}{2}\dS\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\
-    &- \sum_{i=1}^N \sum_{t=1}^T \dS\mathbb{E}[V_r(s_{i,t})] -
-      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-      s')\dS\mathbb{E}[V_r(s')],
-  \end{split}
-\]
-where
-\[ \dS\log|\mathbf{S}| = \mathbf{S}^{-\intercal}, \]
-and
-\[ \dS\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] = \mathbb{E}[\Kuu^{-1}]^\intercal \]
-by \emph{The Matrix Cookbook} \cite{petersen2008matrix}. Then
-\[ \dS\mathbb{E}[V_r(s)] = \iiint V_r(s) q(\mathbf{r}) \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
-  q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda, \]
-where
-\[
-  \begin{split}
-    \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) &=
-    \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\exp \left( -\frac{1}{2}
-        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
-        \mathbf{m})\right)\right] \\
-    &= \dS\left[\frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\right]\exp \left( -\frac{1}{2}
-        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
-        \mathbf{m})\right) \\
-      &+ \frac{1}{(2\pi)^{m/2}|\mathbf{S}|^{1/2}}\dS\left[\exp\left( -\frac{1}{2}
-        (\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
-        \mathbf{m})\right)\right] \\
-    &=
-    \frac{1}{(2\pi)^{m/2}}\dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right]\exp
-    \left( -\frac{1}{2} (\mathbf{u} -
-      \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})\right) \\
-    &- \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})\dS[(\mathbf{u}
-    - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} - \mathbf{m})]. \\
-  \end{split}
-\]
-The two remaining derivatives can be taken with the help of \emph{The Matrix
-  Cookbook} \cite{petersen2008matrix}:
-\begin{align*}
-  \dS\left[\frac{1}{|\mathbf{S}|^{1/2}}\right] =
-  -\frac{1}{2}|\mathbf{S}|^{-3/2}\frac{\partial |\mathbf{S}|}{\partial \mathbf{S}} =
-  -\frac{1}{2}|\mathbf{S}|^{-3/2}|\mathbf{S}|\mathbf{S}^{-\intercal} &=
-  -\frac{1}{2|\mathbf{S}|^{1/2}}\mathbf{S}^{-\intercal}, \\
-  \dS[(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-1}(\mathbf{u} -
-  \mathbf{m})] &= -\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
-  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}.
-\end{align*}
-Plugging them back in gives
-\[ \dS\mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}) =
-  -\frac{1}{2}\mathbf{S}^{-\intercal}\mathcal{N}(\mathbf{u};
-  \mathbf{m}, \mathbf{S}) + \frac{1}{2}\mathcal{N}(\mathbf{u}; \mathbf{m},
-  \mathbf{S})\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
-  \mathbf{m})^\intercal\mathbf{S}^{-\intercal}, \]
-and
-\[
-  \begin{split}
-    \dS\mathbb{E}[V_r(s)] &= \frac{1}{2}\iiint V_r(s)
-    (\mathbf{S}^{-\intercal}(\mathbf{u} - \mathbf{m})(\mathbf{u} -
-    \mathbf{m})^\intercal\mathbf{S}^{-\intercal} - \mathbf{S}^{-\intercal})
-    q(\mathbf{r}) \mathcal{N}(\mathbf{u};
-    \mathbf{m}, \mathbf{S}) q(\bm\lambda)\,d\mathbf{r}\,d\mathbf{u}\,d\bm\lambda
-    \\
-    &= \frac{1}{2}\mathbb{E}[V_r(s)(\mathbf{S}^{-\intercal}(\mathbf{u} -
-    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
-    \mathbf{S}^{-\intercal})].
-  \end{split}
-\]
-Therefore
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\mathbf{S}} &=
-    \frac{1}{2}\mathbf{S}^{-\intercal} -
-    \frac{1}{2}\mathbb{E}[\Kuu^{-1}]^\intercal - \frac{1}{2}
-    \sum_{i=1}^N\sum_{t=1}^T
-    \mathbb{E}[V_r(s_{i,t})(\mathbf{S}^{-\intercal}(\mathbf{u} -
-    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
-    \mathbf{S}^{-\intercal})] \\
-    &- \gamma\sum_{s' \in \mathcal{S}}\mathcal{T}(s_{i,t}, a_{i,t}, s')
-    \mathbb{E}[V_r(s')(\mathbf{S}^{-\intercal}(\mathbf{u} -
-    \mathbf{m})(\mathbf{u} - \mathbf{m})^\intercal\mathbf{S}^{-\intercal} -
-    \mathbf{S}^{-\intercal})].
-  \end{split}
-\]
-
-\subsubsection{$\partial/\partial\alpha_j$}
-
-We begin in the usual way:
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\alpha_j} = &-
-    \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-2}]] - \frac{1}{2}\da\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]] - \frac{1}{2}\da\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] - \frac{1}{2}\da\mathbb{E}[\log|\Kuu|] \\
-    &+ \da\mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] +
-    \da [\alpha_j + \log\Gamma(\alpha_j) + (1 - \alpha_j)\psi(\alpha_j)] \\
-      &- \sum_{i=1}^N \sum_{t=1}^T \da\mathbb{E}[V_r(s_{i,t})] -
-        \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-        s')\da\mathbb{E}[V_r(s')].
-  \end{split}
-\]
-First,
-\[ \da [\alpha_j + \log\Gamma(\alpha_j) + (1 - \alpha_j)\psi(\alpha_j)] = 1 +
-  \psi(\alpha_j) - \psi(\alpha_j) + (1 - \alpha_j)\psi'(\alpha_j) = 1 + (1 -
-  \alpha_j)\psi'(\alpha_j) \]
-by the definition of $\psi$. The remaining terms can all be treated in the same
-way, as they all contain expectations of scalar functions that are independent
-of $\alpha_j$, and $\alpha_j$ only occurs in $\Gamma(\lambda_j; \alpha_j,
-\beta_j)$. Thus we can work with an abstract function as follows:
-\[
-  \begin{split}
-    \da\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \da \iiint f(k_{\bm\lambda},
-    \mathbf{r}) q(\bm\lambda) q(\mathbf{r})
-    q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
-    &= \!\begin{multlined}[t]
-      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1}) \da
-      \left[ \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)} \lambda_j^{\alpha_j -
-          1} \right] e^{-\beta_j\lambda_j} \\
-      q(\lambda_{j+1}) \cdots q(\lambda_d) q(\mathbf{r})
-      q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}.
-    \end{multlined}
-  \end{split}
-\]
-Then
-\[
-  \begin{split}
-    \da \left[ \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)} \lambda_j^{\alpha_j - 1} \right] &=
-    \frac{\da[\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}]\Gamma(\alpha_j) -
-      \beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}\Gamma'(\alpha_j)}{(\Gamma(\alpha_j))^2} \\
-    &= \frac{\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}\da[\alpha_j\log\beta_j + (\alpha_j -
-      1)\log\lambda_j]\Gamma(\alpha_j) - \beta_j^{\alpha_j}\lambda_j^{\alpha_j -
-        1}\Gamma'(\alpha_j)}{(\Gamma(\alpha_j))^2} \\
-    &= \frac{\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}(\log\beta_j + \log\lambda_i)\Gamma(\alpha_j)
-      - \beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}\Gamma'(\alpha_j)}{(\Gamma(\alpha_j))^2} \\
-    &= \frac{\beta_j^{\alpha_j}\lambda_j^{\alpha_j - 1}}{\Gamma(\alpha_j)}\left(\log\beta_j +
-    \log\lambda_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)}\right),
-  \end{split}
-\]
-which means that
-\[
-  \begin{split}
-    \da\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \!\begin{multlined}[t]
-      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
-      \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)}\lambda_j^{\alpha_j -
-        1}e^{-\beta_j\lambda_j} \left(\log\beta_j + \log\lambda_j -
-        \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) \\
-      q(\lambda_{j+1}) \cdots q(\lambda_d)
-      q(\mathbf{r})q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}
-    \end{multlined} \\
-    &= \iiint f(k_{\bm\lambda}, \mathbf{r}) \left(\log\beta_j + \log\lambda_j -
-      \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right) q(\bm\lambda)
-    q(\mathbf{r})q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
-    &= \mathbb{E} \left[ f(k_{\bm\lambda}, \mathbf{r}) \left(\log\beta_j +
-        \log\lambda_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right)
-    \right] \\
-    &= \left( \log\beta_j - \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)}
-    \right)\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] +
-    \mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})\log\lambda_j].
-  \end{split}
-\]
-With these results in mind, we can simplify the initial expression to
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\alpha_j} &= 1 + (1 -
-    \alpha_j)\psi'(\alpha_j) + \left( \log\beta_j -
-      \frac{\Gamma'(\alpha_j)}{\Gamma(\alpha_j)} \right)
-    \left(\vphantom{\sum_{i=1}^N} \right. - \frac{1}{2}
-    \mathbb{E}[\Tr[\Kuu^{-2}]] - \frac{1}{2}\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]]
-    \\
-    &- \frac{1}{2}\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] -
-    \frac{1}{2}\mathbb{E}[\log|\Kuu|] +
-    \mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] \\
-    &- \sum_{i=1}^N \sum_{t=1}^T \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s' \in
-      \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}[V_r(s')] \left.
-      \vphantom{\sum_{i=1}^N} \right) \\
-    & - \frac{1}{2}\mathbb{E}[\Tr[\Kuu^{-2}]\log\lambda_j] -
-    \frac{1}{2}\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]\log\lambda_j] -
-    \frac{1}{2}\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}\log\lambda_j]
-    \\
-    &- \frac{1}{2}\mathbb{E}[\log|\Kuu|\log\lambda_j] +
-    \mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}\log\lambda_j]
-    \\
-    &- \sum_{i=1}^N\sum_{t=1}^T\mathbb{E}[V_r(s_{i,t})\log\lambda_j] -
-    \gamma\sum_{s' \in \mathcal{S}}\mathcal{T}(s_{i,t}, a_{i,t},
-    s')\mathbb{E}[V_r(s')\log\lambda_j].
-  \end{split}
-\]
-
-\subsubsection{$\partial/\partial\beta_j$}
-Finally,
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\beta_j} = &-
-    \frac{1}{2}\db\mathbb{E}[\Tr[\Kuu^{-2}]] -
-    \frac{1}{2}\db\mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}]] -
-    \frac{1}{2}\db\mathbb{E}[\mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] \\
-    &- \frac{1}{2}\db\mathbb{E}[\log|\Kuu|] +
-    \db\mathbb{E}[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}] -
-    \db[\log\beta_j] \\
-    &- \sum_{i=1}^N \sum_{t=1}^T \db \mathbb{E}[V_r(s_{i,t})] - \gamma\sum_{s'
-      \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\db\mathbb{E}[V_r(s')].
-  \end{split}
-\]
-Similarly to the previous section, we can handle all derivatives of expectations
-in the same way:
-\[
-  \begin{split}
-    \db\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \db \iiint f(k_{\bm\lambda},
-    \mathbf{r}) q(\bm\lambda) q(\mathbf{r})
-    q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
-    &= \!\begin{multlined}[t]
-      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
-      \frac{\lambda_j^{\alpha_j - 1}}{\Gamma(\alpha_j)} \db [\beta_j^{\alpha_j}
-      e^{-\beta_j\lambda_j}] \\
-      q(\lambda_{j+1}) \cdots q(\lambda_d) q(\mathbf{r})
-      q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}.
-    \end{multlined}
-  \end{split}
-\]
-Since
-\[ \db [\beta_j^{\alpha_j} e^{-\beta_j\lambda_j}] = \alpha_j\beta_j^{\alpha_j - 1}e^{-\beta_j\lambda_j}
-  - \beta_j^{\alpha_j} e^{-\beta_j\lambda_j}\lambda_j = \beta_j^{\alpha_j}
-  e^{-\beta_j\lambda_j} \left( \frac{\alpha_j}{\beta_j} - \lambda_j \right), \]
-we have that
-\[
-  \begin{split}
-    \db\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] &= \!\begin{multlined}[t]
-      \iiint f(k_{\bm\lambda}, \mathbf{r}) q(\lambda_0) \cdots q(\lambda_{j-1})
-      \frac{\beta_j^{\alpha_j}}{\Gamma(\alpha_j)}\lambda_j^{\alpha_j -
-        1}e^{-\beta_j\lambda_j} \left(\frac{\alpha_j}{\beta_j} - \lambda_j
-      \right) \\
-      q(\lambda_{j+1}) \cdots q(\lambda_d) q(\mathbf{r})
-      q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u}
-    \end{multlined}
-    \\
-    &= \iiint f(k_{\bm\lambda}, \mathbf{r}) \left(\frac{\alpha_j}{\beta_j} -
-      \lambda_j \right) q(\bm\lambda) q(\mathbf{r})
-    q(\mathbf{u})\,d\bm\lambda\,d\mathbf{r}\,d\mathbf{u} \\
-    &= \mathbb{E} \left[ f(k_{\bm\lambda}, \mathbf{r})
-      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] =
-    \frac{\alpha_j}{\beta_j}\mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})] -
-    \mathbb{E}[f(k_{\bm\lambda}, \mathbf{r})\lambda_j].
-  \end{split}
-\]
-This gives us the final expression of $\frac{\partial\mathcal{L}}{\partial\beta_j}$:
-\[
-  \begin{split}
-    \frac{\partial\mathcal{L}}{\partial\beta_j} = &- \frac{1}{\beta_j} -
-    \frac{1}{2}\mathbb{E} \left[ \Tr[\Kuu^{-2}] \left( \frac{\alpha_j}{\beta_j}
-        - \lambda_j \right) \right] - \frac{1}{2}\mathbb{E} \left[
-      \Tr[\Kuu^{-1}\mathbf{S}] \left(\frac{\alpha_j}{\beta_j} - \lambda_j
-      \right) \right] \\
-    &- \frac{1}{2}\mathbb{E} \left[ \mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}
-      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] -
-    \frac{1}{2}\mathbb{E} \left[ \log|\Kuu| \left(\frac{\alpha_j}{\beta_j} -
-        \lambda_j \right) \right] \\
-    &+ \mathbb{E} \left[ \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\mathbf{m}
-      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] \\
-    &- \sum_{i=1}^N \sum_{t=1}^T \mathbb{E} \left[ V_r(s_{i,t})
-      \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right] -
-    \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\mathbb{E}
-    \left[V_r(s') \left(\frac{\alpha_j}{\beta_j} - \lambda_j \right) \right].
-  \end{split}
-\]
-
-\subsection{Variational Inference Algorithms} \label{vi_algs}
-
-% TODO: integrate this into the previous subsection
-% TODO: will need to make this equation more precise, once I know what
-% approximation I want to use
 The typical way to optimise a quantity (the ELBO, in this case) involves
-computing its gradient \cite{blei2017variational}. Due to terms involving
-$\mathcal{D}$ that are computed by solving an MDP and do not resemble a typical
-probability distribution, we turn our attention to the \emph{black box
-  variational inference} \cite{DBLP:conf/aistats/RanganathGB14} paper that
-suggests a convenient trick:
+computing its gradient \cite{blei2017variational}. Unfortunately, some of the
+terms in $\mathcal{L}$ are still left as expected values. \emph{Black box
+  variational inference} (BBVI) \cite{DBLP:conf/aistats/RanganathGB14} suggests
+a way to express the gradient as an expectation without having to take the
+gradient of the posterior:
 \[ \nabla_{\bm\nu} \mathcal{L} = \Eq[\nabla_{\bm\nu} \log \approximation
   (\log \pfull - \log \approximation)]. \]
-With this trick in mind, we only need to evaluate $\log \pfull$ and take the
-gradient of $\log \approximation$. Following the same paper,
-$\nabla_{\bm\nu} \mathcal{L}$ then has an unbiased estimate
+The gradient of the ELBO then has a unbiased estimate
 \[ \nabla_{\bm\nu} \mathcal{L} \approx \frac{1}{S} \sum_{s=1}^S
@@
 \mathbf{r}_s) \sim \approximation$.
+
+The main theorem behind the BBVI trick is the Lebesgue's dominated convergence
+theorem \cite{royden1988real}, which---under some conditions---allows one to
+claim that:
+\[ \lim_{n \to \infty} \int_E f_n = \int_E \lim_{n \to \infty} f_n. \]
+We can use the same idea on parts of $\mathcal{L}$ that we cannot take the
+derivative of otherwise. In this case, we want a gradient instead of a limit,
+and the integral represents an expected value.
 
@@
   the black box VI paper.
-\item <Evaluation>
+\item Evaluation
 \item Experiment with various improvements mentioned in the black box VI paper,
@@
 
-In order to evaluate our approach, we compare it to two related alternatives:
-the original GP IRL paper \cite{DBLP:conf/nips/LevinePK11} and the deep GP
-extension \cite{DBLP:conf/uai/JinDAS17}. To the best of our knowledge,... TODO: continue
-Against what? GPIRL and DGP-IRL.
+We would like to compare our approach to alternatives that support nonlinear
+reward functions. According to a recent survey
+\cite{DBLP:journals/corr/abs-1806-06877}, this leaves us with five algorithms:
+\textsc{LEARCH} \cite{DBLP:journals/arobots/RatliffSB09}, \textsc{MmpBoost}
+\cite{DBLP:conf/nips/RatliffBBC06}, the original GP IRL paper
+\cite{DBLP:conf/nips/LevinePK11}, its deep GP extension
+\cite{DBLP:conf/uai/JinDAS17}, and a neural network reward function
+approximation \cite{wulfmeier2015maximum}. As the first two have already been
+shown to underperform \cite{DBLP:conf/nips/LevinePK11}, we will focus on the
+remaining three.
 
-For what problems? GPIRL paper has a big toolkit, but with MATLAB.
-\begin{itemize}
-\item Object world. Fix one, vary the other: number of colours, number of
-  examples.
-\item Highway driving behaviour: both synthetic and human performance.
-\end{itemize}
+The algorithms will be compared on variations of two commonly used fictional
+scenarios: object world \cite{DBLP:conf/nips/LevinePK11} and highway driving
+behaviour \cite{DBLP:conf/icml/PieterN04,DBLP:conf/nips/LevinePK10}. While more
+recent papers (especially those using deep learning) often use the binary world
+benchmark as well \cite{wulfmeier2015maximum,DBLP:conf/uai/JinDAS17}, our
+approach is not able to construct new features, and thus cannot perform well on
+such a task.
 
-Using what metrics?
-\begin{itemize}
-\item Visualise the (true and learned) rewards as shades of grey.
-\item Also measure expected value difference in four situations:
-  \begin{itemize}
-  \item discrete features
-  \item discrete feature transfer
-  \item continuous features
-  \item continuous feature transfer
-  \end{itemize}
-\item Expected value difference over time rather than #examples (since VI tends
-  to be faster than MCMC).
-\end{itemize}
-
-\section{Notes on papers (to be removed)}
-
-\subsection{Miscellaneous}
-
-(Directed) similarity between MDPs using restricted Boltzmann machines
-\cite{9401f4eeb9a64c77afb3d087261d1080}
-
-Chapter 6 on distance measures \cite{mccune2002analysis}
-
-The PhD thesis behind maximum causal entropy \cite{Ziebart:2010:MPA:2049078}
-
-\subsection{Gaussian Processes}
-
-Simple introduction to GPs for time-series modelling \cite{Roberts2013GaussianPF}
-
-GPs over graphs instead of vectors (haven't actually read)
-\cite{DBLP:journals/corr/abs-1803-05776}
-
-Another introduction from physics (skimmed through) \cite{introduction_to_gps}
-
-Learning a GP from very little data \cite{DBLP:conf/nips/PlattBSWZ01}
-
-One GP for multiple correlated output variables \cite{DBLP:journals/jcphy/BilionisZKL13}
-
-Kernels for categorical and count data \cite{savitsky2011variable}
-
-Scalability/Approximations thesis \cite{kth}
-
-\subsection{Interpretability}
-
-Learning latent factors \cite{DBLP:conf/nips/LiSE17}
-
-The behaviour of Reddit users \cite{DBLP:conf/atal/DasL14}
-
-\subsection{Inverse Reinforcement Learning}
-
-One of the first papers on the topic \cite{DBLP:conf/icml/NgR00}
-
-Bayesian setting \cite{DBLP:conf/ijcai/RamachandranA07}
-
-Learning optimal composite features \cite{DBLP:conf/ijcai/ChoiK13}
-
-A different take on IRL with GPs \cite{DBLP:journals/corr/abs-1208-2112}
-
-IRL for large state spaces (haven't read) \cite{DBLP:journals/jmlr/BoulariasKP11}
-
-Multiple reward functions \cite{DBLP:conf/nips/ChoiK12}
-
-A recent survey \cite{DBLP:journals/corr/abs-1806-06877}
-
-Some not-very-successful method \cite{DBLP:conf/uai/NeuS07}
-
-\subsubsection{Multiple Strategies}
-
-EM clustering \cite{DBLP:conf/icml/BabesMLS11}
-
-Structured priors \cite{DBLP:conf/ewrl/DimitrakakisR11}
-
-There are more, but I haven't gotten to them yet.
-
-\subsection{Variational Inference}
-
-Part IV on probabilities and inference \cite{MacKay:2002:ITI:971143}
-
-Stochastic VI \cite{DBLP:journals/jmlr/HoffmanBWP13}
-
-Structured stochastic VI \cite{DBLP:conf/aistats/HoffmanB15}
-
-Another review of recent advances \cite{DBLP:journals/corr/abs-1711-05597}
-
-Tighter ELBOs are not necessarily better \cite{DBLP:conf/icml/RainforthKLMIWT18}
-
-For details on Lebesgue's dominated convergence theorem \cite{cinlar_probability}
-
-Still looking for the relevant version of the theorem \cite{knill2017probability}
-
-Approximation as a multivariate Gaussian (haven't read) \cite{DBLP:journals/sac/TanN18}
-
-How black box VI can be even more black box \cite{li2016wild}
-
-Evaluating VI \cite{DBLP:conf/icml/YaoVSG18}
-
-\subsubsection{for GPs}
-
-Linear VI for GPs \cite{DBLP:conf/nips/ChengB17}
-
-Sparse VI for GP \cite{DBLP:journals/jmlr/HensmanDS17}
-
-Stochastic VI for sparse spectral GPs (no inducing points, so not that relevant) \cite{DBLP:conf/aaai/HoangHL17}
-
-Sparse GPs 2 \cite{DBLP:conf/nips/GalWR14}
-
-The 3 papers I need to focus on:
-
-SVI for sparse GPs \cite{DBLP:conf/uai/HensmanFL13}
-
-distributed 1 \cite{DBLP:conf/icml/HoangHL15}
-
-distributed 2 (haven't read) \cite{DBLP:conf/icml/PengZZQ17}
-
-The paper at the core of the 3 approaches (approximates posterior) \cite{DBLP:journals/jmlr/Titsias09}
-
-Generalized version? \cite{DBLP:conf/icml/ShethWK15}
+Our main evaluation metric is \emph{expected value difference} (EVD)
+\cite{DBLP:conf/uai/JinDAS17}, calculated as
+\[ \mathbb{E} \left[ \sum_{t=0}^\infty \gamma^tr(s_t) \middle| \pi^* \right] -
+  \mathbb{E} \left[ \sum_{t=0}^\infty \gamma^t r(s_t) \middle| \hat\pi
+  \right], \]
+where $\pi^*$ is the optimal policy, and $\hat\pi$ is the policy generated using
+our reward function. We can measure how EVD changes both over time and as the
+number of demonstrations increase during training as well as use EVD to quantify
+how well the learned GP reward functions transfer to previously unseen yet
+similar environments. The learned reward functions can also be visualised as
+shades of grey on maps, just like in previous papers
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. Lastly, to evaluate how
+well the variational approximation models the real posterior, we can use two
+metrics recently proposed by Yao et al. \cite{DBLP:conf/icml/YaoVSG18}:
+Pareto-smoothed importance sampling that measures overall goodness of fit, and
+variational simulation-based calibration that can detect bias in point
+estimates.
 

### 2018-11-18 mprop/mprop.tex
Importing the IRL toolkit

@@
 \usepackage{booktabs}
+\usepackage[multiple]{footmisc}
 %\usepackage{fancyvrb}
@@
 \end{equation}
-and $V_r(s)$ is a ``soft'' version of the Bellman backup operator, which can be
+and $V_r(s)$ is a `soft' version of the Bellman backup operator, which can be
 obtained by repeatedly applying the following equation until convergence:
@@
 (ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
-  be interpreted as Lebesgue integrals over the entire sample space.}
+  be interpreted as definite integrals over the entire sample space.}
 \cite{DBLP:books/lib/Bishop07,blei2017variational}
@@
 \end{equation}
-% TODO: are they actually Lebesgue? What about notation?
 
@@
 multiple reward functions can explain the same behaviour, and suggested
-heuristics for reward functions that are ``far away'' from reward functions that
+heuristics for reward functions that are `far away' from reward functions that
 do not fit the data.
@@
 in Abbeel and Ng's approach: the algorithm requires features to be
-``appropriately'' scaled, and optimal scaling may not be known. Instead, they
+`appropriately' scaled, and optimal scaling may not be known. Instead, they
 suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
@@
   \caption{Summary of relevant VI approximations to GPs. For both inducing
-    points and hyperparameters, 'fixed' means 'chosen before the algorithm
-    starts' and 'variational' means 'included amongst the variational
+    points and hyperparameters, `fixed' means `chosen before the algorithm
+    starts' and `variational' means `included amongst the variational
     parameters'. Hyperparameters $m_\alpha$ and $m_\beta$ refer to the number of
@@
 
+% TODO: the real reason is that we want to use the prior
+% TODO: the reasoning about inducing points is not clear either
 Table \ref{table:approximations} further summarises some of the recent and/or
@@
 \section{Proposed Approach} \label{sec:proposed_approach}
-
 %state how you propose to solve the software development problem. Show that your
 %proposed approach is feasible, but identify any risks.
+
+In this section we show a feasible way to apply VI to the problem. Section
+\ref{sec:preliminaries} defines several previously-missing parts of the model
+and provides intuition for how the approximating distribution could be
+structured. Section \ref{sec:structure} then defines the approximating
+distribution, where some choices are justified by a body of literature, and some
+are novel. Next, Section \ref{sec:elbo} derives a simplified expression for the
+ELBO. Finally, in Section \ref{vi_algs} we describe our strategy for
+optimising the ELBO.
+
+\subsection{Preliminaries} \label{sec:preliminaries}
 
@@
     \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
-where $\lambda_0$ is the overall ``scale'' factor for how similar or distant the
-states are,
-\[ \bm\Lambda = \diag(\lambda_1, \dots, \lambda_d) \]
-is a diagonal matrix that determines how relevant each feature is (where $d$
-denotes the number of features), $\mathbbm{1}$ is defined as
+where $\lambda_0$ is the overall `scale' factor for how similar or distant the
+states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
+matrix that determines how relevant each feature is (where $d$ denotes the
+number of features), $\mathbbm{1}$ is defined as
 \[ \mathbbm{1}[b] = \begin{cases}
@@
 $p(\mathcal{D} | \bm\lambda, \mathbf{X_u})$ and $p(\mathcal{D} | \bm\lambda,
-\mathbf{X_u}, \mathbf{u})$, which can be computed as follows:
+\mathbf{X_u}, \mathbf{u})$, which can be expressed as follows:
 \begin{align*}
@@
 
-\subsection{Structure of the Approximating Distribution}
+\subsection{Structure of the Approximating Distribution} \label{sec:structure}
 
@@
 
-\subsection{Evidence Lower Bound}
+\subsection{Evidence Lower Bound} \label{sec:elbo}
 
@@
 write\footnote{At this point, we will drop the subscript denoting which
-  variables the expectation is taken over. Also note that throughout the derivation
-  equality is taken to mean ``equality up to an additive constant''.}
+  variables the expectation is taken over. Also note that throughout the
+  derivation equality is taken to mean `equality up to an additive constant'.}
 \[ \mathcal{L}(\bm\nu) = \mathbb{E}[\log\pfull] -
@@
   \nabla_{\bm\nu} \log \approximationS (\log \pfullS - \log
-  \approximationS) \]
+  \approximationS), \]
 computed by drawing $S$ Monte Carlo samples $(\bm\lambda_s, \mathbf{u}_s,
@@
 claim that:
-\[ \lim_{n \to \infty} \int_E f_n = \int_E \lim_{n \to \infty} f_n. \]
-We can use the same idea on parts of $\mathcal{L}$ that we cannot take the
-derivative of otherwise. In this case, we want a gradient instead of a limit,
-and the integral represents an expected value.
+\[ \lim_{n \to \infty} \int_E f_n\,d\mu = \int_E \lim_{n \to \infty}
+  f_n\,d\mu. \]
+We can use the same idea on parts of $\mathcal{L}$, derivatives of which cannot
+be taken otherwise. In our case, we want a gradient instead of a limit, and the
+integral represents an expected value.
+
+% TODO: finish this
+Let us briefly examine the options available in case the theorem is not
+applicable to some of our expected values.
+\[ \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] \quad \text{and} \quad \sum_{i=1}^d
+  \mathbb{E}[\log(\lambda_i + 1)] \]
+come from a chosen prior distribution for $\bm\lambda | \mathbf{X_u}$.
+Therefore, if needed, the distribution can be replaced with an
+easier-to-deal-with alternative.
+
+% what if the parameters of the covariance function are variational?
+
+%mTKuum, log|Kuu|, Tr[KuuS] => p(u | lambda, X_u)
+%tTKruKuum, V_r(s), T(s, a, s')V_r(s) => p(D | r) (DTC?)
+
+We can then take the BBVI algorithm as
+the basis for ours, keeping in mind the numerous improvement possibilities
+provided both by the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14} and many
+others (e.g., the ones covered in Section \ref{literature}).
 
@@
 
+The work plan can be summarised as follows:
 \begin{enumerate}
-\item Prove all applications of the dominated convergence theorem.
-\item Implement the algorithm as a more specialised version of Algorithm 2 in
-  the black box VI paper.
-\item Evaluation
-\item Experiment with various improvements mentioned in the black box VI paper,
-  other VI papers, especially the ones about GPs.
-\item Write/finalise the paper.
+\item Derive expressions for the derivatives of the ELBO w.r.t. the variational
+  parameters. \emph{Deliverables}: full derivations that can be included as
+  supplementary material of the final paper, including formal proofs for
+  applications of the dominated convergence theorem. \emph{Date}: 15~December.
+\item Implement the algorithm as a more specialised version of Algorithm~2 in
+  the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14}. As most of the relevant
+  IRL algorithms have been implemented in
+  MATLAB\footnote{\url{http://graphics.stanford.edu/projects/gpirl/irl_toolkit.zip}}\footnote{\url{https://github.com/jinming99/DGP-IRL}},
+  it makes sense to do the same in order to have accurate time-sensitive
+  comparisons between the algorithms. \emph{Deliverables}: a working
+  implementation that can be integrated into the IRL toolkit by Levine et al.
+  \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
+  code in the paper. \emph{Date}: 1~February.
+\item Evaluate the algorithm by comparing it to multiple alternatives, as
+  detailed in Section \ref{sec:evaluation}. \emph{Deliverables}: multiple plots
+  in an SVG or PDF format, along with code that produces them. \emph{Date}:
+  1~April.
+\item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
+  \emph{Date}: around 23~April, the deadline has not been updated yet.
 \end{enumerate}
 
-\subsection{Evaluation}
+\subsection{Evaluation} \label{sec:evaluation}
 
@@
 recent papers (especially those using deep learning) often use the binary world
-benchmark as well \cite{wulfmeier2015maximum,DBLP:conf/uai/JinDAS17}, our
+benchmark as well \cite{DBLP:conf/uai/JinDAS17,wulfmeier2015maximum}, our
 approach is not able to construct new features, and thus cannot perform well on

### 2018-11-21 mprop/mprop.tex
The literature review is in a readable state now

@@
 handling a wider variety of models, making approximations more accurate, and
-using more complex function approximation techniques (such as neural networks)
+using more complex function approximation techniques (such as NNs)
 to infer local latent variable values without having to calculate them
@@
     Gal et al., 2014 \cite{DBLP:conf/nips/GalWR14} & variational & variational & $\mathcal{O}(nm^2)$ \\
-    Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
+    %Hoang et al., 2015 \cite{DBLP:conf/icml/HoangHL15} & fixed & fixed & $\mathcal{O}(m^3)$ \\
     Cheng and Boots, 2017 \cite{DBLP:conf/nips/ChengB17} & variational & variational & $\mathcal{O}(nm_\alpha + nm_\beta^2)$ \\
@@
 
-% TODO: the real reason is that we want to use the prior
-% TODO: the reasoning about inducing points is not clear either
 Table \ref{table:approximations} further summarises some of the recent and/or
-influential GP approximation approaches that might be suitable for our GP IRL
-model. In order to derive a reliable ELBO, the inducing points should be either
-fixed or modelled by a probability distribution (none of the approaches do
-that), while hyperparameters cannot be variational (the reason will become clear
-in Section \ref{sec:proposed_approach}). We would also like to avoid having the
-hyperparameters fixed, as learning them efficiently would introduce a separate
-problem. As the approach by Hoang et al. does not seem to be easily extendable
-to modelled hyperparameters, it is unsuitable to our needs.
+influential GP approximation approaches that might be relevant to our situation.
+For many of the variables in our model (including the inducing points
+located somewhere in $\mathbb{R}^d$ and the hyperparameters of the covariance
+function), we have three possible ways of handling them:
+\begin{itemize}
+\item Provide a full Bayesian treatment by defining a prior probability
+  distribution. While this option can easily prevent overfitting, the choice of
+  prior can be difficult to justify.
+\item Fix the value of the variable. We can then optimise the value outside of
+  the main algorithm, e.g., by running the algorithm with a set of possible
+  values and keeping the value that produces the best performance. This option
+  can be slow if we are interested in finding the optimal value, but could be
+  appropriate if the variable has little impact on the algorithm's performance.
+\item Treat the variable as a variational parameter. This method will
+  efficiently optimise the value of the variable, but is vulnerable to
+  overfitting. Furthermore, this requires us to find the derivative of the ELBO
+  w.r.t. the variable, which could be difficult.
+\end{itemize}
+
+Titsias \cite{DBLP:journals/jmlr/Titsias09} is the first to suggest a
+variational approximation where inducing points are treated as variational
+parameters. The paper has influenced many of the later works covered in this
+section, and has been reformulated by Gal et al. \cite{DBLP:conf/nips/GalWR14}
+into a distributed algorithm. We will fix the inducing points for the baseline
+version of our algorithm, but keep the idea in mind for possible future work.
+
+Hensman et al. \cite{DBLP:conf/uai/HensmanFL13} reduce the complexity from
+$\mathcal{O}(nm^2)$ to $\mathcal{O}(m^3)$ using a stochastic variational
+inference (SVI) approach. While results such as efficiently computable natural
+gradients and analytically tractable optimal solutions for variational
+parameters are unlikely to transfer to a different problem domain, the overall
+SVI framework and an approximating Gaussian distribution for $\mathbf{u}$ will
+play an important role in our proposal.
+
+Cheng and Boots \cite{DBLP:conf/nips/ChengB17} suggest using different bases in
+the reproducing kernel Hilbert space for the mean and covariance functions.
+The paper makes few assumptions and mostly relies on the work of
+Titsias \cite{DBLP:journals/jmlr/Titsias09}, making it seem transferable to a
+new domain. However, applying the idea to IRL is likely to be beyond the scope
+of the project.
 
@@
 Adams \cite{pmlr-v28-wilson13} is an interesting and promising avenue of work,
-it is likely to be beyond the scope of this project.
+it is likely to be beyond the scope of this project as well.
 
-The derivation of the ELBO in the work of Peng et al.
-\cite{DBLP:conf/icml/PengZZQ17} relies primarily on the fact that the evidence
-for a GP is a Gaussian, while in our case the evidence can be defined in several
-ways depending on the model, but must always involve $\mathcal{D}$, making the
-main idea of the paper inapplicable to IRL. Remaining papers can be rejected for
-similar reasons, as they are all highly dependent on the specific situation of
-approximating a GP regression. Thus we cannot simply apply a GP VI model as a
-whole, although there might be some smaller ideas that we can use (some of which
-will be mentioned in the next section).
+While Peng et al. \cite{DBLP:conf/icml/PengZZQ17} provide a highly efficient
+distributed implementation, the derivation of the ELBO relies primarily on the
+fact that the evidence for a GP is a Gaussian, while in our case the evidence is
+anything but. Hence, it is unlikely that the ideas from this paper could be
+applicable to our IRL model.
 
@@
 functions---parametrised by additional variational parameters---that are applied
-to latent variables \cite{DBLP:conf/icml/RezendeM15}. Unfortunately, this
-parametrisation also means that the gradient of the joint probability
-distribution w.r.t. variational parameters $\nabla_{\bm\nu}\pfull$ is no longer
-zero, making an analytic expression for the ELBO impossible using the usual
-methods. While it may be possible to have an approximating distribution for the
-flow parameters, it is uncertain how such an algorithm would behave, as it would
-have to perform optimisation in a space with significantly more dimensions.
+to latent variables \cite{DBLP:conf/icml/RezendeM15}. Normalising flows, along
+with some of the ideas mentioned previously, could provide great benefits to the
+GP IRL VI model. However, we are forced to keep any non-essential features as
+possible future work in order to make the project feasible in a given time
+frame. Furthermore, note that due to various differences between regression and
+IRL, we cannot simply apply a GP VI model as a whole. Instead, smaller tricks
+and ideas from various papers can (and will) be used throughout the project.
 
@@
 %proposed approach is feasible, but identify any risks.
+
+% TODO: the real reason is that we want to use the prior
 
@@
 \cite{DBLP:conf/nips/LevinePK11}, its deep GP extension
-\cite{DBLP:conf/uai/JinDAS17}, and a neural network reward function
+\cite{DBLP:conf/uai/JinDAS17}, and an NN-based reward function
 approximation \cite{wulfmeier2015maximum}. As the first two have already been

### 2018-11-21 code/singletest.m

@@ -12,9 +12,11 @@ addpaths;
 %   'training_samples',8,'verbosity',1));
 
 % GPIRL's test using DGP-IRL
-test_result = runtest('deepgpirl',struct(),'linearmdp',...
-    'objectworld',struct('n',32,'determinism',0.7,'seed',1,'continuous',0),...
-    struct('training_sample_lengths',8,'training_samples',16,'verbosity',2));
+% TODO: take parameters from the paper
+test_result = runtest('deepgpirl',struct('R', 4),'linearmdp','objectworld',...
+   struct('n',12, 'continuous', 1),...
+   struct('training_sample_lengths',12^2,...
+   'training_samples',64,'verbosity',1));
 
 % Visualize solution.
 printresult(test_result);

### 2018-11-22 mprop/mprop.tex
A bit minimalistic, but readable

@@
   \begin{split}
-    \mathcal{L}(\bm\nu) &= \Eq \left[ \log \frac{\pfull}{\approximation}
+    \mathcal{L} &= \Eq \left[ \log \frac{\pfull}{\approximation}
     \right] \\
@@
 for the reward vector, and fully independent training conditional for the latent
-state (see Section \ref{sec:vi_literature} for more details).
+state (see Section \ref{sec:vi_literature} for details).
 
@@
 influential GP approximation approaches that might be relevant to our situation.
-For many of the variables in our model (including the inducing points
-located somewhere in $\mathbb{R}^d$ and the hyperparameters of the covariance
-function), we have three possible ways of handling them:
+For many of the variables in our model (including the inducing points and the
+hyperparameters of the covariance function), we have three possible ways of
+handling them:
 \begin{itemize}
@@
 
-% TODO: the real reason is that we want to use the prior
-
-In this section we show a feasible way to apply VI to the problem. Section
+In this section we show a feasible way to apply VI to the GP IRL model. Section
 \ref{sec:preliminaries} defines several previously-missing parts of the model
@@
 In order to properly investigate the difference between variational inference
-and maximum likelihood estimation for the model, we keep other parts of the
-model the same. Namely, we set the covariance function to a version of the
-automatic relevance detection kernel
-\cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
+and maximum likelihood estimation, we keep other parts of the model the same.
+Namely, we set the covariance function to a version of the automatic relevance
+detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
 \[ k_{\bm\lambda}(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left(
     -\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i -
-    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr[\bm\Lambda] \right), \]
+    \mathbf{x}_j) - \mathbbm{1}[i \ne j]\sigma^2\Tr(\bm\Lambda) \right), \]
 where $\lambda_0$ is the overall `scale' factor for how similar or distant the
 states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
-matrix that determines how relevant each feature is (where $d$ denotes the
+matrix that determines the relevance of each feature (where $d$ denotes the
 number of features), $\mathbbm{1}$ is defined as
@@
 \begin{equation} \label{eq:theta_prior}
-  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr[\Kuu^{-2}] -
+  p(\bm\lambda | \mathbf{X_u}) = \exp\left( -\frac{1}{2}\Tr(\Kuu^{-2}) -
     \sum_{i=1}^d \log(\lambda_i + 1) \right).
 \end{equation}
+Defining the prior for $\bm\lambda$ means that we commit ourselves to
+a Bayesian treatment of the variable as opposed to, e.g., treating it as a
+variational parameter. While the remainder of the proposal will stick with this
+assumption, the alternative is certainly worth exploring in its own right, and
+will be investigated after the main formulation is implemented.
 
@@
   \begin{split}
-    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u}
-    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{r} | \bm\lambda,
+    \posterior &= p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) p(\mathbf{u}
+    | \bm\lambda, \mathbf{X_u}, \mathcal{D}) p(\mathbf{r} | \bm\lambda,
     \mathbf{X_u}, \mathbf{u}, \mathcal{D}) \\
-    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathbf{u} |
-    \bm\lambda, \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} | r) \times
+    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) p(\mathbf{u} |
+    \bm\lambda, \mathbf{X_u}, \mathcal{D}) p(\mathcal{D} | r)
     p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \\
-    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) \times p(\mathcal{D} |
-    \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda,
-    \mathbf{X_u}) \times p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda,
+    &\propto p(\bm\lambda | \mathbf{X_u}, \mathcal{D}) p(\mathcal{D} |
+    \bm\lambda, \mathbf{X_u}, \mathbf{u}) p(\mathbf{u} | \bm\lambda,
+    \mathbf{X_u}) p(\mathcal{D} | r) p(\mathbf{r} | \bm\lambda,
     \mathbf{X_u}, \mathbf{u}) \\
-    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) \times p(\bm\lambda |
-    \mathbf{X_u}) \times p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
-    \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \times p(\mathcal{D} | r)
-    \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
+    &\propto p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) p(\bm\lambda |
+    \mathbf{X_u}) p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
+    p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) p(\mathcal{D} | r)
+    p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u})
   \end{split}
@@
 \begin{align*}
-  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
-  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) \times p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \times p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
+  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) = \int &p(\mathcal{D} | r) p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) \, d\mathbf{r}, \\
+  p(\mathcal{D} | \bm\lambda, \mathbf{X_u}) = \iint &p(\mathcal{D} | r) p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) \, d\mathbf{u} \, d\mathbf{r}.
 \end{align*}
@@
 
-At this point we are forced to make assumptions about the approximate posterior
+At this point, we are forced to make assumptions about the approximate posterior
 in order to arrive at an implementable solution. Time permitting, ways to relax
@@
 \begin{equation}
-  q(\mathbf{u} | \bm\lambda) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S})
+  q(\mathbf{u} | \bm\lambda) = q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}),
 \end{equation}
-for some variational parameters $\mathbf{m} \in \mathbb{R}^{m}$ and a positive
-semi-definite matrix $\mathbf{S} \in \mathbb{R}^{m \times m}$.
+where $\mathbf{m} \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
+positive semi-definite matrix $\mathbf{S}$ is the covariance matrix
 \cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
 
-In order to have a reasonable way of calculating/approximating the ELBO (see
-Section \ref{vi_algs}), we need to have an approximating distribution for
-$\bm\lambda$, but unfortunately all the papers in Table \ref{table:approximations}
-either fix it or treat it as a variational parameter. Hence we make our first
-assumption without justification from previous literature:
+Next, we need to choose an approximating distribution for $\bm\lambda$, but,
+unfortunately, all papers in Table \ref{table:approximations} either fix it or
+treat it as a variational parameter. Hence, we make our first assumption without
+justification from previous literature:
 \begin{equation} \label{eq:theta}
@@
 variables, we would like the mean to be flexible (i.e., not tied to zero, like
-with the exponential distribution), and the tails to converge to zero as the
+in the exponential distribution), and the tails to converge to zero as the
 value of the random variable moves away from the mean. We might want to support
@@
   derivation equality is taken to mean `equality up to an additive constant'.}
-\[ \mathcal{L}(\bm\nu) = \mathbb{E}[\log\pfull] -
+\[ \mathcal{L} = \mathbb{E}[\log\pfull] -
   \mathbb{E}[\log\approximation]. \]
@@
   \begin{split}
-    \mathcal{L}(\bm\nu) &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\bm\lambda |
+    \mathcal{L} &= \mathbb{E}[\log p(\mathbf{X_u}) + \log p(\bm\lambda |
     \mathbf{X_u}) + \log p(\mathbf{u} | \bm\lambda, \mathbf{X_u}) + \log
@@
 \begin{alignat*}{3}
-  \mathcal{L}(\bm\nu) &= \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] -
+  \mathcal{L} &= \mathbb{E}\left[-\frac{1}{2}\Tr(\Kuu^{-2}) -
     \sum_{i=1}^d \log(\lambda_i + 1)\right] &&+ \mathbb{E}[\log
@@
 \begin{align*}
-  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr[\Kuu^{-2}] - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
+  \mathcal{L} &= \frac{1}{2}\log|\mathbf{S}| + \mathbb{E}\left[-\frac{1}{2}\Tr(\Kuu^{-2}) - \sum_{i=1}^d \log(\lambda_i + 1)\right] \\
   &+ \mathbb{E}\left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
@@
 \[ \mathbb{E}[\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u}] =
-  \mathbb{E}[\Tr[\Kuu^{-1}\mathbf{S}] +
+  \mathbb{E}[\Tr(\Kuu^{-1}\mathbf{S}) +
   \mathbf{m}^\intercal\Kuu^{-1}\mathbf{m}] =
-  \Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] +
+  \Tr(\mathbb{E}[\Kuu^{-1}]\mathbf{S}) +
   \mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}. \]
@@
 \begin{align*}
-  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] - \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] - \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-1}]\mathbf{S}] \\
+  \mathcal{L}(\bm\nu) &= \frac{1}{2}\log|\mathbf{S}| - \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-2}]) - \sum_{i=1}^d \mathbb{E}[\log(\lambda_i + 1)] - \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-1}]\mathbf{S}) \\
   &- \frac{1}{2}\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m} - \frac{1}{2}\mathbb{E}[\log|\Kuu|] + \sum_{i=0}^d \alpha_i - \log\beta_i + \log\Gamma(\alpha_i) + (1 - \alpha_i)\psi(\alpha_i) \\
@@
 
-% TODO: finish this
 Let us briefly examine the options available in case the theorem is not
-applicable to some of our expected values.
-\[ \frac{1}{2}\Tr[\mathbb{E}[\Kuu^{-2}]] \quad \text{and} \quad \sum_{i=1}^d
+applicable to some of our expected values. First,
+\[ \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-2}]) \quad \text{and} \quad \sum_{i=1}^d
   \mathbb{E}[\log(\lambda_i + 1)] \]
 come from a chosen prior distribution for $\bm\lambda | \mathbf{X_u}$.
-Therefore, if needed, the distribution can be replaced with an
-easier-to-deal-with alternative.
+There are two ways to deal with them: replacing the prior with an
+easier-to-handle alternative, and treating $\bm\lambda$ as a variational
+parameter.
 
-% what if the parameters of the covariance function are variational?
+Then,
+\[ \frac{1}{2}\Tr(\mathbb{E}[\Kuu^{-1}]\mathbf{S}), \quad 
+  \frac{1}{2}\mathbf{m}^\intercal\mathbb{E}[\Kuu^{-1}]\mathbf{m}, \quad
+  \text{and} \quad \frac{1}{2}\mathbb{E}[\log|\Kuu|] \]
+all derive from $p(\mathbf{u} | \bm\lambda, \mathbf{X_u})$, which cannot be
+eliminated or replaced without changing the idea of the project. However, notice
+that the expressions inside $\mathbb{E}[\cdot]$ are all functions of $\Kuu$,
+which depends only on $\bm\lambda$ and $\mathbf{X_u}$. Thus, if we make
+$\bm\lambda$ a variational parameter, the expected values are no longer taken
+w.r.t. $\bm\lambda$, and $\mathbb{E}[f(\Kuu)] = f(\Kuu)$ for any function $f$
+that does not involve $\mathbf{u}$ or $\mathbf{r}$. As a result, the three
+terms simplify to expressions without expected values.
 
-%mTKuum, log|Kuu|, Tr[KuuS] => p(u | lambda, X_u)
-%tTKruKuum, V_r(s), T(s, a, s')V_r(s) => p(D | r) (DTC?)
+Lastly,
+$\mathbf{t}^\intercal\mathbb{E}[\Kru^\intercal\Kuu^{-1}]\mathbf{m}$ and both
+instances of $\mathbb{E}[V_r(s)]$ come from $p(\mathcal{D} | r)$. The first
+expected value can be resolved by making $\bm\lambda$ a variational
+parameter. The MDP value function requires more effort. We can start by making
+the deterministic training conditional assumption (as is commonly done in
+previous work \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}), which
+sets the posterior covariance matrix of $\mathbf{r}$ to the zero matrix. Then
+the posterior $\mathbf{r} = \Kru^\intercal\Kuu^{-1}\mathbf{u}$. Hence we still
+have a (complicated) function of $\bm\lambda$ and $\mathbf{u}$ in the integral
+w.r.t. both variables. The integral (or expected value) w.r.t. $\bm\lambda$ can
+be eliminated the usual way, i.e., by making $\bm\lambda$ variational. As for
+$\mathbf{u}$, we can borrow a trick from the deep GP IRL paper
+\cite{DBLP:conf/uai/JinDAS17} and set $q(\mathbf{u}) = \delta(\mathbf{u} -
+\mathbf{m})$, where $\delta$ is the Dirac delta function. The expected value
+then becomes $\mathbb{E}[V_r(s)] = V_{\Kru^\intercal\Kuu^{-1}\mathbf{m}}(s)$.
 
@@
 provided both by the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14} and many
-others (e.g., the ones covered in Section \ref{literature}).
+others (see Section \ref{literature}).
 
@@
   \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
-  code in the paper. \emph{Date}: 1~February.
+  code for the paper. \emph{Date}: 1~February.
 \item Evaluate the algorithm by comparing it to multiple alternatives, as

### 2018-12-05 mprop/mprop.tex
Proposal is ready for submission, but the submission system is not ready for the proposal

@@
 \usepackage[multiple]{footmisc}
+\usepackage[UKenglish]{babel}
+\usepackage[UKenglish]{isodate}
 %\usepackage{fancyvrb}
@@
 \author{Paulius Dilkas}
-\date{Date of submission placed here} % TODO
+\date{4th December 2018}
 \maketitle
@@
 
-%Imagine a scenario where the agent is driving on a three-lane highway. The
-%features of the environment include the agent's speed, lane number, and
-%various distances to different types of vehicles in front of and behind the
-%agent. The agent can regulate its speed and position. The main objective is, of
-%course, to not crash into other vehicles, but one could also add objectives such
-%as staying within the speed limit when close to a police vehicle.
-
 Inverse reinforcement learning (IRL)---a problem proposed by Russell in
 1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
-Markov decision process that best explains a set of given demonstrations. IRL is
-important because reward functions can be hard to define manually
+Markov decision process (MDP) that best explains a set of given demonstrations.
+IRL is important because reward functions can be hard to define manually
 \cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
@@
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
-Our main goal is employ variational inference (VI) for recovering the reward
-function, which can prove useful in two ways:
+The original approach used maximum likelihood estimation
+\cite{DBLP:conf/nips/LevinePK11}, while the goal of this project is to use
+variational inference (VI) instead, which learns approximate posterior
+probability distributions instead of point estimates. This approach can prove
+useful in three major ways:
 \begin{enumerate}
-\item As VI tends to be faster than Markov chain Monte Carlo
-  sampling \cite{blei2017variational}, we should be able to handle more data.
 \item Modelling full posterior distributions for various parameters can result
@@
   information.
+\item Having variance estimates for rewards can direct our choice in what data
+  should be collected next.
+\item An approximate Bayesian treatment of many parameters in the model guards
+  against overfitting \cite{DBLP:conf/uai/JinDAS17}.
 \end{enumerate}
@@
 \begin{definition}
-  A \emph{Markov decision process} (MDP) is a set $\mathcal{M} = \{ \mathcal{S},
+  A \emph{Markov decision process} is a set $\mathcal{M} = \{ \mathcal{S},
   \mathcal{A}, \mathcal{T}, \gamma, r \}$, where $\mathcal{S}$ and
@@
 states $\mathbf{r}$. While the previous paper to consider this IRL model
-computed maximum likelihood estimates for $\bm\lambda$ and $\mathbf{u}$, and
-made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
+computed maximum likelihood (ML) estimates for $\bm\lambda$ and $\mathbf{u}$,
+and made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
 \cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
@@
 \ref{literature}), making the resulting algorithm competitive both in terms of
-running time and model fit. Therefore, we raise the following research
-questions\footnote{For each question, we are primarily comparing the original GP
-IRL approach \cite{DBLP:conf/nips/LevinePK11} with our variational version.}:
-\begin{enumerate}
-\item Does VI allow us to make more accurate reward predictions when given the
-  same amount of training time?
-\item How does the reward prediction accuracy differ between VI and sampling
-  approaches when given the same amount of training data?
-\item How well does our VI model approximate the real posterior distribution?
-\item Is VI better at learning behaviour driven by multiple goals?
-\item Is VI more robust at handling mistakes in demonstrations?
-\end{enumerate}
+running time and model fit.
+
+The project is primarily concerned with investigating how a VI formulation of
+the GP IRL model compares against the original ML approach. Most importantly, we
+aim to compare how the two algorithms converge both over time and as the number
+of demonstrations increases. It would also be interesting to see how close the
+approximation of the posterior distribution is to the real thing---we discuss
+ways to do this in Section \ref{sec:evaluation}. Finally, it is reasonable to
+conjecture that VI can outperform point estimates when dealing with more
+uncertainty, e.g., when experts make mistakes. We can easily investigate this by
+adapting how the evaluation data is generated.
 
@@
 features or even from raw data. The only disadvantage (as demonstrated in the
-paper) is that NNs take longer to learn compared to GPs.
+paper) is that NNs converge slower than GPs (as measured by the number of
+demonstrations).
 
@@
 In order to properly investigate the difference between variational inference
-and maximum likelihood estimation, we keep other parts of the model the same.
+and ML estimation, we keep other parts of the model the same.
 Namely, we set the covariance function to a version of the automatic relevance
@@
 sets the posterior covariance matrix of $\mathbf{r}$ to the zero matrix. Then
-the posterior $\mathbf{r} = \Kru^\intercal\Kuu^{-1}\mathbf{u}$. Hence we still
-have a (complicated) function of $\bm\lambda$ and $\mathbf{u}$ in the integral
-w.r.t. both variables. The integral (or expected value) w.r.t. $\bm\lambda$ can
-be eliminated the usual way, i.e., by making $\bm\lambda$ variational. As for
-$\mathbf{u}$, we can borrow a trick from the deep GP IRL paper
-\cite{DBLP:conf/uai/JinDAS17} and set $q(\mathbf{u}) = \delta(\mathbf{u} -
-\mathbf{m})$, where $\delta$ is the Dirac delta function. The expected value
-then becomes $\mathbb{E}[V_r(s)] = V_{\Kru^\intercal\Kuu^{-1}\mathbf{m}}(s)$.
+the posterior $\mathbf{r}$ becomes $\Kru^\intercal\Kuu^{-1}\mathbf{u}$. But we
+still have a (complicated) function of $\bm\lambda$ and $\mathbf{u}$ in the
+integral w.r.t. both variables. The integral (or expected value) w.r.t.
+$\bm\lambda$ can be eliminated the usual way, i.e., by making $\bm\lambda$
+variational. As for $\mathbf{u}$, we can borrow a trick from the deep GP IRL
+paper \cite{DBLP:conf/uai/JinDAS17} and set $q(\mathbf{u}) = \delta(\mathbf{u} -
+\mathbf{m})$, where $\delta$ is the Dirac delta function. This is essentially a
+different way to impose a zero-variance restriction, modelling $\mathbf{u}$ as a
+random vector with variance tending to zero. The expected value then becomes
+$\mathbb{E}[V_r(s)] = V_{\Kru^\intercal\Kuu^{-1}\mathbf{m}}(s)$.
 
-We can then take the BBVI algorithm as
-the basis for ours, keeping in mind the numerous improvement possibilities
-provided both by the BBVI paper \cite{DBLP:conf/aistats/RanganathGB14} and many
-others (see Section \ref{literature}).
+While the ELBO can now be computed without approximations, we ended up imposing
+two zero-variance restrictions, resulting in a less informative model. Moreover,
+we had to abandon the Bayesian treatment of $\bm\lambda$, making the model more
+vulnerable to overfitting.
+
+Assuming that such workarounds will not be needed (or desired), we can then take
+the BBVI algorithm as the basis for our algorithm, while keeping in mind the
+numerous improvement possibilities provided both by the BBVI paper
+\cite{DBLP:conf/aistats/RanganathGB14} and many others (see Section
+\ref{literature}).
 
@@
 \item Write/finalise the paper. \emph{Deliverable}: a 14-page paper.
-  \emph{Date}: around 23~April, the deadline has not been updated yet.
+  \emph{Date}: 12~April.
 \end{enumerate}

### 2018-12-06 mprop/mprop.tex
Proofs of the helper theorems

@@
   \mathcal{T}(s, a, s')V_r(s'),
-\end{equation}
+\end{equation} % TODO: fix the wording!!!
 and $V_r(s)$ is a `soft' version of the Bellman backup operator, which can be

### 2018-12-13 mprop/mprop.tex
The proof is complete but messy

--- 2018-12-05 mprop/mprop.tex
@@
 \author{Paulius Dilkas}
-\date{4th December 2018}
+\date{12th December 2018}
 \maketitle
@@
 \end{equation}
-and $V_r(s)$ is a `soft' version of the Bellman backup operator, which can be
-obtained by repeatedly applying the following equation until convergence:
+and $V_r : \mathcal{S} \to \mathbb{R}$ is an MDP \emph{value function}, which
+can be obtained by repeatedly applying the 'soft' version of the Bellman backup
+operator until convergence:
 \cite{DBLP:conf/nips/LevinePK11,supplementary_material}
@@
       \mathcal{S}} \mathcal{T}(s, a, s')V_r(s') \right). \]
-The likelihood of the data can then be written down as
+The likelihood of the data can then be expressed as
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
@@
   supplementary material of the final paper, including formal proofs for
-  applications of the dominated convergence theorem. \emph{Date}: 15~December.
+  applications of the dominated convergence theorem. \emph{Date}: 15~January.
 \item Implement the algorithm as a more specialised version of Algorithm~2 in
@@
   \cite{DBLP:conf/nips/LevinePK11}, and the same algorithm written in pseudo
-  code for the paper. \emph{Date}: 1~February.
+  code for the paper. \emph{Date}: 15~February.
 \item Evaluate the algorithm by comparing it to multiple alternatives, as

### 2018-12-25 code/VIGPIRL/vigpirldefaultparams.m
Running but far from fully functional

@@ -1,5 +1,5 @@
 % Fill in default parameters for the GPIRL algorithm.
-function algorithm_params = gpirldefaultparams(algorithm_params)
+function algorithm_params = vigpirldefaultparams(algorithm_params)
 
 % Create default parameters.
 default_params = struct(...
@@ -46,7 +46,8 @@ default_params = struct(...
     'warp_s_init',1,...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
-    'inducing_pts_count',64);
+    'inducing_pts_count',64,...
+    'samples_count',1000);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2018-12-25 code/VIGPIRL/vigpirlgetinducingpoints.m

@@ -1,5 +1,5 @@
 % Get a good set of inducing states under the specified parameters.
-function gp = gpirlgetinducingpoints(gp,~,mu_sa,algorithm_params)
+function gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params)
 
 % Constants.
 states = size(mu_sa,1);

### 2018-12-25 code/VIGPIRL/vigpirlgetstatistics.m

@@ -1,5 +1,5 @@
 % Pre-compute statistics from examples for GPIRL.
-function [mu_sa,init_s] = gpirlgetstatistics(example_samples,mdp_data)
+function [mu_sa,init_s] = vigpirlgetstatistics(example_samples,mdp_data)
 
 % Constants.
 [states,actions,transitions] = size(mdp_data.sa_s);
@@ -21,9 +21,9 @@ for i=1:N,
         % See notes on computing state visitation counts for details of
         % what's going on here.
         % This is the - \sum_{i,t} \gamma P(s_j | s_it, a_it) term.
-        for k=1:transitions,
-            sp = mdp_data.sa_s(s,a,k);
-            init_s(sp) = init_s(sp) - mdp_data.discount*mdp_data.sa_p(s,a,k);
-        end;
+%        for k=1:transitions,
+%            sp = mdp_data.sa_s(s,a,k);
+%            init_s(sp) = init_s(sp) - mdp_data.discount*mdp_data.sa_p(s,a,k);
+%        end;
     end;
 end;

### 2018-12-25 code/VIGPIRL/vigpirlkernel.m

@@ -1,231 +1,110 @@
-% Optimized kernel computation function for DC mode GPIRL.
-function [K_uf,logDetAndPPrior,alpha,invK,dhp,dhpdr] = gpirlkernel(gp,y,Xstar)
+                      % Optimized kernel computation function for DC mode GPIRL.
+function [K_uf, alpha, invK, K_uu, K_ufKinv, K_ff, K_uu_deriv_lambda0, K_uf_deriv_lambda0] = vigpirlkernel(gp,y,Xstar)
 
-% Constants.
-dims = length(gp.inv_widths);
-n = size(gp.X_u,1);
+                                % Constants.
+  dims = length(gp.inv_widths);
+  n = size(gp.X_u,1);
 
-% Undo transforms.
-inv_widths = gpirlhpxform(gp.inv_widths,[],gp.ard_xform,1); % This is \Lambda
-noise_var = gpirlhpxform(gp.noise_var,[],gp.noise_xform,1); % This is 2\sigma^2
-rbf_var = gpirlhpxform(gp.rbf_var,[],gp.rbf_xform,1); % This is \beta
-if gp.warp_x,
+                                                            % Undo transforms.
+  inv_widths = gpirlhpxform(gp.inv_widths,[],gp.ard_xform,1); % This is \Lambda
+  noise_var = gpirlhpxform(gp.noise_var,[],gp.noise_xform,1); % This is 2\sigma^2
+  rbf_var = gpirlhpxform(gp.rbf_var,[],gp.rbf_xform,1); % This is \beta
+  if gp.warp_x,
     warp_c = gpirlhpxform(gp.warp_c,[],gp.warp_c_xform,1); % This is m
     warp_l = gpirlhpxform(gp.warp_l,[],gp.warp_l_xform,1); % This is \ell
     warp_s = gpirlhpxform(gp.warp_s,[],gp.warp_s_xform,1); % This is s
-end;
-inv_widths = min(inv_widths,1e100); % Prevent overflow.
+  end;
+  inv_widths = min(inv_widths,1e100); % Prevent overflow.
 
-% Compute scales.
-iw_sqrt = sqrt(inv_widths);
+                                % Compute scales.
+  iw_sqrt = sqrt(inv_widths);
 
-% Scale positions in feature space.
-if gp.warp_x,
[…]

### 2018-12-25 code/VIGPIRL/vigpirlrun.m

@@ -1,152 +1,117 @@
-% GP-based non-linear IRL algorithm.
-function irl_result = gpirlrun(algorithm_params,mdp_data,mdp_model,...
-    feature_data,example_samples,~,verbosity)
+                                % GP-based non-linear IRL algorithm.
+function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
+                                 feature_data,example_samples,~,verbosity)
 
-% algorithm_params - parameters of the GP IRL algorithm.
-% mdp_data - definition of the MDP to be solved.
-% example_samples - cell array containing examples.
-% irl_result - result of IRL algorithm (see bottom of file).
+                    % algorithm_params - parameters of the GP IRL algorithm.
+                    % mdp_data - definition of the MDP to be solved.
+                    % example_samples - cell array containing examples.
+                    % irl_result - result of IRL algorithm (see bottom of file).
 
-% Fill in default parameters.
-algorithm_params = gpirldefaultparams(algorithm_params);
+                                % Fill in default parameters.
+  algorithm_params = vigpirldefaultparams(algorithm_params);
 
-% Set random seed.
-rand('seed',algorithm_params.seed);
-randn('seed',algorithm_params.seed);
+                                % Set random seed.
+  rand('seed',algorithm_params.seed);
+  randn('seed',algorithm_params.seed);
 
-% Get state-action counts and initial state distributions.
-[mu_sa,init_s] = gpirlgetstatistics(example_samples,mdp_data);
+                      % Get state-action counts and initial state distributions.
+  [mu_sa,init_s] = vigpirlgetstatistics(example_samples,mdp_data);
 
-% Set up optimization options.
-options = struct('Display','iter','LS_init',2,'LS',2,'Method','lbfgs',...
-    'MaxFunEvals',4000,'MaxIter',2000);
-if verbosity < 2,
+                                % Set up optimization options.
+  options = struct('Display','iter','LS_init',2,'LS',2,'Method','lbfgs',...
+                   'MaxFunEvals',4000,'MaxIter',2000);
+  if verbosity < 2,
     options.display = 'none';
-end;
+  end;
 
-% Initialize r.
-if ~isempty(algorithm_params.initial_r),
-    % Use provided initialization.
[…]

### 2018-12-25 code/singletest.m

@@ -12,11 +12,15 @@ addpaths;
 %   'training_samples',8,'verbosity',1));
 
 % GPIRL's test using DGP-IRL
-% TODO: take parameters from the paper
-test_result = runtest('deepgpirl',struct('R', 4),'linearmdp','objectworld',...
-   struct('n',12, 'continuous', 1),...
-   struct('training_sample_lengths',12^2,...
-   'training_samples',64,'verbosity',1));
+%test_result = runtest('deepgpirl',struct('R', 4),'linearmdp','objectworld',...
+%   struct('n',12, 'continuous', 1),...
+%   struct('training_sample_lengths',12^2,...
+%   'training_samples',64,'verbosity',1));
+
+% VIGPIRL
+test_result = runtest('vigpirl',struct(),'linearmdp',...
+    'objectworld',struct('n',32,'determinism',0.7,'seed',1,'continuous',0),...
+    struct('training_sample_lengths',8,'training_samples',16,'verbosity',2));
 
 % Visualize solution.
 printresult(test_result);

### 2018-12-26 code/VIGPIRL/vigpirlrun.m
Restricting lambda0 to positive values (via exp transform) makes it converge to zero...

@@ -20,15 +20,10 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     options.display = 'none';
   end;
 
-                                % Create GP.
-  if ~isempty(algorithm_params.initial_gp),
-    gp = algorithm_params.initial_gp;
-  else
-                                % Create initial GP.
-    gp = gpirlinit(algorithm_params,feature_data);
-                                % Choose inducing points.
-    gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
-  end;
+                              % Create initial GP.
+  gp = gpirlinit(algorithm_params,feature_data);
+                              % Choose inducing points.
+  gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
 
   m = size(gp.X_u, 1);
   d = size(feature_data.splittable, 2);
@@ -36,7 +31,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
                                 % Randomly initialize variational parameters
                                 % TODO: do the optimal thing instead
   mu = rand(m, 1);
-  lambda0 = random('Chisquare', algorithm_params.rbf_init);
   lambda = random('Chisquare', algorithm_params.ard_init, [d, 1]);
 
                  % L is a lower triangular matrix with positive diagonal entries
@@ -45,7 +39,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   L = tril(L);
 
   function fun = estimate_derivative(solution, Kuu_inv, Kuu_derivative, u)
-                                % TODO: make this return a function that takes u
     function derivative = wrapped(u)
       function a = inner(state)
         a = solution.v(state, 1) * trace((Kuu_inv * u * u' * Kuu_inv' - Kuu_inv) * Kuu_derivative);
@@ -65,7 +58,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     fun = wrapped(u);
   end;
 
-  % TODO: lambda should be part of gp
   function answer = estimate_derivative_outer(u, KruKuu, Krr,...
      Kru, Kuu_inv, Kuu_derivative)
     r = mvnrnd(KruKuu * u', Krr - KruKuu * Kru)';
[…]

### 2018-12-26 code/singletest.m

@@ -19,8 +19,8 @@ addpaths;
 
 % VIGPIRL
 test_result = runtest('vigpirl',struct(),'linearmdp',...
-    'objectworld',struct('n',32,'determinism',0.7,'seed',1,'continuous',0),...
-    struct('training_sample_lengths',8,'training_samples',16,'verbosity',2));
+    'objectworld',struct('n',16,'determinism',0.7,'seed',1,'continuous',0),...
+    struct('training_sample_lengths',4,'training_samples',8,'verbosity',2));
 
 % Visualize solution.
 printresult(test_result);

### 2018-12-27 code/VIGPIRL/vigpirlkernel.m
Restructure things to make it a bit faster

@@ -18,9 +18,9 @@ function [K_uf, invK, K_ufKinv, K_ff, K_uu_deriv_lambda0,...
                                 % Scale positions in feature space.
   X_u_warped = gp.X_u;
   X_f_warped = gp.X;
-  if nargin >= 2,
+  if nargin >= 2
     X_s_warped = Xstar;
-  end;
+  end
   X_u_scaled = bsxfun(@times,iw_sqrt,X_u_warped);
   X_f_scaled = bsxfun(@times,iw_sqrt,X_f_warped);
 
@@ -30,7 +30,7 @@ function [K_uf, invK, K_ufKinv, K_ff, K_uu_deriv_lambda0,...
     nconst2 = -0.5 * noise_var;
     nmat = nconst*ones(n) + (1-nconst)*eye(n);
     nmat2 = nconst2*ones(n) + (1-nconst2)*eye(n);
-  end;
+  end
 
   function deriv = K_uu_lambda_i_derivatives(K_uu, X, Y, nmat2, i)
     xi = X(:, i);
@@ -39,7 +39,7 @@ function [K_uf, invK, K_ufKinv, K_ff, K_uu_deriv_lambda0,...
     yi_sq = yi .^ 2;
     inner = repmat(xi_sq', size(Y, 1), 1) + repmat(yi_sq, 1, size(X, 1)) - 2 * yi * xi';
     deriv = K_uu .* (-0.5 * inner + nmat2);
-  end;
+  end
 
   function [K_uu, K_uu_deriv_lambda0, K_uu_deriv_lambda, nconst] = compute_covariance_matrix(X)
     d_uu = bsxfun(@plus,sum(X.^2,2),sum(X.^2,2)') - 2 * X * X';
@@ -47,9 +47,9 @@ function [K_uf, invK, K_ufKinv, K_ff, K_uu_deriv_lambda0,...
     [nmat, nmat2, nconst, nconst2] = construct_noise_matrix(size(X, 1));
     K_uu_deriv_lambda0 = exp(-0.5*d_uu).*nmat;
     K_uu = rbf_var * K_uu_deriv_lambda0;
-    K_uu_deriv_lambda = arrayfun(@(i)...
-      K_uu_lambda_i_derivatives(K_uu, X, X, nmat2, i), 1:size(X, 2), 'Uniform', 0);
-  end;
+    c = arrayfun(@(i) K_uu_lambda_i_derivatives(K_uu, X, X, nmat2, i), 1:size(X, 2), 'Uniform', 0);
[…]

### 2018-12-27 code/VIGPIRL/vigpirlrun.m

@@ -29,34 +29,31 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   L(1:m+1:end) = random('Chisquare', 1, [m, 1]);
   L = tril(L);
 
-  function derivative = estimate_derivative(solution, Kuu_inv, Kuu_derivative, u)
-    % TODO: minimise function calls
-    function a = inner(state)
-      a = solution.v(state, 1) * trace((Kuu_inv * (u * u') * Kuu_inv' - Kuu_inv) * Kuu_derivative);
-    end
-
-    function answer = foo(state_action)
-      s = 0;
-      for next_state = 1:size(mdp_data.sa_p, 3)
-        % TODO: optimise this line
-        s = s + mdp_data.sa_p(state_action(1), state_action(2), next_state) * inner(next_state);
-      end
-      answer = inner(state_action(1)) - mdp_data.discount * s;
-    end
-
-    derivative = sum(sum(cellfun(@foo, example_samples)));
+  function a = extract(mdp, state, action)
+    temp = mdp(state, action, :);
+    a = reshape(temp, size(mdp, 3), 1);
   end
 
-  function answer = estimate_derivative_outer(u, KruKuu, Krr,...
-     Kru, Kuu_inv, Kuu_derivative)
-    % TODO: optimise the two lines.
-    % Could "move" the r sample instead of retaking it for every u
-    r = mvnrnd(KruKuu * u', Krr - KruKuu * Kru)';
-    solution = feval([mdp_model 'solve'], mdp_data, r);
-    answer = estimate_derivative(solution, Kuu_inv, Kuu_derivative, u');
+  function answer = estimate_derivative(u, KruKuu, Krr, Kru, Kuu_inv, Kuu_derivatives)
+    function answer = expected_derivative_for(state_action)
+      state = state_action(1);
+      action = state_action(2);
+      next_states = extract(mdp_data.sa_s, state, action);
+      next_probabilities = extract(mdp_data.sa_p, state, action);
+      trace_part = arrayfun(@(i) trace((Kuu_inv * (u' * u) * Kuu_inv' -...
+        Kuu_inv) * Kuu_derivatives(:, :, i)), 1:size(Kuu_derivatives, 3));
+      s = next_probabilities' * solution.v(next_states, 1);
[…]

### 2018-12-29 code/VIGPIRL/vigpirlkernel.m
Optimising mu (probably incorrectly, though)

@@ -5,6 +5,7 @@ function [K_uf, invK, K_ufKinv, K_ff, K_uu_grad, K_uf_grad] = vigpirlkernel(gp, 
   dims = length(gp.inv_widths);
   n = size(gp.X_u,1);
 
+  % TODO: is this necessary?
                                                             % Undo transforms.
   inv_widths = gpirlhpxform(gp.inv_widths,[],gp.ard_xform,1); % This is \Lambda
   noise_var = gpirlhpxform(gp.noise_var,[],gp.noise_xform,1); % This is 2\sigma^2

### 2018-12-29 code/VIGPIRL/vigpirlpackparam.m

@@ -1,4 +1,4 @@
 % Place parameters from specified GP into parameter vector.
 function x = vigpirlpackparam(gp)
   x = gpirlhpxform(gp.rbf_var, [], 'exp', 3);
-  x = vertcat(x, gpirlhpxform(gp.inv_widths, [], 'exp', 3)');
+  x = vertcat(x, gpirlhpxform(gp.inv_widths, [], 'exp', 3)', gp.mu');

### 2018-12-29 code/VIGPIRL/vigpirlrun.m

@@ -14,20 +14,18 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   [mu_sa,init_s] = vigpirlgetstatistics(example_samples,mdp_data);
 
                               % Create initial GP.
-  gp = gpirlinit(algorithm_params,feature_data);
+  gp = vigpirlinit(algorithm_params,feature_data);
                               % Choose inducing points.
   gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
 
   m = size(gp.X_u, 1);
 
-                                % Randomly initialize variational parameters
-                                % TODO: do the optimal thing instead
-  mu = rand(m, 1);
-
                  % L is a lower triangular matrix with positive diagonal entries
   L = normrnd(0, 1, [m, m]);
   L(1:m+1:end) = random('Chisquare', 1, [m, 1]);
   L = tril(L);
+  Sigma = L * L';
+  Sigma_inv = inv(Sigma);
 
   function a = extract(mdp, state, action)
     temp = mdp(state, action, :);
@@ -40,10 +38,12 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
       action = state_action(2);
       next_states = extract(mdp_data.sa_s, state, action);
       next_probabilities = extract(mdp_data.sa_p, state, action);
-      trace_part = arrayfun(@(i) trace((Kuu_inv * (u' * u) * Kuu_inv' -...
+      unique_lambda_part = arrayfun(@(i) trace((Kuu_inv * (u' * u) * Kuu_inv' -...
         Kuu_inv) * Kuu_derivatives(:, :, i)), 1:size(Kuu_derivatives, 3));
+      unique_mu_part = (Sigma_inv + Sigma_inv') * (u - gp.mu)';
+      unique_part = vertcat(unique_lambda_part', unique_mu_part);
       s = next_probabilities' * solution.v(next_states, 1);
-      answer = trace_part .* (solution.v(state, 1) - mdp_data.discount * s);
+      answer = unique_part * (solution.v(state, 1) - mdp_data.discount * s);
     end
 
     r = mvnrnd(KruKuu * u', Krr - KruKuu * Kru);
[…]

### 2018-12-29 code/VIGPIRL/vigpirlunpackparam.m

@@ -8,5 +8,8 @@ gp.rbf_var = gpirlhpxform(x(endi), [], 'exp', 1);
 endi = endi - 1;
 
 % Read ARD kernel parameters.
-gp.inv_widths = gpirlhpxform(x(endi-length(gp.inv_widths)+1:endi), [], 'exp', 1)';
-endi = endi-length(gp.inv_widths);
+gp.inv_widths = gpirlhpxform(x(endi-length(gp.inv_widths)+1:endi), [], 'exp', 1);
+endi = endi - length(gp.inv_widths);
+
+gp.mu = x(endi-length(gp.mu)+1:endi);
+endi = endi - length(gp.mu);

### 2018-12-30 code/VIGPIRL/estimate_derivative.m
More tests. Surprisingly, all pass

@@ -1,0 +1,28 @@
+function answer = estimate_derivative(u, KruKuu, Kuu_inv, r_covariance_matrix,...
+                                      Kuu_derivatives, mdp_data, Sigma_inv,...
+                                      mu, mdp_model, example_samples)
+             % Uses a point-sample of u to estimate the part of dL/dnu under E[]
+  function answer = expected_derivative_for(state_action)
+    state = state_action(1);
+    action = state_action(2);
+    next_states = extract(mdp_data.sa_s, state, action);
+    next_probabilities = extract(mdp_data.sa_p, state, action);
+    unique_lambda_part = arrayfun(@(i) trace((Kuu_inv * (u' * u) * Kuu_inv' -...
+      Kuu_inv) * Kuu_derivatives(:, :, i)), 1:size(Kuu_derivatives, 3));
+    unique_mu_part = (Sigma_inv + Sigma_inv') * (u - mu)';
+    unique_part = vertcat(unique_lambda_part', unique_mu_part);
+    s = next_probabilities' * solution.v(next_states, 1);
+    answer = unique_part * (solution.v(state, 1) - mdp_data.discount * s);
+  end
+
+  r = mvnrnd(KruKuu * u', r_covariance_matrix);
+  solution = feval([mdp_model 'solve'], mdp_data, r');
+  derivatives_for_each_state_action = cellfun(@expected_derivative_for,...
+                                              example_samples, 'Uniform', 0);
+  answer = sum(cat(3, derivatives_for_each_state_action{:}), 3);
+end
+
+function a = extract(mdp, state, action)
+  temp = mdp(state, action, :);
+  a = reshape(temp, size(mdp, 3), 1);
+end

### 2018-12-30 code/VIGPIRL/full_gradient.m

@@ -1,0 +1,16 @@
+function changes = full_gradient(Sigma, mdp_model,...
+  mdp_data, example_samples, counts, mu, Kru, Kuu_inv, KruKuu, Krr,...
+  Kuu_grad, Kru_grad, z)
+  r_covariance_matrix = Krr - KruKuu * Kru;
+  grad_prediction_for_each_u = arrayfun(@(i) estimate_derivative(z(i, :),...
+    KruKuu, Kuu_inv, r_covariance_matrix, Kuu_grad, mdp_data, inv(Sigma),...
+    mu, mdp_model, example_samples), 1:size(z, 1), 'Uniform', 0);
+  estimated_grad = mean(cat(3, grad_prediction_for_each_u{:}), 3);
+
+  not_estimated_lambda = arrayfun(@(i) -0.5 *...
+    trace(Kuu_inv * Kuu_grad(:, :, i)) + counts' * (Kru_grad(:, :, i)' -...
+    KruKuu * Kuu_grad(:, :, i)) * Kuu_inv * mu', 1:size(Kuu_grad, 3));
+  not_estimated_mu = counts' * Kru' * Kuu_inv;
+  not_estimated = vertcat(not_estimated_lambda', not_estimated_mu');
+  changes = not_estimated - 0.5 * estimated_grad;
+end

### 2018-12-30 code/VIGPIRL/vigpirlinit.m

@@ -22,8 +22,7 @@ gp.noise_prior_wt = algorithm_params.noise_prior_wt;
 gp.rbf_prior_wt = algorithm_params.rbf_prior_wt;
 
 % Initialize hyperparameters.
-gp.noise_var = gpirlhpxform(algorithm_params.noise_init, [],...
-    algorithm_params.noise_xform, 3);
+gp.noise_var = algorithm_params.noise_init;
 gp.rbf_var = random('Chisquare', algorithm_params.rbf_init);
 gp.inv_widths = random('Chisquare', algorithm_params.ard_init,...
     [1, size(feature_data.splittable, 2)]);

### 2018-12-30 code/VIGPIRL/vigpirlkernel.m

@@ -1,6 +1,6 @@
                       % Optimized kernel computation function for DC mode GPIRL.
 function [K_uf, invK, K_ufKinv, K_ff, K_uu_grad, K_uf_grad] = vigpirlkernel(gp, Xstar)
-  inv_widths = min(gp.inv_widths,1e100); % Prevent overflow.
+  inv_widths = min(gp.inv_widths, 1e100); % Prevent overflow.
   iw_sqrt = sqrt(inv_widths); % Compute scales.
 
                                 % Scale positions in feature space.
@@ -9,9 +9,10 @@ function [K_uf, invK, K_ufKinv, K_ff, K_uu_grad, K_uf_grad] = vigpirlkernel(gp, 
   if nargin >= 2
     X_s_warped = Xstar;
   end
-  X_u_scaled = bsxfun(@times,iw_sqrt,X_u_warped);
-  X_f_scaled = bsxfun(@times,iw_sqrt,X_f_warped);
+  X_u_scaled = bsxfun(@times, iw_sqrt, X_u_warped);
+  X_f_scaled = bsxfun(@times, iw_sqrt, X_f_warped);
 
+  % TODO: I think dK/dmu is not used and doesn't have to be computed
   function [K_uu, K_uu_grad, nconst] = compute_covariance_matrix(X)
     d_uu = bsxfun(@plus, sum(X.^2, 2), sum(X.^2, 2)') - 2 * X * X';
     d_uu = max(d_uu, 0);

### 2018-12-30 code/VIGPIRL/vigpirlrun.m

@@ -11,79 +11,39 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   algorithm_params = vigpirldefaultparams(algorithm_params);
 
                       % Get state-action counts and initial state distributions.
-  [mu_sa,init_s] = vigpirlgetstatistics(example_samples,mdp_data);
+  [mu_sa, counts] = vigpirlgetstatistics(example_samples,mdp_data);
 
-                              % Create initial GP.
+                                % Create initial GP.
   gp = vigpirlinit(algorithm_params,feature_data);
-                              % Choose inducing points.
+                                % Choose inducing points.
   gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
 
   m = size(gp.X_u, 1);
 
                  % L is a lower triangular matrix with positive diagonal entries
   L = normrnd(0, 1, [m, m]);
-  L(1:m+1:end) = random('Chisquare', 1, [m, 1]);
+  L(1:m+1:end) = random('Chisquare', 4, [m, 1]);
   L = tril(L);
   Sigma = L * L';
-  Sigma_inv = inv(Sigma);
-
-  function a = extract(mdp, state, action)
-    temp = mdp(state, action, :);
-    a = reshape(temp, size(mdp, 3), 1);
-  end
-
-  function answer = estimate_derivative(u, KruKuu, Krr, Kru, Kuu_inv, Kuu_derivatives)
-    function answer = expected_derivative_for(state_action)
-      state = state_action(1);
-      action = state_action(2);
-      next_states = extract(mdp_data.sa_s, state, action);
-      next_probabilities = extract(mdp_data.sa_p, state, action);
-      unique_lambda_part = arrayfun(@(i) trace((Kuu_inv * (u' * u) * Kuu_inv' -...
-        Kuu_inv) * Kuu_derivatives(:, :, i)), 1:size(Kuu_derivatives, 3));
-      unique_mu_part = (Sigma_inv + Sigma_inv') * (u - gp.mu)';
-      unique_part = vertcat(unique_lambda_part', unique_mu_part);
-      s = next_probabilities' * solution.v(next_states, 1);
-      answer = unique_part * (solution.v(state, 1) - mdp_data.discount * s);
-    end
-
-    r = mvnrnd(KruKuu * u', Krr - KruKuu * Kru);
[…]

### 2019-01-06 code/VIGPIRL/vigpirlgetstatistics.m
Cleanup + describing BB^T+D^2 decomposition + implementing the optimisation of its diagonal entries

@@ -16,14 +16,5 @@ for i=1:N,
         % Add to state action and state expectations.
         mu_sa(s,a) = mu_sa(s,a) + 1;
         init_s(s) = init_s(s) + 1;
-        
-        % Subtract off probabilities of transitioning to other states.
-        % See notes on computing state visitation counts for details of
-        % what's going on here.
-        % This is the - \sum_{i,t} \gamma P(s_j | s_it, a_it) term.
-%        for k=1:transitions,
-%            sp = mdp_data.sa_s(s,a,k);
-%            init_s(sp) = init_s(sp) - mdp_data.discount*mdp_data.sa_p(s,a,k);
-%        end;
     end;
 end;

### 2019-01-06 code/VIGPIRL/vigpirlinit.m

@@ -10,44 +10,8 @@ gp.s_u = 1:size(feature_data.splittable,1);
 gp.X = feature_data.splittable;
 gp.X_u = feature_data.splittable;
 
-% Copy transform and prior information from parameters.
-gp.ard_xform = algorithm_params.ard_xform;
-gp.noise_xform = algorithm_params.noise_xform;
-gp.rbf_xform = algorithm_params.rbf_xform;
-gp.ard_prior = algorithm_params.ard_prior;
-gp.noise_prior = algorithm_params.noise_prior;
-gp.rbf_prior = algorithm_params.rbf_prior;
-gp.ard_prior_wt = algorithm_params.ard_prior_wt;
-gp.noise_prior_wt = algorithm_params.noise_prior_wt;
-gp.rbf_prior_wt = algorithm_params.rbf_prior_wt;
-
 % Initialize hyperparameters.
 gp.noise_var = algorithm_params.noise_init;
 gp.rbf_var = random('Chisquare', algorithm_params.rbf_init);
 gp.inv_widths = random('Chisquare', algorithm_params.ard_init,...
     [1, size(feature_data.splittable, 2)]);
-
-% Specify which values to optimize and how to optimize them.
-gp.learn_noise = algorithm_params.learn_noise;
-gp.learn_rbf = algorithm_params.learn_rbf;
-gp.warp_x = algorithm_params.warp_x;
-gp.gamma_shape = algorithm_params.gamma_shape;
-
-% Initialize warp parameters.
-if gp.warp_x,
-    gp.warp_l = gpirlhpxform(algorithm_params.warp_l_init*ones(1,size(feature_data.splittable,2)),...
-        [],algorithm_params.warp_l_xform,3);
-    gp.warp_l_prior = algorithm_params.warp_l_prior;
-    gp.warp_l_prior_wt = algorithm_params.warp_l_prior_wt;
-    gp.warp_l_xform = algorithm_params.warp_l_xform;
-    gp.warp_c = gpirlhpxform(algorithm_params.warp_c_init*ones(1,size(feature_data.splittable,2)),...
-        [],algorithm_params.warp_c_xform,3);
-    gp.warp_c_prior = algorithm_params.warp_c_prior;
-    gp.warp_c_prior_wt = algorithm_params.warp_c_prior_wt;
-    gp.warp_c_xform = algorithm_params.warp_c_xform;
[…]

### 2019-01-06 code/VIGPIRL/vigpirlpackparam.m

@@ -1,4 +1,5 @@
 % Place parameters from specified GP into parameter vector.
 function x = vigpirlpackparam(gp)
   x = vigpirlhpxform(gp.rbf_var, [], 'exp', 3);
-  x = vertcat(x, vigpirlhpxform(gp.inv_widths, [], 'exp', 3)', gp.mu);
+  x = vertcat(x, vigpirlhpxform(gp.inv_widths, [], 'exp', 3)', gp.mu,...
+    vigpirlhpxform(diag(gp.B), [], 'exp', 3));

### 2019-01-06 code/VIGPIRL/vigpirlrun.m

@@ -20,27 +20,30 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   m = size(gp.X_u, 1);
   gp.mu = rand(1, m)';
 
-                 % L is a lower triangular matrix with positive diagonal entries
-  %L = normrnd(0, 1, [m, m]);
-  %L(1:m+1:end) = random('Chisquare', 4, [m, 1]);
-  %L = tril(L);
-  %Sigma = L * L';
-  Sigma = eye(m);
-  Sigma_inv = inv(Sigma);
+                 % B is a lower triangular matrix with positive diagonal entries
+  gp.B = normrnd(0, 1, [m, m]);
+  gp.B(1:m+1:end) = random('Chisquare', 4, [m, 1]);
+  gp.B = tril(gp.B);
+%  gp.D = diag(normrnd(0, 1));
+  gp.D = zeros(m, m);
+  % TODO: B could be m x p for p << m
+  % TODO: diag() might not work in that case, and need to update the initialisation
 
   d = size(feature_data.splittable, 2);
-  G = zeros(m + d + 1, 1);
+  G = zeros(2 * m + d + 1, 1);
   eta = 1;
 
   tic;
   while true
                  % Draw samples_count samples from the variational approximation
     [Kru, Kuu, Kuu_inv, KruKuu, Krr, Kuu_grad, Kru_grad] = vigpirlkernel(gp);
+    Sigma = gp.B * gp.B' + gp.D * gp.D;
+    Sigma_inv = inv(Sigma);
     z = mvnrnd(gp.mu', Sigma, algorithm_params.samples_count);
+    T = inv(gp.B * gp.B' + (gp.D * gp.D)');
     grad = full_gradient(Sigma, Sigma_inv, mdp_model, mdp_data,...
       example_samples, counts, gp.mu, Kru, Kuu, Kuu_inv, KruKuu, Krr, Kuu_grad,...
-      Kru_grad, z);
-    %disp(grad(1));
+      Kru_grad, z, T, gp.B);
     changes = grad(2:end);
     changes(1) = 0; % TEMP
     G = G + changes .^ 2;
@@ -57,7 +60,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     end
   end
   time = toc;
-                                % TODO: stopping condition
 
                                 % Return corresponding reward function.
   r = KruKuu * gp.mu;
[…]

### 2019-01-06 code/VIGPIRL/vigpirltransfer.m

@@ -1,5 +1,5 @@
 % Transfer learned reward function to a new state space.
-function irl_result = gpirltransfer(prev_result,mdp_data,mdp_model,...
+function irl_result = vigpirltransfer(prev_result,mdp_data,mdp_model,...
     feature_data,~,verbosity)
 
 % To transfer the result to the new state space, we must first compute

### 2019-01-06 code/VIGPIRL/vigpirlunpackparam.m

@@ -3,6 +3,7 @@ function [gp] = vigpirlunpackparam(gp, x)
 
 % Count the last index read.
 endi = length(x);
+m = length(gp.mu);
 
 gp.rbf_var = vigpirlhpxform(x(endi), [], 'exp', 1);
 endi = endi - 1;
@@ -11,5 +12,8 @@ endi = endi - 1;
 gp.inv_widths = vigpirlhpxform(x(endi-length(gp.inv_widths)+1:endi), [], 'exp', 1)';
 endi = endi - length(gp.inv_widths);
 
-gp.mu = x(endi-length(gp.mu)+1:endi);
-endi = endi - length(gp.mu);
+gp.mu = x(endi-m+1:endi);
+endi = endi - m;
+
+gp.B(1:m+1:end) = vigpirlhpxform(x(endi-m+1:endi), [], 'exp', 1)';
+endi = endi - m;

### 2019-01-11 code/VIGPIRL/tests.m
ELBO is increasing in a very simple test

@@ -70,3 +70,9 @@ function test_remaining_vigpirlkernel(testCase)
                        Kru, Kuu, Kuuinv, KruKuu, Krr, Kuu_grad, Kru_grad, u);
   verifyEqual(testCase, grad, [3; 11; -2; 1; 4] - 0.5 * [-14; 35; 0; -28; 42]);
 end
+
+function test_hyperparameter_packing(testCase)
+  gp = struct('rbf_var', 0, 'inv_widths', [1 2],...
+    'B', [3 0 0; 4 5 0; 6 7 8], 'mu', [-1; -2; -3]);
+  verifyEqual(testCase, vigpirlunpackparam(gp, vigpirlpackparam(gp)), gp, 'AbsTol', 1e-10);
+end

### 2019-01-11 code/VIGPIRL/vigpirlgetstatistics.m

@@ -1,5 +1,5 @@
 % Pre-compute statistics from examples for GPIRL.
-function [mu_sa,init_s] = vigpirlgetstatistics(example_samples,mdp_data)
+function [mu_sa, init_s] = vigpirlgetstatistics(example_samples, mdp_data)
 
 % Constants.
 [states,actions,transitions] = size(mdp_data.sa_s);

### 2019-01-11 code/VIGPIRL/vigpirlinit.m

@@ -13,5 +13,7 @@ gp.X_u = feature_data.splittable;
 % Initialize hyperparameters.
 gp.noise_var = algorithm_params.noise_init;
 gp.rbf_var = random('Chisquare', algorithm_params.rbf_init);
+%gp.rbf_var = 1; % TEMP
 gp.inv_widths = random('Chisquare', algorithm_params.ard_init,...
     [1, size(feature_data.splittable, 2)]);
+%gp.inv_widths = ones(1, size(feature_data.splittable, 2));

### 2019-01-11 code/VIGPIRL/vigpirlpackparam.m

@@ -1,5 +1,5 @@
 % Place parameters from specified GP into parameter vector.
 function x = vigpirlpackparam(gp)
-  x = vigpirlhpxform(gp.rbf_var, [], 'exp', 3);
-  x = vertcat(x, vigpirlhpxform(gp.inv_widths, [], 'exp', 3)', gp.mu,...
-    vigpirlhpxform(diag(gp.B), [], 'exp', 3), get_lower_triangle(gp.B));
+  x = vertcat(vigpirlhpxform(gp.rbf_var, [], 'exp', 3),...
+    vigpirlhpxform(gp.inv_widths, [], 'exp', 3)',...
+    vigpirlhpxform(diag(gp.B), [], 'exp', 3), gp.mu, get_lower_triangle(gp.B));

### 2019-01-11 code/VIGPIRL/vigpirlrun.m

@@ -11,12 +11,13 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   algorithm_params = vigpirldefaultparams(algorithm_params);
 
                       % Get state-action counts and initial state distributions.
-  [mu_sa, counts] = vigpirlgetstatistics(example_samples,mdp_data);
+  [mu_sa, counts] = vigpirlgetstatistics(example_samples, mdp_data);
 
                                 % Create initial GP.
-  gp = vigpirlinit(algorithm_params,feature_data);
+  gp = vigpirlinit(algorithm_params, feature_data);
                                 % Choose inducing points.
-  gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
+  %gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
+  gp.X_u = gp.X; % TEMP
   m = size(gp.X_u, 1);
   gp.mu = rand(1, m)';
 
@@ -24,41 +25,85 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   gp.B = normrnd(0, 1, [m, m]);
   gp.B(1:m+1:end) = random('Chisquare', 4, [m, 1]);
   gp.B = tril(gp.B);
-%  gp.D = diag(normrnd(0, 1));
-  gp.D = zeros(m, m);
+  %gp.B = eye(m); % TEMP
+  %gp.D = diag(normrnd(0, 1));
+  gp.D = zeros(m, m); % TEMP
   % TODO: B could be m x p for p << m
-  % TODO: diag() might not work in that case, and need to update the initialisation
+  % diag() might not work in that case, and need to update the initialisation
 
   d = size(feature_data.splittable, 2);
+  elbo_list = [];
+
+  % for AdaGrad
   G = zeros(m + d + 1 + m*(m+1)/2, 1);
-  eta = 1e2;
+  eta = 0.1;
+
+  % for AdaDelta
+  %num_hyperparameters = m + d + 1 + m*(m+1)/2;
+  %E_g = zeros(num_hyperparameters, 1);
+  %E_x = zeros(num_hyperparameters, 1);
+  %epsilon = 1e-6;
+  %rho = 0.1;
 
   tic;
+  i = 0;
   while true
-                 % Draw samples_count samples from the variational approximation
+    % Compute the gradient
[…]

### 2019-01-11 code/VIGPIRL/vigpirlunpackparam.m

@@ -1,16 +1,14 @@
 % Place parameters from parameter vector into GP.
-function [gp] = vigpirlunpackparam(gp, x)
+function gp = vigpirlunpackparam(gp, x)
 
 % Count the last index read.
 endi = length(x);
 m = length(gp.mu);
 
-gp.rbf_var = vigpirlhpxform(x(endi), [], 'exp', 1);
-endi = endi - 1;
-
-% Read ARD kernel parameters.
-gp.inv_widths = vigpirlhpxform(x(endi-length(gp.inv_widths)+1:endi), [], 'exp', 1)';
-endi = endi - length(gp.inv_widths);
+lower_triangle_mask = tril(true(size(gp.B)), -1);
+lower_triangle_size = sum(lower_triangle_mask(:) == true);
+gp.B(lower_triangle_mask) = x(endi-lower_triangle_size+1:endi);
+endi = endi - lower_triangle_size;
 
 gp.mu = x(endi-m+1:endi);
 endi = endi - m;
@@ -18,7 +16,9 @@ endi = endi - m;
 gp.B(1:m+1:end) = vigpirlhpxform(x(endi-m+1:endi), [], 'exp', 1)';
 endi = endi - m;
 
-lower_triangle_mask = tril(true(size(gp.B)), -1);
-lower_triangle_size = sum(lower_triangle_mask(:) == true);
-gp.B(lower_triangle_mask) = x(endi-lower_triangle_size+1:endi);
-endi = endi - lower_triangle_size;
+% Read ARD kernel parameters.
+gp.inv_widths = vigpirlhpxform(x(endi-length(gp.inv_widths)+1:endi), [], 'exp', 1)';
+endi = endi - length(gp.inv_widths);
+
+gp.rbf_var = vigpirlhpxform(x(endi), [], 'exp', 1);
+endi = endi - 1;

### 2019-01-14 code/VIGPIRL/estimate_derivative.m
A talk on theory

@@ -14,7 +14,7 @@ function answer = estimate_derivative(u, KruKuu, Kuu, Kuu_inv, r_covariance_matr
     unique_mu_part = (u' - mu)' * (Sigma_inv + Sigma_inv');
     %unique_B_part = ((adjoint(T) + adjoint(Sigma)) / det(Sigma) + T*U*T +...
     %  Sigma_inv*U*Sigma_inv) * B;
-    unique_B_part = Sigma_inv*U*Sigma_inv*B - adjoint(B*B')*B/det(Sigma);
+    unique_B_part = Sigma_inv*U*Sigma_inv*B - adjoint(Sigma)*B/det(Sigma);
 
     unique_part = vertcat(2, unique_lambda_part', diag(unique_B_part),...
       unique_mu_part', get_lower_triangle(unique_B_part));

### 2019-01-14 code/VIGPIRL/vigpirldefaultparams.m

@@ -7,13 +7,11 @@ default_params = struct(...
     'ard_init',1,...
     'noise_init',1e-2,...
     'rbf_init',5,...
-    'warp_c_init',2.0,...
-    'warp_l_init',1,...
-    'warp_s_init',1,...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
-    'samples_count',100);
+    'samples_count',100,...
+    'learning_rate',1);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2019-01-14 code/VIGPIRL/vigpirlinit.m

@@ -16,4 +16,4 @@ gp.rbf_var = random('Chisquare', algorithm_params.rbf_init);
 %gp.rbf_var = 1; % TEMP
 gp.inv_widths = random('Chisquare', algorithm_params.ard_init,...
     [1, size(feature_data.splittable, 2)]);
-%gp.inv_widths = ones(1, size(feature_data.splittable, 2));
+%gp.inv_widths = ones(1, size(feature_data.splittable, 2)); % TEMP

### 2019-01-14 code/VIGPIRL/vigpirlrun.m

@@ -33,17 +33,18 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
 
   d = size(feature_data.splittable, 2);
   elbo_list = [];
+  grad_history = [];
+  hyperparameter_history = [];
 
   % for AdaGrad
   G = zeros(m + d + 1 + m*(m+1)/2, 1);
-  eta = 0.1;
 
   % for AdaDelta
   %num_hyperparameters = m + d + 1 + m*(m+1)/2;
   %E_g = zeros(num_hyperparameters, 1);
   %E_x = zeros(num_hyperparameters, 1);
   %epsilon = 1e-6;
-  %rho = 0.1;
+  %rho = 0.95;
 
   tic;
   i = 0;
@@ -61,10 +62,11 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     old_hyperparameters = vigpirlpackparam(gp);
     grad = transform_gradient(full_grad(2:end), old_hyperparameters, d, m);
 
-    %grad(1:d+1) = 0; % lambda
-    %grad(d+m+2:d+2*m+1) = 0; % mu
-    %grad(d+2:d+m+1) = 0; % B diagonal
-    %grad(d+2*m+2:end) = 0; % rest of B
+    %grad(1) = 0; % labmda_0
+    grad(2:d+1) = 0; % lambda (except first)
+    grad(d+m+2:d+2*m+1) = 0; % mu
+    grad(d+2:d+m+1) = 0; % B diagonal
+    grad(d+2*m+2:end) = 0; % rest of B
 
     fprintf('Hyperparameters:\n');
     %disp(old_hyperparameters(d+m+2:d+2*m+1));
@@ -73,18 +75,20 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     %disp(grad(d+m+2:d+2*m+1));
     disp(grad);
 
+    hyperparameter_history = horzcat(hyperparameter_history, old_hyperparameters);
+
     % for AdaGrad
     G = G + grad .^ 2;
-    rho = (eta / sqrt(G))';
+    rho = (algorithm_params.learning_rate / sqrt(G))';
 
     % for AdaDelta
     %E_g = rho * E_g + (1 - rho) * grad .^ 2;
     %delta = sqrt(E_x + epsilon) ./ sqrt(E_g + epsilon) .* grad;
[…]

### 2019-01-27 code/VIGPIRL/estimate_derivative.m
mu and lambda seem to be converging (although in a noisy way)

@@ -1,16 +1,26 @@
-function answer = estimate_derivative(u, KruKuu, Kuu, Kuu_inv, r_covariance_matrix,...
-                                      Kuu_derivatives, mdp_data, Sigma, Sigma_inv,...
+function answer = estimate_derivative(u, Krr, Kru, Kuu, Kuu_inv, r_covariance_matrix,...
+                                      Kuu_derivatives, Kru_derivatives, Krr_derivatives, mdp_data, Sigma, Sigma_inv,...
                                       mu, mdp_model, example_samples, counts, T, B)
              % Uses a point-sample of u to estimate the part of dL/dnu under E[]
+
+  function answer = lambda_derivative(i)
+    % Returns the derivative of the ELBO w.r.t. lambda_i.
+    % (Actually, the part of the derivative inside E[] without v)
+
+    Rr = Ss * Kru_derivatives(:, :, i) - Krr_derivatives(:, :, i) +...
+      (Kru_derivatives(:, :, i)' - Ss * Kuu_derivatives(:, :, i)) * Kuu_inv * Kru;
+    t = r' - Ss * u';
+    answer = trace(Rr * adjoint(Gamma)) / det(Gamma) - t' * Gamma_inv * Rr * Gamma_inv * t;
+  end
+
   function answer = expected_derivative_for(state_action)
+    % TODO: check unique_B_part. Doesn't match the PDF. Update the PDF?
     state = state_action(1);
     action = state_action(2);
     next_states = extract(mdp_data.sa_s, state, action);
     next_probabilities = extract(mdp_data.sa_p, state, action);
-    U = (u - mu) * (u - mu)';
 
-    unique_lambda_part = arrayfun(@(i) trace((Kuu_inv * (u' * u) * Kuu_inv' -...
-      Kuu_inv) * Kuu_derivatives(:, :, i)), 1:size(Kuu_derivatives, 3));
+    unique_lambda_part = arrayfun(@lambda_derivative, 1:size(Kuu_derivatives, 3));
     unique_mu_part = (u' - mu)' * (Sigma_inv + Sigma_inv');
     %unique_B_part = ((adjoint(T) + adjoint(Sigma)) / det(Sigma) + T*U*T +...
     %  Sigma_inv*U*Sigma_inv) * B;
[…]

### 2019-01-27 code/VIGPIRL/full_gradient.m

@@ -1,22 +1,25 @@
 function changes = full_gradient(Sigma, Sigma_inv, mdp_model,...
   mdp_data, example_samples, counts, mu, Kru, Kuu, Kuu_inv, KruKuu, Krr,...
-  Kuu_grad, Kru_grad, z, T, B)
+  Kuu_grad, Kru_grad, Krr_grad, z, T, B) % TODO: KruKuu is not needed
 
   r_covariance_matrix = Krr - KruKuu * Kru;
+  disp(r_covariance_matrix);
   grad_prediction_for_each_u = arrayfun(@(i) estimate_derivative(z(i, :),...
-    KruKuu, Kuu, Kuu_inv, r_covariance_matrix, Kuu_grad, mdp_data, Sigma, Sigma_inv,...
+    Krr, Kru, Kuu, Kuu_inv, r_covariance_matrix, Kuu_grad, Kru_grad, Krr_grad, mdp_data, Sigma, Sigma_inv,...
     mu, mdp_model, example_samples, counts, T, B), 1:size(z, 1), 'Uniform', 0);
   estimated_grad = mean(cat(3, grad_prediction_for_each_u{:}), 3);
 
-  not_estimated_lambda = arrayfun(@(i) -0.5 *...
-    trace(Kuu_inv * Kuu_grad(:, :, i)) + counts' * (Kru_grad(:, :, i)' -...
-    KruKuu * Kuu_grad(:, :, i)) * Kuu_inv * mu, 1:size(Kuu_grad, 3));
-  not_estimated_mu = counts' * Kru' * Kuu_inv;
+  not_estimated_lambda = arrayfun(@(i) counts' * (Kru_grad(:, :, i)' -...
+    Kru' * Kuu_inv * Kuu_grad(:, :, i)) * Kuu_inv * mu +...
+    0.5 * (trace(Kuu_inv * Kuu_grad(:, :, i) * Kuu_inv * Sigma) +...
+    mu' * Kuu_inv * Kuu_grad(:, :, i) * Kuu_inv * mu -...
+    trace(Kuu_inv * Kuu_grad(:, :, i))), 1:size(Kuu_grad, 3));
+  not_estimated_mu = (counts' * Kru' * Kuu_inv)' - 0.5 * (Kuu_inv + Kuu_inv') * mu;
   not_estimated_elbo = 0.5 * log(det(Sigma)) - 0.5 * log(det(Kuu)) +...
[…]

### 2019-01-27 code/singletest.m

@@ -9,7 +9,7 @@ addpaths;
 % DGP-IRL original
 %test_result = runtest('deepgpirl',struct(),'linearmdp','binaryworld',...
 %   struct('n',12),struct('training_sample_lengths',12^2,...
-%   'training_samples',8,'verbosity',1));
+%  'training_samples',8,'verbosity',1));
 
 % GPIRL's test using DGP-IRL
 %test_result = runtest('deepgpirl',struct('R', 4),'linearmdp','objectworld',...

### 2019-02-14 code/VIGPIRL/estimate_derivative.m
Updating the derivative test for lambda0. Code corresponds to the theory, but the algorithm still diverges

@@ -7,42 +7,40 @@ function answer = estimate_derivative(u, Krr, Kru, Kuu, Kuu_inv, r_covariance_ma
     % Returns the derivative of the ELBO w.r.t. lambda_i.
     % (Actually, the part of the derivative inside E[] without v)
 
-    Rr = Ss * Kru_derivatives(:, :, i) - Krr_derivatives(:, :, i) +...
-      (Kru_derivatives(:, :, i)' - Ss * Kuu_derivatives(:, :, i)) * Kuu_inv * Kru;
-    t = r' - Ss * u';
-    answer = trace(Rr * adjoint(Gamma)) / det(Gamma) - t' * Gamma_inv * Rr * Gamma_inv * t;
+    R = S * Kru_derivatives(:, :, i) - Krr_derivatives(:, :, i) +...
+      (Kru_derivatives(:, :, i)' - S * Kuu_derivatives(:, :, i)) * Kuu_inv * Kru;
+    t = r' - S * u';
+    answer = trace(R * adjoint(Gamma)) / det(Gamma) - t' * Gamma_inv * R * Gamma_inv * t;
   end
 
-  function answer = expected_derivative_for(state_action)
-    % TODO: check unique_B_part. Doesn't match the PDF. Update the PDF?
+  % The part of v specific to a single state-action pair
+  function elbo_part = expected_derivative_for(state_action)
     state = state_action(1);
     action = state_action(2);
     next_states = extract(mdp_data.sa_s, state, action);
     next_probabilities = extract(mdp_data.sa_p, state, action);
-
-    unique_lambda_part = arrayfun(@lambda_derivative, 1:size(Kuu_derivatives, 3));
-    unique_mu_part = (u' - mu)' * (Sigma_inv + Sigma_inv');
-    %unique_B_part = ((adjoint(T) + adjoint(Sigma)) / det(Sigma) + T*U*T +...
-    %  Sigma_inv*U*Sigma_inv) * B;
-    unique_B_part = Sigma_inv*U*Sigma_inv*B - adjoint(Sigma)*B/det(Sigma);
-
-    unique_part = vertcat(2, unique_lambda_part', diag(unique_B_part),...
-      unique_mu_part', get_lower_triangle(unique_B_part));
     s = next_probabilities' * solution.v(next_states, 1);
     elbo_part = solution.v(state, 1) - mdp_data.discount * s;
[…]

### 2019-02-14 code/VIGPIRL/full_gradient.m

@@ -3,11 +3,19 @@ function changes = full_gradient(Sigma, Sigma_inv, mdp_model,...
   Kuu_grad, Kru_grad, Krr_grad, z, T, B) % TODO: KruKuu is not needed
 
   r_covariance_matrix = Krr - KruKuu * Kru;
-  disp(r_covariance_matrix);
   grad_prediction_for_each_u = arrayfun(@(i) estimate_derivative(z(i, :),...
     Krr, Kru, Kuu, Kuu_inv, r_covariance_matrix, Kuu_grad, Kru_grad, Krr_grad, mdp_data, Sigma, Sigma_inv,...
     mu, mdp_model, example_samples, counts, T, B), 1:size(z, 1), 'Uniform', 0);
   estimated_grad = mean(cat(3, grad_prediction_for_each_u{:}), 3);
+
+  % TEMP FOR TESTING
+  %not_estimated_lambda1 = arrayfun(@(i) counts' * (Kru_grad(:, :, i)' -...
+  %  Kru' * Kuu_inv * Kuu_grad(:, :, i)) * Kuu_inv * mu, 1:size(Kuu_grad, 3));
+  %not_estimated_lambda2 = arrayfun(@(i) 0.5 * (trace(Kuu_inv * Kuu_grad(:, :, i) * Kuu_inv * Sigma) +...
+  %  mu' * Kuu_inv * Kuu_grad(:, :, i) * Kuu_inv * mu -...
+  %  trace(Kuu_inv * Kuu_grad(:, :, i))), 1:size(Kuu_grad, 3));
+  %disp(not_estimated_lambda1);
+  %disp(not_estimated_lambda2);
 
   not_estimated_lambda = arrayfun(@(i) counts' * (Kru_grad(:, :, i)' -...
     Kru' * Kuu_inv * Kuu_grad(:, :, i)) * Kuu_inv * mu +...
@@ -15,11 +23,16 @@ function changes = full_gradient(Sigma, Sigma_inv, mdp_model,...
     mu' * Kuu_inv * Kuu_grad(:, :, i) * Kuu_inv * mu -...
     trace(Kuu_inv * Kuu_grad(:, :, i))), 1:size(Kuu_grad, 3));
   not_estimated_mu = (counts' * Kru' * Kuu_inv)' - 0.5 * (Kuu_inv + Kuu_inv') * mu;
-  not_estimated_elbo = 0.5 * log(det(Sigma)) - 0.5 * log(det(Kuu)) +...
-    counts' * KruKuu * mu;
[…]

### 2019-02-14 code/VIGPIRL/tests.m

@@ -1,78 +1,0 @@
-function tests = tests
-  tests = functiontests(localfunctions);
-end
-
-function test_d_covariance_matrix_d_lambda_i(testCase)
-  answer = d_covariance_matrix_d_lambda_i(ones(3, 2), [2 0; 4 0; 8 0],...
-    [1 0; 3 0], zeros(3, 2), 1);
-  correct = -0.5 * [1 1; 9 1; 49 25];
-  verifyEqual(testCase, answer, correct);
-end
-
-function test_remaining_vigpirlkernel(testCase)
-  function c = k(x, y)
-    c = rbf * exp(-0.5 * 2 * (x - y)^2);
-  end
-
-  Xu = [0 0; 1 3];
-  X = [0 0; 1 0; 2 0];
-  noise = 0;
-  rbf = 0.1;
-  gp = struct('X_u', Xu, 'X', X, 'noise_var', noise, 'rbf_var', rbf,...
-    'inv_widths', [2 0]);
-
-  Kuu = [k(0, 0) k(1, 0); k(0, 1) k(1, 1)];
-  Kuf = [k(0, 0) k(1, 0) k(2, 0); k(1, 0) k(1, 1) k(1, 2)];
-  Kff = [k(0, 0) k(1, 0) k(2, 0); k(0, 1) k(1, 1) k(2, 1); k(0, 2) k(1, 2) k(2, 2)];
-  Kuuinv = inv(Kuu);
-  KruKuu = Kuf' * Kuuinv;
-
-  % Test vigpirlkernel
-  [K_uf, K_uu, invK, K_ufKinv, K_ff] = vigpirlkernel(gp);
-  verifyEqual(testCase, K_uf, Kuf, 'AbsTol', 1e-10);
-  verifyEqual(testCase, K_uu, Kuu, 'AbsTol', 1e-10);
-  verifyEqual(testCase, invK, Kuuinv, 'AbsTol', 1e-10);
-  verifyEqual(testCase, K_ufKinv, KruKuu, 'AbsTol', 1e-10);
-  verifyEqual(testCase, K_ff, Kff, 'AbsTol', 1e-10);
-
-  % Test estimate_derivative
-  u = [1 -2];
-  mdp = struct('sa_s', [2; 3; 1], 'sa_p', [1; 1; 1], 'discount', 0,...
-    'states', 3, 'actions', 1);
-  Kuuinv = eye(2);
-  Kuu = eye(2);
-  KruKuu = [1 2; 3 4; 5 6];
[…]

### 2019-02-14 code/VIGPIRL/vigpirldefaultparams.m

@@ -11,7 +11,9 @@ default_params = struct(...
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
     'samples_count',100,...
-    'learning_rate',1);
+    'learning_rate',0.1,...
+    'B_learning_rate',0.001,...
+    'num_iterations',50);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2019-02-14 code/VIGPIRL/vigpirlrun.m

@@ -24,7 +24,7 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
                  % B is a lower triangular matrix with positive diagonal entries
   gp.B = normrnd(0, 1, [m, m]);
   gp.B(1:m+1:end) = random('Chisquare', 4, [m, 1]);
-  %gp.B = tril(gp.B);
+  gp.B = tril(gp.B);
   gp.B = eye(m); % TEMP
   %gp.D = diag(normrnd(0, 1));
   gp.D = zeros(m, m); % TEMP
@@ -46,8 +46,8 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   %epsilon = 1e-6;
   %rho = 0.95;
 
-  tic;
   i = 0;
+  tic;
   while true
     % Compute the gradient
     [Kru, Kuu, Kuu_inv, KruKuu, Krr, Kuu_grad, Kru_grad, Krr_grad] = vigpirlkernel(gp);
@@ -62,17 +62,15 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     old_hyperparameters = vigpirlpackparam(gp);
     grad = transform_gradient(full_grad(2:end), old_hyperparameters, d, m);
 
-    %grad(1) = 0; % labmda_0
-    %grad(2:d+1) = 0; % lambda (except first)
-    %grad(d+m+2:d+2*m+1) = 0; % mu
+    %grad(1) = 0; % lambda_0
+    grad(2:d+1) = 0; % lambda (except first)
+    grad(d+m+2:d+2*m+1) = 0; % mu
     grad(d+2:d+m+1) = 0; % B diagonal
     grad(d+2*m+2:end) = 0; % rest of B
 
     fprintf('Hyperparameters:\n');
-    %disp(old_hyperparameters(d+m+2:d+2*m+1));
     disp(old_hyperparameters);
     fprintf('Gradient:\n');
-    %disp(grad(d+m+2:d+2*m+1));
     disp(grad);
 
     hyperparameter_history = horzcat(hyperparameter_history, old_hyperparameters);
@@ -86,13 +84,16 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     %delta = sqrt(E_x + epsilon) ./ sqrt(E_g + epsilon) .* grad;
     %E_x = rho * E_x + (1 - rho) * delta .^ 2;
 
+    % Make the derivative of B weaker
[…]

### 2019-02-14 code/singletest.m

@@ -26,8 +26,14 @@ addpaths;
 %printresult(test_result);
 %visualize(test_result);
 
-mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0, 'states', 2, 'actions', 1);
-feature_data = struct('splittable', [1; 2]);
-example_samples = {[1, 1]};
+%mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0.9, 'states', 2, 'actions', 1);
+%feature_data = struct('splittable', [1; 2]);
+%example_samples = {[1, 1]};
+
+mdp_data = struct('discount', 0, 'states', 3, 'actions', 2);
+mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
+mdp_data.sa_p(1:3, 1:2, 1) = 1;
+feature_data = struct('splittable', [1; 2; 3]);
+example_samples = {[1, 1], [3, 2]};
 
 vigpirlrun(struct(), mdp_data, 'linearmdp', feature_data, example_samples);

### 2019-02-14 code/tests.m

@@ -1,0 +1,89 @@
+function tests = tests
+  tests = functiontests(localfunctions);
+end
+
+function test_d_covariance_matrix_d_lambda_i(testCase)
+  answer = d_covariance_matrix_d_lambda_i(ones(3, 2), [2 0; 4 0; 8 0],...
+    [1 0; 3 0], zeros(3, 2), 1);
+  correct = -0.5 * [1 1; 9 1; 49 25];
+  verifyEqual(testCase, answer, correct);
+end
+
+function test_remaining_vigpirlkernel(testCase)
+  function c = k(x, y)
+    c = rbf * exp(-0.5 * 2 * (x - y)^2);
+  end
+
+  %Xu = [0 0; 1 3];
+  %X = [0 0; 1 0; 2 0];
+  %noise = 0;
+  %rbf = 0.1;
+  %gp = struct('X_u', Xu, 'X', X, 'noise_var', noise, 'rbf_var', rbf,...
+  %  'inv_widths', [2 0]);
+
+  %Kuu = [k(0, 0) k(1, 0); k(0, 1) k(1, 1)];
+  %Kuf = [k(0, 0) k(1, 0) k(2, 0); k(1, 0) k(1, 1) k(1, 2)];
+  %Kff = [k(0, 0) k(1, 0) k(2, 0); k(0, 1) k(1, 1) k(2, 1); k(0, 2) k(1, 2) k(2, 2)];
+  %Kuuinv = inv(Kuu);
+  %KruKuu = Kuf' * Kuuinv;
+
+  % Test vigpirlkernel
+  %[K_uf, K_uu, invK, K_ufKinv, K_ff] = vigpirlkernel(gp);
+  %verifyEqual(testCase, K_uf, Kuf, 'AbsTol', 1e-10);
+  %verifyEqual(testCase, K_uu, Kuu, 'AbsTol', 1e-10);
+  %verifyEqual(testCase, invK, Kuuinv, 'AbsTol', 1e-10);
+  %verifyEqual(testCase, K_ufKinv, KruKuu, 'AbsTol', 1e-10);
+  %verifyEqual(testCase, K_ff, Kff, 'AbsTol', 1e-10);
+
+  % Test estimate_derivative
+  example_samples = {[1, 1], [3, 2]};
+  counts = [1; 0; 1];
+  u = [1 -2 3];
+  mdp_data = struct('discount', 0, 'states', 3, 'actions', 2);
[…]

### 2019-03-04 mpaper/paper.tex
The beginning of the paper

@@
 \documentclass{mpaper}
+\usepackage{bm}
+\usepackage{bbm}
 
-\begin{document}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{definition}[theorem]{Definition}
+
+\DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\adj}{adj}
+\DeclareMathOperator{\tr}{tr}
+
+\newcommand{\Kuu}{\mathbf{K}_{\mathbf{u},\mathbf{u}}}
+\newcommand{\Krr}{\mathbf{K}_{\mathbf{r},\mathbf{r}}}
+\newcommand{\Kru}{\mathbf{K}_{\mathbf{r},\mathbf{u}}}
+
+\begin{document} % 14 pages
 
@@
 \matricnum{2146879}
-
 \maketitle
@@
 \section{Introduction}
-
-This paper outlines the standard template for an MSci submission.
-In earlier years, MSci students at the School of Computing
-Science\footnote{\url{http://www.dcs.gla.ac.uk}},
-University of Glasgow, were expected to produce a full-length
-dissertation. Now, the requirement is for MSci students to
-write a paper of up to 14 pages in length, using the supplied
-\texttt{mpaper} \LaTeX style file.
-
-The precise structure of an MSci paper is not mandated, but it should
-probably cover in detail the following aspects of the project.
-\begin{enumerate}
-\item General description of the problem, motivation, relevance
-\item Background information, possibly including a literature survey
-\item Description of approach taken to solve the problem, including
-  high-level design and lower-level implementation details as appropriate
-\item Evaluation, qualitative or quantitative as appropriate
-\item Conclusion, including scope for future work
-\end{enumerate}
+% problem description/statement
 
 \section{Background}
+% motivation, relevance (whatever that means)
+% comprehensive literature survey
 
-This \LaTeX template is based on the ACM \texttt{sig-alternate} class.
-The layout is two-column text. Generally figures and tables only
-extend to one column width, e.g.\ Table \ref{tab-eg},
-but it is possible to make them
-stretch over both columns using the \texttt{figure*} and
-\texttt{table*} environments. For an example, see Figure \ref{fig-eg}.
+\subsection{Linear Algebra and Numerical Analysis}
 
-\begin{table}
-\begin{tabular}{l||c||p{2cm}}
-\emph{Operating System} & \emph{Version} & \emph{Verdict} \\ \hline \hline
-Ubuntu & 12.04 & Everyone's favourite Linux, unless you grew up with
-RedHat \\ \hline
-Slackware & xxx & Pseudo-hacker's Linux, how often do you recompile
-your kernel? \\ \hline
-Mac OS & 10.7 & For people with more money than sense \\ \hline
-\end{tabular}
-\caption{\label{tab-eg}Single column table of figures}
-\end{table}
+\begin{definition}[Norms]
+  For any finite-dimensional vector $\mathbf{x} = (x_1, \dots, x_n)^\intercal$,
+  its \emph{maximum norm} is
+  \[
+    \lVert \mathbf{x} \rVert_\infty = \max_i |x_i|
+  \]
+  whereas its \emph{taxicab} (or \emph{Manhattan}) \emph{norm} is
+  \[
+    \lVert \mathbf{x} \rVert_1 = \sum_{i = 1}^n |x_i|.
+  \]
+  Let $\mathbf{A}$ be a matrix. For any vector norm $\lVert
+  \cdot \rVert_p$, we can also define its \emph{induced norm} for matrices as
+  \[
+    \lVert \mathbf{A} \rVert_p = \sup_{\mathbf{x} \ne \mathbf{0}} \frac{\lVert
+      \mathbf{Ax} \rVert_p}{\lVert \mathbf{x} \rVert_p}.
+  \]
+  In particular, for $p = \infty$, we have
+  \[
+    \lVert \mathbf{A} \rVert_\infty = \max_i \sum_{j} |A_{i,j}|.
+  \]
+\end{definition}
 
-\begin{figure*}
-\begin{center}
-\includegraphics[scale=0.3]{alice.pdf}
-\end{center}
-\caption{\label{fig-eg}An example figure stretching over two columns}
-\end{figure*}
+\begin{lemma}[Perturbation Lemma
+  \cite{layton2014numerical}] \label{prop:condition_number}
+  Let $\lVert \cdot \rVert$ be any matrix norm, and let $\mathbf{A}$ and
+  $\mathbf{E}$ be matrices such that $\mathbf{A}$ is invertible and $\lVert
+  \mathbf{A}^{-1} \rVert \lVert \mathbf{E} \rVert < 1$, then $\mathbf{A} +
+  \mathbf{E}$ is invertible, and
+  \[
+    \lVert (\mathbf{A} + \mathbf{E})^{-1} \rVert \le \frac{\lVert
+      \mathbf{A}^{-1} \rVert}{1 - \lVert \mathbf{A}^{-1} \rVert \lVert
+      \mathbf{E} \rVert}.
+  \]
+\end{lemma}
 
 \section{The WizWoz System}
+% solution
+% TODO: express all lambdas as exponentials (update the derivatives or sth)
 
-Again, Simon Peyton Jones has a lovely description of how to write a
-paper on his
-website\footnote{\url{http://research.microsoft.com/en-us/um/people/simonpj/papers/giving-a-talk/giving-a-talk.htm}}.
-Personally, I put URLs in footnotes and \emph{bona fide} references
-in the bibliography. How many references? Hard to say. Five is not enough, 50 is
-pushing it.
+\subsection{Notation}
 
-\subsection{User Interface}
+For any matrix $\mathbf{A}$, we will use either $A_{i,j}$ or
+$[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
+$j$. Moreover, we use $\tr(\mathbf{A})$ to denote its \emph{trace} and
+$\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}).
 
-Blah blah blah
-Blah blah blah
-Blah blah blah
-Blah blah blah
+For any vector $\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a
+vector space of polynomials with degree at most $d$, where variables are
+elements of $\mathbf{x}$, and coefficients are in $\mathbb{R}$.
 
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-\subsection{Foo}
+\subsection{Preliminaries}
 
-Blah blah blah
-Blah blah blah
-Blah blah blah
+In this paper, all references to measurability are with respect to the Lebesgue
+measure. Similarly, whenever we consider the existence of an integral, we use
+the Lebesgue definition of integration.
 
-\subsubsection{Bar}
-\textsf{Blah} \textit{blah} \textbf{blah}
+As recently suggested by Ong et al. \cite{ong2018gaussian}, we use a
+decomposition $\bm\Sigma = \mathbf{B}\mathbf{B}^\intercal + \mathbf{D}^2$, where
+$\mathbf{B}$ is a lower triangular $m \times p$ matrix with positive diagonal
+entries, and $\mathbf{D} = \diag(d_1, \dots, d_m)$. Typically, we would set $p$
+so that $p \ll m$ to get an efficient approximation, but it is also worth
+pointing out that we can retain full accuracy by setting $p = m$ and $\mathbf{D}
+= \mathbf{O}$. Moreover, we define a few matrices that will simplify expressions
+for the derivatives:
+\begin{align*}
+  \mathbf{U} &= (\mathbf{u} - \bm\mu)(\mathbf{u} - \bm\mu)^\intercal \\
+  \mathbf{S} &= \Kru^\intercal\Kuu^{-1}, \\
+  \bm\Gamma &= \Krr - \mathbf{S}\Kru, \\
+  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru.
+\end{align*}
 
-\section{Evaluation}
+Derivatives such as $\frac{\partial \Kuu}{\partial \lambda_i}$ can be expressed
+as
+\[
+  \frac{\partial \Kuu}{\partial \lambda_i} = \frac{1}{\lambda_i}\Kuu
+\]
 
-Graphs are always good. I recommend getting to grips with Matlab, R or
-gnuplot rather than exporting horribly Excel bitmapped graphs.
+if $i = 0$, and
 
-\begin{quotation}
- The Assyrian came down like the wolf on the fold,
- And his cohorts were gleaming in purple and gold;
- And the sheen of their spears was like stars on the sea,
- When the blue wave rolls nightly on deep Galilee.
+\[
+  \begin{split}
+    \left[ \frac{\partial \Kuu}{\partial \lambda_i} \right]_{j,k} =
+    k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+    \left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(x_{\mathbf{u},j,i} -
+    x_{\mathbf{u},k,i})^2 \\
+    &- \mathbbm{1}[j \ne k]\sigma^2 \left. \vphantom{\frac{1}{2}} \right)
+  \end{split}
+\]
+otherwise.
 
- Like the leaves of the forest when Summer is green,
- That host with their banners at sunset were seen:
- Like the leaves of the forest when Autumn hath blown,
- That host on the morrow lay withered and strown.
-\end{quotation}
+\begin{lemma}[Derivatives of probability
+  distributions] \label{lemma:derivatives}
+  \leavevmode
+  \begin{enumerate}
+  \item $\frac{\partial q(\mathbf{u})}{\partial \bm\mu} =
+    \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} + \bm\Sigma^{-\intercal})(\mathbf{u}
+    - \bm\mu)$.
+  \item
+    \begin{enumerate}
+    \item
+      $\frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} =
+      \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
+      - \bm\Sigma^{-\intercal})$.
+    \item
+      $\frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} =
+      q(\mathbf{u})(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
+      |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}$.
+%    \item
+%      $\frac{\partial q(\mathbf{u})}{\partial \mathbf{D}} =
+%      \frac{1}{2}q(\mathbf{u}) \{ \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1}\mathbf{D}^\intercal +
+%      \mathbf{D}^\intercal\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
+%      |\bm\Sigma|^{-1}(\adj(\bm\Sigma^{-1})\mathbf{D}^\intercal +
+%      \mathbf{D}^\intercal \adj(\bm\Sigma^{-1})) \}$.
+    \end{enumerate}
+  \item For $i = 0, \dots, d$,
+    \[
+      \begin{split}
+        \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &=
+        \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1}
+          \tr(\mathbf{R} \adj(\bm\Gamma)) \\
+          &- (\mathbf{r} -
+          \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r}
+          - \mathbf{Su})),
+      \end{split}
+    \]
+  \end{enumerate}
+\end{lemma}
+
+\section{Evaluation} % showing deep understanding
 
@@
 
-The standard Lorem Ipsum passage, used since the 1500s
+\subsection{Future Work}
 
-``Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.''
-\vskip8pt \noindent
-{\bf Acknowledgments.}
-This is optional; it is a location for you to thank people
-
+%\vskip8pt \noindent
+%{\bf Acknowledgments.}
+%This is optional; it is a location for you to thank people
 \bibliographystyle{abbrv}
 \bibliography{paper}
-
 \end{document}

### 2019-03-04 code/VIGPIRL/full_gradient.m

@@ -48,6 +48,8 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   S = matrices.Kru' * Kuu_inv;
   % we need to round Gamma because otherwise it's not symmetric due to numerical errors
   Gamma = round(matrices.Krr - S * matrices.Kru, 4);
+  %disp(Gamma);
+  %disp(cond(Gamma));
   Gamma_inv = inv(Gamma);
 
   grad_prediction_for_each_u = arrayfun(@(i) estimate_derivative(z(i, :)), 1:size(z, 1), 'Uniform', 0);
@@ -74,8 +76,8 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   kl_derivative = 0.5 * (trace(Kuu_inv * matrices.Kuu_grad(:, :, 1) * Kuu_inv * Sigma) +...
     gp.mu' * Kuu_inv * matrices.Kuu_grad(:, :, 1) * Kuu_inv * gp.mu -...
     trace(Kuu_inv * matrices.Kuu_grad(:, :, 1)));
-  fprintf('d/d_lambda0 of E[log p(D | r)]: %f\n', p_derivative);
-  fprintf('average v: %f\n', 0.5 * estimated_grad(1));
+  %fprintf('d/d_lambda0 of E[log p(D | r)]: %f\n', p_derivative);
+  %fprintf('average v: %f\n', 0.5 * estimated_grad(1));
   %fprintf('d/d_lambda0 of -KL(q(u) || p(u)): %f\n', kl_derivative);
   %fprintf('their sum: %f\n', changes(2));
   %assert(abs(p_derivative + kl_derivative - changes(2)) < 1e-5);

### 2019-03-04 code/VIGPIRL/vigpirldefaultparams.m

@@ -10,10 +10,11 @@ default_params = struct(...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
-    'samples_count',100,...
-    'learning_rate',1,...
+    'samples_count',1000,...
+    'learning_rate',0.1,...
     'B_learning_rate',0.001,...
-    'num_iterations',100);
+    'num_iterations',20,...
+    'required_precision', 0.001);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2019-03-04 code/VIGPIRL/vigpirlrun.m

@@ -22,10 +22,10 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   gp.mu = rand(1, m)';
   %gp.mu = [-2; 3; 3]; % TEMP
                  % B is a lower triangular matrix with positive diagonal entries
-  gp.B = normrnd(0, 1, [m, m]);
-  gp.B(1:m+1:end) = random('Chisquare', 4, [m, 1]);
-  gp.B = tril(gp.B);
-  %gp.B = eye(m); % TEMP
+  %gp.B = normrnd(0, 1, [m, m]);
+  %gp.B(1:m+1:end) = random('Chisquare', 4, [m, 1]);
+  %gp.B = tril(gp.B);
+  gp.B = eye(m); % TEMP
   %gp.D = diag(normrnd(0, 1));
   gp.D = zeros(m, m); % TEMP
 
@@ -48,11 +48,11 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, zz, matrices);
 
     % Disable gradients that don't work
-    %grad(1) = 0; % lambda0
-    %grad(2:d+1) = 0; % lambda (except first)
+    grad(1) = 0; % lambda0
+    grad(2:d+1) = 0; % lambda (except first)
     %grad(d+m+2:d+2*m+1) = 0; % mu
-    %grad(d+2:d+m+1) = 0; % B diagonal
-    %grad(d+2*m+2:end) = 0; % rest of B
+    grad(d+2:d+m+1) = 0; % B diagonal
+    grad(d+2*m+2:end) = 0; % rest of B
     %disp(grad(1));
 
     elbo_list = horzcat(elbo_list, elbo);
@@ -68,32 +68,65 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
 
   parameter_vector = vigpirlpackparam(gp);
 
-  %for lambda1 = 1:100
-  %  parameter_vector(2) = lambda1/10;
-  %  wrapper(parameter_vector);
+  %mu1_log = [];
+  %mu2_log = [];
+  %mu3_log = [];
+  %multiplier = 10;
+  %for mu1 = 1:10
+  %  real_mu1 = multiplier * (mu1 - 5);
+  %  parameter_vector(d+m+2) = real_mu1;
+  %  for mu2 = 1:10
[…]

### 2019-03-06 mpaper/paper.tex
Transferred almost all notes to the paper

@@
 \usepackage{bbm}
+\usepackage{mathtools}
 
@@
 \newtheorem{definition}[theorem]{Definition}
+\newtheorem{remark}[theorem]{Remark}
 
@@
 
+\DeclarePairedDelimiterX{\infdivx}[2]{(}{)}{%
+  #1\;\delimsize\|\;#2%
+}
+\newcommand{\DKL}{D_{\mathrm{KL}}\infdivx}
+\newcommand{\Eq}{\mathbb{E}_{(\mathbf{u}, \mathbf{r}) \sim \approximation}}
+\newcommand{\V}{V_{\mathbf{r}}}
+\newcommand{\dx}{\,d\mathbf{r}\,d\mathbf{u}}
+\newcommand{\vbound}{\frac{\rinf + \log|\mathcal{A}|}{1 - \gamma}}
+\newcommand{\rinf}{\lVert \mathbf{r} \rVert_\infty}
+
+\newcommand{\f}{f(\mathbf{r}, \mathbf{u}, t)}
+\newcommand{\fn}{f_n(\mathbf{r}, \mathbf{u})}
+\newcommand{\ftn}{f(\mathbf{r}, \mathbf{u}, t_n)}
+\newcommand{\g}{g(\mathbf{r}, \mathbf{u})}
+
 \newcommand{\Kuu}{\mathbf{K}_{\mathbf{u},\mathbf{u}}}
@@
 \newcommand{\Kru}{\mathbf{K}_{\mathbf{r},\mathbf{u}}}
+
+\newcommand{\pfull}{p(\mathcal{D}, \mathbf{u}, \mathbf{r})}
+\newcommand{\approximation}{q(\mathbf{u}, \mathbf{r})}
+
+\newcommand{\dm}{\frac{\partial}{\partial\bm\mu}}
+\newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
+\newcommand{\dB}{\frac{\partial}{\partial\mathbf{B}}}
+\newcommand{\dl}{\frac{\partial}{\partial \lambda_i}}
+\newcommand{\dlj}{\frac{\partial}{\partial \lambda_j}}
+\newcommand{\dt}{\frac{\partial}{\partial t}}
+\newcommand{\df}{\left.\frac{\partial f}{\partial t}\right|_{(\mathbf{r},
+    \mathbf{u}, t)}}
 
@@
 elements of $\mathbf{x}$, and coefficients are in $\mathbb{R}$.
+
+We primarily think of rewards as a vector $\mathbf{r} \in
+\mathbb{R}^{|\mathcal{S}|}$, but sometimes we use a function notation $r(s)$ to
+denote the reward of a particular state $s \in \mathcal{S}$.
 
@@
 
+\subsection{Evidence Lower Bound}
+
+\begin{equation} \label{eq:elbo}
+  \begin{split}
+    \mathcal{L} &= \Eq \left[ \log \frac{\pfull}{\approximation}
+    \right] \\
+    &= \iint \approximation \log
+    \frac{\pfull}{\approximation}\dx.
+  \end{split}
+\end{equation}
+
+\begin{equation} \label{eq:full}
+  \pfull = p(\mathbf{u}) \times p(\mathbf{r} \mid \mathbf{u}) \times p(\mathcal{D} \mid \mathbf{r}).
+\end{equation}
+
+\begin{equation} \label{eq:approximation}
+  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}).
+\end{equation}
+
+In this section we derive and simplify the ELBO for this (now fully specified)
+model. In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
+write\footnote{At this point, we will drop the subscript denoting which
+  variables the expectation is taken over. Also note that throughout the
+  derivation equality is taken to mean `equality up to an additive constant'.}
+\[
+  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
+\]
+By substituting in \eqref{eq:full} and \eqref{eq:approximation}, we get
+\[
+  \begin{split}
+    \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid
+    \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
+    &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
+  \end{split}
+\]
+Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
+cancel each other out. Also notice that
+\[
+  \begin{split}
+    \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &=
+    -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
+    &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) +
+      \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
+      &+ \log |\Kuu| - \log |\bm\Sigma|),
+  \end{split}
+\]
+by the definition of KL divergence between two multivariate normal distributions
+\cite{kl}. Hence,
+\[
+  \begin{split}
+    \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T
+      Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
+    &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) +
+      \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
+  \end{split}
+\]
+Using the expressions for $Q_{\mathbf{r}}$ we get
+\[
+  \begin{split}
+  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) -
+    \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
+    s')\V(s') \right] \\
+  &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) +
+    \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
+  \end{split}
+\]
+We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
+$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
+number of times the state associated with the reward $r_i$ has been visited
+across all demonstrations. Then
+\[
+  \begin{split}
+  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &=
+  \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] =
+  \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
+  &=
+  \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] =
+  \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
+  \end{split}
+\]
+This allows us to simplify $\mathcal{L}$ to
+\[
+  \begin{split}
+  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu -
+  \mathbb{E}[v] \\
+  &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) +
+    \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
+  \end{split}
+\]
+where
+\[
+v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in
+  \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
+\]
+
+\subsection{Theoretical Justification}
+
+MDP values are characterised by both a state and a reward function/vector. In
+order to prove the next theorem, we think of the value function as $V :
+\mathcal{S} \to \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a
+state $s \in \mathcal{S}$ and returns a function $V(s) :
+\mathbb{R}^{\mathcal{S}} \to \mathbb{R}$ that takes a reward vector $\mathbf{r}
+\in \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$,
+$\V(s) \in \mathbb{R}$. The function $V(s)$ computes the values of
+all states and returns the value of state $s$.
+
+\begin{proposition} \label{thm:measurability}
+  MDP value functions $V(s) : \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ (for $s
+  \in \mathcal{S}$) are Lebesgue measurable.
+\end{proposition}
+\begin{proof}
+  For any reward vector $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$, the
+  collection of converged value functions $\{ \V(s) \mid s \in
+  \mathcal{S} \}$ satisfy
+  \begin{equation} \label{eq:linear_mdp}
+    \V(s) = \log \sum_{a \in \mathcal{A}}
+    \exp\left( r(s) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a,
+      s')\V(s') \right)
+  \end{equation}
+  for all $s \in \mathcal{S}$. Let $s_0 \in \mathcal{S}$ be an arbitrary state.
+  In order to prove that $V(s_0)$ is measurable, it is enough to show that for
+  any $\alpha \in \mathbb{R}$, the set
+  \[
+    \left\{ \mathbf{r} \in \mathbb{R}^{|\mathcal{S}|} \; \middle|
+    \begin{split}
+      \;&\V(s_0) \in (-\infty, \alpha); \\
+      \;&\V(s) \in \mathbb{R} \text{ for all } s \in \mathcal{S} \setminus \{ s_0 \}; \\
+      \;&\text{ Equation \ref{eq:linear_mdp} is satisfied by all } s \in
+      \mathcal{S}
+    \end{split}
+    \right\}
+  \]
+  is measurable. Since this set can be constructed in Zermelo-Fraenkel set
+  theory \emph{without} the axiom of choice, it is measurable
+  \cite{herrlich2006axiom}, which proves that $V(s)$ is a measurable function
+  for any $s \in \mathcal{S}$.
+\end{proof}
+
+\begin{proposition} \label{thm:bound}
+  If the initial values of the MDP value function satisfy the following
+  bound, then the bound remains satisfied throughout value iteration:
+  \begin{equation} \label{eq:bound}
+    |\V(s)| \le \vbound.
+  \end{equation}
+\end{proposition}
+\begin{proof}
+  We begin by considering \eqref{eq:bound} without taking the absolute value of
+  $\V(s)$, i.e.,
+  \begin{equation} \label{eq:positive_bound}
+    \V(s) \le \vbound,
+  \end{equation}
+  and assuming that the initial values of $\{ \V(s) \mid s \in
+  \mathcal{S} \}$ already satisfy \eqref{eq:positive_bound}. For each $s \in
+  \mathcal{S}$, the value of $\V(s)$ is updated via this rule:
+  \[
+    \V(s) \coloneqq \log \sum_{a \in \mathcal{A}} \exp\left( r(s) +
+      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s')\V(s')
+    \right).
+  \]
+  Note that both $\log$ and $\exp$ are increasing functions, $\gamma > 0$, and
+  the $\mathcal{T}$ function gives a probability (a non-negative number).
+  Thus
+  \[
+    \begin{split}
+      \V(s) &\le \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s'
+          \in \mathcal{S}} \mathcal{T}(s, a, s')\frac{\rinf +
+          \log|\mathcal{A}|}{1 - \gamma} \right) \\
+      &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf +
+          \log|\mathcal{A}|)}{1 - \gamma}\sum_{s' \in \mathcal{S}}
+        \mathcal{T}(s, a, s') \right) \\
+      &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf +
+          \log|\mathcal{A}|)}{1 - \gamma} \right)
+    \end{split}
+  \]
+  by the definition of $\mathcal{T}$. Then
+  \[
+    \begin{split}
+      \V(s) &\le \log \left( |\mathcal{A}| \exp\left( r(s) + \frac{\gamma
+            (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
+      &= \log \left( \exp\left( \log|\mathcal{A}| + r(s) + \frac{\gamma
+            (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
+      &= \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \\
+      &= \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 -
+        \gamma)(\log|\mathcal{A}| + r(s))}{1 - \gamma} \\
+      &\le \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 -
+        \gamma)(\log|\mathcal{A}| + \rinf)}{1 - \gamma} \\
+      &= \vbound
+    \end{split}
+  \]
+  by the definition of $\rinf$.
+
+  The proof for
+  \begin{equation} \label{eq:negative_bound}
+    \V(s) \ge \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1}
+  \end{equation}
+  follows the same argument until we get to
+  \[
+    \begin{split}
+      \V(s) &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma -
+        1)(\log|\mathcal{A}| + r(s))}{\gamma - 1} \\
+      &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma -
+        1)(-\log|\mathcal{A}| -\rinf)}{\gamma - 1} \\
+      &= \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1},
+    \end{split}
+  \]
+  where we use the fact that $r(s) \ge -\rinf - 2\log|\mathcal{A}|$. Combining
+  \eqref{eq:positive_bound} and \eqref{eq:negative_bound} gives
+  \eqref{eq:bound}.
+\end{proof}
+
+\begin{theorem}[Dominated Convergence Theorem
+  \cite{royden2010real}] \label{thm:lebesgue}
+  Let $(X, \mathcal{M}, \mu)$ be a measure space and $\{ f_n \}$ a sequence of
+  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise a.e. on $X$
+  and the function $f$ is measurable. Assume there is a non-negative function
+  $g$ that is integrable over $X$ and dominates the sequence $\{ f_n \}$ on $X$
+  in the sense that
+  \[
+    |f_n| \le g \text{ a.e. on $X$ for all $n$.}
+  \]
+  Then $f$ is integrable over $X$ and
+  \[
+    \lim_{n \to \infty} \int_X f_n\,d\mu = \int_X f\,d\mu.
+  \]
+\end{theorem}
+
+\begin{lemma} \label{lemma:bound1} % TODO: Lemma 1.1 is wrong about the
+                                % derivative. Fix
+  Let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
+  \mathbb{R}$ be an arbitrary bounded function. Then, for $i = 0,
+  \dots, d$,
+  \[
+    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} \right|_{\lambda_i
+      = c(\mathbf{r}, \mathbf{u})}
+  \]
+  has upper and lower bounds of the form $q(\mathbf{r} \mid \mathbf{u})d(\mathbf{u})$, where
+  $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
+\end{lemma}
+\begin{proof}
+  Remember that
+  \[
+    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
+    \frac{1}{2}q(\mathbf{r} \mid \mathbf{u})\tr
+    \left((\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal} - \Kuu^{-1})
+      \frac{\partial \Kuu}{\partial \lambda_i}
+    \right)
+  \]
+  by Lemma \ref{lemma:derivatives}. We begin by producing constant upper and
+  lower bounds for the elements of
+  \[
+    \left. \frac{\partial \Kuu}{\partial \lambda_i} \right|_{\lambda_i =
+      c(\mathbf{r}, \mathbf{u})}.
+  \]
+  If $i = 0$, then each element of $\frac{\partial
+    \Kuu}{\partial \lambda_0}$ is of the form
+  \[
+    \exp \left( -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal
+      \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne
+      k]\sigma^2\tr(\bm\Lambda) \right),
+  \]
+  i.e., without $\lambda_0$, so
+  \[
+    \left. \frac{\partial \Kuu}{\partial \lambda_0} \right|_{\lambda_0 =
+      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \Kuu}{\partial \lambda_0}
+  \]
+  is already independent of $\mathbf{r}$ and $\mathbf{u}$---there is no need
+  for any bounds.
+
+  If $i > 0$, then each element of $\frac{\partial \Kuu}{\partial
+    \lambda_i}$ is a constant multiple of $k_{\bm\lambda}(\mathbf{x}_j,
+  \mathbf{x}_k)$, for some $\mathbf{x}_j$ and $\mathbf{x}_k$. Since
+  $k_{\bm\lambda}(\mathbf{x}_j, \mathbf{x}_k)$ is a decreasing function of
+  $\lambda_i$, and $c(\mathbf{r}, \mathbf{u}) > a$,
+  \[ %TODO: what the hell do I do here?
+    \begin{split}
+      k_{\bm\lambda}(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = 
+        c(\mathbf{r}, \mathbf{u})} &=
+      \begin{multlined}[t]
+        \lambda_0 \exp \left( \vphantom{\sum_{n \in \{ \} \setminus}}
+        \right. -\frac{1}{2}c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2 \\
+        - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) \\
+        - \left. \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}}
+          \frac{1}{2} \lambda_n(x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne
+          k]\sigma^2 \lambda_n \right)
+      \end{multlined} \\
+      &<
+      \begin{multlined}[t]
+        \lambda_0 \exp \left( \vphantom{\sum_{n \in \{ \} \setminus}}
+        \right. -\frac{1}{2}a(x_{j,i} - x_{k,i})^2 -
+        \mathbbm{1}[j \ne k]\sigma^2a \\
+        - \left. \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}}
+          \frac{1}{2} \lambda_n(x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne
+          k]\sigma^2 \lambda_n \right),
+      \end{multlined}
+    \end{split}
+  \]
+  which gives an upper bound on each element of
+  \[
+    \left. \frac{\partial \Kuu}{\partial \lambda_i} \right|_{\lambda_i =
+      c(\mathbf{r}, \mathbf{u})}.
+  \]
+  A similar line of reasoning establishes lower bounds as well.
+
+  Combining the bounds with the observation that
+  every element of
+  $\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal}$ is in
+  $\mathbb{R}_2[\mathbf{u}]$ gives the required result.
+\end{proof}
+
+\begin{remark}
+  In order to find $\frac{\partial q(\mathbf{u})}{\partial t}$,
+  where $t$ is the $i$th element of the vector $\bm\mu$, we can
+  find $\frac{\partial q(\mathbf{u})}{\partial \bm\mu}$ and simply take the
+  $i$th element. A similar line of reasoning applies to matrices as well. Thus,
+  we only need to consider derivatives with respect to $\bm\mu$ and
+  $\bm\Sigma$.
+\end{remark}
+
+\begin{lemma} \label{lemma:bound2}
+  Let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
+  \mathbb{R}$ be an arbitrary bounded function. Then, for $i = 1, \dots, m$,
+  every element of
+  \[
+    \left. \frac{\partial q(\mathbf{u})}{\partial \bm\mu} \right|_{\mu_i =
+      c(\mathbf{r}, \mathbf{u})}
+  \]
+  has upper and lower bounds of the form $q(\mathbf{u})d(\mathbf{u})$,
+  where $d(\mathbf{u}) \in \mathbb{R}_1[\mathbf{u}]$.
+\end{lemma}
+\begin{proof}
+  Using Lemma \ref{lemma:derivatives},
+  \[
+    \left. \frac{\partial q(\mathbf{u})}{\partial \bm\mu} \right|_{\mu_i =
+      c(\mathbf{r}, \mathbf{u})} = \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+    \bm\Sigma^{-\intercal})(\mathbf{u} - \mathbf{c}(\mathbf{r}, \mathbf{u})),
+  \]
+  where $\mathbf{c}(\mathbf{r}, \mathbf{u}) = (\mu_1, \dots, \mu_{i - 1},
+  c(\mathbf{r}, \mathbf{u}), \mu_{i + 1} \dots, \mu_m)^\intercal$. Since
+  $c(\mathbf{r}, \mathbf{u})$ is bounded and $\bm\Sigma^{-1} +
+  \bm\Sigma^{-\intercal}$ is a constant matrix, we can use the bounds on
+  $c(\mathbf{r}, \mathbf{u})$ to manufacture both upper and lower bounds on
+  \[
+     \left. \frac{\partial q(\mathbf{u})}{\partial \bm\mu} \right|_{\mu_i =
+      c(\mathbf{r}, \mathbf{u})}
+  \]
+  of the required form.
+\end{proof}
+
+\begin{lemma} \label{lemma:bound3}
+  Let $i, j = 1, \dots, m$, and let $\epsilon > 0$ be arbitrary. Furthermore,
+  let
+  \[
+    c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\Sigma_{i,j} - \epsilon,
+    \Sigma_{i,j} + \epsilon) \subset \mathbb{R}
+  \]
+  be a function with a codomain arbitrarily close to $\Sigma_{i,j}$. Then every
+  element of
+  \[
+    \left. \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} \right|_{\Sigma_{i,j} =
+    c(\mathbf{r}, \mathbf{u})}
+  \]
+  has upper and lower bounds of the form $q(\mathbf{u})d(\mathbf{u})$, where
+  $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
+\end{lemma}
+\begin{proof}
+  Using Lemma \ref{lemma:derivatives},
+  \[
+    \left. \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} \right|_{\Sigma_{i,j} =
+    c(\mathbf{r}, \mathbf{u})} =
+    \frac{1}{2}q(\mathbf{u})(\mathbf{C}(\mathbf{r},
+    \mathbf{u})^{-\intercal}\mathbf{UC}(\mathbf{r}, \mathbf{u})^{-\intercal} -
+    \mathbf{C}(\mathbf{r}, \mathbf{u})^{-\intercal}),
+  \]
+  where
+  \[
+    [\mathbf{C}(\mathbf{r}, \mathbf{u})]_{k,l} =
+    \begin{cases}
+      c(\mathbf{r}, \mathbf{u}) & \text{if } (k, l) = (i, j), \\
+      \Sigma_{k,l} & \text{otherwise.}
+    \end{cases}
+  \]
+  We can also express $\mathbf{C}(\mathbf{r},\mathbf{u})$ as
+  $\mathbf{C}(\mathbf{r}, \mathbf{u}) = \bm\Sigma + \mathbf{E}(\mathbf{r},
+  \mathbf{u})$, where
+  \[
+    [\mathbf{E}(\mathbf{r}, \mathbf{u})]_{k,l} =
+    \begin{cases}
+      c(\mathbf{r}, \mathbf{u}) - \Sigma_{i,j} & \text{if } (k, l) = (i, j), \\
+      0 & \text{otherwise.}
+    \end{cases}
+  \]
+  We begin by establishing upper and lower bounds on $\mathbf{C}(\mathbf{r},
+  \mathbf{u})^{-1}$. For this, we use the maximum norm $\lVert \cdot
+  \rVert_\infty$ on both vectors and matrices. We can apply Lemma
+  \ref{prop:condition_number} to $\bm\Sigma$ and $\mathbf{E}(\mathbf{r},
+  \mathbf{u})$ since
+  \[
+    \lVert \mathbf{E}(\mathbf{r}, \mathbf{u}) \rVert_\infty = \max_k \sum_l
+    |[\mathbf{E}(\mathbf{r}, \mathbf{u})]_{k,l}| = |c(\mathbf{r}, \mathbf{u}) -
+    \Sigma_{i,j}| < \epsilon
+  \]
+  can be made arbitrarily small so that $\lVert \bm\Sigma^{-1} \rVert_\infty
+  \lVert \mathbf{E}(\mathbf{r}, \mathbf{u}) \rVert_\infty < 1$. Then
+  $\mathbf{C}(\mathbf{r}, \mathbf{u})$ is invertible, and
+  \[
+    \lVert \mathbf{C}(\mathbf{r}, \mathbf{u})^{-1} \rVert_\infty \le
+    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
+      \rVert_\infty \lVert \mathbf{E}(\mathbf{r}, \mathbf{u}) \rVert_\infty} <
+    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
+      \rVert_\infty \epsilon},
+  \]
+  which means that
+  \[
+    \max_k \sum_l \left| [\mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}]_{k,l} \right|
+    < \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
+      \rVert_\infty \epsilon},
+  \]
+  i.e., for any row $k$ and column $l$,
+  \[
+    \left| [\mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}]_{k,l} \right| <
+    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
+      \rVert_\infty \epsilon},
+  \]
+  which bounds all elements of $\mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}$ as
+  required. Since every element of $\mathbf{U} = (\mathbf{u} - \bm\mu)(\mathbf{u} -
+  \bm\mu)^\intercal$ is in $\mathbb{R}_2[\mathbf{u}]$, and the elements of
+  $\mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}$ are bounded, the desired result
+  follows.
+\end{proof}
+
+\begin{lemma} \label{lemma:integral_of_r}
+  \begin{equation} \label{eq:r-inequality}
+    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le a +
+    \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1,
+  \end{equation}
+  where $a$ is a constant independent of $\mathbf{u}$.
+\end{lemma}
+\begin{proof}
+  Since $\rinf \le \lVert \mathbf{r} \rVert_1$,
+  \[
+    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le \int
+    \lVert \mathbf{r} \rVert_1 q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} =
+    \sum_{i=1}^{|\mathcal{S}|} \mathbb{E}[|r_i|].
+  \]
+  As each $\mathbb{E}[|r_i|]$ is a mean of a folded Gaussian distribution,
+  \[
+    \mathbb{E}[|r_i|] = \sigma_i \sqrt{\frac{2}{\pi}} \exp
+    \left(-\frac{\xi_i^2}{2\sigma_i^2} \right) + \xi_i \left( 1 - 2\Phi \left(
+        -\frac{\xi_1}{\sigma_1} \right) \right),
+  \]
+  where $\xi_i = \left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right]_i$, $\sigma_i =
+  \sqrt{[\Krr - \Kru^\intercal\Kuu^{-1}\Kru]_{i,i}}$\footnote{The expression
+    under the square root sign is non-negative because $\Krr -
+    \Kru^\intercal\Kuu^{-1}\Kru$ is a covariance matrix of a Gaussian
+    distribution, hence also positive semi-definite, which means that its
+    diagonal entries are non-negative.}, and $\Phi$ is the cumulative
+  distribution function of the standard normal distribution. Furthermore,
+  \[
+    \mathbb{E}[|r_i|] \le \sigma_i\sqrt{\frac{2}{\pi}} + |\xi_i|,
+  \]
+  as $\sigma_i$ is non-negative, and $\Phi(x) \in [0, 1]$ for all $x$. Since
+  \[ \sum_{i=1}^{|\mathcal{S}|} |\xi_i| = \lVert \Kru^\intercal \Kuu^{-1}
+    \mathbf{u} \rVert_1, \]
+  we can set
+  \[ a = \sum_{i=1}^{|\mathcal{S}|} \sigma_i \sqrt{\frac{2}{\pi}} \]
+  to get \eqref{eq:r-inequality}.
+\end{proof}
+
+Our main theorem is a specialised version of an integral differentiation result
+by Chen \cite{lecture_notes}.
+\begin{theorem} \label{thm:main}
+  Whenever the derivative exists,
+  \[
+    \dt\iint
+    \V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})\dx
+    = \iint
+    \dt[\V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]\dx,
+  \]
+  where $t$ is any scalar part of $\bm\mu$, $\bm\Sigma$, or $\bm\lambda$.
+\end{theorem}
+\begin{proof}
+  Let
+  \begin{align*}
+    \f &= \V(s)q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u}), \\
+    F(t) &= \iint \f\dx,
+  \end{align*}
+  and fix the value of $t$. Let $(t_n)_{n=1}^\infty$ be any sequence such that
+  $\lim_{n \to \infty} t_n = t$, but $t_n \ne t$ for all $n$. We want to show
+  that
+  \begin{equation} \label{eq:to_prove}
+    F'(t) = \lim_{n \to \infty} \frac{F(t_n) - F(t)}{t_n - t} = \iint \df\dx.
+  \end{equation}
+  We have
+  \[
+    \begin{split}
+      \frac{F(t_n) - F(t)}{t_n - t} &= \iint \frac{\ftn - \f}{t_n - t}\dx \\
+      &= \iint \fn\dx,
+    \end{split}
+  \]
+  where
+  \[
+    \fn = \frac{\ftn - \f}{t_n - t}.
+  \]
+  Since
+  \[
+    \lim_{n \to \infty} \fn = \df,
+  \]
+  \eqref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
+  that both $f$ and $f_n$ are measurable and find a non-negative integrable
+  function $g$ such that for all $n$, $\mathbf{r}$, $\mathbf{u}$,
+  \[
+    |\fn| \le \g.
+  \]
+  The MDP value function is measurable by Proposition \ref{thm:measurability}.
+  The result of multiplying or adding measurable functions (e.g., probability
+  density functions (PDFs)) to a measurable function is still measurable. Thus,
+  both $f$ and $f_n$ are measurable.
+
+  It remains to find $g$. For notational simplicity and without loss of
+  generality, we will temporarily assume that $t$ is a parameter of
+  $q(\mathbf{r} \mid \mathbf{u})$. Then
+  \[
+    |\fn| = |\V(s)| \left| \frac{q(\mathbf{r} \mid \mathbf{u})|_{t =
+          t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} \right| q(\mathbf{u})
+  \]
+  since PDFs are non-negative. An upper bound for
+  $|\V(s)|$ is given by Proposition \ref{thm:bound}, while
+  \[
+    \frac{q(\mathbf{r} \mid \mathbf{u})|_{t = t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} = \left.
+      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial t} \right|_{t = c(\mathbf{r},
+      \mathbf{u})}
+  \]
+  for some function $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to
+  (\min\{t, t_n\}, \max\{t, t_n\})$ due to the mean value theorem (since $q$ is
+  a continuous and differentiable function of $t$, regardless of the specific
+  choices of $q$ and $t$).
+
+  We then have that
+  \[
+    |\fn| \le \vbound \left| \left. \frac{\partial
+          q(\mathbf{r} \mid \mathbf{u})}{\partial t} \right|_{t=c(\mathbf{r}, \mathbf{u})}
+    \right| q(\mathbf{u}).
+  \]
+  The bound is clearly non-negative and measurable. It remains to show that it
+  is also integrable. Depending on what $t$ represents, we can use one of the
+  Lemmas \ref{lemma:bound1}, \ref{lemma:bound2}, and \ref{lemma:bound3}, which
+  gives us two polynomials $p_1(\mathbf{u}), p_2(\mathbf{u}) \in
+  \mathbb{R}_2[\mathbf{u}]$ such that
+  \[
+    p_1(\mathbf{u})q(\mathbf{r} \mid \mathbf{u}) < \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial
+        t} \right|_{t=c(\mathbf{r}, \mathbf{u})} < p_2(\mathbf{u})q(\mathbf{r} \mid \mathbf{u}).
+  \]
+  Then
+  \[
+    \left| \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial t}
+      \right|_{t=c(\mathbf{r}, \mathbf{u})} \right| < q(\mathbf{r} \mid \mathbf{u}) \max \{
+    |p_1(\mathbf{u})|, |p_2(\mathbf{u})| \}.
+  \]
+  We can now apply Lemma \ref{lemma:integral_of_r}, which allows us to integrate
+  out $\mathbf{r}$, and we are left with showing the existence of
+  \begin{equation} \label{eq:remaining_integral}
+    \int \left( a + \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1 \right) \max \{|p_1(\mathbf{u})|, |p_2(\mathbf{u})| \} q(\mathbf{u})\,d\mathbf{u},
+  \end{equation}
+  where $a$ is a constant. The integral
+  \[
+    \int \max \left\{
+      \begin{split}
+        &|p_1(\mathbf{u})|, \\
+        &|p_2(\mathbf{u})|
+      \end{split}
+    \right\} q(\mathbf{u})\,d\mathbf{u} = \int \max \left\{
+      \begin{split}
+        &|p_1(\mathbf{u})q(\mathbf{u})|, \\
+        &|p_2(\mathbf{u})q(\mathbf{u})|
+      \end{split}
+    \right\}\,d\mathbf{u}
+  \]
+  exists because $p_1(\mathbf{u})q(\mathbf{u})$ and
+  $p_2(\mathbf{u})q(\mathbf{u})$ are both integrable, hence their absolute
+  values are integrable, and the maximum of two integrable functions is also
+  integrable. Since $\lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1 \in
+  \mathbb{R}_1[\mathbf{u}]$, a similar argument can be applied to the rest of
+  \eqref{eq:remaining_integral} as well.
+\end{proof}
+
+\subsection{Derivatives}
+
+\subsubsection{\texorpdfstring{$\partial/\partial\bm\mu$}{Derivative w.r.t. mu}}
+
+We begin by removing terms independent of $\bm\mu$:
+\[
+  \frac{\partial\mathcal{L}}{\partial\bm\mu} =
+  \dm[\mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu] - \frac{1}{2} \dm
+  \left[ \bm\mu^\intercal \Kuu^{-1} \bm\mu \right] - \dm\mathbb{E}[v].
+\]
+Here
+\[
+  \dm \left[ \bm\mu^\intercal \Kuu^{-1} \bm\mu \right] = (\Kuu^{-1} +
+  \Kuu^{-\intercal}) \bm\mu
+\] % TODO: by the cookbook
+and
+\[
+  \begin{split}
+    \dm\mathbb{E}[\V(s)] &= \dm\iint \V(s) q(\mathbf{r} \mid \mathbf{u})
+    q(\mathbf{u})\dx \\
+    &= \iint \V(s) q(\mathbf{r} \mid \mathbf{u}) \frac{\partial
+      q(\mathbf{u})}{\partial \bm\mu}\dx \\
+    &= \frac{1}{2}\mathbb{E}[\V(s) (\bm\Sigma^{-1} +
+    \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)]
+  \end{split}
+\]
+by Theorem \ref{thm:main} and Lemma \ref{lemma:derivatives}. Hence,
+\[
+  \begin{split}
+    \frac{\partial\mathcal{L}}{\partial\bm\mu} &=
+    \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1} - \frac{1}{2} (\Kuu^{-1} +
+    \Kuu^{-\intercal}) \bm\mu \\
+    &- \frac{1}{2}\mathbb{E} \left[(\bm\Sigma^{-1} +
+      \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu) v \right].
+  \end{split}
+\]
+
+\subsubsection{\texorpdfstring{$\partial/\partial\mathbf{B}$}{Derivative w.r.t. B}}
+
+\[
+  \frac{\partial\mathcal{L}}{\partial\mathbf{B}} =
+  \frac{1}{2} \left( \dB\log|\bm\Sigma| - \dB\tr \left( \Kuu^{-1} \bm\Sigma
+    \right) \right)
+  - \dB\mathbb{E}[v].
+\]
+By Theorem \ref{thm:main},
+\[
+  \dB\mathbb{E}[\V(s)] = \iint \V(s) q(\mathbf{r} \mid \mathbf{u})
+  \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}}\dx.
+\]
+Then, using the aforementioned tool by Laue et al.
+\cite{DBLP:conf/nips/LaueMG18}, we get
+\[
+  \dB\log|\bm\Sigma| = 2\bm\Sigma^{-1}\mathbf{B}, \quad \dB \tr \left( \Kuu^{-1}
+    \bm\Sigma \right) = 2\Kuu^{-1}\mathbf{B},
+\]
+and Lemma \ref{lemma:derivatives} gives
+\begin{gather*}
+  \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} =
+  q(\mathbf{u})(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
+  |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}.
+\end{gather*}
+Therefore,
+\begin{gather*}
+  \frac{\partial \mathcal{L}}{\partial \mathbf{B}} =
+  \left( \bm\Sigma^{-1} - \Kuu^{-1} \right) \mathbf{B} - \mathbb{E}
+  [(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
+  |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}v].
+\end{gather*}
+
+\subsubsection{\texorpdfstring{$\partial/\partial \lambda_j$}{Derivative w.r.t.
+    Lambda}}
+
+For $j = 0, \dots, d$,
+\[
+  \begin{split}
+    \frac{\partial \mathcal{L}}{\partial \lambda_j} &= \mathbf{t}^\intercal\dlj
+    \left[ \Kru^\intercal\Kuu^{-1} \right] \bm\mu - \dlj\mathbb{E}[v] \\
+    &- \frac{1}{2} \left(\dlj \tr \left(\Kuu^{-1}\bm\Sigma \right) +
+      \bm\mu^\intercal \frac{\partial \Kuu^{-1}}{\partial \lambda_j} \bm\mu +
+      \dlj \log |\Kuu| \right),
+  \end{split}
+\]
+where
+\begin{align*}
+  \frac{\partial \Kuu^{-1}}{\partial \lambda_j} &= -\Kuu^{-1}\frac{\partial \Kuu}{\partial \lambda_j}\Kuu^{-1}, \\
+  \dlj \left[ \Kru^\intercal\Kuu^{-1} \right] &= \frac{\partial \Kru^\intercal}{\partial \lambda_j} \Kuu^{-1} + \Kru^\intercal \frac{\partial \Kuu^{-1}}{\partial \lambda_j} \\
+                                                &= \left( \frac{\partial \Kru^\intercal}{\partial \lambda_j} - \Kru^\intercal\Kuu^{-1}\frac{\partial \Kuu}{\partial \lambda_j} \right) \Kuu^{-1}, \\
+  \dlj \tr(\Kuu^{-1}\bm\Sigma) &= \tr \left( \dlj \left[ \Kuu^{-1}\bm\Sigma \right] \right) = \tr \left( \frac{\partial \Kuu^{-1}}{\partial \lambda_j} \bm\Sigma \right) \\
+                                                &= -\tr \left( \Kuu^{-1} \frac{\partial \Kuu}{\partial \lambda_j} \Kuu^{-1} \bm\Sigma \right), \\
+  \dlj\log|\Kuu| &= \tr \left( \Kuu^{-1} \frac{\partial \Kuu}{\partial \lambda_j} \right)
+\end{align*}
+by Petersen and Pedersen \cite{petersen2008matrix}, and
+\[
+  \begin{split}
+    \dlj \mathbb{E}[\V(s)] &= \iint\V(s)\frac{\partial q(\mathbf{r} \mid
+      \mathbf{u})}{\partial \lambda_j}q(\mathbf{u})\dx \\
+    &= \!\begin{multlined}[t]
+      \frac{1}{2}\mathbb{E}[\V(s) (|\bm\Gamma|^{-1} \tr(\mathbf{R}
+          \adj(\bm\Gamma)) \\
+          - (\mathbf{r} -
+          \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r}
+          - \mathbf{Su}))]
+    \end{multlined}
+  \end{split}
+\]
+by Theorem \ref{thm:main} and Lemma \ref{lemma:derivatives}. Thus,
+\[
+  \begin{split}
+    \frac{\partial \mathcal{L}}{\partial \lambda_j} &=
+    \mathbf{t}^\intercal \left( \frac{\partial \Kru^\intercal}{\partial
+        \lambda_j} - \Kru^\intercal\Kuu^{-1}\frac{\partial
+        \Kuu}{\partial \lambda_j} \right) \Kuu^{-1} \bm\mu \\
+    &+ \frac{1}{2} \left[ \begin{split}
+      &\tr \left( \Kuu^{-1} \frac{\partial \Kuu}{\partial
+          \lambda_j} \Kuu^{-1} \bm\Sigma \right) + \bm\mu^\intercal \Kuu^{-1}
+      \frac{\partial \Kuu}{\partial \lambda_j} \Kuu^{-1} \bm\mu \\
+      &- \tr \left(\Kuu^{-1} \frac{\partial \Kuu}{\partial \lambda_j} \right)
+      \end{split} \right]
+    \\
+    &- \frac{1}{2} \mathbb{E} \left[ \left( \begin{split}
+        &|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
+          &- (\mathbf{r} -
+          \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r}
+          - \mathbf{Su})
+    \end{split} \right) v \right],
+  \end{split}
+\]
+where the remaining derivatives can be found in Lemma \ref{lemma:derivatives}.
+
 \section{Evaluation} % showing deep understanding

### 2019-03-06 mpaper/paper.tex
The notes PDF has been eliminated

@@
 \usepackage{mathtools}
+\usepackage[UKenglish]{babel}
+\usepackage[UKenglish]{isodate}
 
@@
 \newcommand{\Kru}{\mathbf{K}_{\mathbf{r},\mathbf{u}}}
+\newcommand{\Luu}{\mathbf{L}_{\mathbf{u},\mathbf{u}}}
 
@@
 \newcommand{\dm}{\frac{\partial}{\partial\bm\mu}}
-\newcommand{\dS}{\frac{\partial}{\partial\mathbf{S}}}
 \newcommand{\dB}{\frac{\partial}{\partial\mathbf{B}}}
@@
 \begin{abstract}
-According to Simon Peyton Jones, an abstract should address
-four key questions. First, what is the problem that this
-paper tackles? Second, why is this an interesting problem?
-Third, what is the solution this paper proposes?
-Finally, why is the proposed solution a good one?
 \end{abstract}
@@
 pointing out that we can retain full accuracy by setting $p = m$ and $\mathbf{D}
-= \mathbf{O}$. Moreover, we define a few matrices that will simplify expressions
-for the derivatives:
+= \mathbf{O}$. Moreover, we define a few variables that will simplify
+expressions for the derivatives:
 \begin{align*}
@@
   \bm\Gamma &= \Krr - \mathbf{S}\Kru, \\
-  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru.
+  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru, \\
+  Q &= (\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} - \bm\mu).
 \end{align*}
@@
 
-\begin{lemma}[Derivatives of probability
-  distributions] \label{lemma:derivatives}
+% TODO: define PDF before this point
+\begin{lemma}[Derivatives of PDFs] \label{lemma:derivatives}
   \leavevmode
@@
       |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}$.
-%    \item
-%      $\frac{\partial q(\mathbf{u})}{\partial \mathbf{D}} =
-%      \frac{1}{2}q(\mathbf{u}) \{ \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1}\mathbf{D}^\intercal +
-%      \mathbf{D}^\intercal\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
-%      |\bm\Sigma|^{-1}(\adj(\bm\Sigma^{-1})\mathbf{D}^\intercal +
-%      \mathbf{D}^\intercal \adj(\bm\Sigma^{-1})) \}$.
     \end{enumerate}
@@
 \end{lemma}
+\begin{proof}
+  \leavevmode
+  \begin{enumerate}
+  \item
+    \[
+      \begin{split}
+        \frac{\partial q(\mathbf{u})}{\partial m} &=
+        q(\mathbf{u})\dm\left[-\frac{Q}{2}\right]
+        \\
+        &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} -
+        \bm\mu] \\
+        &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
+      \end{split}
+    \]
+  \item
+    \begin{align}
+      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\exp \left( -\frac{Q}{2} \right)\right] \nonumber \\
+                                                         &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\right]\exp \left( -\frac{Q}{2} \right) \nonumber \\
+                                                         &+ \frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\dB\left[\exp\left( -\frac{Q}{2} \right)\right] \nonumber \\
+                                                         &= -\frac{1}{2} q(\mathbf{u}) \left( |\bm\Sigma|^{-3/2} \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} + \frac{\partial Q}{\partial \mathbf{B}} \right).
+\label{eq:2_before}
+    \end{align}
+    For derivatives with respect to $\bm\Sigma$, we can refer to Petersen and
+    Pedersen \cite{petersen2008matrix}:
+    \begin{equation} \label{eq:partial_derivatives1}
+      \frac{\partial |\bm\Sigma|}{\partial \bm\Sigma} = |\bm\Sigma|\bm\Sigma^{-\intercal}, \quad \frac{\partial Q}{\partial \mathbf{B}} = -\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal},
+    \end{equation}
+    while we can use an online tool by Laue et
+    al.\footnote{\url{http://www.matrixcalculus.org/}}
+    \cite{DBLP:conf/nips/LaueMG18} for the remaining ones:
+    \begin{equation} \label{eq:partial_derivatives2}
+      \begin{gathered}
+        \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} =
+        (\adj(\mathbf{T}) + \adj(\bm\Sigma))\mathbf{B}, \\
+        \dB[(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} -
+        \bm\mu)] = -(\mathbf{TUT} +
+        \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1})\mathbf{B}.
+      \end{gathered}
+    \end{equation}
+    Substituting results from \eqref{eq:partial_derivatives1} and
+    \eqref{eq:partial_derivatives2} back into \eqref{eq:2_before} gives:
+    \begin{alignat*}{3}
+      \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} &=
+      \frac{1}{2}q(\mathbf{u})(&&\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
+      - \bm\Sigma^{-\intercal}), \\
+      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &=
+      \frac{1}{2}q(\mathbf{u}) \{&&|\bm\Sigma|^{-3/2}(\adj(\mathbf{T}) +
+      \adj(\bm\Sigma)) \\
+      & &&+ \mathbf{TUT} + \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} \} \mathbf{B}.
+    \end{alignat*}
+  \item Using a result by Rasmussen and Williams % TODO: update the proof
+                                % according to the new result
+    \cite{DBLP:books/lib/RasmussenW06},
+    \[
+      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
+      q(\mathbf{r} \mid \mathbf{u}) \dl
+      \left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
+        \frac{1}{2}\log|\Kuu| \right] = \frac{1}{2}q(\mathbf{r} \mid \mathbf{u})\tr
+      \left((\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal} -
+        \Kuu^{-1}) \frac{\partial \Kuu}{\partial \lambda_i} \right).
+    \]
+    The remaining derivative is
+    \[
+      \frac{\partial \Kuu}{\partial \lambda_i} =
+      \begin{cases}
+        \frac{1}{\lambda_i}\Kuu & \text{if } i = 0, \\
+        \Luu & \text{otherwise,}
+      \end{cases}
+    \]
+    where
+    \[
+      \begin{split}
+        [\Luu]_{j,k} &= \dl k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j},
+        \mathbf{x}_{\mathbf{u},k}) \\
+        &= k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+        \dl \left[-\frac{1}{2}(\mathbf{x}_{\mathbf{u},j} -
+          \mathbf{x}_{\mathbf{u},k})^\intercal \bm\Lambda
+          (\mathbf{x}_{\mathbf{u},j} - \mathbf{x}_{\mathbf{u},k}) -
+          \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \right] \\
+        &= k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+        \dl \left[-\frac{1}{2}\sum_{l=1}^d \lambda_l
+          (x_{\mathbf{u},j,l} - x_{\mathbf{u},k,l})^2 -
+          \mathbbm{1}[j \ne k]\sigma^2\sum_{l=1}^d \lambda_l \right] \\
+        &= k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+        \left( -\frac{1}{2}(x_{\mathbf{u},j,i} -
+        x_{\mathbf{u},k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
+      \end{split}
+    \]
+  \end{enumerate}
+\end{proof}
 
@@
   $\lambda_i$, and $c(\mathbf{r}, \mathbf{u}) > a$,
-  \[ %TODO: what the hell do I do here?
-    \begin{split}
-      k_{\bm\lambda}(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = 
-        c(\mathbf{r}, \mathbf{u})} &=
-      \begin{multlined}[t]
-        \lambda_0 \exp \left( \vphantom{\sum_{n \in \{ \} \setminus}}
-        \right. -\frac{1}{2}c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2 \\
-        - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) \\
-        - \left. \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}}
-          \frac{1}{2} \lambda_n(x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne
-          k]\sigma^2 \lambda_n \right)
-      \end{multlined} \\
-      &<
-      \begin{multlined}[t]
-        \lambda_0 \exp \left( \vphantom{\sum_{n \in \{ \} \setminus}}
-        \right. -\frac{1}{2}a(x_{j,i} - x_{k,i})^2 -
-        \mathbbm{1}[j \ne k]\sigma^2a \\
-        - \left. \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}}
-          \frac{1}{2} \lambda_n(x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne
-          k]\sigma^2 \lambda_n \right),
-      \end{multlined}
-    \end{split}
+  \begin{alignat*}{3}
+    k_{\bm\lambda}(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r},
+      \mathbf{u})} &= \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. &&
+    -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2 \\
+    & && - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left.
+      \vphantom{\frac{1}{2}} \right) \\
+    &< \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. && -\frac{1}{2}
+    a(x_{j,i} - x_{k,i})^2 \\
+    & && - \mathbbm{1}[j \ne k]\sigma^2a - S \left. \vphantom{\frac{1}{2}}
+    \right),
+  \end{alignat*}
+  where
+  \[
+    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{1}{2}
+    \lambda_n(x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n,
   \]
@@
   The result of multiplying or adding measurable functions (e.g., probability
-  density functions (PDFs)) to a measurable function is still measurable. Thus,
+  density functions) to a measurable function is still measurable. Thus,
   both $f$ and $f_n$ are measurable.
@@
 
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{elbo_over_mu}
+\end{figure*}
+
 \section{Conclusions}

### 2019-03-07 mpaper/paper.tex
Contour plots and other changes

@@
 \DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator*{\argmax}{arg\,max}
 
@@
 \newcommand{\DKL}{D_{\mathrm{KL}}\infdivx}
-\newcommand{\Eq}{\mathbb{E}_{(\mathbf{u}, \mathbf{r}) \sim \approximation}}
 \newcommand{\V}{V_{\mathbf{r}}}
@@
 \newcommand{\approximation}{q(\mathbf{u}, \mathbf{r})}
+\newcommand{\posterior}{p(\mathbf{u}, \mathbf{r} \mid \mathcal{D})}
 
@@
 
+% TODO: add info from slides/handout
 \section{Introduction}
@@
 
-\section{Background}
+% TODO: will need to rewrite this to avoid plagiarism
+Inverse reinforcement learning (IRL)---a problem proposed by Russell in
+1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
+Markov decision process (MDP) that best explains a set of given demonstrations.
+IRL is important because reward functions can be hard to define manually
+\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
+are not entirely specific to a given environment, allowing one to reuse the same
+reward structure in previously unseen environments
+\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
+Moreover, IRL has seen a wide array of applications in autonomous vehicle
+control \cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} and
+learning to predict another agent's behaviour
+\cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
+Most approaches in the literature (see Section \ref{sec:background}) make a
+convenient yet unjustified assumption that the reward function can be expressed
+as a linear combination of features. One proven way to abandon this assumption
+is by representing the reward function as a Gaussian process (GP)
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
+The original approach used maximum likelihood estimation
+\cite{DBLP:conf/nips/LevinePK11}, while the goal of this project is to use
+variational inference (VI) instead, which learns approximate posterior
+probability distributions instead of point estimates. This approach can prove
+useful in three major ways:
+\begin{enumerate}
+\item Modelling full posterior distributions for various parameters can result
+  in more precise reward predictions, as the model simply holds more
+  information.
+\item Having variance estimates for rewards can direct our choice in what data
+  should be collected next.
+\item An approximate Bayesian treatment of many parameters in the model guards
+  against overfitting \cite{DBLP:conf/uai/JinDAS17}.
+\end{enumerate}
+
+\subsection{Statement of the Problem}
+
+\begin{definition}[MDP]
+  A \emph{Markov decision process} is a set $\mathcal{M} = \{ \mathcal{S},
+  \mathcal{A}, \mathcal{T}, \gamma, r \}$, where $\mathcal{S}$ and
+  $\mathcal{A}$ are sets of states and actions, respectively; $\mathcal{T} :
+  \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is a function
+  defined so that $\mathcal{T}(s, a, s')$ is the probability of moving to state $s'$
+  after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the discount
+  factor (with higher $\gamma$ values, it makes little difference whether a
+  reward is received now or later, while with lower $\gamma$ values the future
+  becomes gradually less and less important); and $r : \mathcal{S} \to
+  \mathbb{R}$ is the reward function.
+\end{definition}
+
+In \emph{inverse reinforcement learning}, one is presented with an MDP without a
+reward function $\mathcal{M} \setminus \{ r \}$ and a set of expert
+demonstrations $\mathcal{D} = \{ \zeta_i \}_{i=1}^N$, where each demonstration
+$\zeta_i = \{ (s_{i,0}, a_{i,0}), \dots, (s_{i,T}, a_{i,T}) \}$ is a multiset of
+state-action pairs representing the actions taken by the expert during a
+particular recorded session. Each state is also characterised by a number of
+features. The goal of IRL is then to find $r$ such that the optimal policy under
+$r$
+\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \left. \sum_{t=0}^\infty \gamma^t r(s_t)
+    \right| \pi \right] \]
+matches the actions in $\mathcal{D}$.
+
+Following previous work on GP IRL
+\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
+entropy IRL model \cite{ziebart2008maximum}, under which we have that
+\[ P(a | s) \propto \exp(Q_{\mathbf{r}}(s, a)), \]
+where
+\[
+  Q_{\mathbf{r}}(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s, a, s')\V(s'),
+\]
+and $\V : \mathcal{S} \to \mathbb{R}$ is an MDP \emph{value function}, which
+can be obtained by repeatedly applying the 'soft' version of the Bellman backup
+operator until convergence:
+\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
+\begin{equation} \label{eq:update_rule}
+  \V(s) \coloneqq \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s'
+      \in \mathcal{S}} \mathcal{T}(s, a, s')\V(s') \right).
+\end{equation}
+The likelihood of the data can then be expressed as
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
+\begin{equation} \label{pDr}
+  \begin{split}
+    p(\mathcal{D} | r) &= \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) \\
+    &= \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right).
+  \end{split}
+\end{equation}
+However, a reward function learned by maximising this likelihood is not
+transferable to new situations
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
+reward structure in a way that would allow reward predictions for previously
+unseen states.
+
+One way to model rewards without assumptions of linearity is with a
+\emph{Gaussian process} (GP). A GP is a collection of random variables, any
+finite combination of which has a joint Gaussian distribution
+\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
+k)$ to say that $r$ is a GP with mean $0$ and covariance function
+$k$, which uses a vector of hyperparameters $\bm\lambda$.
+Covariance functions take two state feature vectors as input and quantify how
+similar the two states are, in a sense that we would expect them to have similar
+rewards.
+
+As training a GP with $n$ data points has a time complexity of
+$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
+methods have been suggested, many of which select a subset of data called
+\emph{inducing points} and focus most of the training effort on them
+\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
+features at inducing states, $\mathbf{u}$ the rewards at those states, and
+$\mathbf{r}$ a vector with $r(\mathcal{S})$ as elements. Then the full joint
+probability distribution can be factorised as
+\[
+  \pfull = p(\mathbf{u}) \times p(\mathbf{r} | \mathbf{u}) \times p(\mathcal{D} | r),
+\]
+where
+\begin{align*}
+  p(\mathbf{u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \\
+                &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \\
+                &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right)
+\end{align*}
+is the GP prior \cite{DBLP:books/lib/RasmussenW06}, where $m \in \mathbb{N}$ is
+the number of inducing points. The GP posterior is a
+multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11}
+\begin{equation} \label{eq:r}
+  p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) =
+  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
+\end{equation}
+and $p(\mathcal{D} | r)$ is as in Equation \ref{pDr}. The matrices such as
+$\Kru$ are called \emph{covariance matrices} and are defined as
+$[\Kru]_{i,j} = k(\mathbf{x}_{\mathbf{r},i},
+\mathbf{x}_{\mathbf{u},j})$, where $\mathbf{x}_{\mathbf{r},i}$ and
+$\mathbf{x}_{\mathbf{u},j}$ denote feature vectors for the $i$th state in
+$\mathcal{S}$ and the $j$th state in $\mathbf{X_u}$, respectively
+\cite{DBLP:conf/uai/JinDAS17}.
+
+Given this model, data $\mathcal{D}$, and inducing feature matrix
+$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
+$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
+states $\mathbf{r}$. While the previous paper to consider this IRL model
+computed maximum likelihood (ML) estimates for $\bm\lambda$ and $\mathbf{u}$,
+and made an assumption that $\mathbf{r}$ in Equation \ref{eq:r} has zero variance
+\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
+VI to approximate the full posterior distribution $\posterior$.
+\emph{Variational inference} is an approximation technique for probability
+densities \cite{blei2017variational}. Let $\approximation$ be our approximating
+family of probability distributions for $\posterior$ with its own hyperparameter
+vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
+to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
+probability distribution and our approximation.  KL divergence (asymmetrically)
+measures how different the two distributions are, and in this case can be
+defined as \cite{blei2017variational}
+\begin{align*}
+  \DKL{\approximation}{\posterior} &= \mathbb{E}[\log\approximation - \log\posterior ] \\
+                                   &= \mathbb{E}[\log\approximation - \log\pfull] \\
+                                   &+ \mathbb{E}[\log p(\mathcal{D}, \mathbf{X_u})].
+\end{align*}
+The last term is both hard to compute and constant w.r.t. $\approximation$
+\cite{blei2017variational}, so we can remove it from our optimisation objective.
+The negation of what remains is often called the \emph{evidence lower bound}
+(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
+  be interpreted as definite integrals over the entire sample space.}
+\cite{DBLP:books/lib/Bishop07,blei2017variational}
+\begin{equation} \label{eq:elbo}
+  \begin{split}
+    \mathcal{L} &= \mathbb{E}\left[ \log \frac{\pfull}{\approximation} \right] \\
+    &= \iiint \log \frac{\pfull}{\approximation} \approximation\dx.
+  \end{split}
+\end{equation}
+
+By considering full probability distributions instead of point estimates---as
+long as the approximations are able to capture important features of the
+posterior---our predictions are likely to be more accurate and rely on fewer
+assumptions. Moreover, we hope to make use of various recent advancements in VI
+for both time complexity and approximation distribution fit (see Section
+\ref{sec:background}), making the resulting algorithm competitive both in terms of
+running time and model fit.
+
+The project is primarily concerned with investigating how a VI formulation of
+the GP IRL model compares against the original ML approach. Most importantly, we
+aim to compare how the two algorithms converge both over time and as the number
+of demonstrations increases. It would also be interesting to see how close the
+approximation of the posterior distribution is to the real thing. Finally, it is
+reasonable to conjecture that VI can outperform point estimates when dealing
+with more uncertainty, e.g., when experts make mistakes. We can easily
+investigate this by adapting how the evaluation data is generated.
+
+\section{Background} \label{sec:background}
 % motivation, relevance (whatever that means)
@@
 
+% TODO: how do we name this?
 \section{The WizWoz System}
@@
     \left[ \frac{\partial \Kuu}{\partial \lambda_i} \right]_{j,k} =
-    k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+    k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
     \left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(x_{\mathbf{u},j,i} -
@@
 \end{lemma}
-\begin{proof}
-  \leavevmode
-  \begin{enumerate}
-  \item
-    \[
-      \begin{split}
-        \frac{\partial q(\mathbf{u})}{\partial m} &=
-        q(\mathbf{u})\dm\left[-\frac{Q}{2}\right]
-        \\
-        &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
-        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} -
-        \bm\mu] \\
-        &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
-        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
-      \end{split}
-    \]
-  \item
-    \begin{align}
-      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\exp \left( -\frac{Q}{2} \right)\right] \nonumber \\
-                                                         &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\right]\exp \left( -\frac{Q}{2} \right) \nonumber \\
-                                                         &+ \frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\dB\left[\exp\left( -\frac{Q}{2} \right)\right] \nonumber \\
-                                                         &= -\frac{1}{2} q(\mathbf{u}) \left( |\bm\Sigma|^{-3/2} \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} + \frac{\partial Q}{\partial \mathbf{B}} \right).
+%\begin{proof}
+%  \leavevmode
+%  \begin{enumerate}
+%  \item
+%    \[
+%      \begin{split}
+%        \frac{\partial q(\mathbf{u})}{\partial m} &=
+%        q(\mathbf{u})\dm\left[-\frac{Q}{2}\right]
+%        \\
+%        &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+%        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} -
+%        \bm\mu] \\
+%        &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+%        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
+%      \end{split}
+%    \]
+%  \item
+%    \begin{align}
+%      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\exp \left( -\frac{Q}{2} \right)\right] \nonumber \\
+%                                                         &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\right]\exp \left( -\frac{Q}{2} \right) \nonumber \\
+%                                                         &+ \frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\dB\left[\exp\left( -\frac{Q}{2} \right)\right] \nonumber \\
+%                                                         &= -\frac{1}{2} q(\mathbf{u}) \left( |\bm\Sigma|^{-3/2} \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} + \frac{\partial Q}{\partial \mathbf{B}} \right).
 \label{eq:2_before}
-    \end{align}
-    For derivatives with respect to $\bm\Sigma$, we can refer to Petersen and
-    Pedersen \cite{petersen2008matrix}:
-    \begin{equation} \label{eq:partial_derivatives1}
-      \frac{\partial |\bm\Sigma|}{\partial \bm\Sigma} = |\bm\Sigma|\bm\Sigma^{-\intercal}, \quad \frac{\partial Q}{\partial \mathbf{B}} = -\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal},
-    \end{equation}
-    while we can use an online tool by Laue et
-    al.\footnote{\url{http://www.matrixcalculus.org/}}
-    \cite{DBLP:conf/nips/LaueMG18} for the remaining ones:
-    \begin{equation} \label{eq:partial_derivatives2}
-      \begin{gathered}
-        \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} =
-        (\adj(\mathbf{T}) + \adj(\bm\Sigma))\mathbf{B}, \\
-        \dB[(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} -
-        \bm\mu)] = -(\mathbf{TUT} +
-        \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1})\mathbf{B}.
-      \end{gathered}
-    \end{equation}
-    Substituting results from \eqref{eq:partial_derivatives1} and
-    \eqref{eq:partial_derivatives2} back into \eqref{eq:2_before} gives:
-    \begin{alignat*}{3}
-      \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} &=
-      \frac{1}{2}q(\mathbf{u})(&&\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
-      - \bm\Sigma^{-\intercal}), \\
-      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &=
-      \frac{1}{2}q(\mathbf{u}) \{&&|\bm\Sigma|^{-3/2}(\adj(\mathbf{T}) +
-      \adj(\bm\Sigma)) \\
-      & &&+ \mathbf{TUT} + \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} \} \mathbf{B}.
-    \end{alignat*}
-  \item Using a result by Rasmussen and Williams % TODO: update the proof
-                                % according to the new result
-    \cite{DBLP:books/lib/RasmussenW06},
-    \[
-      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
-      q(\mathbf{r} \mid \mathbf{u}) \dl
-      \left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
-        \frac{1}{2}\log|\Kuu| \right] = \frac{1}{2}q(\mathbf{r} \mid \mathbf{u})\tr
-      \left((\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal} -
-        \Kuu^{-1}) \frac{\partial \Kuu}{\partial \lambda_i} \right).
-    \]
-    The remaining derivative is
-    \[
-      \frac{\partial \Kuu}{\partial \lambda_i} =
-      \begin{cases}
-        \frac{1}{\lambda_i}\Kuu & \text{if } i = 0, \\
-        \Luu & \text{otherwise,}
-      \end{cases}
-    \]
-    where
-    \[
-      \begin{split}
-        [\Luu]_{j,k} &= \dl k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j},
-        \mathbf{x}_{\mathbf{u},k}) \\
-        &= k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-        \dl \left[-\frac{1}{2}(\mathbf{x}_{\mathbf{u},j} -
-          \mathbf{x}_{\mathbf{u},k})^\intercal \bm\Lambda
-          (\mathbf{x}_{\mathbf{u},j} - \mathbf{x}_{\mathbf{u},k}) -
-          \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \right] \\
-        &= k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-        \dl \left[-\frac{1}{2}\sum_{l=1}^d \lambda_l
-          (x_{\mathbf{u},j,l} - x_{\mathbf{u},k,l})^2 -
-          \mathbbm{1}[j \ne k]\sigma^2\sum_{l=1}^d \lambda_l \right] \\
-        &= k_{\bm\lambda}(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-        \left( -\frac{1}{2}(x_{\mathbf{u},j,i} -
-        x_{\mathbf{u},k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
-      \end{split}
-    \]
-  \end{enumerate}
-\end{proof}
+%    \end{align}
+%    For derivatives with respect to $\bm\Sigma$, we can refer to Petersen and
+%    Pedersen \cite{petersen2008matrix}:
+%    \begin{equation} \label{eq:partial_derivatives1}
+%      \frac{\partial |\bm\Sigma|}{\partial \bm\Sigma} = |\bm\Sigma|\bm\Sigma^{-\intercal}, \quad \frac{\partial Q}{\partial \mathbf{B}} = -\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal},
+%    \end{equation}
+%    while we can use an online tool by Laue et
+%    al.\footnote{\url{http://www.matrixcalculus.org/}}
+%    \cite{DBLP:conf/nips/LaueMG18} for the remaining ones:
+%    \begin{equation} \label{eq:partial_derivatives2}
+%      \begin{gathered}
+%        \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} =
+%        (\adj(\mathbf{T}) + \adj(\bm\Sigma))\mathbf{B}, \\
+%        \dB[(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} -
+%        \bm\mu)] = -(\mathbf{TUT} +
+%        \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1})\mathbf{B}.
+%      \end{gathered}
+%    \end{equation}
+%    Substituting results from Equations \ref{eq:partial_derivatives1} and
+%    \ref{eq:partial_derivatives2} back into Equation \ref{eq:2_before} gives:
+%    \begin{alignat*}{3}
+%      \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} &=
+%      \frac{1}{2}q(\mathbf{u})(&&\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
+%      - \bm\Sigma^{-\intercal}), \\
+%      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &=
+%      \frac{1}{2}q(\mathbf{u}) \{&&|\bm\Sigma|^{-3/2}(\adj(\mathbf{T}) +
+%      \adj(\bm\Sigma)) \\
+%      & &&+ \mathbf{TUT} + \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} \} \mathbf{B}.
+%    \end{alignat*}
+%  \item Using a result by Rasmussen and Williams % TODO: update the proof
+%                                % according to the new result
+%    \cite{DBLP:books/lib/RasmussenW06},
+%    \[
+%      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
+%      q(\mathbf{r} \mid \mathbf{u}) \dl
+%      \left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
+%        \frac{1}{2}\log|\Kuu| \right] = \frac{1}{2}q(\mathbf{r} \mid \mathbf{u})\tr
+%      \left((\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal} -
+%        \Kuu^{-1}) \frac{\partial \Kuu}{\partial \lambda_i} \right).
+%    \]
+%    The remaining derivative is
+%    \[
+%      \frac{\partial \Kuu}{\partial \lambda_i} =
+%      \begin{cases}
+%        \frac{1}{\lambda_i}\Kuu & \text{if } i = 0, \\
+%        \Luu & \text{otherwise,}
+%      \end{cases}
+%    \]
+%    where
+%    \[
+%      \begin{split}
+%        [\Luu]_{j,k} &= \dl k(\mathbf{x}_{\mathbf{u},j},
+%        \mathbf{x}_{\mathbf{u},k}) \\
+%        &= k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+%        \dl \left[-\frac{1}{2}(\mathbf{x}_{\mathbf{u},j} -
+%          \mathbf{x}_{\mathbf{u},k})^\intercal \bm\Lambda
+%          (\mathbf{x}_{\mathbf{u},j} - \mathbf{x}_{\mathbf{u},k}) -
+%          \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \right] \\
+%        &= k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+%        \dl \left[-\frac{1}{2}\sum_{l=1}^d \lambda_l
+%          (x_{\mathbf{u},j,l} - x_{\mathbf{u},k,l})^2 -
+%          \mathbbm{1}[j \ne k]\sigma^2\sum_{l=1}^d \lambda_l \right] \\
+%        &= k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
+%        \left( -\frac{1}{2}(x_{\mathbf{u},j,i} -
+%        x_{\mathbf{u},k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
+%      \end{split}
+%    \]
+%  \end{enumerate}
+%\end{proof}
 
 \subsection{Evidence Lower Bound}
-
-\begin{equation} \label{eq:elbo}
-  \begin{split}
-    \mathcal{L} &= \Eq \left[ \log \frac{\pfull}{\approximation}
-    \right] \\
-    &= \iint \approximation \log
-    \frac{\pfull}{\approximation}\dx.
-  \end{split}
-\end{equation}
 
@@
 In this section we derive and simplify the ELBO for this (now fully specified)
-model. In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
+model. In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
 write\footnote{At this point, we will drop the subscript denoting which
@@
   derivation equality is taken to mean `equality up to an additive constant'.}
+% TODO: move this to the front and say that the expectations are always w.r.t.
+% the full VI approximation unless specified otherwise
 \[
@@
 \]
-By substituting in \eqref{eq:full} and \eqref{eq:approximation}, we get
-\[
-  \begin{split}
-    \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid
-    \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
+By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
+\begin{align*}
+  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
               &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
-  \end{split}
-\]
+\end{align*}
 Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
 cancel each other out. Also notice that
-\[
-  \begin{split}
-    \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &=
-    -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
-    &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) +
-      \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
+\begin{align*}
+  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
+                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
                                                       &+ \log |\Kuu| - \log |\bm\Sigma|),
-  \end{split}
-\]
+\end{align*}
 by the definition of KL divergence between two multivariate normal distributions
 \cite{kl}. Hence,
-\[
-  \begin{split}
-    \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T
-      Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
-    &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) +
-      \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
-  \end{split}
-\]
+\begin{align*}
+  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
+              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
+\end{align*}
 Using the expressions for $Q_{\mathbf{r}}$ we get
-\[
-  \begin{split}
-  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) -
-    \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t},
-    s')\V(s') \right] \\
-  &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) +
-    \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
-  \end{split}
-\]
+\begin{align*}
+  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
+              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
+\end{align*}
 We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
@@
 across all demonstrations. Then
-\[
-  \begin{split}
-  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &=
-  \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] =
-  \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
-  &=
-  \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] =
-  \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
-  \end{split}
-\]
+\begin{align*}
+  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
+                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
+\end{align*}
 This allows us to simplify $\mathcal{L}$ to
-\[
-  \begin{split}
-  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu -
-  \mathbb{E}[v] \\
-  &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) +
-    \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
-  \end{split}
-\]
+\begin{align*}
+  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
+              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
+\end{align*}
 where
 \[
-v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in
-  \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
+  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
 \]
@@
     \left\{ \mathbf{r} \in \mathbb{R}^{|\mathcal{S}|} \; \middle|
-    \begin{split}
+    \begin{aligned}
       \;&\V(s_0) \in (-\infty, \alpha); \\
@@
       \mathcal{S}
-    \end{split}
+    \end{aligned}
     \right\}
@@
 \begin{proof}
-  We begin by considering \eqref{eq:bound} without taking the absolute value of
+  We begin by considering Equation \ref{eq:bound} without taking the absolute value of
   $\V(s)$, i.e.,
@@
   and assuming that the initial values of $\{ \V(s) \mid s \in
-  \mathcal{S} \}$ already satisfy \eqref{eq:positive_bound}. For each $s \in
-  \mathcal{S}$, the value of $\V(s)$ is updated via this rule:
-  \[
-    \V(s) \coloneqq \log \sum_{a \in \mathcal{A}} \exp\left( r(s) +
-      \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s')\V(s')
-    \right).
-  \]
-  Note that both $\log$ and $\exp$ are increasing functions, $\gamma > 0$, and
-  the $\mathcal{T}$ function gives a probability (a non-negative number).
-  Thus
-  \[
-    \begin{split}
-      \V(s) &\le \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s'
-          \in \mathcal{S}} \mathcal{T}(s, a, s')\frac{\rinf +
-          \log|\mathcal{A}|}{1 - \gamma} \right) \\
-      &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf +
-          \log|\mathcal{A}|)}{1 - \gamma}\sum_{s' \in \mathcal{S}}
-        \mathcal{T}(s, a, s') \right) \\
-      &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf +
-          \log|\mathcal{A}|)}{1 - \gamma} \right)
-    \end{split}
-  \]
+  \mathcal{S} \}$ already satisfy Equation \ref{eq:positive_bound}. Recall that for
+  each $s \in \mathcal{S}$, the value of $\V(s)$ is updated by applying Equation
+  \ref{eq:update_rule}. Note that both $\log$ and $\exp$ are increasing
+  functions, $\gamma > 0$, and the $\mathcal{T}$ function gives a probability (a
+  non-negative number). Thus
+  \begin{align*}
+    \V(s) &\le \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s')\frac{\rinf + \log|\mathcal{A}|}{1 - \gamma} \right) \\
+          &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma}\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s') \right) \\
+          &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right)
+  \end{align*}
   by the definition of $\mathcal{T}$. Then
-  \[
-    \begin{split}
-      \V(s) &\le \log \left( |\mathcal{A}| \exp\left( r(s) + \frac{\gamma
-            (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
-      &= \log \left( \exp\left( \log|\mathcal{A}| + r(s) + \frac{\gamma
-            (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
+  \begin{align*}
+    \V(s) &\le \log \left( |\mathcal{A}| \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
+          &= \log \left( \exp\left( \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
           &= \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \\
-      &= \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 -
-        \gamma)(\log|\mathcal{A}| + r(s))}{1 - \gamma} \\
-      &\le \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 -
-        \gamma)(\log|\mathcal{A}| + \rinf)}{1 - \gamma} \\
+          &= \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 - \gamma)(\log|\mathcal{A}| + r(s))}{1 - \gamma} \\
+          &\le \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 - \gamma)(\log|\mathcal{A}| + \rinf)}{1 - \gamma} \\
           &= \vbound
-    \end{split}
-  \]
+  \end{align*}
   by the definition of $\rinf$.
@@
   follows the same argument until we get to
-  \[
-    \begin{split}
-      \V(s) &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma -
-        1)(\log|\mathcal{A}| + r(s))}{\gamma - 1} \\
-      &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma -
-        1)(-\log|\mathcal{A}| -\rinf)}{\gamma - 1} \\
+  \begin{align*}
+    \V(s) &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma - 1)(\log|\mathcal{A}| + r(s))}{\gamma - 1} \\
+          &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma - 1)(-\log|\mathcal{A}| -\rinf)}{\gamma - 1} \\
           &= \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1},
-    \end{split}
-  \]
+  \end{align*}
   where we use the fact that $r(s) \ge -\rinf - 2\log|\mathcal{A}|$. Combining
-  \eqref{eq:positive_bound} and \eqref{eq:negative_bound} gives
-  \eqref{eq:bound}.
+  Equations \ref{eq:positive_bound} and \ref{eq:negative_bound} gives
+  Equation \ref{eq:bound}.
 \end{proof}
@@
   If $i > 0$, then each element of $\frac{\partial \Kuu}{\partial
-    \lambda_i}$ is a constant multiple of $k_{\bm\lambda}(\mathbf{x}_j,
+    \lambda_i}$ is a constant multiple of $k(\mathbf{x}_j,
   \mathbf{x}_k)$, for some $\mathbf{x}_j$ and $\mathbf{x}_k$. Since
-  $k_{\bm\lambda}(\mathbf{x}_j, \mathbf{x}_k)$ is a decreasing function of
+  $k(\mathbf{x}_j, \mathbf{x}_k)$ is a decreasing function of
   $\lambda_i$, and $c(\mathbf{r}, \mathbf{u}) > a$,
   \begin{alignat*}{3}
-    k_{\bm\lambda}(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r},
+    k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r},
       \mathbf{u})} &= \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. &&
@@
 \begin{lemma} \label{lemma:integral_of_r}
-  \begin{equation} \label{eq:r-inequality}
+  \[
     \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le a +
     \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1,
-  \end{equation}
+  \]
   where $a$ is a constant independent of $\mathbf{u}$.
@@
   \[ a = \sum_{i=1}^{|\mathcal{S}|} \sigma_i \sqrt{\frac{2}{\pi}} \]
-  to get \eqref{eq:r-inequality}.
+  to get the desired result.
 \end{proof}
@@
   \]
-  \eqref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
+  Equation \ref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
   that both $f$ and $f_n$ are measurable and find a non-negative integrable
@@
     \int \max \left\{
-      \begin{split}
+      \begin{aligned}
         &|p_1(\mathbf{u})|, \\
         &|p_2(\mathbf{u})|
-      \end{split}
+      \end{aligned}
     \right\} q(\mathbf{u})\,d\mathbf{u} = \int \max \left\{
-      \begin{split}
+      \begin{aligned}
         &|p_1(\mathbf{u})q(\mathbf{u})|, \\
         &|p_2(\mathbf{u})q(\mathbf{u})|
-      \end{split}
+      \end{aligned}
     \right\}\,d\mathbf{u}
@@
   \mathbb{R}_1[\mathbf{u}]$, a similar argument can be applied to the rest of
-  \eqref{eq:remaining_integral} as well.
+  Equation \ref{eq:remaining_integral} as well.
 \end{proof}
@@
   \centering
-  \includegraphics[width=\textwidth]{elbo_over_mu}
+  \includegraphics[width=\textwidth]{elbo_contour_plots}
 \end{figure*}

### 2019-03-07 code/VIGPIRL/vigpirlrun.m

@@ -72,7 +72,7 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   mu2_log = [];
   mu3_log = [];
   multiplier = 10;
-  for mu1 = 5:5
+  for mu1 = 1:10
     real_mu1 = multiplier * (mu1 - 5);
     parameter_vector(d+m+2) = real_mu1;
     for mu2 = 1:10
@@ -99,12 +99,24 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   %end
   %[red, green, blue] = arrayfun(@(x) choose_color(x, max_negative, max_positive), elbo_list);
 
+  % 2D plots
+  for mu1 = 1:10
+    indices = 100*(mu1-1)+1:100*mu1;
+    other_mu = multiplier * ((1:10) - 5);
+    subplot(5, 2, mu1);
+    contourf(other_mu, other_mu, reshape(elbo_list(indices), [10, 10]), 30);
+    xlabel('$\mu_2$', 'Interpreter', 'latex');
+    ylabel('$\mu_3$', 'Interpreter', 'latex');
+    t = ['$\mu_1 = ', num2str(multiplier * (mu1 - 5)), '$'];
+    title(t, 'Interpreter', 'latex');
+  end
+
   % 3D Plots
   %scatter3(mu1_log, mu2_log, mu3_log, 100, elbo_list, 'filled');
   %colorbar();
   %q = quiver3(mu1_log, mu2_log, mu3_log, grad_history(d+m+2, :), grad_history(d+m+3, :), grad_history(d+m+4, :));
   %q.LineWidth = 2;
-  quiver(mu2_log, mu3_log, grad_history(d+m+3, :), grad_history(d+m+4, :));
+  %quiver(mu2_log, mu3_log, grad_history(d+m+3, :), grad_history(d+m+4, :));
   %xlabel('$\mu_1$', 'Interpreter', 'latex');
   %ylabel('$\mu_2$', 'Interpreter', 'latex');
   %zlabel('$\mu_3$', 'Interpreter', 'latex');

### 2019-03-10 mpaper/paper.tex
Dependency graph for the theoretical results

@@
 \usepackage{mathtools}
+\usepackage{thmtools, thm-restate}
 \usepackage[UKenglish]{babel}
 \usepackage[UKenglish]{isodate}
+\usepackage{tikz}
+
+\usetikzlibrary{shapes,bayesnet,arrows.meta}
+\tikzstyle{arrow} = [-{Latex[length=2mm,width=2mm]}]
 
@@
 
+\newcommand{\approximationS}{q(\mathbf{u}_s, \mathbf{r}_s)}
+\newcommand{\pfullS}{p(\mathcal{D}, \mathbf{u}_s, \mathbf{r}_s)}
+
 \newcommand{\f}{f(\mathbf{r}, \mathbf{u}, t)}
@@
 
-% TODO: add info from slides/handout
 \section{Introduction}
@@
 
-% TODO: will need to rewrite this to avoid plagiarism
 Inverse reinforcement learning (IRL)---a problem proposed by Russell in
@@
   A \emph{Markov decision process} is a set $\mathcal{M} = \{ \mathcal{S},
-  \mathcal{A}, \mathcal{T}, \gamma, r \}$, where $\mathcal{S}$ and
+  \mathcal{A}, \mathcal{T}, \gamma, \mathbf{r} \}$, where $\mathcal{S}$ and
   $\mathcal{A}$ are sets of states and actions, respectively; $\mathcal{T} :
@@
   after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the discount
-  factor (with higher $\gamma$ values, it makes little difference whether a
-  reward is received now or later, while with lower $\gamma$ values the future
-  becomes gradually less and less important); and $r : \mathcal{S} \to
-  \mathbb{R}$ is the reward function.
+  factor; and $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ is the reward
+  vector\footnote{Depending on the situation, we will sometimes represent
+    rewards as a function $r : \mathcal{S} \to \mathbb{R}$.}.
 \end{definition}
 
-In \emph{inverse reinforcement learning}, one is presented with an MDP without a
-reward function $\mathcal{M} \setminus \{ r \}$ and a set of expert
-demonstrations $\mathcal{D} = \{ \zeta_i \}_{i=1}^N$, where each demonstration
-$\zeta_i = \{ (s_{i,0}, a_{i,0}), \dots, (s_{i,T}, a_{i,T}) \}$ is a multiset of
-state-action pairs representing the actions taken by the expert during a
-particular recorded session. Each state is also characterised by a number of
-features. The goal of IRL is then to find $r$ such that the optimal policy under
-$r$
-\[ \pi^* = \argmax_\pi \mathbb{E}\left[ \left. \sum_{t=0}^\infty \gamma^t r(s_t)
-    \right| \pi \right] \]
-matches the actions in $\mathcal{D}$.
-
-Following previous work on GP IRL
-\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, we use a maximum
-entropy IRL model \cite{ziebart2008maximum}, under which we have that
-\[ P(a | s) \propto \exp(Q_{\mathbf{r}}(s, a)), \]
-where
-\[
-  Q_{\mathbf{r}}(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
-  \mathcal{T}(s, a, s')\V(s'),
-\]
-and $\V : \mathcal{S} \to \mathbb{R}$ is an MDP \emph{value function}, which
-can be obtained by repeatedly applying the 'soft' version of the Bellman backup
-operator until convergence:
-\cite{DBLP:conf/nips/LevinePK11,supplementary_material}
+\begin{definition}[IRL]
+  Given an MDP without rewards $\mathcal{M} \setminus \{ \mathbf{r} \}$, an
+  $|\mathcal{S}| \times d$ feature matrix $\mathbf{X}$ (where $d$ is the number
+  of features), and a set of expert demonstrations $\mathcal{D} = \{\zeta_i
+  \}_{i=1}^N$, where each demonstration $\zeta_i = \{ (s_{i,t}, a_{i,t})
+  \}_{t=1}^T$ is a multiset of state-action pairs representing optimal actions
+  executed by an expert, find the reward function that maximises the probability
+  of observing the demonstrations, i.e.,
+  \[
+    \argmax_{\mathbf{r}} p(\mathcal{D} \mid \mathbf{r}).
+  \]
+\end{definition}
+
+The optimal \emph{policy} $\pi : \mathcal{S} \to \mathcal{A}$ (i.e., a choice of
+actions for each state that maximises reward over time) is usually constructed
+by defining a \emph{value (utility) function} $\V : \mathcal{S} \to \mathbb{R}$
+that measures how good a state is based on the reward $\mathbf{r}$ as well as
+the structure of the MDP. One can then find $\V$ by applying the Bellman backup
+operator until convergence to every $s \in \mathcal{S}$ (the technique is known
+as \emph{value iteration}) \cite{DBLP:books/daglib/0023820}:
+\[
+  V_{\mathbf{r}}(s) \coloneqq r(s) + \gamma \max_{a \in \mathcal{A}} \sum_{s' \in
+    \mathcal{S}} \mathcal{T}(s, a, s')V_{\mathbf{r}}(s').
+\]
+
+However, we follow previous work on GP IRL
+\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, and use a
+\emph{linearly solvable} (or \emph{maximum causal entropy}) MDP with stochastic
+policy that define probability distributions over actions (instead of
+suggesting a single action for each state) \cite{ziebart2008maximum}. This type
+of MDP can be solved by applying the 'soft' version of the operator
+\cite{DBLP:conf/nips/LevinePK11,supplementary_material}:
 \begin{equation} \label{eq:update_rule}
@@
 \end{equation}
-The likelihood of the data can then be expressed as
+With this model, we can express the likelihood as
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
@@
   \begin{split}
-    p(\mathcal{D} | r) &= \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} | s_{i,t}) \\
-    &= \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right).
+    p(\mathcal{D} \mid \mathbf{r}) &= \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} \mid s_{i,t}) \\
+    &= \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right),
   \end{split}
 \end{equation}
+where
+\[
+  Q_{\mathbf{r}}(s, a) = r(s) + \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s, a, s')\V(s').
+\]
+
 However, a reward function learned by maximising this likelihood is not
@@
 k)$ to say that $r$ is a GP with mean $0$ and covariance function
-$k$, which uses a vector of hyperparameters $\bm\lambda$.
-Covariance functions take two state feature vectors as input and quantify how
-similar the two states are, in a sense that we would expect them to have similar
-rewards.
+$k$. Covariance functions take two state feature vectors as input and quantify
+how similar the two states are, in a sense that we would expect them to have
+similar rewards.
 
@@
 \cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
-features at inducing states, $\mathbf{u}$ the rewards at those states, and
-$\mathbf{r}$ a vector with $r(\mathcal{S})$ as elements. Then the full joint
-probability distribution can be factorised as
-\[
-  \pfull = p(\mathbf{u}) \times p(\mathbf{r} | \mathbf{u}) \times p(\mathcal{D} | r),
-\]
+features at inducing states, $\mathbf{u}$ the rewards at those states. Then the
+full joint probability distribution can be factorised as
+\begin{equation} \label{eq:full}
+  \pfull = p(\mathbf{u}) \times p(\mathbf{r} \mid \mathbf{u}) \times
+  p(\mathcal{D} \mid \mathbf{r}),
+\end{equation}
 where
@@
 \end{align*}
-is the GP prior \cite{DBLP:books/lib/RasmussenW06}, where $m \in \mathbb{N}$ is
-the number of inducing points. The GP posterior is a
-multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11}
+is the GP prior \cite{DBLP:books/lib/RasmussenW06}, and $m \in \mathbb{N}$ is
+the number of inducing points. The GP posterior is then a multivariate Gaussian
+\cite{DBLP:conf/nips/LevinePK11}
 \begin{equation} \label{eq:r}
-  p(\mathbf{r} | \bm\lambda, \mathbf{X_u}, \mathbf{u}) =
+  p(\mathbf{r} \mid \mathbf{u}) =
   \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
 \end{equation}
-and $p(\mathcal{D} | r)$ is as in Equation \ref{pDr}. The matrices such as
+and $p(\mathcal{D} \mid \mathbf{r})$ is as in Equation \ref{pDr}. The matrices such as
 $\Kru$ are called \emph{covariance matrices} and are defined as
@@
 Given this model, data $\mathcal{D}$, and inducing feature matrix
-$\mathbf{X_u}$, our goal is then to find optimal values of hyperparameters
+$\mathbf{X_u}$, our goal is then to find optimal values of parameters
 $\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
@@
 densities \cite{blei2017variational}. Let $\approximation$ be our approximating
-family of probability distributions for $\posterior$ with its own hyperparameter
-vector $\bm\nu$. Then the job of VI algorithms is to optimise $\bm\nu$ in order
-to minimise the \emph{Kullback-Leibler} (KL) divergence between the original
-probability distribution and our approximation.  KL divergence (asymmetrically)
-measures how different the two distributions are, and in this case can be
-defined as \cite{blei2017variational}
+family of probability distributions for $\posterior$. Then the job of VI is to
+optimise the parameters of the approximating distribution in order to minimise
+the \emph{Kullback-Leibler} (KL) divergence between the original probability
+distribution and our approximation.  KL divergence (asymmetrically) measures how
+different the two distributions are, and can be defined as
+\cite{blei2017variational}
 \begin{align*}
@@
                                    &= \mathbb{E}[\log\approximation - \log\pfull] \\
-                                   &+ \mathbb{E}[\log p(\mathcal{D}, \mathbf{X_u})].
+                                   &+ \mathbb{E}[\log p(\mathcal{D})].
 \end{align*}
@@
 
-\subsection{Linear Algebra and Numerical Analysis}
+\subsection{Linear Algebra and Analysis}
 
@@
 
-% TODO: how do we name this?
-\section{The WizWoz System}
-% solution
+\begin{theorem}[Dominated Convergence Theorem
+  \cite{royden2010real}] \label{thm:lebesgue}
+  Let $(X, \mathcal{M}, \mu)$ be a measure space and $\{ f_n \}$ a sequence of
+  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise a.e. on $X$
+  and the function $f$ is measurable. Assume there is a non-negative function
+  $g$ that is integrable over $X$ and dominates the sequence $\{ f_n \}$ on $X$
+  in the sense that
+  \[
+    |f_n| \le g \text{ a.e. on $X$ for all $n$.}
+  \]
+  Then $f$ is integrable over $X$ and
+  \[
+    \lim_{n \to \infty} \int_X f_n\,d\mu = \int_X f\,d\mu.
+  \]
+\end{theorem}
+
+\section{Solution}
 % TODO: express all lambdas as exponentials (update the derivatives or sth)
@@
 
-For any matrix $\mathbf{A}$, we will use either $A_{i,j}$ or
-$[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
-$j$. Moreover, we use $\tr(\mathbf{A})$ to denote its \emph{trace} and
+We use $\tr(\mathbf{A})$ to denote its \emph{trace} and
 $\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}).
@@
 elements of $\mathbf{x}$, and coefficients are in $\mathbb{R}$.
-
-We primarily think of rewards as a vector $\mathbf{r} \in
-\mathbb{R}^{|\mathcal{S}|}$, but sometimes we use a function notation $r(s)$ to
-denote the reward of a particular state $s \in \mathcal{S}$.
 
@@
 
-Derivatives such as $\frac{\partial \Kuu}{\partial \lambda_i}$ can be expressed
-as
-\[
-  \frac{\partial \Kuu}{\partial \lambda_i} = \frac{1}{\lambda_i}\Kuu
-\]
-
-if $i = 0$, and
-
-\[
-  \begin{split}
-    \left[ \frac{\partial \Kuu}{\partial \lambda_i} \right]_{j,k} =
-    k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-    \left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(x_{\mathbf{u},j,i} -
-    x_{\mathbf{u},k,i})^2 \\
-    &- \mathbbm{1}[j \ne k]\sigma^2 \left. \vphantom{\frac{1}{2}} \right)
-  \end{split}
-\]
-otherwise.
-
 % TODO: define PDF before this point
-\begin{lemma}[Derivatives of PDFs] \label{lemma:derivatives}
+\begin{restatable}[Derivatives of PDFs]{lemma}{derivatives} \label{lemma:derivatives}
   \leavevmode
@@
   \item For $i = 0, \dots, d$,
-    \[
-      \begin{split}
-        \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &=
-        \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1}
-          \tr(\mathbf{R} \adj(\bm\Gamma)) \\
-          &- (\mathbf{r} -
-          \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r}
-          - \mathbf{Su})),
-      \end{split}
+    \begin{enumerate}
+    \item
+      \begin{align*}
+        \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
+                                                                          &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
+      \end{align*}
+    \item For any covariance matrix $\mathbf{K}$,
+      \[
+        \frac{\partial \mathbf{K}}{\partial \lambda_i} =
+        \begin{cases}
+          \frac{1}{\lambda_i}\mathbf{K} & \text{if } i = 0, \\
+          \mathbf{L} & \text{otherwise,}
+        \end{cases}
+      \]
+      where
+      \[
+        L_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( -\frac{1}{2}(x_{j,i} -
+          x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
       \]
     \end{enumerate}
-\end{lemma}
-%\begin{proof}
-%  \leavevmode
-%  \begin{enumerate}
-%  \item
-%    \[
-%      \begin{split}
-%        \frac{\partial q(\mathbf{u})}{\partial m} &=
-%        q(\mathbf{u})\dm\left[-\frac{Q}{2}\right]
-%        \\
-%        &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
-%        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} -
-%        \bm\mu] \\
-%        &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
-%        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
-%      \end{split}
-%    \]
-%  \item
-%    \begin{align}
-%      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\exp \left( -\frac{Q}{2} \right)\right] \nonumber \\
-%                                                         &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\right]\exp \left( -\frac{Q}{2} \right) \nonumber \\
-%                                                         &+ \frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\dB\left[\exp\left( -\frac{Q}{2} \right)\right] \nonumber \\
-%                                                         &= -\frac{1}{2} q(\mathbf{u}) \left( |\bm\Sigma|^{-3/2} \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} + \frac{\partial Q}{\partial \mathbf{B}} \right).
-\label{eq:2_before}
-%    \end{align}
-%    For derivatives with respect to $\bm\Sigma$, we can refer to Petersen and
-%    Pedersen \cite{petersen2008matrix}:
-%    \begin{equation} \label{eq:partial_derivatives1}
-%      \frac{\partial |\bm\Sigma|}{\partial \bm\Sigma} = |\bm\Sigma|\bm\Sigma^{-\intercal}, \quad \frac{\partial Q}{\partial \mathbf{B}} = -\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal},
-%    \end{equation}
-%    while we can use an online tool by Laue et
-%    al.\footnote{\url{http://www.matrixcalculus.org/}}
-%    \cite{DBLP:conf/nips/LaueMG18} for the remaining ones:
-%    \begin{equation} \label{eq:partial_derivatives2}
-%      \begin{gathered}
-%        \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} =
-%        (\adj(\mathbf{T}) + \adj(\bm\Sigma))\mathbf{B}, \\
-%        \dB[(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} -
-%        \bm\mu)] = -(\mathbf{TUT} +
-%        \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1})\mathbf{B}.
-%      \end{gathered}
-%    \end{equation}
-%    Substituting results from Equations \ref{eq:partial_derivatives1} and
-%    \ref{eq:partial_derivatives2} back into Equation \ref{eq:2_before} gives:
-%    \begin{alignat*}{3}
-%      \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} &=
-%      \frac{1}{2}q(\mathbf{u})(&&\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
-%      - \bm\Sigma^{-\intercal}), \\
-%      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &=
-%      \frac{1}{2}q(\mathbf{u}) \{&&|\bm\Sigma|^{-3/2}(\adj(\mathbf{T}) +
-%      \adj(\bm\Sigma)) \\
-%      & &&+ \mathbf{TUT} + \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} \} \mathbf{B}.
-%    \end{alignat*}
-%  \item Using a result by Rasmussen and Williams % TODO: update the proof
-%                                % according to the new result
-%    \cite{DBLP:books/lib/RasmussenW06},
-%    \[
-%      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
-%      q(\mathbf{r} \mid \mathbf{u}) \dl
-%      \left[-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} -
-%        \frac{1}{2}\log|\Kuu| \right] = \frac{1}{2}q(\mathbf{r} \mid \mathbf{u})\tr
-%      \left((\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal} -
-%        \Kuu^{-1}) \frac{\partial \Kuu}{\partial \lambda_i} \right).
-%    \]
-%    The remaining derivative is
-%    \[
-%      \frac{\partial \Kuu}{\partial \lambda_i} =
-%      \begin{cases}
-%        \frac{1}{\lambda_i}\Kuu & \text{if } i = 0, \\
-%        \Luu & \text{otherwise,}
-%      \end{cases}
-%    \]
-%    where
-%    \[
-%      \begin{split}
-%        [\Luu]_{j,k} &= \dl k(\mathbf{x}_{\mathbf{u},j},
-%        \mathbf{x}_{\mathbf{u},k}) \\
-%        &= k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-%        \dl \left[-\frac{1}{2}(\mathbf{x}_{\mathbf{u},j} -
-%          \mathbf{x}_{\mathbf{u},k})^\intercal \bm\Lambda
-%          (\mathbf{x}_{\mathbf{u},j} - \mathbf{x}_{\mathbf{u},k}) -
-%          \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \right] \\
-%        &= k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-%        \dl \left[-\frac{1}{2}\sum_{l=1}^d \lambda_l
-%          (x_{\mathbf{u},j,l} - x_{\mathbf{u},k,l})^2 -
-%          \mathbbm{1}[j \ne k]\sigma^2\sum_{l=1}^d \lambda_l \right] \\
-%        &= k(\mathbf{x}_{\mathbf{u},j}, \mathbf{x}_{\mathbf{u},k})
-%        \left( -\frac{1}{2}(x_{\mathbf{u},j,i} -
-%        x_{\mathbf{u},k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
-%      \end{split}
-%    \]
-%  \end{enumerate}
-%\end{proof}
-
-\subsection{Evidence Lower Bound}
-
-\begin{equation} \label{eq:full}
-  \pfull = p(\mathbf{u}) \times p(\mathbf{r} \mid \mathbf{u}) \times p(\mathcal{D} \mid \mathbf{r}).
-\end{equation}
-
+  \end{enumerate}
+\end{restatable}
+
+\subsection{Details of the Model} \label{sec:details}
+
+In order to properly investigate the difference between variational inference
+and ML estimation, we keep other parts of the model the same.
+Namely, we set the covariance function to a version of the automatic relevance
+detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
+\begin{align*}
+  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
+                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
+\end{align*}
+where $\lambda_0$ is the overall `scale' factor for how similar or distant the
+states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
+matrix that determines the relevance of each feature (where $d$ denotes the
+number of features), $\mathbbm{1}$ is defined as
+\[ \mathbbm{1}[b] = \begin{cases}
+    1 & \mbox{if $b$ is true} \\
+    0 & \mbox{otherwise},
+  \end{cases} \]
+and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
+the value makes little difference to the performance of the algorithm
+\cite{DBLP:conf/nips/LevinePK11}). Our vector of parameters for the
+covariance function is then $\bm\lambda = (\lambda_0, \dots,
+\lambda_d)^\intercal$. % TODO: write about lambda having to be variational
+
+TODO: introductory sentence for the following assumption:
 \begin{equation} \label{eq:approximation}
-  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}).
-\end{equation}
+  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
+\end{equation}
+where
+\begin{equation}
+  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}),
+\end{equation}
+where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
+positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
+\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
+
+  \begin{figure}
+    \centering
+    \begin{tikzpicture}
+      \node[latent] (u) {$\mathbf{u}$};
+      \node[latent, right=of u] (r) {$\mathbf{r}$};
+      \node[obs, right=of r] (D) {$\mathcal{D}$};
+      \node[const, below=of r] (lambda) {$\bm\lambda$};
+      \node[const, below=of u] (mu) {$\bm\mu$};
+      \node[const, left=of mu] (Sigma) {$\bm\Sigma$};
+      \edge {mu, Sigma} {u};
+      \edge {lambda, u} {r};
+      \edge {r} {D};
+    \end{tikzpicture}
+    \label{fig:graphical_model}
+  \end{figure}
+
+This gives us our variational parameters $\bm\nu = (\bm\mu, \mathbf{B},
+\bm\lambda)^\intercal$. TODO: reference the figure
+
+\subsection{Evidence Lower Bound} \label{sec:elbo}
 
@@
 
+\subsection{Variational Inference} \label{vi_algs}
+
+The typical way to optimise a quantity (the ELBO, in this case) involves
+computing its gradient \cite{blei2017variational}. Unfortunately, some of the
+terms in $\mathcal{L}$ are still left as expected values. \emph{Black box
+  variational inference} (BBVI) \cite{DBLP:conf/aistats/RanganathGB14} suggests
+a way to express the gradient as an expectation without having to take the
+gradient of the posterior:
+\[ \nabla_{\bm\nu} \mathcal{L} = \mathbb{E}[\nabla_{\bm\nu} \log \approximation
+  (\log \pfull - \log \approximation)]. \]
+The gradient of the ELBO then has a unbiased estimate
+\[
+  \nabla_{\bm\nu} \mathcal{L} \approx \frac{1}{S} \sum_{s=1}^S \nabla_{\bm\nu} \log \approximationS (\log \pfullS - \log \approximationS),
+\]
+computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s, \mathbf{r}_s) \sim
+\approximation$.
+
+The main theorem behind the BBVI trick is the Lebesgue's dominated convergence
+theorem \cite{royden1988real}, which---under some conditions---allows one to
+claim that:
+\[ \lim_{n \to \infty} \int_E f_n\,d\mu = \int_E \lim_{n \to \infty}
+  f_n\,d\mu. \]
+We can use the same idea on parts of $\mathcal{L}$, derivatives of which cannot
+be taken otherwise. In our case, we want a gradient instead of a limit, and the
+integral represents an expected value.
+
+%While the ELBO can now be computed without approximations, we ended up imposing
+%two zero-variance restrictions, resulting in a less informative model. Moreover,
+%we had to abandon the Bayesian treatment of $\bm\lambda$, making the model more
+%vulnerable to overfitting.
+
 \subsection{Theoretical Justification}
+
+% TODO: reference in text
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \node [rectangle, draw, fill=lightgray] (perturbation) at (2, 0) {\textsc{Perturbation Lemma}};
+    \node [rectangle, draw, fill=lightgray] (lebesgue) at (-1, -3) {\textsc{Dominated Convergence Theorem}};
+    \node [rectangle, draw] (derivatives) at (-2, 0) {\textsc{Derivatives of PDFs}};
+    \node [rectangle, draw] (measurability) at (-3, -2) {\textsc{Measurability}};
+    \node [rectangle, draw] (boundedness) at (3, -2) {\textsc{Boundedness}};
+    \node [rectangle, draw] (lemma1) at (-2, -1) {\textsc{Lemma \ref{lemma:bound1}}};
+    \node [rectangle, draw] (lemma2) at (0, -1) {\textsc{Lemma \ref{lemma:bound2}}};
+    \node [rectangle, draw] (lemma3) at (2, -1) {\textsc{Lemma \ref{lemma:bound3}}};
+    \node [rectangle, draw] (lemma4) at (3, -3) {\textsc{Lemma \ref{lemma:integral_of_r}}};
+    \node [rectangle, draw] (theorem) at (0, -2) {\textsc{Theorem \ref{thm:main}}};
+
+    \draw [arrow] (perturbation) edge (lemma3);
+    \draw [arrow] (lebesgue) edge (theorem);
+    \draw [arrow] (derivatives) edge (lemma1);
+    \draw [arrow] (derivatives) edge (lemma2);
+    \draw [arrow] (derivatives) edge (lemma3);
+    \draw [arrow] (measurability) edge (theorem);
+    \draw [arrow] (boundedness) edge (theorem);
+    \draw [arrow] (lemma1) edge (theorem);
+    \draw [arrow] (lemma2) edge (theorem);
+    \draw [arrow] (lemma3) edge (theorem);
+    \draw [arrow] (lemma4) edge (theorem);
+  \end{tikzpicture}
+  \caption{A graphical representation of dependencies between our theoretical
+    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
+    Results from the literature are in gray.}
+\end{figure}
 
@@
 
-\begin{proposition} \label{thm:measurability}
+\begin{proposition}[Measurability] \label{thm:measurability}
   MDP value functions $V(s) : \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ (for $s
@@
 
-\begin{proposition} \label{thm:bound}
+\begin{proposition}[Boundedness] \label{thm:bound}
   If the initial values of the MDP value function satisfy the following
@@
 
-\begin{theorem}[Dominated Convergence Theorem
-  \cite{royden2010real}] \label{thm:lebesgue}
-  Let $(X, \mathcal{M}, \mu)$ be a measure space and $\{ f_n \}$ a sequence of
-  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise a.e. on $X$
-  and the function $f$ is measurable. Assume there is a non-negative function
-  $g$ that is integrable over $X$ and dominates the sequence $\{ f_n \}$ on $X$
-  in the sense that
-  \[
-    |f_n| \le g \text{ a.e. on $X$ for all $n$.}
-  \]
-  Then $f$ is integrable over $X$ and
-  \[
-    \lim_{n \to \infty} \int_X f_n\,d\mu = \int_X f\,d\mu.
-  \]
-\end{theorem}
-
-\begin{lemma} \label{lemma:bound1} % TODO: Lemma 1.1 is wrong about the
-                                % derivative. Fix
+\begin{lemma} \label{lemma:bound1}
   Let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
@@
 \end{lemma}
+% TODO: Lemma 1.1 is wrong about the % derivative. Fix (my handwritten notes
+% have the correct idea)
 \begin{proof}
   Remember that
-  \[
-    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
-    \frac{1}{2}q(\mathbf{r} \mid \mathbf{u})\tr
-    \left((\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal} - \Kuu^{-1})
-      \frac{\partial \Kuu}{\partial \lambda_i}
-    \right)
-  \]
+  \begin{align*}
+    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
+                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
+  \end{align*}
   by Lemma \ref{lemma:derivatives}. We begin by producing constant upper and
@@
   \[
-    \left. \frac{\partial \Kuu}{\partial \lambda_i} \right|_{\lambda_i =
-      c(\mathbf{r}, \mathbf{u})}.
-  \]
-  If $i = 0$, then each element of $\frac{\partial
-    \Kuu}{\partial \lambda_0}$ is of the form
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i =
+      c(\mathbf{r}, \mathbf{u})},
+  \]
+  where $\mathbf{K}$ is any covariance matrix.
+
+  If $i = 0$, then each element of $\frac{\partial \mathbf{K}}{\partial
+    \lambda_0}$ is of the form
   \[
@@
   \[
-    \left. \frac{\partial \Kuu}{\partial \lambda_0} \right|_{\lambda_0 =
-      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \Kuu}{\partial \lambda_0}
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
+      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \mathbf{K}}{\partial
+      \lambda_0}
   \]
@@
 
-  If $i > 0$, then each element of $\frac{\partial \Kuu}{\partial
+  If $i > 0$, then each element of $\frac{\partial \mathbf{K}}{\partial
     \lambda_i}$ is a constant multiple of $k(\mathbf{x}_j,
@@
   \[
-    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{1}{2}
-    \lambda_n(x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n,
+    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
+    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n,
   \]
@@
   \[
-    \left. \frac{\partial \Kuu}{\partial \lambda_i} \right|_{\lambda_i =
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i =
       c(\mathbf{r}, \mathbf{u})}.
@@
 
-\subsection{Derivatives}
-
-\subsubsection{\texorpdfstring{$\partial/\partial\bm\mu$}{Derivative w.r.t. mu}}
+\section{Evaluation} % showing deep understanding
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{elbo_contour_plots}
+\end{figure*}
+
+\section{Conclusions}
+
+\subsection{Future Work}
+
+%\vskip8pt \noindent
+%{\bf Acknowledgments.}
+%This is optional; it is a location for you to thank people
+\bibliographystyle{abbrv}
+\bibliography{paper}
+
+\appendix
+\section{Proofs}
+
+\derivatives*
+\begin{proof}
+  \leavevmode
+  \begin{enumerate}
+  \item
+    \[
+      \begin{split}
+        \frac{\partial q(\mathbf{u})}{\partial m} &=
+        q(\mathbf{u})\dm\left[-\frac{Q}{2}\right]
+        \\
+        &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} -
+        \bm\mu] \\
+        &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
+        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
+      \end{split}
+    \]
+  \item
+    \begin{align}
+      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\exp \left( -\frac{Q}{2} \right)\right] \nonumber \\
+                                                         &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\right]\exp \left( -\frac{Q}{2} \right) \nonumber \\
+                                                         &+ \frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\dB\left[\exp\left( -\frac{Q}{2} \right)\right] \nonumber \\
+                                                         &= -\frac{1}{2} q(\mathbf{u}) \left( |\bm\Sigma|^{-3/2} \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} + \frac{\partial Q}{\partial \mathbf{B}} \right).
+\label{eq:2_before}
+    \end{align}
+    For derivatives with respect to $\bm\Sigma$, we can refer to Petersen and
+    Pedersen \cite{petersen2008matrix}:
+    \begin{equation} \label{eq:partial_derivatives1}
+      \frac{\partial |\bm\Sigma|}{\partial \bm\Sigma} = |\bm\Sigma|\bm\Sigma^{-\intercal}, \quad \frac{\partial Q}{\partial \mathbf{B}} = -\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal},
+    \end{equation}
+    while we can use an online tool by Laue et
+    al.\footnote{\url{http://www.matrixcalculus.org/}}
+    \cite{DBLP:conf/nips/LaueMG18} for the remaining ones:
+    \begin{equation} \label{eq:partial_derivatives2}
+      \begin{gathered}
+        \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} =
+        (\adj(\mathbf{T}) + \adj(\bm\Sigma))\mathbf{B}, \\
+        \dB[(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} -
+        \bm\mu)] = -(\mathbf{TUT} +
+        \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1})\mathbf{B}.
+      \end{gathered}
+    \end{equation}
+    Substituting results from Equations \ref{eq:partial_derivatives1} and
+    \ref{eq:partial_derivatives2} back into Equation \ref{eq:2_before} gives:
+    \begin{alignat*}{3}
+      \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} &=
+      \frac{1}{2}q(\mathbf{u})(&&\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
+      - \bm\Sigma^{-\intercal}), \\
+      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &=
+      \frac{1}{2}q(\mathbf{u}) \{&&|\bm\Sigma|^{-3/2}(\adj(\mathbf{T}) +
+      \adj(\bm\Sigma)) \\
+      & &&+ \mathbf{TUT} + \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} \} \mathbf{B}.
+    \end{alignat*}
+  \item
+    \begin{enumerate}
+    \item Since
+    \begin{align*}
+      q(\mathbf{r} \mid \mathbf{u}) &= \mathcal{N}(\mathbf{r};
+      \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal \Kuu^{-1}\Kru) \\
+      &= \mathcal{N}(\mathbf{r}; \mathbf{Su}, \bm\Gamma),
+    \end{align*}
+    we have
+    \[
+      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
+      -\frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) \dl[(\mathbf{r} -
+      \mathbf{Su})^\intercal\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su}) +
+      \log|\bm\Gamma|].
+    \]
+    The same online tool can be used to show that
+    \[
+      \dl \log|\bm\Gamma| = -|\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma)),
+    \]
+    and
+    \[
+      \dl \bm\Gamma^{-1} = \bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}.
+    \]
+  \item
+    If $i=0$, then
+    \[
+      \frac{\partial \mathbf{K}}{\partial \lambda_i} =
+      \frac{1}{\lambda_i}\mathbf{K}
+    \]
+    by the structure of each element of $\mathbf{K}$. If $i \ne 0$, then each
+    element of $\frac{\partial \mathbf{K}}{\partial \lambda_i}$ is
+    \begin{align*}
+      L_{j,k} &= \frac{\partial k(\mathbf{x}_j, \mathbf{x}_k)}{\partial \lambda_i} \\
+      &\begin{aligned}
+        = k(\mathbf{x}_j, \mathbf{x}_k) \dl \left[ \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) \\
+        &- \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right]
+      \end{aligned} \\
+      &\begin{aligned}
+        = k(\mathbf{x}_j, \mathbf{x}_k) \dl \left[ \vphantom{\sum_{k=1}^d} \right. &-\frac{1}{2} \sum_{l=1}^d \lambda_l (x_{j,l} - x_{k,l})^2 \\
+        &- \mathbbm{1}[j \ne k]\sigma^2\sum_{l=1}^d \lambda_l \left. \vphantom{\sum_{l=1}^d} \right]
+      \end{aligned} \\
+      &= k(\mathbf{x}_j, \mathbf{x}_k) \left( -\frac{1}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
+    \end{align*}
+    \end{enumerate}
+  \end{enumerate}
+\end{proof}
+
+\section{Derivatives}
+
+\subsection{\texorpdfstring{$\partial/\partial\bm\mu$}{Derivative w.r.t. mu}}
 
@@
   \Kuu^{-\intercal}) \bm\mu
-\] % TODO: by the cookbook
-and
+\]
+by Petersen and Pedersen \cite{petersen2008matrix}, and
 \[
@@
 
-\subsubsection{\texorpdfstring{$\partial/\partial\mathbf{B}$}{Derivative w.r.t. B}}
+\subsection{\texorpdfstring{$\partial/\partial\mathbf{B}$}{Derivative w.r.t. B}}
 
@@
 
-\subsubsection{\texorpdfstring{$\partial/\partial \lambda_j$}{Derivative w.r.t.
+\subsection{\texorpdfstring{$\partial/\partial \lambda_j$}{Derivative w.r.t.
     Lambda}}
@@
 by Theorem \ref{thm:main} and Lemma \ref{lemma:derivatives}. Thus,
-\[
-  \begin{split}
+\begin{align*}
   \frac{\partial \mathcal{L}}{\partial \lambda_j} &=
@@
       \Kuu}{\partial \lambda_j} \right) \Kuu^{-1} \bm\mu \\
-    &+ \frac{1}{2} \left[ \begin{split}
-      &\tr \left( \Kuu^{-1} \frac{\partial \Kuu}{\partial
-          \lambda_j} \Kuu^{-1} \bm\Sigma \right) + \bm\mu^\intercal \Kuu^{-1}
-      \frac{\partial \Kuu}{\partial \lambda_j} \Kuu^{-1} \bm\mu \\
-      &- \tr \left(\Kuu^{-1} \frac{\partial \Kuu}{\partial \lambda_j} \right)
-      \end{split} \right]
-    \\
-    &- \frac{1}{2} \mathbb{E} \left[ \left( \begin{split}
-        &|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
-          &- (\mathbf{r} -
-          \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r}
-          - \mathbf{Su})
-    \end{split} \right) v \right],
-  \end{split}
-\]
+  &\begin{aligned}
+    + \frac{1}{2} \left[ \vphantom{\frac{\partial \Kuu}{\partial \lambda_j}}
+    \right. &\tr \left( \Kuu^{-1} \frac{\partial \Kuu}{\partial \lambda_j}
+      \Kuu^{-1} \bm\Sigma \right) + \bm\mu^\intercal \Kuu^{-1} \frac{\partial
+      \Kuu}{\partial \lambda_j} \Kuu^{-1} \bm\mu \\
+    &- \tr \left. \left(\Kuu^{-1} \frac{\partial \Kuu}{\partial \lambda_j}
+      \right) \right]
+  \end{aligned} \\
+  &\begin{aligned}
+    - \frac{1}{2} \mathbb{E} [ (&|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
+    &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})) v],
+  \end{aligned} 
+\end{align*}
 where the remaining derivatives can be found in Lemma \ref{lemma:derivatives}.
 
-\section{Evaluation} % showing deep understanding
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\textwidth]{elbo_contour_plots}
-\end{figure*}
-
-\section{Conclusions}
-
-\subsection{Future Work}
-
-%\vskip8pt \noindent
-%{\bf Acknowledgments.}
-%This is optional; it is a location for you to thank people
-\bibliographystyle{abbrv}
-\bibliography{paper}
 \end{document}

### 2019-03-10 mpaper/paper.tex
Introduction to the theoretical results

@@
 
+% TODO: mention in text
+%While the ELBO can now be computed without approximations, we ended up imposing
+%two zero-variance restrictions, resulting in a less informative model. Moreover,
+%we had to abandon the Bayesian treatment of $\bm\lambda$, making the model more
+%vulnerable to overfitting.
+
 \subsection{Evidence Lower Bound} \label{sec:elbo}
@@
 
-\subsection{Variational Inference} \label{vi_algs}
-
-The typical way to optimise a quantity (the ELBO, in this case) involves
-computing its gradient \cite{blei2017variational}. Unfortunately, some of the
-terms in $\mathcal{L}$ are still left as expected values. \emph{Black box
-  variational inference} (BBVI) \cite{DBLP:conf/aistats/RanganathGB14} suggests
-a way to express the gradient as an expectation without having to take the
-gradient of the posterior:
-\[ \nabla_{\bm\nu} \mathcal{L} = \mathbb{E}[\nabla_{\bm\nu} \log \approximation
-  (\log \pfull - \log \approximation)]. \]
-The gradient of the ELBO then has a unbiased estimate
-\[
-  \nabla_{\bm\nu} \mathcal{L} \approx \frac{1}{S} \sum_{s=1}^S \nabla_{\bm\nu} \log \approximationS (\log \pfullS - \log \approximationS),
-\]
-computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s, \mathbf{r}_s) \sim
-\approximation$.
-
-The main theorem behind the BBVI trick is the Lebesgue's dominated convergence
-theorem \cite{royden1988real}, which---under some conditions---allows one to
-claim that:
-\[ \lim_{n \to \infty} \int_E f_n\,d\mu = \int_E \lim_{n \to \infty}
-  f_n\,d\mu. \]
-We can use the same idea on parts of $\mathcal{L}$, derivatives of which cannot
-be taken otherwise. In our case, we want a gradient instead of a limit, and the
-integral represents an expected value.
-
-%While the ELBO can now be computed without approximations, we ended up imposing
-%two zero-variance restrictions, resulting in a less informative model. Moreover,
-%we had to abandon the Bayesian treatment of $\bm\lambda$, making the model more
-%vulnerable to overfitting.
-
 \subsection{Theoretical Justification}
 
-% TODO: reference in text
 \begin{figure}
@@
     Results from the literature are in gray.}
+  \label{fig:theorems}
 \end{figure}
 
+The typical way to optimise a quantity (the ELBO, in this case) involves
+computing its gradient. Unfortunately, some of the terms in $\mathcal{L}$ are
+still left as expected values. The goal of this section is to show how Theorem
+\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
+anyway\footnote{This technique is inspired by black box variational inference
+  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
+  problem and requires significantly more work to prove correctness.}. After
+showing that the theorem applies to our situation, given any term expressed as
+an expected value $\mathbb{E}[f(\mathbf{u}, \mathbf{r})]$, we can estimate its
+gradient as
+\[
+  \nabla\mathbb{E}[f(\mathbf{u}, \mathbf{r})] = \mathbb{E}[\nabla f(\mathbf{u},
+  \mathbf{r})] \approx \frac{1}{S} \sum_{s=1}^S \nabla f(\mathbf{u}_s,
+  \mathbf{r}_s),
+\]
+which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
+\mathbf{r}_s) \sim \approximation$.
+
+Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
+inside the integral. In order to prove it, we use a number of intermediate
+results. We start by providing a sketch proof of the measurability of MDP value
+functions, which is non-obvious due to their non-trivial definition. We then
+prove that the value functions are bounded, and establish four helpful lemmas,
+three of which provide well-behaved bounds on derivatives of PDFs. See Figure
+\ref{fig:theorems} for an overview of how the results fit together.
+
+\begin{remark}
   MDP values are characterised by both a state and a reward function/vector. In
-order to prove the next theorem, we think of the value function as $V :
-\mathcal{S} \to \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a
-state $s \in \mathcal{S}$ and returns a function $V(s) :
-\mathbb{R}^{\mathcal{S}} \to \mathbb{R}$ that takes a reward vector $\mathbf{r}
-\in \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$,
-$\V(s) \in \mathbb{R}$. The function $V(s)$ computes the values of
+  this section, we think of the value function as $V : \mathcal{S} \to
+  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a state $s \in
+  \mathcal{S}$ and returns a function $V(s) : \mathbb{R}^{|\mathcal{S}|} \to
+  \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
+  \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$, $\V(s) \in
+  \mathbb{R}$. Given a reward vector, the function $V(s)$ computes the values of
   all states and returns the value of state $s$.
+\end{remark}
 
@@
 \end{lemma}
-% TODO: Lemma 1.1 is wrong about the % derivative. Fix (my handwritten notes
-% have the correct idea)
 \begin{proof}
@@
   \end{align*}
-  by Lemma \ref{lemma:derivatives}. We begin by producing constant upper and
-  lower bounds for the elements of
-  \[
-    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i =
-      c(\mathbf{r}, \mathbf{u})},
-  \]
-  where $\mathbf{K}$ is any covariance matrix.
+  by Lemma \ref{lemma:derivatives}.
+
+  %We begin by producing constant upper and
+  %lower bounds for the elements of
+  %\[
+  %  \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i =
+  %    c(\mathbf{r}, \mathbf{u})},
+  %\]
+  %where $\mathbf{K}$ is any covariance matrix.
 
@@
 
+  % TODO: Lemma 1.1 is wrong about the % derivative. Fix (my handwritten notes
+  % have the correct idea)
   Combining the bounds with the observation that
@@
 \begin{remark}
-  In order to find $\frac{\partial q(\mathbf{u})}{\partial t}$,
-  where $t$ is the $i$th element of the vector $\bm\mu$, we can
-  find $\frac{\partial q(\mathbf{u})}{\partial \bm\mu}$ and simply take the
-  $i$th element. A similar line of reasoning applies to matrices as well. Thus,
-  we only need to consider derivatives with respect to $\bm\mu$ and
-  $\bm\Sigma$.
+  In order to find a derivative such as $\frac{\partial q(\mathbf{u})}{\partial
+    \mu_i}$, we can find $\frac{\partial q(\mathbf{u})}{\partial \bm\mu}$ and
+  simply take the $i$th element. A similar line of reasoning applies to matrices
+  as well. Thus, we only need to consider derivatives with respect to $\bm\mu$
+  and $\bm\Sigma$.
 \end{remark}

### 2019-03-11 mpaper/paper.tex
Minor readability improvements

@@
 
-\subsection{Linear Algebra and Analysis}
+\subsection{Linear Algebra, Numerical Analysis, and Measure Theory}
+
+Here we introduce a few definitions and results that will be used later in the
+paper. Namely, we will use several different vector and matrix norms, consider
+how an inverse of a matrix changes with a small perturbation, and use Lebesgue's
+dominated convergence theorem in order to justify the validity of our approach.
 
@@
 % TODO: express all lambdas as exponentials (update the derivatives or sth)
-
-\subsection{Notation}
-
-We use $\tr(\mathbf{A})$ to denote its \emph{trace} and
-$\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}).
-
-For any vector $\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a
-vector space of polynomials with degree at most $d$, where variables are
-elements of $\mathbf{x}$, and coefficients are in $\mathbb{R}$.
-
-\subsection{Preliminaries}
+% TODO: give an overview of this section. Also, consider renaming it or
+% splitting it into multiple sections
+
+For any matrix $\mathbf{A}$, we will use either $A_{i,j}$ or
+$[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
+$j$. Moreover, we use $\tr(\mathbf{A})$ to denote its \emph{trace} and
+$\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}). For
+any vector $\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a vector
+space of polynomials with degree at most $d$, where variables are elements of
+$\mathbf{x}$, and coefficients are in $\mathbb{R}$.
 
@@
 
-As recently suggested by Ong et al. \cite{ong2018gaussian}, we use a
-decomposition $\bm\Sigma = \mathbf{B}\mathbf{B}^\intercal + \mathbf{D}^2$, where
-$\mathbf{B}$ is a lower triangular $m \times p$ matrix with positive diagonal
-entries, and $\mathbf{D} = \diag(d_1, \dots, d_m)$. Typically, we would set $p$
-so that $p \ll m$ to get an efficient approximation, but it is also worth
-pointing out that we can retain full accuracy by setting $p = m$ and $\mathbf{D}
-= \mathbf{O}$. Moreover, we define a few variables that will simplify
-expressions for the derivatives:
-\begin{align*}
-  \mathbf{U} &= (\mathbf{u} - \bm\mu)(\mathbf{u} - \bm\mu)^\intercal \\
+\subsection{Details of the Model}
+
+In order to properly investigate the difference between VI
+and ML estimation, we keep other parts of the model the same.
+Namely, we set the covariance function to a version of the automatic relevance
+detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
+\begin{align*}
+  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
+                                                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
+\end{align*}
+where $\lambda_0$ is the overall `scale' factor for how similar or distant the
+states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
+matrix that determines the relevance of each feature (where $d$ denotes the
+number of features), $\mathbbm{1}$ is defined as
+\[
+  \mathbbm{1}[b] = \begin{cases}
+    1 & \mbox{if $b$ is true} \\
+    0 & \mbox{otherwise},
+  \end{cases}
+\]
+and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
+the value makes little difference to the performance of the algorithm
+\cite{DBLP:conf/nips/LevinePK11}). We will write $\bm\lambda = (\lambda_0,
+\dots, \lambda_d)^\intercal$ to refer to both $\lambda_0$ and $\bm\Lambda$ at
+the same time.
+
+% TODO: write about lambda having to be variational
+% TODO: mention the open problem of finding the expected value of the inverse of
+% a matrix, where each element is a non-trivial function (w.r.t. a parameter of
+% the function)
+%TODO: introductory sentence for the following assumption:
+\begin{equation} \label{eq:approximation}
+  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
+\end{equation}
+where
+\begin{equation}
+  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma),
+\end{equation}
+where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
+positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
+\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
+
+Ong et al. \cite{ong2018gaussian} have recently suggested that, in order to make
+variational approximation of a multivariate Gaussian more scalable, the
+covariance matrix should be decomposed as $\bm\Sigma =
+\mathbf{B}\mathbf{B}^\intercal + \mathbf{D}^2$, where $\mathbf{B}$ is a lower
+triangular $m \times p$ matrix with positive diagonal entries, and $\mathbf{D}$
+is a diagonal matrix. Typically, we would set $p$ so that $p \ll m$ to get an
+efficient approximation. However, as our goal is precision rather than
+scalability, we will set $p = m$ and $\mathbf{D} = \mathbf{O}_m$ in order to
+retain full covariance structure.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \node[latent] (u) {$\mathbf{u}$};
+    \node[latent, right=of u] (r) {$\mathbf{r}$};
+    \node[obs, right=of r] (D) {$\mathcal{D}$};
+    \node[const, below=of r] (lambda) {$\bm\lambda$};
+    \node[const, below=of u] (mu) {$\bm\mu$};
+    \node[const, left=of mu] (B) {$\mathbf{B}$};
+    \edge {mu, B} {u};
+    \edge {lambda, u} {r};
+    \edge {r} {D};
+  \end{tikzpicture}
+  \caption{Our VI problem expressed as a (simplified)
+    Bayesian network. The only observed variable (representing the
+    demonstrations) is in a gray circle, modelled latent variables are in white
+    circles, and the variational parameters are at the bottom.}
+  \label{fig:graphical_model}
+\end{figure}
+
+The resulting model is summarised in Figure \ref{fig:graphical_model}. We rely
+on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and the model.
+Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
+and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
+parameters of the kernel, $\bm\lambda$. The two remaining dependencies are quite
+clear, given that the approximating distribution for $\mathbf{u}$ is
+$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$. % TODO: concluding sentence
+
+\subsection{Evidence Lower Bound} \label{sec:elbo}
+
+In this section, we derive and simplify the ELBO for this (now fully specified)
+model. Note that in order to keep the derivation simple, we drop all constant
+terms in the expression of $\mathcal{L}$, i.e., equality is taken to mean
+`equality up to an additive constant'. Also note that all expected values are
+with respect to $(\mathbf{u}, \mathbf{r}) \sim q(\mathbf{u}, \mathbf{r})$.
+
+In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
+write
+\[
+  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
+\]
+By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
+\begin{align*}
+  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
+              &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
+\end{align*}
+Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
+cancel each other out. Also notice that
+\begin{align*}
+  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
+                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
+                                                      &+ \log |\Kuu| - \log |\bm\Sigma|),
+\end{align*}
+by the definition of KL divergence between two multivariate Gaussians
+\cite{kl}. Hence,
+\begin{align*}
+  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
+              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
+\end{align*}
+Using the expressions for $Q_{\mathbf{r}}$ we get
+\begin{align*}
+  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
+              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
+\end{align*}
+We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
+$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
+number of times the state associated with the reward $r_i$ has been visited
+across all demonstrations. Then
+\begin{align*}
+  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
+                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
+\end{align*}
+This allows us to simplify $\mathcal{L}$ to
+\begin{align*}
+  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
+              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
+\end{align*}
+where
+\[
+  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
+  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
+\]
+
+\subsection{Theoretical Justification}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \node [rectangle, draw, fill=lightgray] (perturbation) at (2, 0) {\textsc{Perturbation Lemma}};
+    \node [rectangle, draw, fill=lightgray] (lebesgue) at (-1, -3) {\textsc{Dominated Convergence Theorem}};
+    \node [rectangle, draw] (derivatives) at (-2, 0) {\textsc{Derivatives of PDFs}};
+    \node [rectangle, draw] (measurability) at (-3, -2) {\textsc{Measurability}};
+    \node [rectangle, draw] (boundedness) at (3, -2) {\textsc{Boundedness}};
+    \node [rectangle, draw] (lemma1) at (-2, -1) {\textsc{Lemma \ref{lemma:bound1}}};
+    \node [rectangle, draw] (lemma2) at (0, -1) {\textsc{Lemma \ref{lemma:bound2}}};
+    \node [rectangle, draw] (lemma3) at (2, -1) {\textsc{Lemma \ref{lemma:bound3}}};
+    \node [rectangle, draw] (lemma4) at (3, -3) {\textsc{Lemma \ref{lemma:integral_of_r}}};
+    \node [rectangle, draw] (theorem) at (0, -2) {\textsc{Theorem \ref{thm:main}}};
+
+    \draw [arrow] (perturbation) edge (lemma3);
+    \draw [arrow] (lebesgue) edge (theorem);
+    \draw [arrow] (derivatives) edge (lemma1);
+    \draw [arrow] (derivatives) edge (lemma2);
+    \draw [arrow] (derivatives) edge (lemma3);
+    \draw [arrow] (measurability) edge (theorem);
+    \draw [arrow] (boundedness) edge (theorem);
+    \draw [arrow] (lemma1) edge (theorem);
+    \draw [arrow] (lemma2) edge (theorem);
+    \draw [arrow] (lemma3) edge (theorem);
+    \draw [arrow] (lemma4) edge (theorem);
+  \end{tikzpicture}
+  \caption{A graphical representation of dependencies between our theoretical
+    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
+    Results from the literature are in gray.}
+  \label{fig:theorems}
+\end{figure}
+
+The typical way to optimise a quantity (the ELBO, in this case) involves
+computing its gradient. Unfortunately, some of the terms in $\mathcal{L}$ are
+still left as expected values. The goal of this section is to show how Theorem
+\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
+anyway\footnote{This technique is inspired by black box VI
+  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
+  problem and requires significantly more work to prove correctness.}. After
+showing that the theorem applies to our situation, given any term expressed as
+an expected value $\mathbb{E}[f(\mathbf{u}, \mathbf{r})]$, we can estimate its
+gradient as
+\[
+  \nabla\mathbb{E}[f(\mathbf{u}, \mathbf{r})] = \mathbb{E}[\nabla f(\mathbf{u},
+  \mathbf{r})] \approx \frac{1}{S} \sum_{s=1}^S \nabla f(\mathbf{u}_s,
+  \mathbf{r}_s),
+\]
+which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
+\mathbf{r}_s) \sim \approximation$.
+
+Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
+inside the integral. In order to prove it, we use a number of intermediate
+results. We start by stating a few derivatives of probability density functions
+(PDFs) and covariance matrices, and bound their values with some
+easy-to-deal-with polynomials. We then provide a sketch proof of the
+measurability of MDP value functions, which is non-obvious due to their
+non-trivial definition. Afterwards, we establish bounds for the value functions,
+and, after another quick lemma, tackle the main proof of this paper. See Figure
+\ref{fig:theorems} for an overview of how these results fit together.
+
+Before that, however, we define a few extra variables in order to simplify
+expressions of derivatives:
+\begin{align*}
+  \mathbf{U} &= (\mathbf{u} - \bm\mu)(\mathbf{u} - \bm\mu)^\intercal, \\
   \mathbf{S} &= \Kru^\intercal\Kuu^{-1}, \\
@@
 
-% TODO: define PDF before this point
 \begin{restatable}[Derivatives of PDFs]{lemma}{derivatives} \label{lemma:derivatives}
@@
 
-\subsection{Details of the Model} \label{sec:details}
-
-In order to properly investigate the difference between variational inference
-and ML estimation, we keep other parts of the model the same.
-Namely, we set the covariance function to a version of the automatic relevance
-detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
-\begin{align*}
-  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
-                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
-\end{align*}
-where $\lambda_0$ is the overall `scale' factor for how similar or distant the
-states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
-matrix that determines the relevance of each feature (where $d$ denotes the
-number of features), $\mathbbm{1}$ is defined as
-\[ \mathbbm{1}[b] = \begin{cases}
-    1 & \mbox{if $b$ is true} \\
-    0 & \mbox{otherwise},
-  \end{cases} \]
-and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
-the value makes little difference to the performance of the algorithm
-\cite{DBLP:conf/nips/LevinePK11}). Our vector of parameters for the
-covariance function is then $\bm\lambda = (\lambda_0, \dots,
-\lambda_d)^\intercal$. % TODO: write about lambda having to be variational
-
-TODO: introductory sentence for the following assumption:
-\begin{equation} \label{eq:approximation}
-  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
-\end{equation}
-where
-\begin{equation}
-  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}),
-\end{equation}
-where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
-positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
-\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
-
-  \begin{figure}
-    \centering
-    \begin{tikzpicture}
-      \node[latent] (u) {$\mathbf{u}$};
-      \node[latent, right=of u] (r) {$\mathbf{r}$};
-      \node[obs, right=of r] (D) {$\mathcal{D}$};
-      \node[const, below=of r] (lambda) {$\bm\lambda$};
-      \node[const, below=of u] (mu) {$\bm\mu$};
-      \node[const, left=of mu] (Sigma) {$\bm\Sigma$};
-      \edge {mu, Sigma} {u};
-      \edge {lambda, u} {r};
-      \edge {r} {D};
-    \end{tikzpicture}
-    \label{fig:graphical_model}
-  \end{figure}
-
-This gives us our variational parameters $\bm\nu = (\bm\mu, \mathbf{B},
-\bm\lambda)^\intercal$. TODO: reference the figure
-
-% TODO: mention in text
-%While the ELBO can now be computed without approximations, we ended up imposing
-%two zero-variance restrictions, resulting in a less informative model. Moreover,
-%we had to abandon the Bayesian treatment of $\bm\lambda$, making the model more
-%vulnerable to overfitting.
-
-\subsection{Evidence Lower Bound} \label{sec:elbo}
-
-In this section we derive and simplify the ELBO for this (now fully specified)
-model. In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
-write\footnote{At this point, we will drop the subscript denoting which
-  variables the expectation is taken over. Also note that throughout the
-  derivation equality is taken to mean `equality up to an additive constant'.}
-% TODO: move this to the front and say that the expectations are always w.r.t.
-% the full VI approximation unless specified otherwise
-\[
-  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
-\]
-By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
-\begin{align*}
-  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
-              &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
-\end{align*}
-Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
-cancel each other out. Also notice that
-\begin{align*}
-  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
-                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
-                                                      &+ \log |\Kuu| - \log |\bm\Sigma|),
-\end{align*}
-by the definition of KL divergence between two multivariate normal distributions
-\cite{kl}. Hence,
-\begin{align*}
-  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
-              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
-\end{align*}
-Using the expressions for $Q_{\mathbf{r}}$ we get
-\begin{align*}
-  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
-              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
-\end{align*}
-We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
-$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
-number of times the state associated with the reward $r_i$ has been visited
-across all demonstrations. Then
-\begin{align*}
-  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
-                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
-\end{align*}
-This allows us to simplify $\mathcal{L}$ to
-\begin{align*}
-  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
-              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
-\end{align*}
-where
-\[
-  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
-  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
-\]
-
-\subsection{Theoretical Justification}
-
-\begin{figure}
-  \centering
-  \begin{tikzpicture}
-    \node [rectangle, draw, fill=lightgray] (perturbation) at (2, 0) {\textsc{Perturbation Lemma}};
-    \node [rectangle, draw, fill=lightgray] (lebesgue) at (-1, -3) {\textsc{Dominated Convergence Theorem}};
-    \node [rectangle, draw] (derivatives) at (-2, 0) {\textsc{Derivatives of PDFs}};
-    \node [rectangle, draw] (measurability) at (-3, -2) {\textsc{Measurability}};
-    \node [rectangle, draw] (boundedness) at (3, -2) {\textsc{Boundedness}};
-    \node [rectangle, draw] (lemma1) at (-2, -1) {\textsc{Lemma \ref{lemma:bound1}}};
-    \node [rectangle, draw] (lemma2) at (0, -1) {\textsc{Lemma \ref{lemma:bound2}}};
-    \node [rectangle, draw] (lemma3) at (2, -1) {\textsc{Lemma \ref{lemma:bound3}}};
-    \node [rectangle, draw] (lemma4) at (3, -3) {\textsc{Lemma \ref{lemma:integral_of_r}}};
-    \node [rectangle, draw] (theorem) at (0, -2) {\textsc{Theorem \ref{thm:main}}};
-
-    \draw [arrow] (perturbation) edge (lemma3);
-    \draw [arrow] (lebesgue) edge (theorem);
-    \draw [arrow] (derivatives) edge (lemma1);
-    \draw [arrow] (derivatives) edge (lemma2);
-    \draw [arrow] (derivatives) edge (lemma3);
-    \draw [arrow] (measurability) edge (theorem);
-    \draw [arrow] (boundedness) edge (theorem);
-    \draw [arrow] (lemma1) edge (theorem);
-    \draw [arrow] (lemma2) edge (theorem);
-    \draw [arrow] (lemma3) edge (theorem);
-    \draw [arrow] (lemma4) edge (theorem);
-  \end{tikzpicture}
-  \caption{A graphical representation of dependencies between our theoretical
-    results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
-    Results from the literature are in gray.}
-  \label{fig:theorems}
-\end{figure}
-
-The typical way to optimise a quantity (the ELBO, in this case) involves
-computing its gradient. Unfortunately, some of the terms in $\mathcal{L}$ are
-still left as expected values. The goal of this section is to show how Theorem
-\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
-anyway\footnote{This technique is inspired by black box variational inference
-  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
-  problem and requires significantly more work to prove correctness.}. After
-showing that the theorem applies to our situation, given any term expressed as
-an expected value $\mathbb{E}[f(\mathbf{u}, \mathbf{r})]$, we can estimate its
-gradient as
-\[
-  \nabla\mathbb{E}[f(\mathbf{u}, \mathbf{r})] = \mathbb{E}[\nabla f(\mathbf{u},
-  \mathbf{r})] \approx \frac{1}{S} \sum_{s=1}^S \nabla f(\mathbf{u}_s,
-  \mathbf{r}_s),
-\]
-which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
-\mathbf{r}_s) \sim \approximation$.
-
-Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
-inside the integral. In order to prove it, we use a number of intermediate
-results. We start by providing a sketch proof of the measurability of MDP value
-functions, which is non-obvious due to their non-trivial definition. We then
-prove that the value functions are bounded, and establish four helpful lemmas,
-three of which provide well-behaved bounds on derivatives of PDFs. See Figure
-\ref{fig:theorems} for an overview of how the results fit together.
-
-\begin{remark}
-  MDP values are characterised by both a state and a reward function/vector. In
-  this section, we think of the value function as $V : \mathcal{S} \to
-  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a state $s \in
-  \mathcal{S}$ and returns a function $V(s) : \mathbb{R}^{|\mathcal{S}|} \to
-  \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
-  \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$, $\V(s) \in
-  \mathbb{R}$. Given a reward vector, the function $V(s)$ computes the values of
-  all states and returns the value of state $s$.
-\end{remark}
-
-\begin{proposition}[Measurability] \label{thm:measurability}
-  MDP value functions $V(s) : \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ (for $s
-  \in \mathcal{S}$) are Lebesgue measurable.
-\end{proposition}
-\begin{proof}
-  For any reward vector $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$, the
-  collection of converged value functions $\{ \V(s) \mid s \in
-  \mathcal{S} \}$ satisfy
-  \begin{equation} \label{eq:linear_mdp}
-    \V(s) = \log \sum_{a \in \mathcal{A}}
-    \exp\left( r(s) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a,
-      s')\V(s') \right)
-  \end{equation}
-  for all $s \in \mathcal{S}$. Let $s_0 \in \mathcal{S}$ be an arbitrary state.
-  In order to prove that $V(s_0)$ is measurable, it is enough to show that for
-  any $\alpha \in \mathbb{R}$, the set
-  \[
-    \left\{ \mathbf{r} \in \mathbb{R}^{|\mathcal{S}|} \; \middle|
-    \begin{aligned}
-      \;&\V(s_0) \in (-\infty, \alpha); \\
-      \;&\V(s) \in \mathbb{R} \text{ for all } s \in \mathcal{S} \setminus \{ s_0 \}; \\
-      \;&\text{ Equation \ref{eq:linear_mdp} is satisfied by all } s \in
-      \mathcal{S}
-    \end{aligned}
-    \right\}
-  \]
-  is measurable. Since this set can be constructed in Zermelo-Fraenkel set
-  theory \emph{without} the axiom of choice, it is measurable
-  \cite{herrlich2006axiom}, which proves that $V(s)$ is a measurable function
-  for any $s \in \mathcal{S}$.
-\end{proof}
-
-\begin{proposition}[Boundedness] \label{thm:bound}
-  If the initial values of the MDP value function satisfy the following
-  bound, then the bound remains satisfied throughout value iteration:
-  \begin{equation} \label{eq:bound}
-    |\V(s)| \le \vbound.
-  \end{equation}
-\end{proposition}
-\begin{proof}
-  We begin by considering Equation \ref{eq:bound} without taking the absolute value of
-  $\V(s)$, i.e.,
-  \begin{equation} \label{eq:positive_bound}
-    \V(s) \le \vbound,
-  \end{equation}
-  and assuming that the initial values of $\{ \V(s) \mid s \in
-  \mathcal{S} \}$ already satisfy Equation \ref{eq:positive_bound}. Recall that for
-  each $s \in \mathcal{S}$, the value of $\V(s)$ is updated by applying Equation
-  \ref{eq:update_rule}. Note that both $\log$ and $\exp$ are increasing
-  functions, $\gamma > 0$, and the $\mathcal{T}$ function gives a probability (a
-  non-negative number). Thus
-  \begin{align*}
-    \V(s) &\le \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s')\frac{\rinf + \log|\mathcal{A}|}{1 - \gamma} \right) \\
-          &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma}\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s') \right) \\
-          &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right)
-  \end{align*}
-  by the definition of $\mathcal{T}$. Then
-  \begin{align*}
-    \V(s) &\le \log \left( |\mathcal{A}| \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
-          &= \log \left( \exp\left( \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
-          &= \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \\
-          &= \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 - \gamma)(\log|\mathcal{A}| + r(s))}{1 - \gamma} \\
-          &\le \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 - \gamma)(\log|\mathcal{A}| + \rinf)}{1 - \gamma} \\
-          &= \vbound
-  \end{align*}
-  by the definition of $\rinf$.
-
-  The proof for
-  \begin{equation} \label{eq:negative_bound}
-    \V(s) \ge \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1}
-  \end{equation}
-  follows the same argument until we get to
-  \begin{align*}
-    \V(s) &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma - 1)(\log|\mathcal{A}| + r(s))}{\gamma - 1} \\
-          &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma - 1)(-\log|\mathcal{A}| -\rinf)}{\gamma - 1} \\
-          &= \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1},
-  \end{align*}
-  where we use the fact that $r(s) \ge -\rinf - 2\log|\mathcal{A}|$. Combining
-  Equations \ref{eq:positive_bound} and \ref{eq:negative_bound} gives
-  Equation \ref{eq:bound}.
-\end{proof}
-
 \begin{lemma} \label{lemma:bound1}
@@
 
+\begin{remark}
+  MDP values are characterised by both a state and a reward function/vector. In
+  this section, we think of the value function as $V : \mathcal{S} \to
+  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a state $s \in
+  \mathcal{S}$ and returns a function $V(s) : \mathbb{R}^{|\mathcal{S}|} \to
+  \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
+  \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$, $\V(s) \in
+  \mathbb{R}$. Given a reward vector, the function $V(s)$ computes the values of
+  all states and returns the value of state $s$.
+\end{remark}
+
+\begin{proposition}[Measurability] \label{thm:measurability}
+  MDP value functions $V(s) : \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ (for $s
+  \in \mathcal{S}$) are Lebesgue measurable.
+\end{proposition}
+\begin{proof}
+  For any reward vector $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$, the
+  collection of converged value functions $\{ \V(s) \mid s \in
+  \mathcal{S} \}$ satisfy
+  \begin{equation} \label{eq:linear_mdp}
+    \V(s) = \log \sum_{a \in \mathcal{A}}
+    \exp\left( r(s) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a,
+      s')\V(s') \right)
+  \end{equation}
+  for all $s \in \mathcal{S}$. Let $s_0 \in \mathcal{S}$ be an arbitrary state.
+  In order to prove that $V(s_0)$ is measurable, it is enough to show that for
+  any $\alpha \in \mathbb{R}$, the set
+  \[
+    \left\{ \mathbf{r} \in \mathbb{R}^{|\mathcal{S}|} \; \middle|
+    \begin{aligned}
+      \;&\V(s_0) \in (-\infty, \alpha); \\
+      \;&\V(s) \in \mathbb{R} \text{ for all } s \in \mathcal{S} \setminus \{ s_0 \}; \\
+      \;&\text{ Equation \ref{eq:linear_mdp} is satisfied by all } s \in
+      \mathcal{S}
+    \end{aligned}
+    \right\}
+  \]
+  is measurable. Since this set can be constructed in Zermelo-Fraenkel set
+  theory \emph{without} the axiom of choice, it is measurable
+  \cite{herrlich2006axiom}, which proves that $V(s)$ is a measurable function
+  for any $s \in \mathcal{S}$.
+\end{proof}
+
+\begin{proposition}[Boundedness] \label{thm:bound}
+  If the initial values of the MDP value function satisfy the following
+  bound, then the bound remains satisfied throughout value iteration:
+  \begin{equation} \label{eq:bound}
+    |\V(s)| \le \vbound.
+  \end{equation}
+\end{proposition}
+\begin{proof}
+  We begin by considering Equation \ref{eq:bound} without taking the absolute value of
+  $\V(s)$, i.e.,
+  \begin{equation} \label{eq:positive_bound}
+    \V(s) \le \vbound,
+  \end{equation}
+  and assuming that the initial values of $\{ \V(s) \mid s \in
+  \mathcal{S} \}$ already satisfy Equation \ref{eq:positive_bound}. Recall that for
+  each $s \in \mathcal{S}$, the value of $\V(s)$ is updated by applying Equation
+  \ref{eq:update_rule}. Note that both $\log$ and $\exp$ are increasing
+  functions, $\gamma > 0$, and the $\mathcal{T}$ function gives a probability (a
+  non-negative number). Thus
+  \begin{align*}
+    \V(s) &\le \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s')\frac{\rinf + \log|\mathcal{A}|}{1 - \gamma} \right) \\
+          &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma}\sum_{s' \in \mathcal{S}} \mathcal{T}(s, a, s') \right) \\
+          &= \log \sum_{a \in \mathcal{A}} \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right)
+  \end{align*}
+  by the definition of $\mathcal{T}$. Then
+  \begin{align*}
+    \V(s) &\le \log \left( |\mathcal{A}| \exp\left( r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
+          &= \log \left( \exp\left( \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \right) \right) \\
+          &= \log|\mathcal{A}| + r(s) + \frac{\gamma (\rinf + \log|\mathcal{A}|)}{1 - \gamma} \\
+          &= \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 - \gamma)(\log|\mathcal{A}| + r(s))}{1 - \gamma} \\
+          &\le \frac{\gamma (\rinf + \log|\mathcal{A}|) + (1 - \gamma)(\log|\mathcal{A}| + \rinf)}{1 - \gamma} \\
+          &= \vbound
+  \end{align*}
+  by the definition of $\rinf$.
+
+  The proof for
+  \begin{equation} \label{eq:negative_bound}
+    \V(s) \ge \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1}
+  \end{equation}
+  follows the same argument until we get to
+  \begin{align*}
+    \V(s) &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma - 1)(\log|\mathcal{A}| + r(s))}{\gamma - 1} \\
+          &\ge \frac{\gamma(\rinf + \log|\mathcal{A}|) + (\gamma - 1)(-\log|\mathcal{A}| -\rinf)}{\gamma - 1} \\
+          &= \frac{\rinf + \log|\mathcal{A}|}{\gamma - 1},
+  \end{align*}
+  where we use the fact that $r(s) \ge -\rinf - 2\log|\mathcal{A}|$. Combining
+  Equations \ref{eq:positive_bound} and \ref{eq:negative_bound} gives
+  Equation \ref{eq:bound}.
+\end{proof}
+
 \begin{lemma} \label{lemma:integral_of_r}
@@
     diagonal entries are non-negative.}, and $\Phi$ is the cumulative
-  distribution function of the standard normal distribution. Furthermore,
+  distribution function of the standard Gaussian. Furthermore,
   \[
@@
 \section{Evaluation} % showing deep understanding
+% Could measure how algorithms scale with the size of the map (number of states)
+% Could add plots that show how the ELBO changes with different parameter values
+% (the inverse parabolas)
+% And plot you get if you start with mu = 0 and move to negative infinity (hint:
+% it increases)
 
@@
 
+More variables/information. Fewer assumptions.
+
+We show how to avoid the deterministic training conditional assumption.
+
 \subsection{Future Work}
+
+\begin{itemize}
+\item Let the agent take uncertainty (variation estimates) into account when
+  choosing what to do next (i.e., choosing what to learn).
+\item MDP solver that works with distributions of rewards.
+\item Stress how solving MDPs is the performance bottleneck: num_samples MDPs
+  have to be solved for each iteration.
+  \begin{itemize}
+  \item Integration with MDP solvers that approximate the value function
+    sufficiently well for training.
+  \item Or perhaps there is a way to ``modify'' a solution to one MDP in order
+    to get a solution for a similar MDP (there was a FATA talk about that, but
+    for a different problem of course).
+  \item Perhaps value iteration converges faster if the values are initialised
+    to the optimal values of the previous iteration.
+  \end{itemize}
+\end{itemize}
 
@@
 \appendix
-\section{Proofs}
+\section{Proofs} \label{appendix:proofs}
 

### 2019-03-13 mpaper/paper.tex
More writing (and a new plot)

@@
 \usepackage{tikz}
-
-\usetikzlibrary{shapes,bayesnet,arrows.meta}
-\tikzstyle{arrow} = [-{Latex[length=2mm,width=2mm]}]
+\usepackage{layouts}
+
+\usetikzlibrary{shapes,bayesnet,arrows.meta,arrows}
+%\tikzstyle{myarrow} = [-{Latex[length=2mm,width=2mm]}]
 
@@
 \newtheorem{remark}[theorem]{Remark}
+\newtheorem{problem}[theorem]{Open Problem}
 
@@
 The original approach used maximum likelihood estimation
-\cite{DBLP:conf/nips/LevinePK11}, while the goal of this project is to use
-variational inference (VI) instead, which learns approximate posterior
-probability distributions instead of point estimates. This approach can prove
-useful in three major ways:
+\cite{DBLP:conf/nips/LevinePK11}, whereas we use variational inference (VI)
+instead, which learns approximate posterior probability distributions instead of
+point estimates. This approach can prove useful in three major ways:
 \begin{enumerate}
@@
 k)$ to say that $r$ is a GP with mean $0$ and covariance function
-$k$. Covariance functions take two state feature vectors as input and quantify
-how similar the two states are, in a sense that we would expect them to have
-similar rewards.
+$k$. \emph{Covariance functions} (also known as \emph{kernels}) take two state
+feature vectors as input and quantify how similar the two states are, in a sense
+that we would expect them to have similar rewards.
 
@@
 \end{equation}
-and $p(\mathcal{D} \mid \mathbf{r})$ is as in Equation \ref{pDr}. The matrices such as
+and $p(\mathcal{D} \mid \mathbf{r})$ is as in \eqref{pDr}. The matrices such as
 $\Kru$ are called \emph{covariance matrices} and are defined as
@@
 computed maximum likelihood (ML) estimates for $\bm\lambda$ and $\mathbf{u}$,
-and made an assumption that $\mathbf{r}$ in Equation \ref{eq:r} has zero variance
+and made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
 \cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
@@
 
-The project is primarily concerned with investigating how a VI formulation of
-the GP IRL model compares against the original ML approach. Most importantly, we
-aim to compare how the two algorithms converge both over time and as the number
-of demonstrations increases. It would also be interesting to see how close the
-approximation of the posterior distribution is to the real thing. Finally, it is
-reasonable to conjecture that VI can outperform point estimates when dealing
-with more uncertainty, e.g., when experts make mistakes. We can easily
-investigate this by adapting how the evaluation data is generated.
-
 \section{Background} \label{sec:background}
@@
 \section{Solution}
-% TODO: express all lambdas as exponentials (update the derivatives or sth)
 % TODO: give an overview of this section. Also, consider renaming it or
@@
 
-In order to properly investigate the difference between VI
-and ML estimation, we keep other parts of the model the same.
-Namely, we set the covariance function to a version of the automatic relevance
-detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
+We keep the covariance function the same as in the work by Levine et al.
+\cite{DBLP:conf/nips/LevinePK11}, which is a version of the automatic relevance
+detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}:
 \begin{align*}
   k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
-                                                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
-\end{align*}
-where $\lambda_0$ is the overall `scale' factor for how similar or distant the
+                                                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right).
+\end{align*}
+Here, $\lambda_0$ is the overall `scale' factor for how similar or distant the
 states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
@@
 
-% TODO: write about lambda having to be variational
-% TODO: mention the open problem of finding the expected value of the inverse of
-% a matrix, where each element is a non-trivial function (w.r.t. a parameter of
-% the function)
-%TODO: introductory sentence for the following assumption:
+Ideally, we would like to model $\bm\lambda$ with an approximating distribution.
+However, due to how $p(\mathbf{u})$ has $\Kuu^{-1}$ in its expression, and the
+ELBO is defined as an expectation, we are unable to show that the ELBO is
+well-defined. More generally, we pose the following problem, which is open to
+the best of our knowledge:
+\begin{problem}
+  Let $\mathbf{A}$ be a $n \times n$ matrix of coefficients, $X$ be a random
+  variable, and $\mathbf{M}$ be an $n \times n$ matrix such that $M_{i,j} = f(X,
+  A_{i,j})$, where $f$ is an arbitrary function. Under what circumstances does
+  $\mathbb{E}[\mathbf{M}^{-1}]$ exist?
+\end{problem}
+While there are some obvious examples of when the required expected value
+exists (e.g., $f(X, A_{i,j}) = A_{i,j}X$ for an invertible $\mathbf{A}$ and many
+distributions of $X$), it would be particularly interesting to know whether the
+answer is `always'. A proof of such a result would allow us to model
+$\bm\lambda$ instead of treating it as a variational parameter, and would thus
+guard against overfitting. For now, $\bm\lambda$ will have to be treated as a
+variational parameter.
+
+It remains to decide on the model for $\mathbf{u}$ and $\mathbf{r}$. As is
+commonly done when applying VI to GPs, we set
 \begin{equation} \label{eq:approximation}
-  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
+  \approximation = q(\mathbf{u}) q(\mathbf{r} \mid \mathbf{u}),
 \end{equation}
-where
-\begin{equation}
-  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma),
-\end{equation}
-where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
-positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
-\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.
+where $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$ and
+$q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma)$
+\cite{DBLP:conf/nips/ChengB17}.
 
@@
 and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
-parameters of the kernel, $\bm\lambda$. The two remaining dependencies are quite
-clear, given that the approximating distribution for $\mathbf{u}$ is
-$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$. % TODO: concluding sentence
+parameters of the kernel, $\bm\lambda$. The two remaining dependencies stem from
+the fact that the approximating distribution for $\mathbf{u}$ is
+$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$.
+
+As we want to restrict some parameters (namely, $\bm\lambda$ and the diagonal of
+$\mathbf{B}$) to positive values, we express them as exponentials and later
+adjust their derivatives accordingly. Specifically, we can set $\lambda_i =
+e^{\lambda'_i}$ and optimise $\lambda'_i$ using the chain rule:
+\[
+  \frac{\partial \mathcal{L}}{\partial \lambda'_i} = e^{\lambda'_i}
+  \frac{\partial \mathcal{L}}{\partial \lambda_i}.
+\]
+This way, we restrict $\lambda_i$ to positive values while allowing $\lambda'_i$
+to range over $\mathbb{R}$.
+
+Finally, the parameters are initialised as follows:
+\begin{alignat*}{3}
+  \mu_i &\sim \mathcal{U}(0, 1) \quad &&\text{for } i = 1, \dots, m, \\
+  \lambda_0 &\sim \chi^2_5, && \\
+  \lambda_i &\sim \chi^2_1 \quad &&\text{for } i = 1, \dots, d, \\
+  \diag(\mathbf{B}) &\sim \chi^2_4, && \\
+  \text{the rest of } \mathbf{B} &\sim \mathcal{N}(0, 1). &&
+\end{alignat*}
+The initialisation of $\bm\mu$ mirrors the initialisation of $\mathbf{r}$ in
+previous work by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. While they have
+constant initial values for $\bm\lambda$, we sample from $\chi^2$ distributions
+centred around those values ($5$ for $\lambda_0$ and $1$ for any other
+$\lambda_i$). The distributions for initial values of $\mathbf{B}$ are simply
+set to provide a reasonable spread of positive values for the diagonal, and both
+positive and negative values for all other entries in the matrix.
 
@@
 
-In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
+In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
 write
@@
 \]
-By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
+By substituting in \eqref{eq:full} and \eqref{eq:approximation}, we get
 \begin{align*}
@@
 
-    \draw [arrow] (perturbation) edge (lemma3);
-    \draw [arrow] (lebesgue) edge (theorem);
-    \draw [arrow] (derivatives) edge (lemma1);
-    \draw [arrow] (derivatives) edge (lemma2);
-    \draw [arrow] (derivatives) edge (lemma3);
-    \draw [arrow] (measurability) edge (theorem);
-    \draw [arrow] (boundedness) edge (theorem);
-    \draw [arrow] (lemma1) edge (theorem);
-    \draw [arrow] (lemma2) edge (theorem);
-    \draw [arrow] (lemma3) edge (theorem);
-    \draw [arrow] (lemma4) edge (theorem);
+    \draw [->] (perturbation) edge (lemma3);
+    \draw [->] (lebesgue) edge (theorem);
+    \draw [->] (derivatives) edge (lemma1);
+    \draw [->] (derivatives) edge (lemma2);
+    \draw [->] (derivatives) edge (lemma3);
+    \draw [->] (measurability) edge (theorem);
+    \draw [->] (boundedness) edge (theorem);
+    \draw [->] (lemma1) edge (theorem);
+    \draw [->] (lemma2) edge (theorem);
+    \draw [->] (lemma3) edge (theorem);
+    \draw [->] (lemma4) edge (theorem);
   \end{tikzpicture}
@@
       \;&\V(s) \in \mathbb{R} \text{ for all } s \in \mathcal{S} \setminus \{ s_0 \}; \\
-      \;&\text{ Equation \ref{eq:linear_mdp} is satisfied by all } s \in
+      \;&\text{\eqref{eq:linear_mdp} is satisfied by all } s \in
       \mathcal{S}
@@
 \begin{proof}
-  We begin by considering Equation \ref{eq:bound} without taking the absolute value of
+  We begin by considering \eqref{eq:bound} without taking the absolute value of
   $\V(s)$, i.e.,
@@
   \end{equation}
-  and assuming that the initial values of $\{ \V(s) \mid s \in
-  \mathcal{S} \}$ already satisfy Equation \ref{eq:positive_bound}. Recall that for
-  each $s \in \mathcal{S}$, the value of $\V(s)$ is updated by applying Equation
-  \ref{eq:update_rule}. Note that both $\log$ and $\exp$ are increasing
+  and assuming that the initial values of $\{ \V(s) \mid s \in \mathcal{S} \}$
+  already satisfy \eqref{eq:positive_bound}. Recall that for each $s \in
+  \mathcal{S}$, the value of $\V(s)$ is updated by applying
+  \eqref{eq:update_rule}. Note that both $\log$ and $\exp$ are increasing
   functions, $\gamma > 0$, and the $\mathcal{T}$ function gives a probability (a
@@
   where we use the fact that $r(s) \ge -\rinf - 2\log|\mathcal{A}|$. Combining
-  Equations \ref{eq:positive_bound} and \ref{eq:negative_bound} gives
-  Equation \ref{eq:bound}.
+  \eqref{eq:positive_bound} and \eqref{eq:negative_bound} gives
+  \eqref{eq:bound}.
 \end{proof}
@@
   \]
-  Equation \ref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
+  \eqref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
   that both $f$ and $f_n$ are measurable and find a non-negative integrable
@@
   \mathbb{R}_1[\mathbf{u}]$, a similar argument can be applied to the rest of
-  Equation \ref{eq:remaining_integral} as well.
+  \eqref{eq:remaining_integral} as well.
 \end{proof}
@@
 % Could measure how algorithms scale with the size of the map (number of states)
-% Could add plots that show how the ELBO changes with different parameter values
-% (the inverse parabolas)
-% And plot you get if you start with mu = 0 and move to negative infinity (hint:
-% it increases)
+% Mention all the constants I've used.
+% How well does the variance prediction work?
+
+%\begin{figure} % this is more confusing than helpful
+%  \centering
+%  \begin{tikzpicture}
+%    \node [ellipse,draw] (s1) at (-1, -1) {$s_1$};
+%    \node [ellipse,draw] (s2) at (0, 0) {$s_2$};
+%    \node [ellipse,draw] (s3) at (1, -1) {$s_3$};
+%    \draw [->,-to,bend left] (s1) edge (s2);
+%    \draw [->,-latex,dashed] (s1) edge (s3);
+%    \draw [->,-to] (s2) edge (s1);
+%    \draw [->,-latex,bend left,dashed] (s2) edge (s3);
+%    \draw [->,-to,bend left,dashed] (s3) edge (s1);
+%    \draw [->,-latex] (s3) edge (s2);
+%  \end{tikzpicture}
+%\end{figure}
+
+%\printinunitsof{cm}\prntlen{\textwidth}
+%\printinunitsof{cm}\prntlen{\textheight}
+%\printinunitsof{cm}\prntlen{\columnwidth}
+
+In order to fully understand the model's behaviour, we focus on a three-state
+MDP where the agent can deterministically move from any state to any other
+state. More formally, we set $\mathcal{S} = \{ s_1, s_2, s_3 \}$, $\mathcal{A} =
+\{ a_1, a_2 \}$,
+\begin{align*}
+  &\mathcal{T}(s_1, a_1, s_2) = 1, \quad \mathcal{T}(s_1, a_2, s_3) = 1, \\
+  &\mathcal{T}(s_2, a_1, s_1) = 1, \quad \mathcal{T}(s_1, a_2, s_3) = 1, \\
+  &\mathcal{T}(s_3, a_1, s_1) = 1, \quad \mathcal{T}(s_1, a_2, s_2) = 1,
+\end{align*}
+all other values of $\mathcal{T}$ to zero, and vary the value of $\gamma$. We
+also set the inducing points to be equal to the three states in $\mathcal{S}$,
+add a single feature $f : \mathcal{S} \to \mathbb{R}$ such that
+\[
+  f(s_1) = 1, \quad f(s_2) = 2, \quad f(s_3) = 3,
+\]
+and create two demonstrations: $\zeta_1 = \{ (s_1, a_1) \}$ and $\zeta_2 = \{
+(s_3, a_2) \}$. In other words, the demonstrations move from $s_1$ and $s_3$ to
+$s_2$. Therefore, we would expect the reward (and value) of $s_2$ to be the
+highest.
 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{elbo_over_gamma}
+  \caption{Changes in $\mathcal{L}$ as a function of $\bm\mu$, where $\mu_1 =
+    \mu_2 = \mu_3$, and $\gamma = 0, 0.1, \dots, 0.9$. Each of the ten curves
+    plots a different value of $\gamma$, ranging from $0$ in dark blue to $0.9$
+    in dark red.}
+\end{figure}
+
 \section{Conclusions}
@@
 
+An interesting extension to our work would be to consider IRL in the context
+of a reinforcement learning (RL) agent. Suppose we have an agent whose purpose
+is to learn optimal behaviour from observing other agents using IRL. It could
+then take reward variance estimates into account when choosing what states to
+visit next. It would have to handle the balance between exploration and
+exploitation similarly to many RL agents, but the information about rewards
+would come from observing (presumably near-optimal) behaviour exhibited by other
+agents rather than directly from the environment.
+
+%Our work also poses a new challenge to the usual methods of determining an
+%optimal strategy in a given MDP:
+%\begin{problem}
+%  Let $\mathcal{M} = \{ \mathcal{S}, \mathcal{A}, \mathcal{T}, \gamma,
+%  r \}$ be an MDP where the reward $r : \mathcal{S} \to
+%  \textsf{Prob}(\mathbb{R})$ is expressed as a function from the set of states
+%  to the set of probability distributions with support over $\mathbb{R}$. Find
+%  the optimal policy $\pi^*$ in $\mathcal{M}$, i.e., % TODO: argmax expression
+%\end{problem}
+%Of course, in many situations,... % TOOD: gaussians, pdf's, something simpler
+
 \begin{itemize}
-\item Let the agent take uncertainty (variation estimates) into account when
-  choosing what to do next (i.e., choosing what to learn).
 \item MDP solver that works with distributions of rewards.
-\item Stress how solving MDPs is the performance bottleneck: num_samples MDPs
+\item Stress how solving MDPs is the performance bottleneck: numSamples MDPs
   have to be solved for each iteration.
@@
     \end{equation}
-    Substituting results from Equations \ref{eq:partial_derivatives1} and
-    \ref{eq:partial_derivatives2} back into Equation \ref{eq:2_before} gives:
+    Substituting results from \eqref{eq:partial_derivatives1} and
+    \eqref{eq:partial_derivatives2} back into \eqref{eq:2_before} gives:
     \begin{alignat*}{3}
@@
 
-\section{Derivatives}
+\section{Derivatives of the ELBO}
 

### 2019-03-13 code/VIGPIRL/full_gradient.m

@@ -69,6 +69,9 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
     diag(not_estimated_B), not_estimated_mu, get_lower_triangle(not_estimated_B));
   changes = not_estimated - 0.5 * estimated_grad;
 
+  disp(not_estimated_elbo);
+  disp(estimated_grad(1));
+
   % TESTING
   %fprintf('Not estimated: %f, estimated: %f, total: %f\n', not_estimated(2), estimated_grad(2), changes(2));
   p_derivative = -0.5 * estimated_grad(2) + counts' * (matrices.Kru_grad(:, :, 1)' -...

### 2019-03-13 code/VIGPIRL/vigpirlinit.m

@@ -12,8 +12,8 @@ gp.X_u = feature_data.splittable;
 
 % Initialize hyperparameters.
 gp.noise_var = algorithm_params.noise_init;
-gp.lambda0 = random('Chisquare', algorithm_params.rbf_init);
-%gp.lambda0 = 1; % TEMP
+%gp.lambda0 = random('Chisquare', algorithm_params.rbf_init);
+gp.lambda0 = 1; % TEMP
 gp.lambda = random('Chisquare', algorithm_params.ard_init,...
     [1, size(feature_data.splittable, 2)]);
 %gp.lambda = ones(1, size(feature_data.splittable, 2)); % TEMP

### 2019-03-13 code/singletest.m

@@ -19,20 +19,20 @@ addpaths;
 
 % VIGPIRL
 %test_result = runtest('vigpirl',struct(),'linearmdp',...
-%    'gridworld',struct('n',8,'determinism',0.7,'seed',1,'continuous',0),...
-%    struct('training_sample_lengths',8,'training_samples',32,'verbosity',2));
+%    'gridworld',struct('n',8,'determinism',1,'seed',1,'continuous',0),...
+%    struct('training_sample_lengths',4,'training_samples',6,'verbosity',2));
 
 % Visualize solution.
 %printresult(test_result);
 %visualize(test_result);
 
 % Two states -- works well
-%mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0.9, 'states', 2, 'actions', 1);
+%mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0, 'states', 2, 'actions', 1);
 %feature_data = struct('splittable', [1; 2]);
 %example_samples = {[1, 1]};
 
 % Three states -- doesn't work
-mdp_data = struct('discount', 0.9, 'states', 3, 'actions', 2);
+mdp_data = struct('discount', 0.7, 'states', 3, 'actions', 2);
 mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
 mdp_data.sa_p(1:3, 1:2, 1) = 1;
 feature_data = struct('splittable', [1; 2; 3]);

### 2019-03-13 mpaper/paper.tex
Committing before an attempt at fixing problematic numerical errors

@@
 \section{Evaluation} % showing deep understanding
-% Could measure how algorithms scale with the size of the map (number of states)
 % Mention all the constants I've used.
-% How well does the variance prediction work?
-
-%\begin{figure} % this is more confusing than helpful
-%  \centering
-%  \begin{tikzpicture}
-%    \node [ellipse,draw] (s1) at (-1, -1) {$s_1$};
-%    \node [ellipse,draw] (s2) at (0, 0) {$s_2$};
-%    \node [ellipse,draw] (s3) at (1, -1) {$s_3$};
-%    \draw [->,-to,bend left] (s1) edge (s2);
-%    \draw [->,-latex,dashed] (s1) edge (s3);
-%    \draw [->,-to] (s2) edge (s1);
-%    \draw [->,-latex,bend left,dashed] (s2) edge (s3);
-%    \draw [->,-to,bend left,dashed] (s3) edge (s1);
-%    \draw [->,-latex] (s3) edge (s2);
-%  \end{tikzpicture}
-%\end{figure}
-
-%\printinunitsof{cm}\prntlen{\textwidth}
-%\printinunitsof{cm}\prntlen{\textheight}
-%\printinunitsof{cm}\prntlen{\columnwidth}
 
@@
 
+ % TODO: need to redo this to get better labels, more precision, wider range of values
 \begin{figure*}
@@
 
+% How does policy depend on rewards? Choose state and action, generate an array
+% of contour plots showing the probability of taking that action in an optimal policy
+Observation from a plot not worth including here. The policy behaves as
+expected: if we fix $r(s_1)$, $\pi(s_1, a_1)$ increases with higher $r(s_2)$ as
+well as lower $r(s_3)$ values.
+
+% How well does the variance prediction work?
+
 \section{Conclusions}
@@
 
-%Our work also poses a new challenge to the usual methods of determining an
-%optimal strategy in a given MDP:
-%\begin{problem}
-%  Let $\mathcal{M} = \{ \mathcal{S}, \mathcal{A}, \mathcal{T}, \gamma,
-%  r \}$ be an MDP where the reward $r : \mathcal{S} \to
-%  \textsf{Prob}(\mathbb{R})$ is expressed as a function from the set of states
-%  to the set of probability distributions with support over $\mathbb{R}$. Find
-%  the optimal policy $\pi^*$ in $\mathcal{M}$, i.e., % TODO: argmax expression
-%\end{problem}
-%Of course, in many situations,... % TOOD: gaussians, pdf's, something simpler
-
+It is also worth noting the approach presented in this paper requires solving
+$S$ MDPs for every iteration of optimising the parameters (where $S$ is the
+number of samples drawn from $\approximation$). There are at least two ways to
+reduce or eliminate this performance bottleneck:
 \begin{itemize}
-\item MDP solver that works with distributions of rewards.
-\item Stress how solving MDPs is the performance bottleneck: numSamples MDPs
-  have to be solved for each iteration.
-  \begin{itemize}
-  \item Integration with MDP solvers that approximate the value function
-    sufficiently well for training.
-  \item Or perhaps there is a way to ``modify'' a solution to one MDP in order
-    to get a solution for a similar MDP (there was a FATA talk about that, but
-    for a different problem of course).
-  \item Perhaps value iteration converges faster if the values are initialised
-    to the optimal values of the previous iteration.
-  \end{itemize}
+\item The MDP value function could be approximated, allowing for some minor
+  mistakes in the resulting policy.
+\item Perhaps there is a good way to use information about previously computed
+  value functions for similar rewards to hasten the current computation. One
+  simple way to do this would be by initialising the current values to the
+  optimal values of the previous MDP value function computation.
 \end{itemize}

### 2019-03-13 code/VIGPIRL/full_gradient.m

@@ -69,9 +69,6 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
     diag(not_estimated_B), not_estimated_mu, get_lower_triangle(not_estimated_B));
   changes = not_estimated - 0.5 * estimated_grad;
 
-  disp(not_estimated_elbo);
-  disp(estimated_grad(1));
-
   % TESTING
   %fprintf('Not estimated: %f, estimated: %f, total: %f\n', not_estimated(2), estimated_grad(2), changes(2));
   p_derivative = -0.5 * estimated_grad(2) + counts' * (matrices.Kru_grad(:, :, 1)' -...
@@ -87,6 +84,15 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
   elbo = changes(1);
   grad = transform_gradient(changes(2:end), vigpirlpackparam(gp), size(gp.lambda), size(gp.mu));
+
+  % If our rewards go too far towards negative infinity, and probability of a
+  % state-action pair is over 1, freeze the reward of current state
+  elbo_data_part = counts' * S * gp.mu - 0.5 * estimated_grad(1);
+  fprintf('Should be negative: %f\n', elbo_data_part);
+  if (elbo_data_part > 0)
+    elbo = elbo - elbo_data_part;
+    grad(d+m+2+)
+  end
 
   %fprintf('Not estimated grad:');
   %disp(not_estimated(2));

### 2019-03-13 code/VIGPIRL/vigpirldefaultparams.m

@@ -13,8 +13,8 @@ default_params = struct(...
     'samples_count',100,...
     'learning_rate',0.1,...
     'B_learning_rate',0.001,...
-    'num_iterations',20,...
-    'required_precision', 0.001);
+    'num_iterations',50,...
+    'required_precision', 0.01);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2019-03-13 code/VIGPIRL/vigpirlrun.m

@@ -68,10 +68,45 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
 
   parameter_vector = vigpirlpackparam(gp);
 
+  % ELBO as a function of mu over 10 plots
+  mu1_log = [];
+  mu2_log = [];
+  mu3_log = [];
+  multiplier = 10;
+  for mu1 = 1:10
+    real_mu1 = multiplier * (mu1 - 5);
+    parameter_vector(d+m+2) = real_mu1;
+    for mu2 = 1:10
+      real_mu2 = multiplier * (mu2 - 5);
+      parameter_vector(d+m+3) = real_mu2;
+      for mu3 = 1:10
+        real_mu3 = multiplier * (mu3 - 5)
+        parameter_vector(d+m+4) = real_mu3;
+        wrapper(parameter_vector);
+        mu1_log = [mu1_log real_mu1];
+        mu2_log = [mu2_log real_mu2];
+        mu3_log = [mu3_log real_mu3];
+      end
+    end
+  end
+  for mu1 = 1:10
+    indices = 100*(mu1-1)+1:100*mu1;
+    other_mu = multiplier * ((1:10) - 5);
+    subplot(5, 2, mu1);
+    contourf(other_mu, other_mu, reshape(elbo_list(indices), [10, 10]), 30);
+    xlabel('$\mu_2$', 'Interpreter', 'latex');
+    ylabel('$\mu_3$', 'Interpreter', 'latex');
+    t = ['$\mu_1 = ', num2str(multiplier * (mu1 - 5)), '$'];
+    title(t, 'Interpreter', 'latex');
+  end
+  return;
+
+  % Policy as a function of rewards over 10 plots
   %mu1_log = [];
   %mu2_log = [];
   %mu3_log = [];
   %multiplier = 10;
+  %policy_log = [];
   %for mu1 = 1:10
   %  real_mu1 = multiplier * (mu1 - 5);
   %  parameter_vector(d+m+2) = real_mu1;
@@ -81,75 +116,55 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   %    for mu3 = 1:10
   %      real_mu3 = multiplier * (mu3 - 5)
   %      parameter_vector(d+m+4) = real_mu3;
-  %      wrapper(parameter_vector);
+  %      solution = feval([mdp_model 'solve'], mdp_data, [mu1; mu2; mu3]);
[…]

### 2019-03-13 code/singletest.m

@@ -32,7 +32,7 @@ addpaths;
 %example_samples = {[1, 1]};
 
 % Three states -- doesn't work
-mdp_data = struct('discount', 0.7, 'states', 3, 'actions', 2);
+mdp_data = struct('discount', 0.5, 'states', 3, 'actions', 2);
 mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
 mdp_data.sa_p(1:3, 1:2, 1) = 1;
 feature_data = struct('splittable', [1; 2; 3]);

### 2019-03-15 mpaper/paper.tex
Visualising the incorrect derivative

@@
 
-  %We begin by producing constant upper and
-  %lower bounds for the elements of
-  %\[
-  %  \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i =
-  %    c(\mathbf{r}, \mathbf{u})},
-  %\]
-  %where $\mathbf{K}$ is any covariance matrix.
+  Given any covariance matrix $\mathbf{K}$ and $i = 0, \dots, d$, we can easily
+  establish constant upper and lower bounds on $\mathbf{K}|_{\lambda_i =
+    c(\mathbf{r}, \mathbf{u})}$ by the boundedness of $c$ and continuity of the
+  covariance function. Similar reasoning combined with the expressions for
+  derivatives of covariance matrices in Lemma \ref{lemma:derivatives} gives
+  constant upper and lower bounds on the elements of $\left.
+    \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
+    c(\mathbf{r}, \mathbf{u})}$ as well. We can also use Lemma ?? to show
+  constant upper and lower bounds on $\mathbf{K}^{-1}|_{\lambda_i =
+    c(\mathbf{r}, \mathbf{u})}$. This is enough to prove constant upper and
+  lower bounds on $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$, which means that
+  $(\mathbf{r} -
+  \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} -
+  \mathbf{Su})$ has upper and lower bounds in $\mathbb{R}_2[\mathbf{u}]$.
+  % TODO: write the lemma
 
@@
 
-  % TODO: Lemma 1.1 is wrong about the % derivative. Fix (my handwritten notes
+  % TODO: Lemma 1.1 is wrong about the derivative. Fix (my handwritten notes
   % have the correct idea)
@@
 
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{ELBO_and_gradient_over_B}
+  \caption{ELBO in blue, derivative in red. The first row is for diagonal values
+  of B, the second row is for non-diagonal values.}
+\end{figure*}
+
 % How does policy depend on rewards? Choose state and action, generate an array
@@
     \]
-  \item
-    \begin{align}
-      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\exp \left( -\frac{Q}{2} \right)\right] \nonumber \\
-                                                         &= \dB\left[\frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\right]\exp \left( -\frac{Q}{2} \right) \nonumber \\
-                                                         &+ \frac{1}{(2\pi)^{m/2}|\bm\Sigma|^{1/2}}\dB\left[\exp\left( -\frac{Q}{2} \right)\right] \nonumber \\
-                                                         &= -\frac{1}{2} q(\mathbf{u}) \left( |\bm\Sigma|^{-3/2} \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} + \frac{\partial Q}{\partial \mathbf{B}} \right).
-\label{eq:2_before}
-    \end{align}
-    For derivatives with respect to $\bm\Sigma$, we can refer to Petersen and
-    Pedersen \cite{petersen2008matrix}:
-    \begin{equation} \label{eq:partial_derivatives1}
-      \frac{\partial |\bm\Sigma|}{\partial \bm\Sigma} = |\bm\Sigma|\bm\Sigma^{-\intercal}, \quad \frac{\partial Q}{\partial \mathbf{B}} = -\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal},
-    \end{equation}
-    while we can use an online tool by Laue et
+  \item An online tool by Laue et
     al.\footnote{\url{http://www.matrixcalculus.org/}}
-    \cite{DBLP:conf/nips/LaueMG18} for the remaining ones:
-    \begin{equation} \label{eq:partial_derivatives2}
-      \begin{gathered}
-        \frac{\partial |\bm\Sigma|}{\partial \mathbf{B}} =
-        (\adj(\mathbf{T}) + \adj(\bm\Sigma))\mathbf{B}, \\
-        \dB[(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} -
-        \bm\mu)] = -(\mathbf{TUT} +
-        \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1})\mathbf{B}.
-      \end{gathered}
-    \end{equation}
-    Substituting results from \eqref{eq:partial_derivatives1} and
-    \eqref{eq:partial_derivatives2} back into \eqref{eq:2_before} gives:
-    \begin{alignat*}{3}
-      \frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} &=
-      \frac{1}{2}q(\mathbf{u})(&&\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
-      - \bm\Sigma^{-\intercal}), \\
-      \frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} &=
-      \frac{1}{2}q(\mathbf{u}) \{&&|\bm\Sigma|^{-3/2}(\adj(\mathbf{T}) +
-      \adj(\bm\Sigma)) \\
-      & &&+ \mathbf{TUT} + \bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} \} \mathbf{B}.
-    \end{alignat*}
+    \cite{DBLP:conf/nips/LaueMG18} can be used to find both derivatives.
   \item

### 2019-03-15 code/VIGPIRL/full_gradient.m

@@ -22,7 +22,7 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
       next_states = extract(mdp_data.sa_s, state, action);
       next_probabilities = extract(mdp_data.sa_p, state, action);
       s = next_probabilities' * solution.v(next_states, 1);
-      elbo_part = solution.v(state, 1) + mdp_data.discount * s;
+      elbo_part = solution.v(state, 1) - mdp_data.discount * s;
     end
 
     U = (u - gp.mu) * (u - gp.mu)';
@@ -43,6 +43,8 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   end
 
   Sigma = gp.B * gp.B';
+  disp(gp.B);
+  disp(Sigma);
   Sigma_inv = inv(Sigma);
   Kuu_inv = inv(matrices.Kuu);
   S = matrices.Kru' * Kuu_inv;
@@ -88,10 +90,9 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   % If our rewards go too far towards negative infinity, and probability of a
   % state-action pair is over 1, freeze the reward of current state
   elbo_data_part = counts' * S * gp.mu - 0.5 * estimated_grad(1);
-  fprintf('Should be negative: %f\n', elbo_data_part);
   if (elbo_data_part > 0)
-    elbo = elbo - elbo_data_part;
-    grad(d+m+2+)
+    fprintf('Should be negative: %f\n', elbo_data_part);
+  %  elbo = elbo - elbo_data_part;
   end
 
   %fprintf('Not estimated grad:');

### 2019-03-15 code/VIGPIRL/vigpirldefaultparams.m

@@ -10,11 +10,11 @@ default_params = struct(...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
-    'samples_count',100,...
+    'samples_count',1000,...
     'learning_rate',0.1,...
     'B_learning_rate',0.001,...
-    'num_iterations',50,...
-    'required_precision', 0.01);
+    'num_iterations',400,...
+    'required_precision', 0.001);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2019-03-15 code/VIGPIRL/vigpirlrun.m

@@ -37,7 +37,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   % wrapper: vector of parameters -> scalar value * gradient vector
   function [elbo, grad] = wrapper(parameter_vector)
     gp = vigpirlunpackparam(gp, parameter_vector);
-    %gp.lambda0 = exp(parameter_vector);
 
     %disp(gp);
     %disp(gp.B * gp.B');
@@ -48,12 +47,11 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, zz, matrices);
 
     % Disable gradients that don't work
-    grad(1) = 0; % lambda0
-    grad(2:d+1) = 0; % lambda (except first)
+    %grad(1) = 0; % lambda0
+    %grad(2:d+1) = 0; % lambda (except first)
     %grad(d+m+2:d+2*m+1) = 0; % mu
-    grad(d+2:d+m+1) = 0; % B diagonal
-    grad(d+2*m+2:end) = 0; % rest of B
-    %disp(grad(1));
+    %grad(d+2:d+m+1) = 0; % B diagonal
+    %grad(d+2*m+2:end) = 0; % rest of B
 
     elbo_list = horzcat(elbo_list, elbo);
     hyperparameter_history = horzcat(hyperparameter_history, parameter_vector);
@@ -62,44 +60,42 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
 
     % We want to maximise the ELBO, while fmincon wants to minimize...
     grad = -grad;
-    %grad = -grad(1);
     elbo = -elbo;
   end
 
   parameter_vector = vigpirlpackparam(gp);
 
   % ELBO as a function of mu over 10 plots
-  mu1_log = [];
-  mu2_log = [];
-  mu3_log = [];
-  multiplier = 10;
-  for mu1 = 1:10
-    real_mu1 = multiplier * (mu1 - 5);
-    parameter_vector(d+m+2) = real_mu1;
-    for mu2 = 1:10
-      real_mu2 = multiplier * (mu2 - 5);
-      parameter_vector(d+m+3) = real_mu2;
-      for mu3 = 1:10
-        real_mu3 = multiplier * (mu3 - 5)
-        parameter_vector(d+m+4) = real_mu3;
[…]

### 2019-03-15 code/singletest.m

@@ -26,12 +26,12 @@ addpaths;
 %printresult(test_result);
 %visualize(test_result);
 
-% Two states -- works well
+% Two states
 %mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0, 'states', 2, 'actions', 1);
 %feature_data = struct('splittable', [1; 2]);
 %example_samples = {[1, 1]};
 
-% Three states -- doesn't work
+% Three states
 mdp_data = struct('discount', 0.5, 'states', 3, 'actions', 2);
 mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
 mdp_data.sa_p(1:3, 1:2, 1) = 1;

### 2019-03-15 mpaper/paper.tex
A bit of progress on the proof

@@
 \DeclareMathOperator{\adj}{adj}
+%\DeclareMathOperator{\dett}{det}
 \DeclareMathOperator{\tr}{tr}
@@
 
+% TODO: add this lemma to the graph
+\begin{lemma} \label{lemma:matrix_inverse}
+  TODO: constant bounds on the inverse
+\end{lemma}
+
 \begin{lemma} \label{lemma:bound1}
-  Let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
-  \mathbb{R}$ be an arbitrary bounded function. Then, for $i = 0,
-  \dots, d$,
-  \[
-    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} \right|_{\lambda_i
-      = c(\mathbf{r}, \mathbf{u})}
-  \]
-  has upper and lower bounds of the form $q(\mathbf{r} \mid \mathbf{u})d(\mathbf{u})$, where
-  $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
+  Let $i \in \{ 0, \dots, d \}$ and $\epsilon > 0$ be arbitrary. Furthermore,
+  let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i -
+  \epsilon, \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a
+  codomain arbitrarily close to $\lambda_i$. Then
+  \[
+    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
+    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
+  \]
+  has upper and lower bounds of the form $q(\mathbf{r} \mid
+  \mathbf{u})d(\mathbf{u})$, where $d(\mathbf{u}) \in \mathbb{R}_2[\mathbf{u}]$.
 \end{lemma}
@@
 
-  Given any covariance matrix $\mathbf{K}$ and $i = 0, \dots, d$, we can easily
-  establish constant upper and lower bounds on $\mathbf{K}|_{\lambda_i =
-    c(\mathbf{r}, \mathbf{u})}$ by the boundedness of $c$ and continuity of the
-  covariance function. Similar reasoning combined with the expressions for
-  derivatives of covariance matrices in Lemma \ref{lemma:derivatives} gives
-  constant upper and lower bounds on the elements of $\left.
-    \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
-    c(\mathbf{r}, \mathbf{u})}$ as well. We can also use Lemma ?? to show
-  constant upper and lower bounds on $\mathbf{K}^{-1}|_{\lambda_i =
-    c(\mathbf{r}, \mathbf{u})}$. This is enough to prove constant upper and
-  lower bounds on $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$, which means that
-  $(\mathbf{r} -
+  Given any covariance matrix $\mathbf{K}$, we can easily establish constant
+  upper and lower bounds on $\mathbf{K}|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$ by the boundedness of $c$ and continuity of the covariance
+  function. Similar reasoning combined with the expressions for derivatives of
+  covariance matrices in Lemma \ref{lemma:derivatives} gives constant upper and
+  lower bounds on the elements of
+  \[
+    \left. \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
+      c(\mathbf{r}, \mathbf{u})}
+  \]
+  as well. We can also use Lemma \ref{lemma:matrix_inverse} to show constant
+  upper and lower bounds on $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$. This is enough to prove constant upper and lower bounds on
+  $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$ (all with $\lambda_i$ replaced
+  with $c(\mathbf{r}, \mathbf{u})$), which means that $(\mathbf{r} -
   \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} -
-  \mathbf{Su})$ has upper and lower bounds in $\mathbb{R}_2[\mathbf{u}]$.
-  % TODO: write the lemma
+  \mathbf{Su})|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ has upper and lower
+  bounds in $\mathbb{R}_2[\mathbf{u}]$.
+
+  Since $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is bounded, so is
+  $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Assuming that
+  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists, we
+  can make $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ close enough to
+  $\bm\Gamma$ to ensure that
+  \[
+    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0.
+  \]
+  Moreover, boundedness of $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
+  gives boundedness of $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$.
+
+  Combined with the fact that $ab > 0$ (i.e., both endpoints together are either
+  positive or negative), this gives boundedness of $\det(\bm\Gamma)|_{\lambda_i
+    = c(\mathbf{r}, \mathbf{u})}^{-1}$.
+
+  TODO: Show that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is
+  bounded.
+
+  Since we already know that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$ and $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
+  are bounded, so is $\adj(\bm\Gamma) =
+  \det(\bm\Gamma)\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Thus,
+  we have constant bounds on
+  $|\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma))|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$, which means that
+  \[
+    \left. \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i}
+    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
+  \]
+  has the required bounds.
 

### 2019-03-16 mpaper/paper.tex
One more step towards completing the proof

@@
 \begin{abstract}
+  We prove the validity of performing variational inference on the Gaussian
+  process-based inverse reinforcement learning model that preserves and
+  approximates the full joint probability distribution of rewards across states
+  of the Markov decision process.
 \end{abstract}
@@
 \begin{lemma}[Perturbation Lemma
-  \cite{layton2014numerical}] \label{prop:condition_number}
+  \cite{layton2014numerical}] \label{lemma:perturbation}
   Let $\lVert \cdot \rVert$ be any matrix norm, and let $\mathbf{A}$ and
@@
 
-% TODO: add this lemma to the graph
-\begin{lemma} \label{lemma:matrix_inverse}
-  TODO: constant bounds on the inverse
-\end{lemma}
-
 \begin{lemma} \label{lemma:bound1}
@@
   \]
-  as well. We can also use Lemma \ref{lemma:matrix_inverse} to show constant
-  upper and lower bounds on $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
-    \mathbf{u})}$. This is enough to prove constant upper and lower bounds on
+  as well.
+
+  First, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$ has constant bounds. If $i = 0$, then $\mathbf{K}|_{\lambda_i
+    = c(\mathbf{r}, \mathbf{u})} = c(\mathbf{r}, \mathbf{u})\mathbf{A}$, where
+  $\mathbf{A}$ is a constant matrix defined by
+  \[
+    \mathbf{A} = \frac{1}{\lambda_0}\mathbf{K},
+  \]
+  and, thus, is invertible.
+  Therefore,
+  \[
+    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
+    \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1}.
+  \]
+  Since $\lambda_0$ cannot be zero, we can make $\epsilon$ small enough so that
+  either $\lambda_0 - \epsilon > 0$ or $\lambda_0 + \epsilon < 0$, i.e., the
+  values of $c(\mathbf{r}, \mathbf{u})$ are either all positive or all negative.
+  Then, by continuity of $x \mapsto \frac{1}{x}$, $\mathbf{K}^{-1}|_{\lambda_i =
+    c(\mathbf{r}, \mathbf{u})}$ has constant bounds.
+
+  Let
+  \[
+    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
+    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n
+  \]
+  and $\delta = c(\mathbf{r}, \mathbf{u}) - \lambda_i$ so that $c(\mathbf{r},
+  \mathbf{u}) = \lambda_i + \delta$. If $i > 0$, then
+  \begin{align*}
+    &k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \!\begin{multlined}[t]
+      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2\\
+      - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left. \vphantom{\frac{1}{2}} \right)
+    \end{multlined} \\
+    &= \!\begin{multlined}[t]
+      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} (\lambda_i + \delta)(x_{j,i} - x_{k,i})^2 \\
+      - \mathbbm{1}[j \ne k]\sigma^2(\lambda_i + \delta) - S \left. \vphantom{\frac{1}{2}} \right)
+    \end{multlined} \\
+    &= \!\begin{multlined}[t]
+      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \\
+      - \frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right)
+    \end{multlined} \\
+    &= k(\mathbf{x}_j, \mathbf{x}_k)\exp \left( \vphantom{\frac{1}{2}} \right. -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right) \\
+    &= \!\begin{multlined}[t]
+      k(\mathbf{x}_j, \mathbf{x}_k) + k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp
+        \left( -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. \\
+      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right)
+    \end{multlined}
+  \end{align*}
+  Hence, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
+  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
+  \mathbf{E}$, where
+  \[
+    E_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp \left(
+        -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. -\mathbbm{1}[j
+    \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1
+    \right).
+  \]
+  Note that $E_{j,k} \to 0$ as $\delta \to 0$ and $\delta$ can be made
+  arbitrarily small via $\epsilon$. Then, since $\mathbf{K}$ is invertible, we
+  can make $\delta$ small enough so that $\lVert \mathbf{K}^{-1} \rVert_\infty
+  \lVert \mathbf{E} \rVert_\infty < 1$, and Lemma \ref{lemma:perturbation} shows
+  the existence of $\mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ and
+  gives upper and lower bounds on all of its elements.
+
+  This is enough to prove constant upper and lower bounds on
   $\mathbf{S}$, $\bm\Gamma$, and $\mathbf{R}$ (all with $\lambda_i$ replaced
@@
   TODO: Show that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is
-  bounded.
+  bounded. Might need something similar to the matrix bound lemma, since the
+  definition of $\bm\Gamma$ involves addition.
 
@@
   has the required bounds.
-
-  If $i = 0$, then each element of $\frac{\partial \mathbf{K}}{\partial
-    \lambda_0}$ is of the form
-  \[
-    \exp \left( -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal
-      \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne
-      k]\sigma^2\tr(\bm\Lambda) \right),
-  \]
-  i.e., without $\lambda_0$, so
-  \[
-    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
-      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \mathbf{K}}{\partial
-      \lambda_0}
-  \]
-  is already independent of $\mathbf{r}$ and $\mathbf{u}$---there is no need
-  for any bounds.
-
-  If $i > 0$, then each element of $\frac{\partial \mathbf{K}}{\partial
-    \lambda_i}$ is a constant multiple of $k(\mathbf{x}_j,
-  \mathbf{x}_k)$, for some $\mathbf{x}_j$ and $\mathbf{x}_k$. Since
-  $k(\mathbf{x}_j, \mathbf{x}_k)$ is a decreasing function of
-  $\lambda_i$, and $c(\mathbf{r}, \mathbf{u}) > a$,
-  \begin{alignat*}{3}
-    k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r},
-      \mathbf{u})} &= \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. &&
-    -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2 \\
-    & && - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left.
-      \vphantom{\frac{1}{2}} \right) \\
-    &< \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. && -\frac{1}{2}
-    a(x_{j,i} - x_{k,i})^2 \\
-    & && - \mathbbm{1}[j \ne k]\sigma^2a - S \left. \vphantom{\frac{1}{2}}
-    \right),
-  \end{alignat*}
-  where
-  \[
-    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
-    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n,
-  \]
-  which gives an upper bound on each element of
-  \[
-    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i =
-      c(\mathbf{r}, \mathbf{u})}.
-  \]
-  A similar line of reasoning establishes lower bounds as well.
-
-  % TODO: Lemma 1.1 is wrong about the derivative. Fix (my handwritten notes
-  % have the correct idea)
-  Combining the bounds with the observation that
-  every element of
-  $\Kuu^{-1}\mathbf{u}\mathbf{u}^\intercal\Kuu^{-\intercal}$ is in
-  $\mathbb{R}_2[\mathbf{u}]$ gives the required result.
 \end{proof}
@@
   \rVert_\infty$ on both vectors and matrices. We can apply Lemma
-  \ref{prop:condition_number} to $\bm\Sigma$ and $\mathbf{E}(\mathbf{r},
+  \ref{lemma:perturbation} to $\bm\Sigma$ and $\mathbf{E}(\mathbf{r},
   \mathbf{u})$ since

### 2019-03-16 mpaper/paper.tex
All the important information should be there, but in a very messy form

@@
 \newcommand{\Luu}{\mathbf{L}_{\mathbf{u},\mathbf{u}}}
+\newcommand{\Euu}{\mathbf{E}_{\mathbf{u},\mathbf{u}}}
+\newcommand{\Err}{\mathbf{E}_{\mathbf{r},\mathbf{r}}}
+\newcommand{\Eru}{\mathbf{E}_{\mathbf{r},\mathbf{u}}}
 
@@
 \subsection{Theoretical Justification}
+% TODO: in this section, 'constant' means 'constant w.r.t. u and r'
 
@@
 
-  TODO: Show that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is
-  bounded. Might need something similar to the matrix bound lemma, since the
-  definition of $\bm\Gamma$ involves addition.
+  % TODO: actually demonstrate this for the inverse
+  Recall that
+  \[
+    \bm\Gamma = \Krr - \mathbf{S}\Kru = \Krr - \Kru^\intercal\Kuu^{-1}\Kru.
+  \]
+  We have already demonstrated that for $\mathbf{K} \in \{ \Krr, \Kru, \Kuu^{-1}
+  \}$, $\lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
+  = \mathbf{K}$. We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
+  then $f(x) = f + g(x)$ for some function $g$ such that $\lim_{x \to 0} g(x) =
+  0$ to define matrices $\Err$, $\Eru$, and $\Euu$ such that
+  \begin{alignat*}{3}
+    \Krr|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Krr + \Err, &&\Err \to
+    \mathbf{O}_{|\mathcal{S}|}, \\
+    \Kru|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kru + \Eru, \quad
+    \text{and} \quad &&\Eru \to \mathbf{O}_{|\mathcal{S}|, m}, \\
+    \Kuu^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kuu^{-1} + \Euu,
+    &&\Euu \to \mathbf{O}_{m}.
+  \end{alignat*}
+  as $\epsilon \to 0$. Then, $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
+  = \bm\Gamma + \mathbf{E}$, where
+  \begin{align*}
+    \mathbf{E} &= \Err - \Kru^\intercal\Kuu^{-1}\Eru - \Kru^\intercal\Euu(\Kru + \Eru) \\
+               &- \Eru^\intercal(\Kuu^{-1} + \Euu)(\Kru + \Eru) \\
+               &\to \mathbf{O}_{|\mathcal{S}|}
+  \end{align*}
+  as $\epsilon \to 0$. Thus, Lemma \ref{lemma:perturbation} shows that
+  $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ exists and provides
+  constant bounds on its elements.
 
@@
 \end{proof}
+% TOOD: epsilon should not be treated as a constant. Revise the whole theory
+% section.
 

### 2019-03-17 mpaper/paper.tex
I localised the bug, but still don't see anything wrong

@@
   \end{align*}
-  by Lemma \ref{lemma:derivatives}.
-
-  Given any covariance matrix $\mathbf{K}$, we can easily establish constant
-  upper and lower bounds on $\mathbf{K}|_{\lambda_i = c(\mathbf{r},
-    \mathbf{u})}$ by the boundedness of $c$ and continuity of the covariance
-  function. Similar reasoning combined with the expressions for derivatives of
-  covariance matrices in Lemma \ref{lemma:derivatives} gives constant upper and
-  lower bounds on the elements of
+  by Lemma \ref{lemma:derivatives}. Let $\mathbf{K}$ be any covariance
+  matrix and
+  \[
+    \mathbf{A} = \frac{1}{\lambda_0} \mathbf{K}.
+  \]
+
+  First, we will show that
+  \[
+    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to 0 \quad \text{and}
+    \quad \left. \frac{\mathbf{\partial K}}{\partial \lambda_i}
+    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to 0
+  \]
+  as $\epsilon \to 0$.
+  We can easily establish constant upper and lower bounds on
+  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ by the boundedness of
+  $c$ and continuity of the covariance function. Similar reasoning combined with
+  the expressions for derivatives of covariance matrices in Lemma
+  \ref{lemma:derivatives} gives constant upper and lower bounds on the elements
+  of
   \[
@@
 
-  First, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
-    \mathbf{u})}$ has constant bounds. If $i = 0$, then $\mathbf{K}|_{\lambda_i
-    = c(\mathbf{r}, \mathbf{u})} = c(\mathbf{r}, \mathbf{u})\mathbf{A}$, where
-  $\mathbf{A}$ is a constant matrix defined by
-  \[
-    \mathbf{A} = \frac{1}{\lambda_0}\mathbf{K},
-  \]
-  and, thus, is invertible.
-  Therefore,
-  \[
-    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
-    \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1}.
-  \]
-  Since $\lambda_0$ cannot be zero, we can make $\epsilon$ small enough so that
-  either $\lambda_0 - \epsilon > 0$ or $\lambda_0 + \epsilon < 0$, i.e., the
-  values of $c(\mathbf{r}, \mathbf{u})$ are either all positive or all negative.
-  Then, by continuity of $x \mapsto \frac{1}{x}$, $\mathbf{K}^{-1}|_{\lambda_i =
-    c(\mathbf{r}, \mathbf{u})}$ has constant bounds.
-
-  Let
+  % Inverse of a covariance matrix
+  Now, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
+    \mathbf{u})}$ exists and
+  \[
+    \lim_{\epsilon \to 0} \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
+      \mathbf{u})} = \mathbf{K}.
+  \]
+  If $i = 0$, then $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
+  c(\mathbf{r}, \mathbf{u})\mathbf{A}$. Therefore\footnote{Note that since
+    $\lambda_0 \ne 0$, $c(\mathbf{r}, \mathbf{u}) \ne 0$ for small enough
+    $\epsilon$.},
+  \begin{align*}
+    \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &=
+                                                               \frac{1}{c(\mathbf{r}, \mathbf{u})}\mathbf{A}^{-1} \\
+                                                             &\to \frac{1}{\lambda_0}\mathbf{A}^{-1} = \frac{1}{\lambda_0} \left( \frac{1}{\lambda_0}\mathbf{K} \right)^{-1} = \mathbf{K}^{-1}
+  \end{align*}
+  as $\epsilon \to 0$.
+
+  We will now show the same result for $\lambda_i > 0$. Let
   \[
@@
   and $\delta = c(\mathbf{r}, \mathbf{u}) - \lambda_i$ so that $c(\mathbf{r},
-  \mathbf{u}) = \lambda_i + \delta$. If $i > 0$, then
+  \mathbf{u}) = \lambda_i + \delta$, and $\lim_{\epsilon \to 0} \delta = 0$.
+  Then,
   \begin{align*}
@@
   $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
-  \mathbf{E}$, where
+  \mathbf{E}$, where $\mathbf{E}$ is defined as
   \[
@@
   \]
-  Note that $E_{j,k} \to 0$ as $\delta \to 0$ and $\delta$ can be made
-  arbitrarily small via $\epsilon$. Then, since $\mathbf{K}$ is invertible, we
-  can make $\delta$ small enough so that $\lVert \mathbf{K}^{-1} \rVert_\infty
-  \lVert \mathbf{E} \rVert_\infty < 1$, and Lemma \ref{lemma:perturbation} shows
+  By this definition,
+  \[
+    \lim_{\epsilon \to 0} E_{j,k} = 0.
+  \]
+  Then, since $\mathbf{K}$ is invertible, Lemma \ref{lemma:perturbation} shows
   the existence of $\mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ and
@@
 
-  Since $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ is bounded, so is
-  $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Assuming that
-  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists, we
-  can make $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ close enough to
-  $\bm\Gamma$ to ensure that
-  \[
-    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0.
-  \]
-  Moreover, boundedness of $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
-  gives boundedness of $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r},
-    \mathbf{u})}$.
-
-  Combined with the fact that $ab > 0$ (i.e., both endpoints together are either
-  positive or negative), this gives boundedness of $\det(\bm\Gamma)|_{\lambda_i
-    = c(\mathbf{r}, \mathbf{u})}^{-1}$.
-
-  % TODO: actually demonstrate this for the inverse
+  Since
+  \[
+    \lim_{\epsilon \to 0} \bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
+    \bm\Gamma,
+  \]
+  we also have that
+  \[
+    \lim_{\epsilon \to 0} \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
+    \det(\bm\Gamma).
+  \]
+  Assuming that
+  $\bm\Gamma$ is invertible so that $q(\mathbf{r} \mid \mathbf{u})$ exists,
+  \[
+    \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \ne 0
+  \]
+  for small enough $\epsilon$, and, thus, $\det(\bm\Gamma)^{-1}|_{\lambda_i =
+    c(\mathbf{r}, \mathbf{u})}$ exists and has constant bounds.
+
   Recall that
@@
   We have already demonstrated that for $\mathbf{K} \in \{ \Krr, \Kru, \Kuu^{-1}
-  \}$, $\lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
-  = \mathbf{K}$. We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
+  \}$,
+  \[
+    \lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
+    \mathbf{K}.
+  \]
+  We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
   then $f(x) = f + g(x)$ for some function $g$ such that $\lim_{x \to 0} g(x) =
@@
 
+% TODO: a section about the algorithm
+
 \section{Evaluation} % showing deep understanding

### 2019-03-17 code/VIGPIRL/full_gradient.m

@@ -1,6 +1,4 @@
 function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, matrices)
-% TODO: will need to make this return optimal r as a function of mu so that
-% vigpirlrun can construct an optimal solution (just like before)
 % NOTE: returns the gradient for parameters in vector form
 
   function answer = estimate_derivative(u)
@@ -35,7 +33,7 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
     unique_lambda_part = arrayfun(@lambda_derivative, 1:size(matrices.Kuu_grad, 3));
     unique_mu_part = (u' - gp.mu)' * (Sigma_inv + Sigma_inv');
-    unique_B_part = 2*(Sigma_inv*U*Sigma_inv - adjoint(Sigma)/det(Sigma)) * gp.B;
+    unique_B_part = 2 * (Sigma_inv * U * Sigma_inv - Sigma_inv) * gp.B;
 
     unique_part = vertcat(2, unique_lambda_part', diag(unique_B_part),...
       unique_mu_part', get_lower_triangle(unique_B_part));
@@ -43,8 +41,8 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   end
 
   Sigma = gp.B * gp.B';
-  disp(gp.B);
-  disp(Sigma);
+  %disp(gp.B);
+  %disp(Sigma);
   Sigma_inv = inv(Sigma);
   Kuu_inv = inv(matrices.Kuu);
   S = matrices.Kru' * Kuu_inv;
@@ -70,6 +68,7 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   not_estimated = vertcat(not_estimated_elbo, not_estimated_lambda',...
     diag(not_estimated_B), not_estimated_mu, get_lower_triangle(not_estimated_B));
   changes = not_estimated - 0.5 * estimated_grad;
+  %changes = -0.5 * estimated_grad;
 
   % TESTING
   %fprintf('Not estimated: %f, estimated: %f, total: %f\n', not_estimated(2), estimated_grad(2), changes(2));

### 2019-03-17 code/VIGPIRL/vigpirldefaultparams.m

@@ -10,11 +10,11 @@ default_params = struct(...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
-    'samples_count',1000,...
+    'samples_count',100,...
     'learning_rate',0.1,...
-    'B_learning_rate',0.001,...
-    'num_iterations',400,...
-    'required_precision', 0.001);
+    'B_learning_rate',0.1,...
+    'num_iterations',300,...
+    'required_precision', 0.01);
 
 % Set parameters.
 algorithm_params = filldefaultparams(algorithm_params,default_params);

### 2019-03-17 code/VIGPIRL/vigpirlrun.m

@@ -179,12 +179,13 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     parameter_vector(index) = 0;
     first_index = size(grad_history, 2) - length(mu_log) + 1;
     subplot(2, 3, i);
+    yyaxis left;
     plot(mu_log, elbo_list(first_index:end));
     hold on;
+    yyaxis right;
     plot(mu_log, grad_history(index, first_index:end));
     hold off;
   end
-
   mu_from = -5;
   mu_to = 5;
   for i = 1:3
@@ -198,8 +199,10 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     parameter_vector(index) = 0;
     first_index = size(grad_history, 2) - length(mu_log) + 1;
     subplot(2, 3, i + 3);
+    yyaxis left;
     plot(mu_log, elbo_list(first_index:end));
     hold on;
+    yyaxis right;
     plot(mu_log, grad_history(index,first_index:end));
     hold off;
   end
@@ -248,11 +251,11 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
 
     old_hyperparameters = vigpirlpackparam(gp);
 
-    grad(1) = 0; % lambda0
-    grad(2:d+1) = 0; % lambda (except first)
-    grad(d+m+2:d+2*m+1) = 0; % mu
-    grad(d+m+3) = 0; % mu (second)
-    grad(d+m+4) = 0; % mu (third)
+    %grad(1) = 0; % lambda0
+    %grad(2:d+1) = 0; % lambda (except first)
+    %grad(d+m+2:d+2*m+1) = 0; % mu
+    %grad(d+m+3) = 0; % mu (second)
+    %grad(d+m+4) = 0; % mu (third)
     %grad(d+2:d+m+1) = 0; % B diagonal
     %grad(d+2*m+2:end) = 0; % rest of B
 

### 2019-03-17 code/singletest.m

@@ -20,7 +20,7 @@ addpaths;
 % VIGPIRL
 %test_result = runtest('vigpirl',struct(),'linearmdp',...
 %    'gridworld',struct('n',8,'determinism',1,'seed',1,'continuous',0),...
-%    struct('training_sample_lengths',4,'training_samples',6,'verbosity',2));
+%    struct('training_sample_lengths',32,'training_samples',16,'verbosity',2));
 
 % Visualize solution.
 %printresult(test_result);
@@ -32,7 +32,7 @@ addpaths;
 %example_samples = {[1, 1]};
 
 % Three states
-mdp_data = struct('discount', 0.5, 'states', 3, 'actions', 2);
+mdp_data = struct('discount', 0.9, 'states', 3, 'actions', 2);
 mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
 mdp_data.sa_p(1:3, 1:2, 1) = 1;
 feature_data = struct('splittable', [1; 2; 3]);

### 2019-03-18 mpaper/paper.tex
Sage plots of q(u) and its derivatives

@@
 % problem description/statement
+% TODO: rewrite and make it good
 
@@
 % comprehensive literature survey
+% TODO: need to add a literature survey here
 
@@
 \subsection{Theoretical Justification}
-% TODO: in this section, 'constant' means 'constant w.r.t. u and r'
 
@@
 The typical way to optimise a quantity (the ELBO, in this case) involves
-computing its gradient. Unfortunately, some of the terms in $\mathcal{L}$ are
-still left as expected values. The goal of this section is to show how Theorem
+computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
+complicates the situation. The goal of this section is to show how Theorem
 \ref{thm:lebesgue} can be applied to our model in order to derive the gradient
@@
   problem and requires significantly more work to prove correctness.}. After
-showing that the theorem applies to our situation, given any term expressed as
-an expected value $\mathbb{E}[f(\mathbf{u}, \mathbf{r})]$, we can estimate its
-gradient as
-\[
-  \nabla\mathbb{E}[f(\mathbf{u}, \mathbf{r})] = \mathbb{E}[\nabla f(\mathbf{u},
-  \mathbf{r})] \approx \frac{1}{S} \sum_{s=1}^S \nabla f(\mathbf{u}_s,
-  \mathbf{r}_s),
-\]
+showing that the theorem applies to our situation, we can express
+$\nabla\mathbb{E}[v]$ as
+\begin{align*}
+  \nabla\mathbb{E}[v] &= \nabla \iint q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
+                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
+                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
+                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})},
+\end{align*}
 which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
@@
 \end{align*}
+Also note that throughout this section the word `constant' means `constant with
+respect to $\mathbf{u}$ and $\mathbf{r}$'.
 
@@
       $\frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} =
-      \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
-      - \bm\Sigma^{-\intercal})$.
+      \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
+      \bm\Sigma^{-1})$.
     \item
@@
       q(\mathbf{u})(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
-      |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}$.
+      \bm\Sigma^{-1})\mathbf{B}$.
     \end{enumerate}
@@
 
-% TODO: a section about the algorithm
+% Maybe?
+%\subsection{Implementation}
 
@@
 
+How well does the variance prediction work? It doesn't, but why?
+TODO: Sage should consider 3x3 matrices
+TODO: 4 plots---two my MATLAB, two my Sage; two for B[1, 1], two for B[2, 1].
+The MATLAB plots show E[v] and dE[v]/dB, the Sage plots show q(u) and
+dq(u)/dB. Function in blue, derivative in red, over a range of values.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{sage1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{sage2}
+\end{figure}
+
 % How does policy depend on rewards? Choose state and action, generate an array
@@
 well as lower $r(s_3)$ values.
-
-% How well does the variance prediction work?
 

### 2019-03-18 mpaper/paper.tex
Second round of plots for describing the part that doesn't work

@@
 \usepackage{layouts}
+%\usepackage{subcaption}
 
@@
 \DeclareMathOperator{\adj}{adj}
-%\DeclareMathOperator{\dett}{det}
 \DeclareMathOperator{\tr}{tr}
@@
 
-% Maybe?
-%\subsection{Implementation}
-
 \section{Evaluation} % showing deep understanding
@@
 
- % TODO: need to redo this to get better labels, more precision, wider range of values
 \begin{figure*}
   \centering
-  \includegraphics[width=\textwidth]{elbo_contour_plots}
-\end{figure*}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{elbo_over_gamma}
-  \caption{Changes in $\mathcal{L}$ as a function of $\bm\mu$, where $\mu_1 =
-    \mu_2 = \mu_3$, and $\gamma = 0, 0.1, \dots, 0.9$. Each of the ten curves
-    plots a different value of $\gamma$, ranging from $0$ in dark blue to $0.9$
-    in dark red.}
-\end{figure}
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\textwidth]{ELBO_and_gradient_over_B}
-  \caption{ELBO in blue, derivative in red. The first row is for diagonal values
-  of B, the second row is for non-diagonal values.}
-\end{figure*}
-
-How well does the variance prediction work? It doesn't, but why?
-TODO: Sage should consider 3x3 matrices
-TODO: 4 plots---two my MATLAB, two my Sage; two for B[1, 1], two for B[2, 1].
-The MATLAB plots show E[v] and dE[v]/dB, the Sage plots show q(u) and
-dq(u)/dB. Function in blue, derivative in red, over a range of values.
-
-\begin{figure}
+  \begin{minipage}{.5\textwidth}
     \centering
     \includegraphics[width=\columnwidth]{sage1}
-\end{figure}
-
-\begin{figure}
+    \includegraphics[width=\columnwidth]{sage2}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
     \centering
-  \includegraphics[width=\columnwidth]{sage2}
-\end{figure}
-
+    \includegraphics[width=\columnwidth]{elbo_over_B_short}
+  \end{minipage}
+  \caption{How a well-defined problem with a correct solutions becomes a
+    well-defined problem with an incorrect solution. In each plot, the function
+    we are trying to optimise is in blue, and its derivative is in red. The
+    plots on the left are for $q(\mathbf{u})$ and its derivatives with respect
+    to $B_{1,1}$ (at the top) and $B_{2,1}$ (at the bottom), whereas the plots
+    on the right are for $\mathbb{E}[v]$ and its corresponding derivatives.}
+\end{figure*}
+% TODO: describe the figure
+
+% TODO: evaluate by setting dE[v]/dB to zero
 % How does policy depend on rewards? Choose state and action, generate an array
 % of contour plots showing the probability of taking that action in an optimal policy
-Observation from a plot not worth including here. The policy behaves as
-expected: if we fix $r(s_1)$, $\pi(s_1, a_1)$ increases with higher $r(s_2)$ as
-well as lower $r(s_3)$ values.
+
+% Observation from a plot not worth including here. The policy behaves as
+% expected: if we fix $r(s_1)$, $\pi(s_1, a_1)$ increases with higher $r(s_2)$ as
+% well as lower $r(s_3)$ values.
 

### 2019-03-18 code/VIGPIRL/vigpirldefaultparams.m

@@ -10,7 +10,7 @@ default_params = struct(...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
-    'samples_count',100,...
+    'samples_count',1000,...
     'learning_rate',0.1,...
     'B_learning_rate',0.1,...
     'num_iterations',300,...

### 2019-03-18 code/VIGPIRL/vigpirlrun.m

@@ -17,17 +17,16 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   gp = vigpirlinit(algorithm_params, feature_data);
                                 % Choose inducing points.
   %gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
-  gp.X_u = gp.X; % TEMP
+  gp.X_u = gp.X;
   m = size(gp.X_u, 1);
-  gp.mu = rand(1, m)';
-  %gp.mu = [-2; 3; 3]; % TEMP
-                 % B is a lower triangular matrix with positive diagonal entries
+  %gp.mu = rand(1, m)';
+  gp.mu = [-5; 0; 4]; % TEMP
+
+  % B is a lower triangular matrix with positive diagonal entries
   %gp.B = normrnd(0, 1, [m, m]);
   %gp.B(1:m+1:end) = random('Chisquare', 4, [m, 1]);
   %gp.B = tril(gp.B);
   gp.B = eye(m); % TEMP
-  %gp.D = diag(normrnd(0, 1));
-  gp.D = zeros(m, m); % TEMP
 
   d = size(feature_data.splittable, 2);
   elbo_list = [];
@@ -165,47 +164,94 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   % ELBO and derivative over some element of B
   % NOTE: this is the log of the actual value
   % TODO: titles, fix values on the x-axis, add two axes and their names to the y-aixs
-  accuracy = 0.1;
-  mu_from = -3;
-  mu_to = 3;
-  for i = 1:3
-    mu_log = [];
-    index = d + 1 + i;
-    for mu = mu_from:accuracy:mu_to
-      parameter_vector(index) = mu;
-      wrapper(parameter_vector);
-      mu_log = [mu_log mu];
-    end
-    parameter_vector(index) = 0;
-    first_index = size(grad_history, 2) - length(mu_log) + 1;
-    subplot(2, 3, i);
-    yyaxis left;
-    plot(mu_log, elbo_list(first_index:end));
-    hold on;
-    yyaxis right;
-    plot(mu_log, grad_history(index, first_index:end));
-    hold off;
+  %accuracy = 0.1;
+  %mu_from = -3;
+  %mu_to = 3;
+  %for i = 1:3
+  %  mu_log = [];
+  %  index = d + 1 + i;
[…]

### 2019-03-19 mpaper/paper.tex
Big chunk of evaluation done

@@
 \usepackage{layouts}
-%\usepackage{subcaption}
+\usepackage{xcolor}
 
@@
 \section{Introduction}
-% problem description/statement
 % TODO: rewrite and make it good
+% 1. Problem (with examples, intuitive)
+% 2. Its importance/relevance/motivation
+% 3. My idea
+% 4. List of contributions
+% Forward-reference which section deals with what
 
@@
 \cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
-Most approaches in the literature (see Section \ref{sec:background}) make a
-convenient yet unjustified assumption that the reward function can be expressed
-as a linear combination of features. One proven way to abandon this assumption
-is by representing the reward function as a Gaussian process (GP)
+Most approaches in the literature make a convenient yet unjustified assumption
+that the reward function can be expressed as a linear combination of features.
+One proven way to abandon this assumption is by representing the reward function
+as a Gaussian process (GP)
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
@@
 
-\subsection{Statement of the Problem}
+\section{The Problem}
 
@@
 assumptions. Moreover, we hope to make use of various recent advancements in VI
-for both time complexity and approximation distribution fit (see Section
-\ref{sec:background}), making the resulting algorithm competitive both in terms of
-running time and model fit.
-
-\section{Background} \label{sec:background}
-% motivation, relevance (whatever that means)
-% comprehensive literature survey
-% TODO: need to add a literature survey here
-
-\subsection{Linear Algebra, Numerical Analysis, and Measure Theory}
-
-Here we introduce a few definitions and results that will be used later in the
-paper. Namely, we will use several different vector and matrix norms, consider
-how an inverse of a matrix changes with a small perturbation, and use Lebesgue's
-dominated convergence theorem in order to justify the validity of our approach.
+for both time complexity and approximation distribution fit, making the
+resulting algorithm competitive both in terms of running time and model fit.
+
+\section{Background}
+
+Here we introduce a few definitions and results from linear algebra, numerical
+analysis, and measure theory that will be used later in the paper. Namely, we
+will use several different vector and matrix norms, consider how an inverse of a
+matrix changes with a small perturbation, and use Lebesgue's dominated
+convergence theorem in order to justify the validity of our approach.
 
@@
 
+% TODO: need some editing
 \begin{lemma} \label{lemma:bound1}
@@
 \section{Evaluation} % showing deep understanding
-% Mention all the constants I've used.
 
@@
 \end{align*}
-all other values of $\mathcal{T}$ to zero, and vary the value of $\gamma$. We
+all other values of $\mathcal{T}$ to zero, and $\gamma = 0.9$. We
 also set the inducing points to be equal to the three states in $\mathcal{S}$,
@@
 \end{figure*}
-% TODO: describe the figure
-
-% TODO: evaluate by setting dE[v]/dB to zero
-% How does policy depend on rewards? Choose state and action, generate an array
-% of contour plots showing the probability of taking that action in an optimal policy
-
-% Observation from a plot not worth including here. The policy behaves as
-% expected: if we fix $r(s_1)$, $\pi(s_1, a_1)$ increases with higher $r(s_2)$ as
-% well as lower $r(s_3)$ values.
+
+TODO: describe the problem with B.
+
+\begin{figure*}
+  \centering
+  \begin{minipage}{.5\textwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{convergence1}
+  \end{minipage}%
+  \begin{minipage}{.5\textwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{convergence2}
+  \end{minipage}
+  \caption{The convergence of $\mathcal{L}$ (at the top) and several example
+    policies (at the bottom) over a number of iterations for Scenario~1 on the
+    left and Scenario~2 on the right.}
+  \label{fig:convergence1}
+\end{figure*}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{parameter_convergence1}
+  \includegraphics[width=\textwidth]{parameter_convergence2}
+  \caption{Convergence of all optimised parameters for Scenario~1 at the top and
+  Scenario~2 at the bottom. In order to represent different variables on the
+  same scale, some variables have been $\log$-transformed. Colours denote which
+  vector or matrix each scalar comes from: black for $\bm\lambda$,
+  \textcolor{blue}{blue} for $\bm\mu$, \textcolor{red}{red} for diagonal
+  elements of $\mathbf{B}$, and \textcolor{green}{green} for its non-diagonal
+  elements.}
+  \label{fig:convergence2}
+\end{figure*}
+
+Because of this issue, we look at the convergence of our model using a constant
+step size in two different scenarios:
+\begin{itemize}
+\item In \textbf{Scenario 1}, we remove $\frac{\partial \mathbb{E}[v]}{\partial
+    \mathbf{B}}$ from $\frac{\partial \mathcal{L}}{\partial \mathbf{B}}$. This
+  essentially optimises $\mathbf{B}$ to match $\Kuu$, because then
+  $\frac{\partial \mathcal{L}}{\partial \mathbf{B}}$ becomes
+  $-\frac{\partial}{\partial \mathbf{B}} \DKL{q(\mathbf{u})}{p(\mathbf{u})}$,
+  i.e., we are optimising $\mathbf{B}$ to minimise the difference between the
+  prior and the posterior of $\mathbf{u}$.
+\item In \textbf{Scenario 2}, we set $\mathbf{B} = \mathbf{I}_m$, and do not
+  optimise it at all.
+\end{itemize}
+% TODO: use \ell when introducing the norms
+
+We plot how $\mathcal{L}$ as well as policies $\pi(a_1 \mid s_1)$, $\pi(a_2 \mid
+s_3)$, and $\pi(a_1 \mid s_2)$ converge over a number of iterations in Figure
+\ref{fig:convergence1}. The first two policies correspond to actions taken in
+the set of demonstrations $\mathcal{D}$, so we would expect to see their
+probabilities converge to values above $0.5$. Note that due to the stochastic
+nature of our MDP model, we do not expect to see any probability reach exactly
+$1$. The third policy, however, has no relevant data in $\mathcal{D}$, so the
+maximum causal entropy framework would put the probability at around $0.5$.
+
+As we ran the algorithm with two stopping conditions---$300$ iterations and the
+$\ell_1$-norm of the change in parameter values being below $0.01$---note that
+the algorithm terminated early in Scenario~2 but not in Scenario~1, although
+there is clear convergence in both cases. An important difference is that
+$\pi(a_1 \mid s_1)$ and $\pi(a_2 \mid s_3)$ converge closer and closer to $1$
+with the slightly-restricted model of Scenario~1, but stabilise at $0.6$ in
+Scenario~2 due to an important part of the model being completely fixed. In both
+cases, $\pi(a_1 \mid s_2)$ converges to $0.5$ as expected.
+
+Figure \ref{fig:convergence2} shows how the parameters of the model converge in
+both scenarios. The algorithm seems to converge fine in Scenario~2, but most of
+the parameters fail to stabilise in Scenario~1. Most importantly, the diagonal
+values of the $\mathbf{B}$ matrix diverge to positive infinity, leading to higher
+variance in ELBO estimates seen in Figure \ref{fig:convergence1}. The
+non-diagonal values, however, converge just fine. Both $\lambda_0$ and
+$\lambda_1$ also diverge to positive infinity, while the elements of $\bm\mu$
+diverge, but in `the right' direction: as $\mu_2$ increases while $\mu_1$ and
+$\mu_3$ decrease, the policies in Figure \ref{fig:convergence1} converge to
+their optimal values.
+
+This leaves us with two models: one that converges to reasonable-but-suboptimal
+values, and one that diverges to infinite variance but also provides correct
+policies.
+
+\section{Related Work}
+% TODO: comprehensive literature survey
 
 \section{Conclusions}
-
+% TODO: write
+
+Reasonable results in convergence.
 More variables/information. Fewer assumptions.
@@
 
-\subsection{Future Work}
+\subsection{Further Work}
 

### 2019-03-19 code/VIGPIRL/full_gradient.m

@@ -33,7 +33,8 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
     unique_lambda_part = arrayfun(@lambda_derivative, 1:size(matrices.Kuu_grad, 3));
     unique_mu_part = (u' - gp.mu)' * (Sigma_inv + Sigma_inv');
-    unique_B_part = 2 * (Sigma_inv * U * Sigma_inv - Sigma_inv) * gp.B;
+    %unique_B_part = 2 * (Sigma_inv * U * Sigma_inv - Sigma_inv) * gp.B;
+    unique_B_part = zeros(length(gp.mu));
 
     unique_part = vertcat(2, unique_lambda_part', diag(unique_B_part),...
       unique_mu_part', get_lower_triangle(unique_B_part));
@@ -42,7 +43,10 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
   Sigma = gp.B * gp.B';
   %disp(gp.B);
-  %disp(Sigma);
+  fprintf('Kuu:\n');
+  disp(matrices.Kuu);
+  fprintf('Sigma:\n');
+  disp(Sigma);
   Sigma_inv = inv(Sigma);
   Kuu_inv = inv(matrices.Kuu);
   S = matrices.Kru' * Kuu_inv;

### 2019-03-19 code/VIGPIRL/vigpirlinit.m

@@ -12,8 +12,8 @@ gp.X_u = feature_data.splittable;
 
 % Initialize hyperparameters.
 gp.noise_var = algorithm_params.noise_init;
-%gp.lambda0 = random('Chisquare', algorithm_params.rbf_init);
-gp.lambda0 = 1;
-%gp.lambda = random('Chisquare', algorithm_params.ard_init,...
-%    [1, size(feature_data.splittable, 2)]);
-gp.lambda = ones(1, size(feature_data.splittable, 2));
+gp.lambda0 = random('Chisquare', algorithm_params.rbf_init);
+%gp.lambda0 = 1;
+gp.lambda = random('Chisquare', algorithm_params.ard_init,...
+    [1, size(feature_data.splittable, 2)]);
+%gp.lambda = ones(1, size(feature_data.splittable, 2));

### 2019-03-19 code/VIGPIRL/vigpirlrun.m

@@ -19,8 +19,8 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   %gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
   gp.X_u = gp.X;
   m = size(gp.X_u, 1);
-  %gp.mu = rand(1, m)';
-  gp.mu = [-5; 0; 4]; % TEMP
+  gp.mu = rand(1, m)';
+  %gp.mu = [-5; 0; 4]; % TEMP
 
   % B is a lower triangular matrix with positive diagonal entries
   %gp.B = normrnd(0, 1, [m, m]);
@@ -32,6 +32,7 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   elbo_list = [];
   grad_history = [];
   hyperparameter_history = [];
+  policy_history = [];
 
   % wrapper: vector of parameters -> scalar value * gradient vector
   function [elbo, grad] = wrapper(parameter_vector)
@@ -100,34 +101,29 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   %mu1_log = [];
   %mu2_log = [];
   %mu3_log = [];
-  %multiplier = 10;
   %policy_log = [];
-  %for mu1 = 1:10
-  %  real_mu1 = multiplier * (mu1 - 5);
-  %  parameter_vector(d+m+2) = real_mu1;
-  %  for mu2 = 1:10
-  %    real_mu2 = multiplier * (mu2 - 5);
-  %    parameter_vector(d+m+3) = real_mu2;
-  %    for mu3 = 1:10
-  %      real_mu3 = multiplier * (mu3 - 5)
-  %      parameter_vector(d+m+4) = real_mu3;
+  %mu1_range = -5:5;
+  %mu_range = -5:0.5:5;
+  %num_points = length(mu_range) * length(mu_range);
+  %for mu1 = mu1_range
+  %  for mu2 = mu_range
+  %    for mu3 = mu_range
   %      solution = feval([mdp_model 'solve'], mdp_data, [mu1; mu2; mu3]);
   %      policy = solution.p(1, 1); % state 1, action 1
   %      policy_log = [policy_log policy];
-  %      mu1_log = [mu1_log real_mu1];
-  %      mu2_log = [mu2_log real_mu2];
-  %      mu3_log = [mu3_log real_mu3];
+  %      mu1_log = [mu1_log mu1];
[…]

### 2019-03-19 mpaper/paper.tex
Restructuring + importing the related work section

@@
 \usepackage{xcolor}
+\usepackage{booktabs}
 
@@
 
-\section{Introduction}
+\section{Introduction} % first page
 % TODO: rewrite and make it good
@@
   For any finite-dimensional vector $\mathbf{x} = (x_1, \dots, x_n)^\intercal$,
-  its \emph{maximum norm} is
+  its \emph{maximum norm} (\emph{$\ell_\infty$-norm}) is
   \[
@@
   \]
-  whereas its \emph{taxicab} (or \emph{Manhattan}) \emph{norm} is
+  whereas its \emph{$\ell_1$-norm} is
   \[
@@
 
-\section{Solution}
-% TODO: give an overview of this section. Also, consider renaming it or
-% splitting it into multiple sections
-
+\section{The Model}
+
+% TODO: make this fit the narrative
 For any matrix $\mathbf{A}$, we will use either $A_{i,j}$ or
@@
 the Lebesgue definition of integration.
-
-\subsection{Details of the Model}
 
@@
 
-\subsection{Theoretical Justification}
+\section{Theoretical Justification}
 
@@
 \]
-and create two demonstrations: $\zeta_1 = \{ (s_1, a_1) \}$ and $\zeta_2 = \{
-(s_3, a_2) \}$. In other words, the demonstrations move from $s_1$ and $s_3$ to
-$s_2$. Therefore, we would expect the reward (and value) of $s_2$ to be the
-highest.
+and create two demonstrations $\zeta_1 = \{ (s_1, a_1) \}$ and $\zeta_2 = \{
+(s_3, a_2) \}$ that correspond to moving from $s_1$ and $s_3$ to $s_2$.
+Therefore, we would expect the reward of $s_2$ to be higher than the other two
+rewards in order to reflect this.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{sage1}
-    \includegraphics[width=\columnwidth]{sage2}
+    \includegraphics[width=0.8\columnwidth]{sage1}
+    \includegraphics[width=0.8\columnwidth]{sage2}
   \end{minipage}%
@@
     \centering
-    \includegraphics[width=\columnwidth]{elbo_over_B_short}
+    \includegraphics[width=0.8\columnwidth]{elbo_over_B_short}
   \end{minipage}
@@
     on the right are for $\mathbb{E}[v]$ and its corresponding derivatives.}
+  \label{fig:B_gradient}
 \end{figure*}
-
-TODO: describe the problem with B.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{convergence1}
+    \includegraphics[width=0.8\columnwidth]{convergence1}
   \end{minipage}%
@@
     \centering
-    \includegraphics[width=\columnwidth]{convergence2}
+    \includegraphics[width=0.8\columnwidth]{convergence2}
   \end{minipage}
@@
 
-Because of this issue, we look at the convergence of our model using a constant
-step size in two different scenarios:
+Unfortunately, we are not able to use $\frac{\partial \mathcal{L}}{\partial
+  \mathbf{B}}$ in order to optimise $\mathbf{B}$. We illustrate the problem in
+Figure \ref{fig:B_gradient}. On the left side of the figure, we plot how
+$q(\mathbf{u})$ behaves as a function of a diagonal and a non-diagonal element
+of $\mathbf{B}$. Both functions have maximum values that can be attained by
+following the corresponding derivatives. However, when these derivatives are
+used to estimate $\frac{\partial \mathbb{E}[v]}{\partial \mathbf{B}}$, the
+resulting derivatives no longer match their corresponding functions, although
+the functions themselves still have optimal values: at or below $1$ for
+$B_{1,1}$ and at $0$ for $B_{2,1}$. This leads us to consider two restrictions
+of the model in order to investigate its convergence behaviour:
 \begin{itemize}
@@
 \end{itemize}
-% TODO: use \ell when introducing the norms
 
@@
 
-\section{Related Work}
-% TODO: comprehensive literature survey
+\section{Related Work} % 0.5-1 page
+% TODO: rewrite and make this section smaller
+
+As mentioned in the introduction, most IRL algorithms assume that the reward
+function can be represented as a linear combination of features. This assumption
+originated in one of the earliest papers on the topic by Ng and Russell
+\cite{DBLP:conf/icml/NgR00}, which introduced several linear programming
+approaches to the problem. The authors also noticed that often
+multiple reward functions can explain the same behaviour, and suggested
+heuristics for reward functions that are `far away' from reward functions that
+do not fit the data.
+
+A few years later, Abbeel and Ng \cite{DBLP:conf/icml/PieterN04} developed an
+algorithm for the same formulation of the problem with a guarantee to converge
+quickly. Neu and Szepesv\'ari \cite{DBLP:conf/uai/NeuS07} identified a weakness
+in Abbeel and Ng's approach: the algorithm requires features to be
+`appropriately' scaled, and optimal scaling may not be known. Instead, they
+suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
+supervised learning task for optimal \emph{policy} recovery (whereas IRL focuses
+on recovering the reward function).
+
+Ramachandran and Amir \cite{DBLP:conf/ijcai/RamachandranA07} were the first to
+formulate IRL in terms of Bayesian learning. While the model is easily
+interpretable and able to handle experts that make mistakes, the algorithm can
+only handle small state spaces and requires Monte Carlo Markov Chain (MCMC)
+sampling for inference.
+
+Ziebart et al. \cite{ziebart2008maximum} keep the linearity assumption, but
+introduce an influential idea: resolving the ambiguity when multiple reward
+functions explain the data by appealing to the maximum entropy principle.
+
+Choi and Kim \cite{DBLP:conf/ijcai/ChoiK13} extend the Bayesian model to learn
+good features as well as the reward function, trying to overcome the limitation
+of linearity. However, the approach is quite limiting: all features are assumed
+to have Boolean values, and the algorithm simply learns their conjunctions.
+
+Levine et al. \cite{DBLP:conf/nips/LevinePK11} are the first to suggest a way to
+learn nonlinear reward functions without harsh restrictions on the problem
+domain by using GPs. We base our work primarily on their paper, and the
+weaknesses we hope to address have already been covered in the previous section.
+A recent extension to their work by Jin et al. \cite{DBLP:conf/uai/JinDAS17}
+aims to harness the power of deep learning by using several layers of GPs,
+making the model less dependent on being provided good features. They also use
+VI, but with a few simplifying assumptions: deterministic training conditional
+for the reward vector, and fully independent training conditional for the latent
+state.
+
+Finally, instead of using GPs to model nonlinear reward functions, one can use a
+neural network (NN), as demonstrated by Wulfmeier et al.
+\cite{wulfmeier2015maximum}. Their approach benefits from constant time
+inference and the ability to learn complex features either from already-given
+features or even from raw data. The only disadvantage (as demonstrated in the
+paper) is that NNs converge slower than GPs (as measured by the number of
+demonstrations).
+
+\subsection{Variational Inference}
+
+Variational inference has seen a recent increase in interest among academics,
+with different approaches focusing on different goals: better time complexity,
+handling a wider variety of models, making approximations more accurate, and
+using more complex function approximation techniques (such as NNs)
+to infer local latent variable values without having to calculate them
+individually for each data point \cite{DBLP:journals/corr/abs-1711-05597}. As
+our IRL model is based on a GP, we will begin by reviewing some of the VI
+approaches applied specifically to GP regression. Based on a recent review of
+scalable GPs \cite{DBLP:journals/corr/abs-1711-05597}, we will concentrate on
+\emph{stochastic variational sparse approximations}, as they have achieved
+modelling accuracy close to that of the full GP with no approximations, with
+many methods providing a time complexity of $\mathcal{O}(m^3)$. Below we provide
+a short overview of various assumptions that have been used in approximating
+\emph{sparse} GPs (i.e., GPs that use inducing points), following on a paper by
+Qui\~{n}onero-Candela and Rasmussen \cite{DBLP:journals/jmlr/CandelaR05}.
+
+Table ?? further summarises some of the recent and/or
+influential GP approximation approaches that might be relevant to our situation.
+For many of the variables in our model (including the inducing points and the
+hyperparameters of the covariance function), we have three possible ways of
+handling them:
+\begin{itemize}
+\item Provide a full Bayesian treatment by defining a prior probability
+  distribution. While this option can easily prevent overfitting, the choice of
+  prior can be difficult to justify.
+\item Fix the value of the variable. We can then optimise the value outside of
+  the main algorithm, e.g., by running the algorithm with a set of possible
+  values and keeping the value that produces the best performance. This option
+  can be slow if we are interested in finding the optimal value, but could be
+  appropriate if the variable has little impact on the algorithm's performance.
+\item Treat the variable as a variational parameter. This method will
+  efficiently optimise the value of the variable, but is vulnerable to
+  overfitting. Furthermore, this requires us to find the derivative of the ELBO
+  w.r.t. the variable, which could be difficult.
+\end{itemize}
+
+Titsias \cite{DBLP:journals/jmlr/Titsias09} is the first to suggest a
+variational approximation where inducing points are treated as variational
+parameters. The paper has influenced many of the later works covered in this
+section, and has been reformulated by Gal et al. \cite{DBLP:conf/nips/GalWR14}
+into a distributed algorithm. We will fix the inducing points for the baseline
+version of our algorithm, but keep the idea in mind for possible future work.
+
+Hensman et al. \cite{DBLP:conf/uai/HensmanFL13} reduce the complexity from
+$\mathcal{O}(nm^2)$ to $\mathcal{O}(m^3)$ using a stochastic variational
+inference (SVI) approach. While results such as efficiently computable natural
+gradients and analytically tractable optimal solutions for variational
+parameters are unlikely to transfer to a different problem domain, the overall
+SVI framework and an approximating Gaussian distribution for $\mathbf{u}$ will
+play an important role in our proposal.
+
+Cheng and Boots \cite{DBLP:conf/nips/ChengB17} suggest using different bases in
+the reproducing kernel Hilbert space for the mean and covariance functions.
+The paper makes few assumptions and mostly relies on the work of
+Titsias \cite{DBLP:journals/jmlr/Titsias09}, making it seem transferable to a
+new domain. However, applying the idea to IRL is likely to be beyond the scope
+of the project.
+
+The variational Fourier features (VFF) algorithm by Hensman et al.
+\cite{DBLP:journals/jmlr/HensmanDS17} is only defined for Mat\'ern kernels,
+which is likely to be too restrictive for our situation (the original paper on
+using GPs for IRL \cite{DBLP:conf/nips/LevinePK11} used the automatic relevance
+detection kernel that has weights controlling how important each feature is).
+While extending VFF to support a flexible class of kernels defined by Wilson and
+Adams \cite{pmlr-v28-wilson13} is an interesting and promising avenue of work,
+it is likely to be beyond the scope of this project as well.
+
+While Peng et al. \cite{DBLP:conf/icml/PengZZQ17} provide a highly efficient
+distributed implementation, the derivation of the ELBO relies primarily on the
+fact that the evidence for a GP is a Gaussian, while in our case the evidence is
+anything but. Hence, it is unlikely that the ideas from this paper could be
+applicable to our IRL model.
+
+Finally, since we expect $\posterior$ to be highly irregular, we would like our
+approximation to be capable of representing a wide range of possible probability
+distributions. The primary way of representing complex posteriors in VI is by
+using \emph{normalising flows}, i.e., a collection of invertible
+functions---parametrised by additional variational parameters---that are applied
+to latent variables \cite{DBLP:conf/icml/RezendeM15}. Normalising flows, along
+with some of the ideas mentioned previously, could provide great benefits to the
+GP IRL VI model. However, we are forced to keep any non-essential features as
+possible future work in order to make the project feasible in a given time
+frame. Furthermore, note that due to various differences between regression and
+IRL, we cannot simply apply a GP VI model as a whole. Instead, smaller tricks
+and ideas from various papers can (and will) be used throughout the project.
 
@@
 We show how to avoid the deterministic training conditional assumption.
+
+%Why doesn't it work? Perhaps a stupid bug.
 

### 2019-03-20 mpaper/paper.tex
Introduction!

@@
   approximates the full joint probability distribution of rewards across states
-  of the Markov decision process.
+  of the Markov decision process.%TODO: finish this.
+  According to Simon Peyton Jones, an abstract should address
+  four key questions. First, what is the problem that this
+  paper tackles? Second, why is this an interesting problem?
+  Third, what is the solution this paper proposes?
+  Finally, why is the proposed solution a good one?
 \end{abstract}
@@
 \section{Introduction} % first page
-% TODO: rewrite and make it good
 % 1. Problem (with examples, intuitive)
@@
 
-Inverse reinforcement learning (IRL)---a problem proposed by Russell in
-1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
-Markov decision process (MDP) that best explains a set of given demonstrations.
-IRL is important because reward functions can be hard to define manually
-\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
-are not entirely specific to a given environment, allowing one to reuse the same
-reward structure in previously unseen environments
+Imagine using a machine learning (ML) algorithm to teach a robot how to move
+around people so that it learns to predict where people are going and adjust its
+path accordingly. The ML algorithm would use data about various possible
+situations. But do we have enough data to ensure reasonably optimal behaviour?
+Perhaps the robot behaves well in most situations, but fails in less common
+scenarios. Can the ML model itself describe its weaknesses so that we could
+ensure it is exposed to sufficiently many uncommon or difficult situations?
+
+This learning problem
+\cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} as well as
+many others have benefited from an approach called \emph{inverse reinforcement
+  learning} (IRL) (also known as inverse optimal control). IRL proposes a way to
+learn behaviour from \emph{demonstrations} that typically come from human
+actions. More formally, the IRL problem asks us to find a reward function for a
+Markov decision process (MDP), where demonstrations are encoded as sets of
+state-action pairs.
+
+IRL is an important problem because adjusting the reward function by hand is
+often unwieldy, since human behaviour often depends on many factors in
+complicated ways
+\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}. Moreover,
+learning the reward function rather than the policy itself makes the model more
+transferable to new environments---a minor change in the environment can
+reorganise the whole policy but only have a local effect in the reward structure
 \cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
-Moreover, IRL has seen a wide array of applications in autonomous vehicle
-control \cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} and
-learning to predict another agent's behaviour
-\cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
-Most approaches in the literature make a convenient yet unjustified assumption
-that the reward function can be expressed as a linear combination of features.
-One proven way to abandon this assumption is by representing the reward function
-as a Gaussian process (GP)
-\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
-The original approach used maximum likelihood estimation
-\cite{DBLP:conf/nips/LevinePK11}, whereas we use variational inference (VI)
-instead, which learns approximate posterior probability distributions instead of
-point estimates. This approach can prove useful in three major ways:
+IRL has been used to teach helicopters how to perform tricks
+\cite{DBLP:conf/nips/AbbeelCQN06}, predict taxi destinations
+\cite{DBLP:conf/huc/ZiebartMDB08}, and make driving safer and more efficient by
+predicting pedestrian movement \cite{DBLP:conf/iros/ZiebartRGMPBHDS09} and the
+driver's intentions \cite{DBLP:conf/aaai/VogelRGR12}.
+
+% TODO: need to rewrite the first sentence
+However, most IRL models in the literature make a convenient yet unjustified
+assumption that the reward function can be expressed as a linear combination of
+features
+\cite{DBLP:conf/icml/PieterN04,DBLP:conf/uai/NeuS07,DBLP:conf/icml/NgR00,ziebart2008maximum}.
+This assumption makes the models unable to represent many reward structures. Out
+of the non-linear models proposed to date, none can answer the questions posed
+in the first paragraph. Quite often, the models assume that rewards have no
+variance \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}. In this paper,
+we show how that assumption can be lifted by switching from maximum likelihood
+estimation to \emph{variational inference} (VI), i.e., we approximate the
+posterior distribution of the model by optimising the parameters of a simpler
+distribution to make it similar to the posterior. This approach can prove useful
+in four major ways:
 \begin{enumerate}
-\item Modelling full posterior distributions for various parameters can result
-  in more precise reward predictions, as the model simply holds more
-  information.
-\item Having variance estimates for rewards can direct our choice in what data
-  should be collected next.
-\item An approximate Bayesian treatment of many parameters in the model guards
-  against overfitting \cite{DBLP:conf/uai/JinDAS17}.
+\item By working with full distributions instead of point estimates, we can
+  expect more precise reward predictions.
+\item Variance estimates can be used to guide what data should be collected
+  next, i.e., if the rewards of some states have abnormally high variance, we
+  might want to expose the model to more data visiting those and surrounding
+  states.
+\item Variances estimates can also be used to judge whether we can trust the
+  predictions of the model or, perhaps, the model could benefit from some
+  adjustments or more data.
+\item By adopting a more Bayesian approach, we automatically incorporate Occam's
+  razor into the model that guards against overfitting
+  \cite{DBLP:conf/uai/JinDAS17}.
 \end{enumerate}
+
+Our main contribution is a lengthy proof in Section \ref{sec:proof} that shows
+how VI can be applied to the maximum-entropy IRL model with Gaussian processes
+(GPs) proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. We describe
+how we adapt the model and set up the VI problem in Section \ref{sec:model}.
+Finally, in Section \ref{sec:evaluation} we examine the convergence properties
+of our model and its ability to deduce optimal policies in practice.
 
@@
 
-\section{The Model}
+\section{The Model} \label{sec:model}
 
@@
 
-\section{Theoretical Justification}
+\section{Theoretical Justification} \label{sec:proof}
 
@@
   problem and requires significantly more work to prove correctness.}. After
-showing that the theorem applies to our situation, we can express
-$\nabla\mathbb{E}[v]$ as
+showing that the theorem applies to our situation, we can estimate
+$\nabla\mathbb{E}[v]$ with
 \begin{align*}
@@
                       &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
-                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})},
-\end{align*}
-which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
-\mathbf{r}_s) \sim \approximation$.
+                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)]}{q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)},
+\end{align*}
+which can be computed by drawing $S$ samples $\{(\mathbf{u}_s,
+\mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.
 
@@
 
-\section{Evaluation} % showing deep understanding
+\section{Evaluation} \label{sec:evaluation}
+% showing deep understanding
 
@@
 
+% apprenticeship learning, imitation learning
+%a problem proposed by Russell in 1998 \cite{DBLP:conf/colt/Russell98}
+
 As mentioned in the introduction, most IRL algorithms assume that the reward
@@
 
-Table ?? further summarises some of the recent and/or
-influential GP approximation approaches that might be relevant to our situation.
 For many of the variables in our model (including the inducing points and the
@@
 
+% Show how we achieved our claimed contributions
+
 Reasonable results in convergence.
@@
 \subsection{Further Work}
+% How much does a reward have to change to change the policy? Could be an
+% interesting use case for variance predictions.
 

### 2019-03-21 mpaper/paper.tex
Relevant work -- quite short, but I only kept the interesting bits

@@
 complicated ways
-\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}. Moreover,
+\cite{DBLP:conf/icml/PieterN04}. Moreover,
 learning the reward function rather than the policy itself makes the model more
@@
 reorganise the whole policy but only have a local effect in the reward structure
-\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
 IRL has been used to teach helicopters how to perform tricks
@@
 features
-\cite{DBLP:conf/icml/PieterN04,DBLP:conf/uai/NeuS07,DBLP:conf/icml/NgR00,ziebart2008maximum}.
+\cite{DBLP:conf/icml/PieterN04,DBLP:conf/icml/NgR00,ziebart2008maximum}.
 This assumption makes the models unable to represent many reward structures. Out
@@
 \cite{DBLP:conf/nips/LevinePK11}, which is a version of the automatic relevance
-detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}:
+detection kernel \cite{DBLP:conf/nips/LevinePK11}:
 \begin{align*}
@@
 It remains to decide on the model for $\mathbf{u}$ and $\mathbf{r}$. As is
-commonly done when applying VI to GPs, we set
+commonly done when applying VI to GPs \cite{DBLP:conf/nips/ChengB17}, we set
 \begin{equation} \label{eq:approximation}
@@
 $q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma)$
-\cite{DBLP:conf/nips/ChengB17}.
 
@@
     \draw [->] (perturbation) edge (lemma3);
+    \draw [->] (perturbation) edge (lemma1);
     \draw [->] (lebesgue) edge (theorem);
@@
 Our main theorem is a specialised version of an integral differentiation result
-by Chen \cite{lecture_notes}.
+by Timoney \cite{lecture_notes}.
 \begin{theorem} \label{thm:main}
@@
 \section{Related Work} % 0.5-1 page
-% TODO: rewrite and make this section smaller
-
-% apprenticeship learning, imitation learning
-%a problem proposed by Russell in 1998 \cite{DBLP:conf/colt/Russell98}
-
-As mentioned in the introduction, most IRL algorithms assume that the reward
-function can be represented as a linear combination of features. This assumption
-originated in one of the earliest papers on the topic by Ng and Russell
-\cite{DBLP:conf/icml/NgR00}, which introduced several linear programming
-approaches to the problem. The authors also noticed that often
-multiple reward functions can explain the same behaviour, and suggested
-heuristics for reward functions that are `far away' from reward functions that
-do not fit the data.
-
-A few years later, Abbeel and Ng \cite{DBLP:conf/icml/PieterN04} developed an
-algorithm for the same formulation of the problem with a guarantee to converge
-quickly. Neu and Szepesv\'ari \cite{DBLP:conf/uai/NeuS07} identified a weakness
-in Abbeel and Ng's approach: the algorithm requires features to be
-`appropriately' scaled, and optimal scaling may not be known. Instead, they
-suggest a way to combine IRL with \emph{apprenticeship learning}, i.e., a
-supervised learning task for optimal \emph{policy} recovery (whereas IRL focuses
-on recovering the reward function).
-
-Ramachandran and Amir \cite{DBLP:conf/ijcai/RamachandranA07} were the first to
-formulate IRL in terms of Bayesian learning. While the model is easily
-interpretable and able to handle experts that make mistakes, the algorithm can
-only handle small state spaces and requires Monte Carlo Markov Chain (MCMC)
-sampling for inference.
-
-Ziebart et al. \cite{ziebart2008maximum} keep the linearity assumption, but
-introduce an influential idea: resolving the ambiguity when multiple reward
-functions explain the data by appealing to the maximum entropy principle.
-
-Choi and Kim \cite{DBLP:conf/ijcai/ChoiK13} extend the Bayesian model to learn
-good features as well as the reward function, trying to overcome the limitation
-of linearity. However, the approach is quite limiting: all features are assumed
-to have Boolean values, and the algorithm simply learns their conjunctions.
-
-Levine et al. \cite{DBLP:conf/nips/LevinePK11} are the first to suggest a way to
-learn nonlinear reward functions without harsh restrictions on the problem
-domain by using GPs. We base our work primarily on their paper, and the
-weaknesses we hope to address have already been covered in the previous section.
-A recent extension to their work by Jin et al. \cite{DBLP:conf/uai/JinDAS17}
-aims to harness the power of deep learning by using several layers of GPs,
-making the model less dependent on being provided good features. They also use
-VI, but with a few simplifying assumptions: deterministic training conditional
-for the reward vector, and fully independent training conditional for the latent
-state.
-
-Finally, instead of using GPs to model nonlinear reward functions, one can use a
-neural network (NN), as demonstrated by Wulfmeier et al.
-\cite{wulfmeier2015maximum}. Their approach benefits from constant time
-inference and the ability to learn complex features either from already-given
-features or even from raw data. The only disadvantage (as demonstrated in the
-paper) is that NNs converge slower than GPs (as measured by the number of
-demonstrations).
+
+The IRL problem itself was originally proposed by Russell in 1998
+\cite{DBLP:conf/colt/Russell98}. Most of the early approaches had the
+aforementioned reward linearity assumption. One of the first papers on the
+subject by Ng and Russell \cite{DBLP:conf/icml/NgR00} introduced several linear
+programming algorithms and identified an important issue: there are typically
+many reward functions that can explain the data equally well. This problem was
+solved by Ziebart et al. \cite{ziebart2008maximum} with the introduction of IRL
+based on the principles of maximum causal entropy in a linearly-solvable MDP.
+
+Levine et al. \cite{DBLP:conf/nips/LevinePK11} were the first to lift the
+linearity assumption without imposing additional restrictions on the problem.
+They do, however, model rewards as having no variance---our work removes this
+restriction without any compromises.
+
+Recently, Jin et al. \cite{DBLP:conf/uai/JinDAS17} have adapted the model
+proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11} to use deep GPs,
+harnessing the power of deep learning to make the model less dependent on what
+features are provided. Although they use VI, their approximating distribution
+for rewards at inducing points is simply the Dirac $\delta$ function, which is
+essentially equivalent to the assumption of no variance.
+
+An alternative to GPs for modelling nonlinear functions is, of course, neural
+networks. Wulfmeier et al. \cite{wulfmeier2015maximum} have shown how they can
+be used in the IRL setting. While this approach benefits from constant-time
+inference and the ability to learn complex features from data, neural
+networks often need significantly more data for the weights across all layers to
+stabilise.
 
 \subsection{Variational Inference}
-
-Variational inference has seen a recent increase in interest among academics,
-with different approaches focusing on different goals: better time complexity,
-handling a wider variety of models, making approximations more accurate, and
-using more complex function approximation techniques (such as NNs)
-to infer local latent variable values without having to calculate them
-individually for each data point \cite{DBLP:journals/corr/abs-1711-05597}. As
-our IRL model is based on a GP, we will begin by reviewing some of the VI
-approaches applied specifically to GP regression. Based on a recent review of
-scalable GPs \cite{DBLP:journals/corr/abs-1711-05597}, we will concentrate on
-\emph{stochastic variational sparse approximations}, as they have achieved
-modelling accuracy close to that of the full GP with no approximations, with
-many methods providing a time complexity of $\mathcal{O}(m^3)$. Below we provide
-a short overview of various assumptions that have been used in approximating
-\emph{sparse} GPs (i.e., GPs that use inducing points), following on a paper by
-Qui\~{n}onero-Candela and Rasmussen \cite{DBLP:journals/jmlr/CandelaR05}.
-
-For many of the variables in our model (including the inducing points and the
-hyperparameters of the covariance function), we have three possible ways of
-handling them:
-\begin{itemize}
-\item Provide a full Bayesian treatment by defining a prior probability
-  distribution. While this option can easily prevent overfitting, the choice of
-  prior can be difficult to justify.
-\item Fix the value of the variable. We can then optimise the value outside of
-  the main algorithm, e.g., by running the algorithm with a set of possible
-  values and keeping the value that produces the best performance. This option
-  can be slow if we are interested in finding the optimal value, but could be
-  appropriate if the variable has little impact on the algorithm's performance.
-\item Treat the variable as a variational parameter. This method will
-  efficiently optimise the value of the variable, but is vulnerable to
-  overfitting. Furthermore, this requires us to find the derivative of the ELBO
-  w.r.t. the variable, which could be difficult.
-\end{itemize}
-
-Titsias \cite{DBLP:journals/jmlr/Titsias09} is the first to suggest a
-variational approximation where inducing points are treated as variational
-parameters. The paper has influenced many of the later works covered in this
-section, and has been reformulated by Gal et al. \cite{DBLP:conf/nips/GalWR14}
-into a distributed algorithm. We will fix the inducing points for the baseline
-version of our algorithm, but keep the idea in mind for possible future work.
-
-Hensman et al. \cite{DBLP:conf/uai/HensmanFL13} reduce the complexity from
-$\mathcal{O}(nm^2)$ to $\mathcal{O}(m^3)$ using a stochastic variational
-inference (SVI) approach. While results such as efficiently computable natural
-gradients and analytically tractable optimal solutions for variational
-parameters are unlikely to transfer to a different problem domain, the overall
-SVI framework and an approximating Gaussian distribution for $\mathbf{u}$ will
-play an important role in our proposal.
-
-Cheng and Boots \cite{DBLP:conf/nips/ChengB17} suggest using different bases in
-the reproducing kernel Hilbert space for the mean and covariance functions.
-The paper makes few assumptions and mostly relies on the work of
-Titsias \cite{DBLP:journals/jmlr/Titsias09}, making it seem transferable to a
-new domain. However, applying the idea to IRL is likely to be beyond the scope
-of the project.
-
-The variational Fourier features (VFF) algorithm by Hensman et al.
-\cite{DBLP:journals/jmlr/HensmanDS17} is only defined for Mat\'ern kernels,
-which is likely to be too restrictive for our situation (the original paper on
-using GPs for IRL \cite{DBLP:conf/nips/LevinePK11} used the automatic relevance
-detection kernel that has weights controlling how important each feature is).
-While extending VFF to support a flexible class of kernels defined by Wilson and
-Adams \cite{pmlr-v28-wilson13} is an interesting and promising avenue of work,
-it is likely to be beyond the scope of this project as well.
-
-While Peng et al. \cite{DBLP:conf/icml/PengZZQ17} provide a highly efficient
-distributed implementation, the derivation of the ELBO relies primarily on the
-fact that the evidence for a GP is a Gaussian, while in our case the evidence is
-anything but. Hence, it is unlikely that the ideas from this paper could be
-applicable to our IRL model.
-
-Finally, since we expect $\posterior$ to be highly irregular, we would like our
-approximation to be capable of representing a wide range of possible probability
-distributions. The primary way of representing complex posteriors in VI is by
-using \emph{normalising flows}, i.e., a collection of invertible
-functions---parametrised by additional variational parameters---that are applied
-to latent variables \cite{DBLP:conf/icml/RezendeM15}. Normalising flows, along
-with some of the ideas mentioned previously, could provide great benefits to the
-GP IRL VI model. However, we are forced to keep any non-essential features as
-possible future work in order to make the project feasible in a given time
-frame. Furthermore, note that due to various differences between regression and
-IRL, we cannot simply apply a GP VI model as a whole. Instead, smaller tricks
-and ideas from various papers can (and will) be used throughout the project.
+% none of these approaches really work in our situation because they tend to
+% depend on properties of how GPs are used in regression that are no longer true
+% in IRL
+
+Since the focus of our work was on proving feasibility rather than ensuring
+performance, we simply used a combination of Gaussians for our variational
+approximation. However, while VI was initially focused on approximating
+distributions using simplistic models where all variables are independent
+\cite{blei2017variational}, the last few years have brought many advances in
+approximating more complex distributions and greatly reducing the computational
+complexity of the task. Adapting some of them to IRL is a nontrivial but highly
+valuable undertaking.
+
+For approximating complex distributions, Rezende and Mohamed
+\cite{DBLP:conf/icml/RezendeM15} suggest using \emph{normalising flows}, i.e., a
+collection of invertible functions---parametrised by additional variational
+parameters---that are applied to latent variables. A major challenge in applying
+their work to IRL is related to variational parameters being used to compute the
+MDP value function, i.e., how can we take the derivative of the value function
+with respect to such a variational parameter? Alternatively, perhaps one can
+construct a different model that would make this question moot.
+
+Another approach to flexibility in modelling could come from considering
+different GP kernels. For instance, Wilson and Adams \cite{pmlr-v28-wilson13}
+show how all \emph{stationary} (i.e., invariant to translations) kernels can be
+generated (or at least approximated) from a mixture of Gaussians in their
+spectral representation using Bochner's theorem
+\cite{bochner1959lectures,DBLP:journals/technometrics/Woodard00}. It looks
+promising to combine these kernels with the variational Fourier features
+approach by Hensman et al. \cite{DBLP:journals/jmlr/HensmanDS17} that leverages
+the same spectral representations for efficient VI.
 
@@
 \subsection{Further Work}
-% How much does a reward have to change to change the policy? Could be an
-% interesting use case for variance predictions.
 
@@
 \end{itemize}
+
+Finally, building on the idea that variance estimates can be used to judge
+whether the model has learned optimal policy, an interesting question for
+MDP (or, perhaps, dynamical systems) research would be: how much does a reward
+have to change in order to affect the deterministic policy? A simple answer to
+this question would allow us to use variance estimates in order to quantify the
+model's confidence regarding optimal behaviour.
 

### 2019-03-21 mpaper/paper.tex
The abstract

@@
 \begin{abstract}
-  We prove the validity of performing variational inference on the Gaussian
-  process-based inverse reinforcement learning model that preserves and
-  approximates the full joint probability distribution of rewards across states
-  of the Markov decision process.%TODO: finish this.
-  According to Simon Peyton Jones, an abstract should address
-  four key questions. First, what is the problem that this
-  paper tackles? Second, why is this an interesting problem?
-  Third, what is the solution this paper proposes?
-  Finally, why is the proposed solution a good one?
+  The inverse reinforcement learning (IRL) problem asks us to find a reward
+  function of a Markov decision process that explains observed behaviour. Many
+  approaches are only able to construct reward functions as linear combinations
+  of state features. Out of those that can handle nonlinearity, none can provide
+  a full posterior distribution of rewards. Providing variance estimates for
+  rewards would allow one to judge how well the model has learned its policy and
+  discover any weak spots the model may have. We show how to perform variational
+  inference (VI) on a Gaussian process-based IRL model in order to approximate
+  the posterior distribution of rewards. We prove the correctness of the
+  approach and demonstrate the model's behaviour in practice. Being able to
+  provide full posterior probability distributions in IRL unlocks many new
+  research frontiers ranging from integrating recent developments in VI to make
+  the models more efficient and flexible, to developing complex reinforcement
+  learning agents that can explicitly search for opportunities to fix their
+  weaknesses.
 \end{abstract}

### 2019-03-21 mpaper/paper.tex
The proof is now correct, complete, although could be cleaned up a bit

@@
 \section{Theoretical Justification} \label{sec:proof}
+% TODO: Possible improvements:
+% 1. Use limits more often.
+% 2. Perturbation lemma can be replaced by continuity of A |-> A^-1.
 
@@
 
-% TODO: need some editing
+% TODO: too long, lemma could be replaced by the continuity argument (and arrow
+% removed from the figure), needs more explanations. Otherwise quite tolerable.
 \begin{lemma} \label{lemma:bound1}
@@
 
-  First, we will show that
-  \[
-    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to 0 \quad \text{and}
-    \quad \left. \frac{\mathbf{\partial K}}{\partial \lambda_i}
-    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to 0
-  \]
-  as $\epsilon \to 0$.
-  We can easily establish constant upper and lower bounds on
-  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ by the boundedness of
-  $c$ and continuity of the covariance function. Similar reasoning combined with
-  the expressions for derivatives of covariance matrices in Lemma
-  \ref{lemma:derivatives} gives constant upper and lower bounds on the elements
-  of
-  \[
-    \left. \frac{\mathbf{\partial K}}{\partial \lambda_i} \right|_{\lambda_i =
+  % K and its derivative
+  First, we can easily deduce that
+  \begin{equation} \label{eq:convergence1}
+    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}
+  \end{equation}
+  and
+  \begin{equation} \label{eq:convergence2}
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \frac{\partial \mathbf{K}}{\partial \lambda_i}
+  \end{equation}
+  as $\epsilon \to 0$. For \eqref{eq:convergence1}, the result is obvious if $i
+  = 0$. Otherwise, it follows from the continuity of the exponential function.
+  For \eqref{eq:convergence2}, observe that if $i=0$, then
+  \[
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
       c(\mathbf{r}, \mathbf{u})}
   \]
-  as well.
+  as an expression has no $\lambda_0$ in it, so
+  \[
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_0} \right|_{\lambda_0 =
+      c(\mathbf{r}, \mathbf{u})} = \frac{\partial \mathbf{K}}{\partial
+      \lambda_0}.
+  \]
+  Finally, if $i>0$, then each element of 
+  \[
+    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i}
+    \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
+  \]
+  is a constant multiple of the corresponding element of
+  \[
+    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})},
+  \]
+  so the same reasoning as for \eqref{eq:convergence1} applies.
 
@@
   \end{align*}
-  as $\epsilon \to 0$.
-
-  We will now show the same result for $\lambda_i > 0$. Let
+  as $\epsilon \to 0$. For $i > 0$, let
   \[
@@
         \left( -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. \\
-      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right)
+      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right).
     \end{multlined}
   \end{align*}
-  Hence, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
+  Thus, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
   $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
-  \mathbf{E}$, where $\mathbf{E}$ is defined as
+  \mathbf{E}$, where $\mathbf{E}$ is defined by
   \[
@@
   \]
-  Then, since $\mathbf{K}$ is invertible, Lemma \ref{lemma:perturbation} shows
-  the existence of $\mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ and
-  gives upper and lower bounds on all of its elements.
+  Then, since $\mathbf{K}$ is invertible,
+  \[
+    \mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}^{-1}
+  \]
+  by continuity of $\mathbf{A} \mapsto \mathbf{A}^{-1}$.
 
@@
   \mathbf{Su})|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ has upper and lower
-  bounds in $\mathbb{R}_2[\mathbf{u}]$.
-
-  Since
+  bounds in $\mathbb{R}_2[\mathbf{u}]$. Furthermore, having convergence results
+  for arbitrary covariance matrices and their inverses means that
   \[
     \lim_{\epsilon \to 0} \bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
-    \bm\Gamma,
-  \]
-  we also have that
+    \bm\Gamma.
+  \]
+  Hence,
   \[
@@
   for small enough $\epsilon$, and, thus, $\det(\bm\Gamma)^{-1}|_{\lambda_i =
-    c(\mathbf{r}, \mathbf{u})}$ exists and has constant bounds.
+    c(\mathbf{r}, \mathbf{u})}$ exists and is bounded.
 
@@
 \end{proof}
-% TOOD: epsilon should not be treated as a constant. Revise the whole theory
-% section.
 

### 2019-03-22 mpaper/paper.tex
Cleaned up the second section

@@
 
+In this section, we introduce definitions and mathematical details relevant to
+the problem. We begin by formally defining what the problem is.
+
 \begin{definition}[MDP]
@@
   \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is a function
-  defined so that $\mathcal{T}(s, a, s')$ is the probability of moving to state $s'$
-  after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the discount
+  defined so that $\mathcal{T}(s, a, s')$ is the probability of moving to state
+  $s'$ after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the discount
   factor; and $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ is the reward
@@
 
-The optimal \emph{policy} $\pi : \mathcal{S} \to \mathcal{A}$ (i.e., a choice of
-actions for each state that maximises reward over time) is usually constructed
-by defining a \emph{value (utility) function} $\V : \mathcal{S} \to \mathbb{R}$
-that measures how good a state is based on the reward $\mathbf{r}$ as well as
-the structure of the MDP. One can then find $\V$ by applying the Bellman backup
-operator until convergence to every $s \in \mathcal{S}$ (the technique is known
-as \emph{value iteration}) \cite{DBLP:books/daglib/0023820}:
+The optimal (deterministic) \emph{policy} $\pi : \mathcal{S} \to \mathcal{A}$
+(i.e., a choice of actions for each state that maximises reward over time) is
+usually constructed by defining a \emph{value (utility) function} $\V :
+\mathcal{S} \to \mathbb{R}$ that measures how good a state is based on the
+reward $\mathbf{r}$ as well as the structure of the MDP. One can then find $\V$
+by applying the Bellman backup operator until convergence to every $s \in
+\mathcal{S}$ (the technique is known as \emph{value iteration})
+\cite{DBLP:books/daglib/0023820}:
 \[
@@
 \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, and use a
-\emph{linearly solvable} (or \emph{maximum causal entropy}) MDP with stochastic
-policy that define probability distributions over actions (instead of
-suggesting a single action for each state) \cite{ziebart2008maximum}. This type
-of MDP can be solved by applying the 'soft' version of the operator
+\emph{linearly solvable} (or \emph{maximum causal entropy}) MDP with a
+stochastic policy that defines probability distributions over actions (instead
+of suggesting a single action for each state) \cite{ziebart2008maximum}. This
+type of MDP can be solved by applying the `soft' version of the operator
 \cite{DBLP:conf/nips/LevinePK11,supplementary_material}:
@@
 \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}
-\begin{equation} \label{pDr}
-  \begin{split}
+\begin{align*}
   p(\mathcal{D} \mid \mathbf{r}) &= \prod_{i=1}^N \prod_{t=1}^T p(a_{i,t} \mid s_{i,t}) \\
                                  &= \exp\left( \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right),
-  \end{split}
-\end{equation}
+\end{align*}
 where
@@
 
-However, a reward function learned by maximising this likelihood is not
-transferable to new situations
-\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}. One needs to model the
-reward structure in a way that would allow reward predictions for previously
-unseen states.
-
-One way to model rewards without assumptions of linearity is with a
-\emph{Gaussian process} (GP). A GP is a collection of random variables, any
-finite combination of which has a joint Gaussian distribution
-\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0,
-k)$ to say that $r$ is a GP with mean $0$ and covariance function
-$k$. \emph{Covariance functions} (also known as \emph{kernels}) take two state
-feature vectors as input and quantify how similar the two states are, in a sense
-that we would expect them to have similar rewards.
-
-As training a GP with $n$ data points has a time complexity of
-$\mathcal{O}(n^3)$ \cite{DBLP:books/lib/RasmussenW06}, numerous approximation
-methods have been suggested, many of which select a subset of data called
-\emph{inducing points} and focus most of the training effort on them
-\cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be the matrix of
-features at inducing states, $\mathbf{u}$ the rewards at those states. Then the
-full joint probability distribution can be factorised as
+% TODO: rewrite the 'we write' sentence and the sentence before that
+As we want learned rewards to generalise to previously unseen states and
+for states to have a notion of similarity, the IRL definition also includes $d$
+features associated with each state. In this paper, we will focus on modelling
+rewards as a function from feature space to $\mathbb{R}$ using a \emph{Gaussian
+  process}. A GP is defined as a collection of random variables, any finite
+number of which has a joint Gaussian distribution
+\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0, k)$ to say
+that $r$ is a GP with mean $0$ and covariance function $k$. \emph{Covariance
+  functions} (also known as \emph{kernels}) take two state feature vectors as
+input and quantify how similar the two states are, in a sense that we would
+expect high covariance scores to be associated with similar rewards.
+
+A common way to scale GPs to larger data sets is by selecting $m$ points in the
+feature space--called \emph{inducing points}---and focus most of the training
+effort on them \cite{DBLP:journals/corr/abs-1807-01065}. Let $\mathbf{X_u}$ be
+the $m \times d$ matrix of features at inducing points, and let $\mathbf{u}$ be
+the rewards at those states. Then the full joint probability distribution can be
+factorised as
 \begin{equation} \label{eq:full}
@@
 \end{align*}
-is the GP prior \cite{DBLP:books/lib/RasmussenW06}, and $m \in \mathbb{N}$ is
-the number of inducing points. The GP posterior is then a multivariate Gaussian
-\cite{DBLP:conf/nips/LevinePK11}
+is the GP prior \cite{DBLP:books/lib/RasmussenW06}. The GP posterior is then a
+multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11} defined as
 \begin{equation} \label{eq:r}
   p(\mathbf{r} \mid \mathbf{u}) =
-  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru),
+  \mathcal{N}(\mathbf{r}; \Kru^\intercal\Kuu^{-1}\mathbf{u}, \Krr - \Kru^\intercal\Kuu^{-1}\Kru).
 \end{equation}
-and $p(\mathcal{D} \mid \mathbf{r})$ is as in \eqref{pDr}. The matrices such as
-$\Kru$ are called \emph{covariance matrices} and are defined as
+The matrices such as
+$\Kru$ are called \emph{covariance matrices} (also \emph{kernel matrices} and
+\emph{Gram matrices}) and are defined as
 $[\Kru]_{i,j} = k(\mathbf{x}_{\mathbf{r},i},
@@
 $\mathbf{x}_{\mathbf{u},j}$ denote feature vectors for the $i$th state in
-$\mathcal{S}$ and the $j$th state in $\mathbf{X_u}$, respectively
+$\mathcal{S}$ and the $j$th inducing point, respectively
 \cite{DBLP:conf/uai/JinDAS17}.
 
-Given this model, data $\mathcal{D}$, and inducing feature matrix
-$\mathbf{X_u}$, our goal is then to find optimal values of parameters
-$\bm\lambda$, inducing rewards $\mathbf{u}$, and the rewards for all relevant
-states $\mathbf{r}$. While the previous paper to consider this IRL model
-computed maximum likelihood (ML) estimates for $\bm\lambda$ and $\mathbf{u}$,
-and made an assumption that $\mathbf{r}$ in \eqref{eq:r} has zero variance
-\cite{DBLP:conf/nips/LevinePK11}, we aim to avoid this assumption and use
-VI to approximate the full posterior distribution $\posterior$.
-\emph{Variational inference} is an approximation technique for probability
-densities \cite{blei2017variational}. Let $\approximation$ be our approximating
-family of probability distributions for $\posterior$. Then the job of VI is to
-optimise the parameters of the approximating distribution in order to minimise
-the \emph{Kullback-Leibler} (KL) divergence between the original probability
-distribution and our approximation.  KL divergence (asymmetrically) measures how
-different the two distributions are, and can be defined as
-\cite{blei2017variational}
-\begin{align*}
-  \DKL{\approximation}{\posterior} &= \mathbb{E}[\log\approximation - \log\posterior ] \\
-                                   &= \mathbb{E}[\log\approximation - \log\pfull] \\
-                                   &+ \mathbb{E}[\log p(\mathcal{D})].
-\end{align*}
-The last term is both hard to compute and constant w.r.t. $\approximation$
-\cite{blei2017variational}, so we can remove it from our optimisation objective.
-The negation of what remains is often called the \emph{evidence lower bound}
-(ELBO) and is defined as\footnote{Throughout the proposal, all integrals should
-  be interpreted as definite integrals over the entire sample space.}
+Our goal is then to use VI in order to approximate $\posterior$.
+Let $\approximation$ denote our approximation. VI aims to optimise this
+approximation by minimising the \emph{Kullback-Leibler} (KL) divergence between
+the original probability distribution and our approximation. KL divergence
+(asymmetrically) measures the difference between two probability distributions
+and can be defined as \cite{blei2017variational}
+\begin{align*}
+  \DKL{\approximation}{\posterior} &= \mathbb{E}_{\approximation}[\log\approximation - \log\posterior ] \\
+                                   &= \mathbb{E}_{\approximation}[\log\approximation - \log\pfull] \\
+                                   &+ \mathbb{E}_{\approximation}[\log p(\mathcal{D})].
+\end{align*}
+The last term is both hard to compute and constant with respect to
+$\approximation$ \cite{blei2017variational}, so we can remove it from our
+optimisation objective. The negation of what remains is known as the
+\emph{evidence lower bound} (ELBO) and is defined as
 \cite{DBLP:books/lib/Bishop07,blei2017variational}
@@
 \end{equation}
-
-By considering full probability distributions instead of point estimates---as
-long as the approximations are able to capture important features of the
-posterior---our predictions are likely to be more accurate and rely on fewer
-assumptions. Moreover, we hope to make use of various recent advancements in VI
-for both time complexity and approximation distribution fit, making the
-resulting algorithm competitive both in terms of running time and model fit.
+Thus, instead of minimising KL divergence, we focus on maximising $\mathcal{L}$
+by optimising the values of parameters to be defined in Section \ref{sec:model}.
+Note that in this paper all integrals should be interpreted as definite
+integrals over the entire sample space. Also, all expected values are with
+respect to $\approximation$.
 
@@
 \section{The Model} \label{sec:model}
+% TODO: matrix notation should be at the start of The Problem
 

### 2019-03-22 mpaper/paper.tex
Everything up to the theory section should be quite readable

@@
 
-\section{The Problem}
+\paragraph{Notation and Conventions}
+
+For any matrix $\mathbf{A}$, we use either $A_{i,j}$ or
+$[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
+$j$. We use $\tr(\mathbf{A})$ to denote its \emph{trace} and $\adj(\mathbf{A})$
+for its \emph{adjugate} (or \emph{classical adjoint}). For any vector
+$\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a vector space of
+polynomials with degree at most $d$, where variables are elements of
+$\mathbf{x}$, and coefficients are in $\mathbb{R}$.
+
+Throughout the paper, all integrals should be interpreted as definite integrals
+over the entire sample space. When referencing measurability, we assume Lebesgue
+measure in a Euclidean space with a suitable number of dimensions. Similarly,
+whenever we consider the existence of an integral, we use the Lebesgue
+definition of integration.
+
+\section{The Problem} \label{sec:problem}
 
@@
   \begin{split}
-    \mathcal{L} &= \mathbb{E}\left[ \log \frac{\pfull}{\approximation} \right] \\
+    \mathcal{L} &= \mathbb{E}_{\approximation}\left[ \log \frac{\pfull}{\approximation} \right] \\
     &= \iiint \log \frac{\pfull}{\approximation} \approximation\dx.
@@
 by optimising the values of parameters to be defined in Section \ref{sec:model}.
-Note that in this paper all integrals should be interpreted as definite
-integrals over the entire sample space. Also, all expected values are with
-respect to $\approximation$.
+Note that hereafter we drop the subscript notation, as all expected values will
+be with respect to $\approximation$.
 
@@
 
-Here we introduce a few definitions and results from linear algebra, numerical
+We introduce a few definitions and results from linear algebra, numerical
 analysis, and measure theory that will be used later in the paper. Namely, we
 will use several different vector and matrix norms, consider how an inverse of a
-matrix changes with a small perturbation, and use Lebesgue's dominated
-convergence theorem in order to justify the validity of our approach.
+matrix changes with a small perturbation, and introduce Lebesgue's dominated
+convergence theorem which plays a major role in our theoretical results in
+Section \ref{sec:proof}.
 
@@
 \section{The Model} \label{sec:model}
-% TODO: matrix notation should be at the start of The Problem
-
-% TODO: make this fit the narrative
-For any matrix $\mathbf{A}$, we will use either $A_{i,j}$ or
-$[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
-$j$. Moreover, we use $\tr(\mathbf{A})$ to denote its \emph{trace} and
-$\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}). For
-any vector $\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a vector
-space of polynomials with degree at most $d$, where variables are elements of
-$\mathbf{x}$, and coefficients are in $\mathbb{R}$.
-
-In this paper, all references to measurability are with respect to the Lebesgue
-measure. Similarly, whenever we consider the existence of an integral, we use
-the Lebesgue definition of integration.
-
-We keep the covariance function the same as in the work by Levine et al.
-\cite{DBLP:conf/nips/LevinePK11}, which is a version of the automatic relevance
-detection kernel \cite{DBLP:conf/nips/LevinePK11}:
+
+In order to have a fully functional model, we still need to make several key
+design decisions. In this section, we describe the covariance function used to
+define the GP, consider what parameters should be used to optimise $\mathcal{L}$
+and how they should be initialised, and in Section \ref{sec:elbo}, we derive our
+final expression for $\mathcal{L}$.
+
+We stick with the same covariance function as in the work by Levine et al.
+\cite{DBLP:conf/nips/LevinePK11}, which is a version of the \emph{automatic
+  relevance detection kernel}:
 \begin{align*}
@@
 \end{align*}
-Here, $\lambda_0$ is the overall `scale' factor for how similar or distant the
-states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
-matrix that determines the relevance of each feature (where $d$ denotes the
-number of features), $\mathbbm{1}$ is defined as
+Here, $\lambda_0$ is the overall `scale' factor, $\bm\Lambda = \diag(\lambda_1,
+\dots, \lambda_d)$ is a diagonal matrix that determines the importance of each
+feature, $\mathbbm{1}$ is defined as
 \[
@@
 \]
-and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
-the value makes little difference to the performance of the algorithm
-\cite{DBLP:conf/nips/LevinePK11}). We will write $\bm\lambda = (\lambda_0,
-\dots, \lambda_d)^\intercal$ to refer to both $\lambda_0$ and $\bm\Lambda$ at
-the same time.
+and $\sigma^2$ is fixed as $10^{-2}/2$, since this value has little influence on
+the behaviour of the algorithm and is here as a noise factor used to avoid
+singular covariance matrices \cite{DBLP:conf/nips/LevinePK11}. We will write
+$\bm\lambda = (\lambda_0, \dots, \lambda_d)^\intercal$ to refer to both
+$\lambda_0$ and $\bm\Lambda$ at the same time.
 
 Ideally, we would like to model $\bm\lambda$ with an approximating distribution.
-However, due to how $p(\mathbf{u})$ has $\Kuu^{-1}$ in its expression, and the
-ELBO is defined as an expectation, we are unable to show that the ELBO is
-well-defined. More generally, we pose the following problem, which is open to
-the best of our knowledge:
+However, due to how the prior probability density function (PDF) for
+$\mathbf{u}$ involves $\Kuu^{-1}$, and $\mathcal{L}$ has an expected value that
+cannot be eliminated (see Section \ref{sec:elbo}), we are unable to show that
+such an $\mathcal{L}$ is well-defined. More generally, we pose the following
+problem, which, to the best of our knowledge, is currently open:
 \begin{problem}
@@
 \end{problem}
-While there are some obvious examples of when the required expected value
-exists (e.g., $f(X, A_{i,j}) = A_{i,j}X$ for an invertible $\mathbf{A}$ and many
-distributions of $X$), it would be particularly interesting to know whether the
-answer is `always'. A proof of such a result would allow us to model
-$\bm\lambda$ instead of treating it as a variational parameter, and would thus
-guard against overfitting. For now, $\bm\lambda$ will have to be treated as a
-variational parameter.
-
-It remains to decide on the model for $\mathbf{u}$ and $\mathbf{r}$. As is
-commonly done when applying VI to GPs \cite{DBLP:conf/nips/ChengB17}, we set
+While there are some obvious examples where the expected value exists (e.g.,
+$f(X, A_{i,j}) = A_{i,j}X$ for an invertible $\mathbf{A}$ and many distributions
+of $X$), it would be particularly interesting to know whether the answer is
+`always'. A proof of such a result would allow us to model $\bm\lambda$ instead
+of treating it as a variational parameter, and would help to guard against
+overfitting. For now, $\bm\lambda$ will have to be treated as a variational
+parameter.
+
+It remains to decide on the approximation distribution for $\mathbf{u}$ and
+$\mathbf{r}$. As is commonly done when applying VI to GPs
+\cite{DBLP:conf/nips/ChengB17}, we set
 \begin{equation} \label{eq:approximation}
@@
 where $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$ and
-$q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma)$
+$q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma)$.
 
@@
 is a diagonal matrix. Typically, we would set $p$ so that $p \ll m$ to get an
-efficient approximation. However, as our goal is precision rather than
-scalability, we will set $p = m$ and $\mathbf{D} = \mathbf{O}_m$ in order to
-retain full covariance structure.
+efficient approximation, but in this case we will simply set $p = m$ and
+$\mathbf{D} = \mathbf{O}_m$ in order to retain full covariance structure.
 
@@
 The resulting model is summarised in Figure \ref{fig:graphical_model}. We rely
-on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and the model.
+on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and our model.
 Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
@@
 the fact that the approximating distribution for $\mathbf{u}$ is
-$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$.
+$\mathcal{N}(\bm\mu, \mathbf{BB}^\intercal)$.
 
@@
 
-Finally, the parameters are initialised as follows:
+Finally, the parameters can be initialised as follows\footnote{In practice, it
+  is often easier to start with a fixed $\mathbf{B} = \mathbf{I}_m$, as random
+  values of $\mathbf{B}$ can often result in a near-singular $\bm\Sigma$.}:
 \begin{alignat*}{3}
@@
 The initialisation of $\bm\mu$ mirrors the initialisation of $\mathbf{r}$ in
-previous work by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. While they have
-constant initial values for $\bm\lambda$, we sample from $\chi^2$ distributions
-centred around those values ($5$ for $\lambda_0$ and $1$ for any other
-$\lambda_i$). The distributions for initial values of $\mathbf{B}$ are simply
-set to provide a reasonable spread of positive values for the diagonal, and both
-positive and negative values for all other entries in the matrix.
+previous work by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. In contrast,
+while they have constant initial values for $\bm\lambda$, we sample from
+$\chi^2$ distributions centred around those values ($5$ for $\lambda_0$ and $1$
+for any other $\lambda_i$). The distributions for initial values of $\mathbf{B}$
+are simply set to provide a reasonable spread of positive values for the
+diagonal, and both positive and negative values for all other entries in the
+matrix.
 
@@
 
-In this section, we derive and simplify the ELBO for this (now fully specified)
-model. Note that in order to keep the derivation simple, we drop all constant
-terms in the expression of $\mathcal{L}$, i.e., equality is taken to mean
-`equality up to an additive constant'. Also note that all expected values are
-with respect to $(\mathbf{u}, \mathbf{r}) \sim q(\mathbf{u}, \mathbf{r})$.
-
-In order to derive the ELBO, let us go back to \eqref{eq:elbo} and
-write
+It remains to express $\mathcal{L}$ for our (now fully specified) model. Note
+that in order to keep the derivation simple, we drop all constant terms in the
+expression of $\mathcal{L}$, i.e., equality is taken to mean `equality up to an
+additive constant'.
+
+Firstly, let us return to \eqref{eq:elbo} and write
 \[
@@
 $\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
-number of times the state associated with the reward $r_i$ has been visited
-across all demonstrations. Then
+number of times the state associated with reward $r_i$ has been visited
+across all demonstrations. Then,
 \begin{align*}
@@
 inside the integral. In order to prove it, we use a number of intermediate
-results. We start by stating a few derivatives of probability density functions
-(PDFs) and covariance matrices, and bound their values with some
-easy-to-deal-with polynomials. We then provide a sketch proof of the
-measurability of MDP value functions, which is non-obvious due to their
-non-trivial definition. Afterwards, we establish bounds for the value functions,
-and, after another quick lemma, tackle the main proof of this paper. See Figure
-\ref{fig:theorems} for an overview of how these results fit together.
+results. We start by stating a few derivatives of PDFs and covariance matrices,
+and bound their values with some easy-to-deal-with polynomials. We then provide
+a sketch proof of the measurability of MDP value functions, which is non-obvious
+due to their non-trivial definition. Afterwards, we establish bounds for the
+value functions, and, after another quick lemma, tackle the main proof of this
+paper. See Figure \ref{fig:theorems} for an overview of how these results fit
+together.
 
@@
 
-\section{Conclusions}
+\section{Conclusion}
 % TODO: write
-
 % Show how we achieved our claimed contributions
 
-Reasonable results in convergence.
-More variables/information. Fewer assumptions.
-
-We show how to avoid the deterministic training conditional assumption.
-
-%Why doesn't it work? Perhaps a stupid bug.
+\begin{itemize}
+\item Why my theoretical contributions are important...
+\item Reasonable results with convergence: policies do converge to optimal
+  values, even at the extremes (i.e., $0$ and $1$).
+\item Shown how to eliminate the deterministic training conditional
+  assumption---a common weakness of previous approaches.
+\item Being able to model rewards using full probability distributions without
+  limiting assumptions opens up many new research directions.
+\item My excuse for the buggy implementation...
+\end{itemize}
 

### 2019-03-23 mpaper/paper.tex
It's working (I think)!

@@
 % 2. Perturbation lemma can be replaced by continuity of A |-> A^-1.
+% 3. epsilon is treated as an arbitrary constant.
+
+The typical way to optimise a quantity (the ELBO, in this case) involves
+computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
+complicates the situation. The goal of this section is to show how Theorem
+\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
+anyway\footnote{This technique is inspired by black box VI
+  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
+  problem and requires significantly more work to prove correctness.}. After
+showing that the theorem applies to our situation, we can estimate
+$\nabla\mathbb{E}[v]$ with
+\begin{align*}
+  \nabla\mathbb{E}[v] &= \nabla \iint v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
+                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
+                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
+                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)]}{q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)},
+\end{align*}
+which can be computed by drawing $S$ samples $\{(\mathbf{u}_s,
+\mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.
 
@@
     \draw [->] (perturbation) edge (lemma3);
-    \draw [->] (perturbation) edge (lemma1);
     \draw [->] (lebesgue) edge (theorem);
@@
 
-The typical way to optimise a quantity (the ELBO, in this case) involves
-computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
-complicates the situation. The goal of this section is to show how Theorem
-\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
-anyway\footnote{This technique is inspired by black box VI
-  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
-  problem and requires significantly more work to prove correctness.}. After
-showing that the theorem applies to our situation, we can estimate
-$\nabla\mathbb{E}[v]$ with
-\begin{align*}
-  \nabla\mathbb{E}[v] &= \nabla \iint q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
-                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
-                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
-                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)]}{q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)},
-\end{align*}
-which can be computed by drawing $S$ samples $\{(\mathbf{u}_s,
-\mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.
-
 Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
@@
 and bound their values with some easy-to-deal-with polynomials. We then provide
-a sketch proof of the measurability of MDP value functions, which is non-obvious
-due to their non-trivial definition. Afterwards, we establish bounds for the
-value functions, and, after another quick lemma, tackle the main proof of this
-paper. See Figure \ref{fig:theorems} for an overview of how these results fit
-together.
+a sketch proof of the measurability of MDP value functions as well as new upper
+and lower bounds on their values. After another quick lemma, we tackle the main
+proof of this paper. See Figure \ref{fig:theorems} for an overview of how these
+results fit together.
 
@@
 
-% TODO: too long, lemma could be replaced by the continuity argument (and arrow
-% removed from the figure), needs more explanations. Otherwise quite tolerable.
 \begin{lemma} \label{lemma:bound1}
-  Let $i \in \{ 0, \dots, d \}$ and $\epsilon > 0$ be arbitrary. Furthermore,
-  let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i -
-  \epsilon, \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a
-  codomain arbitrarily close to $\lambda_i$. Then
+  Let $i \in \{ 0, \dots, d \}$ be arbitrary, and let $c :
+  \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i - \epsilon,
+  \lambda_i + \epsilon) \subset \mathbb{R}$ be a function with a codomain
+  arbitrarily close to $\lambda_i$. Then
   \[
@@
 \begin{proof}
-  Remember that
-  \begin{align*}
-    \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
-                                                                      &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
-  \end{align*}
-  by Lemma \ref{lemma:derivatives}. Let $\mathbf{K}$ be any covariance
-  matrix and
+  Let $\mathbf{K}$ be any covariance matrix and set
   \[
@@
   % K and its derivative
-  First, we can easily deduce that
+  First, we can easily deduce that\footnote{Note that we often switch between
+    proving limits and bounds. We are fundamentally interested in proving
+    constant bounds, but sometimes it is more convenient to show convergence
+    instead. In this situation, convergence as $\epsilon \to 0$ implies constant
+    upper and lower bounds.}
   \begin{equation} \label{eq:convergence1}
@@
   \]
-  as an expression has no $\lambda_0$ in it, so
+  as an expression contains no $\lambda_0$, so
   \[
@@
   \]
-  so the same reasoning as for \eqref{eq:convergence1} applies.
+  so the same reasoning applies as in case of \eqref{eq:convergence1}.
 
   % Inverse of a covariance matrix
-  Now, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
+  Next, we will show that $\mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
     \mathbf{u})}$ exists and
-  \[
+  \begin{equation} \label{eq:limit3}
     \lim_{\epsilon \to 0} \mathbf{K}^{-1}|_{\lambda_i = c(\mathbf{r},
-      \mathbf{u})} = \mathbf{K}.
-  \]
+      \mathbf{u})} = \mathbf{K}^{-1}.
+  \end{equation}
   If $i = 0$, then $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
   c(\mathbf{r}, \mathbf{u})\mathbf{A}$. Therefore\footnote{Note that since
-    $\lambda_0 \ne 0$, $c(\mathbf{r}, \mathbf{u}) \ne 0$ for small enough
+    $\lambda_0 > 0$, $c(\mathbf{r}, \mathbf{u}) \ne 0$ for small enough
     $\epsilon$.},
@@
   \end{align*}
-  as $\epsilon \to 0$. For $i > 0$, let
-  \[
-    S = \sum_{n \in \{ 1, \dots, d \} \setminus \{ i \}} \frac{\lambda_n}{2}
-    (x_{j,n} - x_{k,n})^2 + \mathbbm{1}[j \ne k]\sigma^2 \lambda_n
-  \]
-  and $\delta = c(\mathbf{r}, \mathbf{u}) - \lambda_i$ so that $c(\mathbf{r},
-  \mathbf{u}) = \lambda_i + \delta$, and $\lim_{\epsilon \to 0} \delta = 0$.
-  Then,
-  \begin{align*}
-    &k(\mathbf{x}_j, \mathbf{x}_k)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \!\begin{multlined}[t]
-      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} c(\mathbf{r}, \mathbf{u})(x_{j,i} - x_{k,i})^2\\
-      - \mathbbm{1}[j \ne k]\sigma^2c(\mathbf{r}, \mathbf{u}) - S \left. \vphantom{\frac{1}{2}} \right)
-    \end{multlined} \\
-    &= \!\begin{multlined}[t]
-      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2} (\lambda_i + \delta)(x_{j,i} - x_{k,i})^2 \\
-      - \mathbbm{1}[j \ne k]\sigma^2(\lambda_i + \delta) - S \left. \vphantom{\frac{1}{2}} \right)
-    \end{multlined} \\
-    &= \!\begin{multlined}[t]
-      \lambda_0 \exp \left( \vphantom{\frac{1}{2}} \right. -\frac{1}{2}(\mathbf{x}_j - \mathbf{x}_k)^\intercal \bm\Lambda (\mathbf{x}_j - \mathbf{x}_k) - \mathbbm{1}[j \ne k]\sigma^2\tr(\bm\Lambda) \\
-      - \frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right)
-    \end{multlined} \\
-    &= k(\mathbf{x}_j, \mathbf{x}_k)\exp \left( \vphantom{\frac{1}{2}} \right. -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2\delta \left. \vphantom{\frac{1}{2}} \right) \\
-    &= \!\begin{multlined}[t]
-      k(\mathbf{x}_j, \mathbf{x}_k) + k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp
-        \left( -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. \\
-      -\mathbbm{1}[j \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1 \right).
-    \end{multlined}
-  \end{align*}
-  Thus, we can express $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ as
-  $\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K} +
-  \mathbf{E}$, where $\mathbf{E}$ is defined by
-  \[
-    E_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( \exp \left(
-        -\frac{\delta}{2}(x_{j,i} - x_{k,i})^2 \right. \right. -\mathbbm{1}[j
-    \ne k]\sigma^2\delta \left. \left. \vphantom{\frac{1}{2}} \right) - 1
-    \right).
-  \]
-  By this definition,
-  \[
-    \lim_{\epsilon \to 0} E_{j,k} = 0.
-  \]
-  Then, since $\mathbf{K}$ is invertible,
-  \[
-    \mathbf{K}^{-1}_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}^{-1}
-  \]
-  by continuity of $\mathbf{A} \mapsto \mathbf{A}^{-1}$.
+  as $\epsilon \to 0$. For $i > 0$, by \eqref{eq:convergence1} and continuity of
+  $\mathbf{A} \mapsto \mathbf{A}^{-1}$ we immediate get \eqref{eq:limit3}.
 
@@
   for arbitrary covariance matrices and their inverses means that
-  \[
+  \begin{equation} \label{eq:gamma_convergence}
     \lim_{\epsilon \to 0} \bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
-    \bm\Gamma.
-  \]
-  Hence,
-  \[
+    \bm\Gamma
+  \end{equation}
+  and, by continuity of the determinant function,
+  \begin{equation} \label{eq:det_gamma}
     \lim_{\epsilon \to 0} \det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
     \det(\bm\Gamma).
-  \]
+  \end{equation}
   Assuming that
@@
 
-  Recall that
-  \[
-    \bm\Gamma = \Krr - \mathbf{S}\Kru = \Krr - \Kru^\intercal\Kuu^{-1}\Kru.
-  \]
-  We have already demonstrated that for $\mathbf{K} \in \{ \Krr, \Kru, \Kuu^{-1}
-  \}$,
-  \[
-    \lim_{\epsilon \to 0}\mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
-    \mathbf{K}.
-  \]
-  We can use a general fact that if $\lim_{x \to 0} f(x) = f$,
-  then $f(x) = f + g(x)$ for some function $g$ such that $\lim_{x \to 0} g(x) =
-  0$ to define matrices $\Err$, $\Eru$, and $\Euu$ such that
-  \begin{alignat*}{3}
-    \Krr|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Krr + \Err, &&\Err \to
-    \mathbf{O}_{|\mathcal{S}|}, \\
-    \Kru|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kru + \Eru, \quad
-    \text{and} \quad &&\Eru \to \mathbf{O}_{|\mathcal{S}|, m}, \\
-    \Kuu^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} &= \Kuu^{-1} + \Euu,
-    &&\Euu \to \mathbf{O}_{m}.
-  \end{alignat*}
-  as $\epsilon \to 0$. Then, $\bm\Gamma|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}
-  = \bm\Gamma + \mathbf{E}$, where
-  \begin{align*}
-    \mathbf{E} &= \Err - \Kru^\intercal\Kuu^{-1}\Eru - \Kru^\intercal\Euu(\Kru + \Eru) \\
-               &- \Eru^\intercal(\Kuu^{-1} + \Euu)(\Kru + \Eru) \\
-               &\to \mathbf{O}_{|\mathcal{S}|}
-  \end{align*}
-  as $\epsilon \to 0$. Thus, Lemma \ref{lemma:perturbation} shows that
-  $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$ exists and provides
-  constant bounds on its elements.
-
-  Since we already know that $\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r},
-    \mathbf{u})}$ and $\det(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$
-  are bounded, so is $\adj(\bm\Gamma) =
-  \det(\bm\Gamma)\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})}$. Thus,
-  we have constant bounds on
-  $|\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma))|_{\lambda_i = c(\mathbf{r},
-    \mathbf{u})}$, which means that
+  From \eqref{eq:gamma_convergence} and continuity of $\mathbf{A} \mapsto
+  \mathbf{A}^{-1}$ we can immediately deduce that
+  \begin{equation} \label{eq:gamma_inverse}
+    \lim_{\epsilon \to 0} \bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r},
+      \mathbf{u})} = \bm\Gamma^{-1}.
+  \end{equation}
+  Similarly, from \eqref{eq:det_gamma} and \eqref{eq:gamma_inverse} we get that
+  \[
+    \adj(\bm\Gamma)|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} =
+    \det(\bm\Gamma)\bm\Gamma^{-1}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to
+    \det(\bm\Gamma)\bm\Gamma^{-1}
+  \]
+  as $\epsilon \to 0$. This gives us constant bounds on
+  \[
+    |\bm\Gamma|^{-1}\tr(\mathbf{R}\adj(\bm\Gamma))|_{\lambda_i = c(\mathbf{r},
+      \mathbf{u})},
+  \]
+  which completes the proof that
   \[
@@
   \]
-  has the required bounds.
+  has the required quadratic bounds.
 \end{proof}
@@
 \begin{lemma} \label{lemma:bound3}
-  Let $i, j = 1, \dots, m$, and let $\epsilon > 0$ be arbitrary. Furthermore,
-  let
+  Let $i, j = 1, \dots, m$, and let
   \[
@@
     \frac{1}{2}q(\mathbf{u})(\mathbf{C}(\mathbf{r},
-    \mathbf{u})^{-\intercal}\mathbf{UC}(\mathbf{r}, \mathbf{u})^{-\intercal} -
-    \mathbf{C}(\mathbf{r}, \mathbf{u})^{-\intercal}),
+    \mathbf{u})^{-1}\mathbf{UC}(\mathbf{r}, \mathbf{u})^{-1} -
+    \mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}),
   \]
@@
   \]
-  We begin by establishing upper and lower bounds on $\mathbf{C}(\mathbf{r},
-  \mathbf{u})^{-1}$. For this, we use the maximum norm $\lVert \cdot
-  \rVert_\infty$ on both vectors and matrices. We can apply Lemma
-  \ref{lemma:perturbation} to $\bm\Sigma$ and $\mathbf{E}(\mathbf{r},
-  \mathbf{u})$ since
+  We will show how we can establish bounds on $\mathbf{C}(\mathbf{r},
+  \mathbf{u})^{-1}$ without using the continuity of matrix inversion. For this,
+  we use the maximum norm $\lVert \cdot \rVert_\infty$ on both vectors and
+  matrices. We can apply Lemma \ref{lemma:perturbation} to $\bm\Sigma$ and
+  $\mathbf{E}(\mathbf{r}, \mathbf{u})$ since
   \[
@@
     |[\mathbf{E}(\mathbf{r}, \mathbf{u})]_{k,l}| = |c(\mathbf{r}, \mathbf{u}) -
-    \Sigma_{i,j}| < \epsilon
-  \]
-  can be made arbitrarily small so that $\lVert \bm\Sigma^{-1} \rVert_\infty
+    \Sigma_{i,j}| < \epsilon \to 0,
+  \]
+  ensuring that $\lVert \bm\Sigma^{-1} \rVert_\infty
   \lVert \mathbf{E}(\mathbf{r}, \mathbf{u}) \rVert_\infty < 1$. Then
@@
       \rVert_\infty \lVert \mathbf{E}(\mathbf{r}, \mathbf{u}) \rVert_\infty} <
-    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
-      \rVert_\infty \epsilon},
+    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \epsilon \lVert
+      \bm\Sigma^{-1} \rVert_\infty},
   \]
@@
     \max_k \sum_l \left| [\mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}]_{k,l} \right|
-    < \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
-      \rVert_\infty \epsilon},
+    < \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \epsilon \lVert
+      \bm\Sigma^{-1} \rVert_\infty},
   \]
@@
     \left| [\mathbf{C}(\mathbf{r}, \mathbf{u})^{-1}]_{k,l} \right| <
-    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \lVert \bm\Sigma^{-1}
-      \rVert_\infty \epsilon},
+    \frac{\lVert \bm\Sigma^{-1} \rVert_\infty}{1 - \epsilon \lVert
+      \bm\Sigma^{-1} \rVert_\infty},
   \]
@@
 \begin{remark}
-  MDP values are characterised by both a state and a reward function/vector. In
-  this section, we think of the value function as $V : \mathcal{S} \to
-  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a state $s \in
-  \mathcal{S}$ and returns a function $V(s) : \mathbb{R}^{|\mathcal{S}|} \to
-  \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
-  \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$, $\V(s) \in
-  \mathbb{R}$. Given a reward vector, the function $V(s)$ computes the values of
-  all states and returns the value of state $s$.
+  MDP values are characterised by both a state and a reward function/vector. For
+  the rest of this section, we will think of the value function as $V :
+  \mathcal{S} \to \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a
+  state $s \in \mathcal{S}$ and returns a function $V(s) :
+  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ that takes a reward vector
+  $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state
+  $s$, $\V(s) \in \mathbb{R}$. Given a reward vector, the function $V(s)$
+  computes the values of all states and returns the value of state $s$.
 \end{remark}
@@
   For any reward vector $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$, the
-  collection of converged value functions $\{ \V(s) \mid s \in
+  set of converged value functions $\{ \V(s) \mid s \in
   \mathcal{S} \}$ satisfy
@@
   \mathcal{S}$, the value of $\V(s)$ is updated by applying
-  \eqref{eq:update_rule}. Note that both $\log$ and $\exp$ are increasing
-  functions, $\gamma > 0$, and the $\mathcal{T}$ function gives a probability (a
-  non-negative number). Thus
+  \eqref{eq:update_rule}. Note that both the natural logarithm and the
+  exponential function are increasing, $\gamma > 0$, and the $\mathcal{T}$
+  function gives a probability (a non-negative number). Thus,
   \begin{align*}
@@
   The MDP value function is measurable by Proposition \ref{thm:measurability}.
-  The result of multiplying or adding measurable functions (e.g., probability
-  density functions) to a measurable function is still measurable. Thus,
-  both $f$ and $f_n$ are measurable.
+  The result of multiplying or adding measurable functions (e.g., PDFs) to a
+  measurable function is still measurable. Thus, both $f$ and $f_n$ are
+  measurable.
 
@@
     |\fn| = |\V(s)| \left| \frac{q(\mathbf{r} \mid \mathbf{u})|_{t =
-          t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} \right| q(\mathbf{u})
+          t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} \right| q(\mathbf{u}),
   \]

### 2019-03-23 code/VIGPIRL/full_gradient.m

@@ -1,5 +1,6 @@
-function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, matrices)
-% NOTE: returns the gradient for parameters in vector form
+function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, matrices, deterministic_r)
+  % z is a sample from q(u), deterministic_r is used for testing
+  % NOTE: returns the gradient for parameters in vector form
 
   function answer = estimate_derivative(u)
    % Uses a point-sample of u to estimate the part of dL/dnu under E[]
@@ -23,8 +24,13 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
       elbo_part = solution.v(state, 1) - mdp_data.discount * s;
     end
 
-    U = (u - gp.mu) * (u - gp.mu)';
+    U = (u' - gp.mu) * (u' - gp.mu)';
+    %if (nargin('full_gradient') > 6)
+    %  fprintf('Deterministic mode is on!\n');
+    %  r = (S * u')';
+    %else
     r = mvnrnd(S * u', Gamma);
+    %end
     solution = linearmdpsolve(mdp_data, r');
     v_for_each_state_action = cellfun(@expected_derivative_for,...
       demonstrations, 'Uniform', 0);
@@ -33,8 +39,8 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
     unique_lambda_part = arrayfun(@lambda_derivative, 1:size(matrices.Kuu_grad, 3));
     unique_mu_part = (u' - gp.mu)' * (Sigma_inv + Sigma_inv');
-    %unique_B_part = 2 * (Sigma_inv * U * Sigma_inv - Sigma_inv) * gp.B;
-    unique_B_part = zeros(length(gp.mu));
+    unique_B_part = 2 * (Sigma_inv * U * Sigma_inv - Sigma_inv) * gp.B;
+    %unique_B_part = zeros(length(gp.mu));
 
     unique_part = vertcat(2, unique_lambda_part', diag(unique_B_part),...
       unique_mu_part', get_lower_triangle(unique_B_part));
[…]

### 2019-03-23 code/VIGPIRL/vigpirldefaultparams.m

@@ -10,10 +10,10 @@ default_params = struct(...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
-    'samples_count',1000,...
+    'samples_count',100,...
     'learning_rate',0.1,...
     'B_learning_rate',0.1,...
-    'num_iterations',300,...
+    'num_iterations',500,...
     'required_precision', 0.01);
 
 % Set parameters.

### 2019-03-23 code/VIGPIRL/vigpirlrun.m

@@ -296,8 +296,8 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     %grad(1) = 0; % lambda0
     %grad(2:d+1) = 0; % lambda (except first)
     %grad(d+m+2:d+2*m+1) = 0; % mu
-    grad(d+2:d+m+1) = 0; % B diagonal
-    grad(d+2*m+2:end) = 0; % rest of B
+    %grad(d+2:d+m+1) = 0; % B diagonal
+    %grad(d+2*m+2:end) = 0; % rest of B
 
     %fprintf('Hyperparameters:\n');
     %disp(old_hyperparameters);
@@ -355,32 +355,32 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   plot_history(policy_history);
   xlabel('number of iterations');
   legend('$\pi(a_1 \mid s_1)$', '$\pi(a_1 \mid s_2)$', '$\pi(a_2 \mid s_3)$', 'Interpreter', 'latex', 'Location', 'east');
-  print('../mpaper/convergence2', '-depsc2');
+  %print('../mpaper/convergence2', '-depsc2');
 
   figure('Units', 'centimeters', 'Position', [0 0 15 5], 'PaperPositionMode', 'auto');
-  %plot(hyperparameter_history(1,:), 'k-');
-  %hold on;
-  %plot(exp(hyperparameter_history(2,:)), 'k--');
-  %plot(hyperparameter_history(6,:), 'b');
-  %plot(hyperparameter_history(7,:), 'b--');
-  %plot(hyperparameter_history(8,:), 'b-.');
-  %plot(exp(hyperparameter_history(3,:)), 'r');
-  %plot(exp(hyperparameter_history(4,:)), 'r--');
-  %plot(exp(hyperparameter_history(5,:)), 'r-.');
-  %plot(hyperparameter_history(9,:), 'g');
-  %plot(hyperparameter_history(10,:), 'g--');
-  %plot(hyperparameter_history(11,:), 'g-.');
-  %legend('$\log \lambda_0$', '$\lambda_1$', '$\mu_1$', '$\mu_2$', '$\mu_3$', '$B_{1,1}$', '$B_{2,2}$', '$B_{3,3}$', '$B_{2,1}$', '$B_{3,1}$', '$B_{3,2}$', 'Interpreter', 'latex', 'Location', 'westoutside');
[…]

### 2019-03-23 code/singletest.m

@@ -36,5 +36,5 @@ mdp_data = struct('discount', 0.9, 'states', 3, 'actions', 2);
 mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
 mdp_data.sa_p(1:3, 1:2, 1) = 1;
 feature_data = struct('splittable', [1; 2; 3]);
-example_samples = {[1, 1], [3, 2]};
+example_samples = {[1, 1], [3, 2], [1, 1], [1, 1]};
 vigpirlrun(struct(), mdp_data, 'linearmdp', feature_data, example_samples);

### 2019-03-23 code/tests.m

@@ -2,38 +2,10 @@ function tests = tests
   tests = functiontests(localfunctions);
 end
 
-function test_d_covariance_matrix_d_lambda_i(testCase)
-  answer = d_covariance_matrix_d_lambda_i(ones(3, 2), [2 0; 4 0; 8 0],...
-    [1 0; 3 0], zeros(3, 2), 1);
-  correct = -0.5 * [1 1; 9 1; 49 25];
-  verifyEqual(testCase, answer, correct);
-end
-
-function test_remaining_vigpirlkernel(testCase)
+function test_full_gradient(testCase)
   function c = k(x, y)
     c = rbf * exp(-0.5 * 2 * (x - y)^2);
   end
-
-  %Xu = [0 0; 1 3];
-  %X = [0 0; 1 0; 2 0];
-  %noise = 0;
-  %rbf = 0.1;
-  %gp = struct('X_u', Xu, 'X', X, 'noise_var', noise, 'rbf_var', rbf,...
-  %  'inv_widths', [2 0]);
-
-  %Kuu = [k(0, 0) k(1, 0); k(0, 1) k(1, 1)];
-  %Kuf = [k(0, 0) k(1, 0) k(2, 0); k(1, 0) k(1, 1) k(1, 2)];
-  %Kff = [k(0, 0) k(1, 0) k(2, 0); k(0, 1) k(1, 1) k(2, 1); k(0, 2) k(1, 2) k(2, 2)];
-  %Kuuinv = inv(Kuu);
-  %KruKuu = Kuf' * Kuuinv;
-
-  % Test vigpirlkernel
-  %[K_uf, K_uu, invK, K_ufKinv, K_ff] = vigpirlkernel(gp);
-  %verifyEqual(testCase, K_uf, Kuf, 'AbsTol', 1e-10);
-  %verifyEqual(testCase, K_uu, Kuu, 'AbsTol', 1e-10);
-  %verifyEqual(testCase, invK, Kuuinv, 'AbsTol', 1e-10);
-  %verifyEqual(testCase, K_ufKinv, KruKuu, 'AbsTol', 1e-10);
-  %verifyEqual(testCase, K_ff, Kff, 'AbsTol', 1e-10);
 
   demonstrations = {[1, 1], [3, 2]};
   counts = [1; 0; 1];
@@ -58,13 +30,26 @@ function test_remaining_vigpirlkernel(testCase)
   r = S * u';
   mdp_solution = linearmdpsolve(mdp_data, r);
[…]

### 2019-03-23 mpaper/paper.tex
Fancy evaluation plots: version 1

@@
 \usepackage{booktabs}
-
-\usetikzlibrary{shapes,bayesnet,arrows.meta,arrows}
+\usepackage{csvsimple}
+\usepackage{pgfplots}
+
+\usetikzlibrary{shapes,bayesnet,arrows.meta,arrows,calc,pgfplots.colormaps}
 %\tikzstyle{myarrow} = [-{Latex[length=2mm,width=2mm]}]
@@
 
+\begin{figure}
+  \begin{tikzpicture}[
+    /pgfplots/colormap/violet,
+    ellipse2/.style={
+      draw,
+      ellipse,
+      /utils/exec={
+        \pgfplotscolormapdefinemappedcolor{#1}},
+      fill=mapped color},
+    edge2/.style={
+      ultra thick,
+      /utils/exec={
+        \pgfplotscolormapdefinemappedcolor{#1}},
+      draw=mapped color}]
+    \csvreader[
+    no head,
+    column count=11
+    ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
+      \node[ellipse2=1000] (s1) at (3*\x, 2*\y) {$s_1$};
+      \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
+      \node[ellipse2=\Sthreethree] (s3) at (3*\x+2, 2*\y) {$s_3$};
+      \draw[edge2=\Stwoone] (s1) -- (s2);
+      \draw[edge2=\Sthreeone] (s1) to (s3);
+      \draw[edge2=\Sthreetwo] (s2) to (s3);
+    }
+    \draw [<->] (2.5, 5.5) node (yaxis) [above] {$\zeta_2$}
+    |- (8.5, 1.5) node (xaxis) [right] {$\zeta_1$};
+    \foreach \x in {1, 2} {
+      \node (x\x) at (1+3*\x, 1.3) {$\x$};
+    }
+    \foreach \y in {1, 2} {
+      \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
+    }
+  \end{tikzpicture}
+\end{figure}
+
 \section{Related Work} % 0.5-1 page

### 2019-03-23 code/VIGPIRL/full_gradient.m

@@ -53,14 +53,28 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
   %disp(matrices.Kuu);
   %fprintf('Sigma:\n');
   %disp(Sigma);
-  Sigma_inv = inv(Sigma);
+
   Kuu_inv = inv(matrices.Kuu);
+  warning('');
+  Sigma_inv = inv(Sigma);
+  [warning_message, ~] = lastwarn;
+  if ~isempty(warning_message)
+    grad = [];
+    elbo = 0;
+    return;
+  end
+
   S = matrices.Kru' * Kuu_inv;
   % we need to round Gamma because otherwise it's not symmetric due to numerical errors
   Gamma = round(matrices.Krr - S * matrices.Kru, 4);
-  %disp(Gamma);
-  %disp(cond(Gamma));
+
   Gamma_inv = inv(Gamma);
+  [warning_message, ~] = lastwarn;
+  if ~isempty(warning_message)
+    grad = [];
+    elbo = 0;
+    return;
+  end
 
   grad_prediction_for_each_u = arrayfun(@(i) estimate_derivative(z(i, :)), 1:size(z, 1), 'Uniform', 0);
   estimated_grad = mean(cat(3, grad_prediction_for_each_u{:}), 3);

### 2019-03-23 code/VIGPIRL/vigpirldefaultparams.m

@@ -13,7 +13,7 @@ default_params = struct(...
     'samples_count',100,...
     'learning_rate',0.1,...
     'B_learning_rate',0.1,...
-    'num_iterations',500,...
+    'num_iterations',300,...
     'required_precision', 0.01);
 
 % Set parameters.

### 2019-03-23 code/VIGPIRL/vigpirlrun.m

@@ -56,7 +56,7 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     elbo_list = horzcat(elbo_list, elbo);
     hyperparameter_history = horzcat(hyperparameter_history, parameter_vector);
     grad_history = horzcat(grad_history, grad);
-    fprintf('elbo: %f\n', elbo);
+    %fprintf('elbo: %f\n', elbo);
 
     % We want to maximise the ELBO, while fmincon wants to minimize...
     grad = -grad;
@@ -286,10 +286,14 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   i = 0;
   tic;
   while true
+    fprintf('.');
     % Compute the gradient
     matrices = vigpirlkernel(gp);
     z = mvnrnd(gp.mu', gp.B * gp.B', algorithm_params.samples_count);
     [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, matrices);
+    if (isempty(grad))
+      break;
+    end
 
     old_hyperparameters = vigpirlpackparam(gp);
 
@@ -325,7 +329,7 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     %hyperparameters = old_hyperparameters + rho .* grad;
     gp = vigpirlunpackparam(gp, hyperparameters);
 
-    fprintf('ELBO: %f\n', elbo);
+    %fprintf('ELBO: %f\n', elbo);
     %fprintf('----------\n');
     elbo_list = horzcat(elbo_list, elbo);
     grad_history = horzcat(grad_history, grad);
@@ -343,34 +347,35 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
       break;
     end
   end
+  fprintf('\n');
   time = toc;
 
-  figure('Units', 'centimeters', 'Position', [0 0 15 20], 'PaperPositionMode', 'auto');
-  subplot(2, 1, 1);
-  stem(elbo_list);
-  xlabel('number of iterations');
-  ylabel('$\mathcal{L}$', 'Interpreter', 'latex');
+  %figure('Units', 'centimeters', 'Position', [0 0 15 20], 'PaperPositionMode', 'auto');
[…]

### 2019-03-23 mpaper/transform_csv.py

@@ -1,0 +1,14 @@
+import pandas
+
+def normalise(data):
+    max_value = data.transform(lambda x: abs(x)).max().max()
+    return data.transform(lambda x: 500 * x / max_value + 500)
+
+df = pandas.read_csv('covariance_and_policy.csv', header=None)
+indices = df.iloc[:,:2]
+diagonal = df.iloc[:,[2,6,10]]
+nondiagonal = df.iloc[:,[3,4,7]]
+policy = df.iloc[:,[11,12,16]]
+
+together = pandas.concat([indices, normalise(diagonal), normalise(nondiagonal), policy], axis=1)
+together.to_csv('covariance_and_policy_transformed.csv', header=None, index=False)

### 2019-03-24 mpaper/paper.tex
Fancy policy plots

@@
 
+% Makes the legend have a single rectangle per colour
+\pgfplotsset{
+  /pgfplots/ybar legend/.style={
+    /pgfplots/legend image code/.code={%
+      \draw[##1,/tikz/.cd,bar width=3pt,yshift=-0.2em,bar shift=0pt]
+      plot coordinates {(0cm,0.8em)};},
+  },
+}
+
 \newtheorem{theorem}{Theorem}[section]
@@
     \mathbf{u}, t)}}
+
+\newcommand{\pione}{\pi(a_1 \mid s_1)}
+\newcommand{\pitwo}{\pi(a_1 \mid s_2)}
+\newcommand{\pithree}{\pi(a_2 \mid s_3)}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \begin{tikzpicture}[
-    /pgfplots/colormap/violet,
+    /pgfplots/colormap/viridis,
     ellipse2/.style={
@@
     ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
-      \node[ellipse2=1000] (s1) at (3*\x, 2*\y) {$s_1$};
+      \node[ellipse2=\Soneone] (s1) at (3*\x, 2*\y) {$s_1$};
       \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
@@
     }
-    \draw [<->] (2.5, 5.5) node (yaxis) [above] {$\zeta_2$}
-    |- (8.5, 1.5) node (xaxis) [right] {$\zeta_1$};
-    \foreach \x in {1, 2} {
+    \draw [<->] (2.5, 7.5) node (yaxis) [above] {$\zeta_2$}
+    |- (11.5, 1.5) node (xaxis) [right] {$\zeta_1$};
+    \foreach \x in {1, 2, 3} {
       \node (x\x) at (1+3*\x, 1.3) {$\x$};
     }
-    \foreach \y in {1, 2} {
+    \foreach \y in {1, 2, 3} {
       \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
@@
   \end{tikzpicture}
-\end{figure}
+  \begin{tikzpicture}
+    \csvreader[
+    no head,
+    column count=11
+    ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
+      \begin{axis}[
+        ybar=0,
+        symbolic x coords={a,b,c},
+        ticks=none,
+        ymax=1,
+        ymin=0,
+        x=20,
+        bar width=20,
+        enlarge x limits=0.25,
+        at={(350*\x, 110*\y)},
+        height=100,
+        bar shift=0pt,
+        %legend pos=outer south,
+        legend style={
+          legend columns=-1,
+          at={(1.7,-0.4)}
+        }
+        ]
+        \addplot coordinates {(a, \ponetwo)};
+        \addplot coordinates {(b, \ptwoone)};
+        \addplot coordinates {(c, \pthreetwo)};
+        \draw[dashed] (-50,50) -- (270,50);
+        \ifthenelse{\x=2 \AND \y=1}{
+          \legend{$\pione$,$\pitwo$,$\pithree$};
+        }{}
+      \end{axis}
+    }
+    \draw [<->] (2.3, 8.5) node (yaxis) [above] {$\zeta_2$}
+    |- (10, 2) node (xaxis) [right] {$\zeta_1$};
+    \foreach \x in {1, 2, 3} {
+      \node (x\x) at (1+2.5*\x, 1.7) {$\x$};
+    }
+    \foreach \y in {1, 2, 3} {
+      \node (y\y) at (2.1, 1+2.1*\y) {$\y$};
+    }
+  \end{tikzpicture}
+\end{figure*}
 

### 2019-03-24 code/VIGPIRL/full_gradient.m

@@ -113,11 +113,11 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
   % If our rewards go too far towards negative infinity, and probability of a
   % state-action pair is over 1, freeze the reward of current state
-  elbo_data_part = counts' * S * gp.mu - 0.5 * estimated_grad(1);
-  if (elbo_data_part > 0)
-    fprintf('Should be negative: %f\n', elbo_data_part);
+  %elbo_data_part = counts' * S * gp.mu - 0.5 * estimated_grad(1);
+  %if (elbo_data_part > 0)
+    %fprintf('Should be negative: %f\n', elbo_data_part);
   %  elbo = elbo - elbo_data_part;
-  end
+  %end
 
   %fprintf('Not estimated grad:');
   %disp(not_estimated(2));

### 2019-03-24 code/singletest.m

@@ -41,7 +41,7 @@ feature_data = struct('splittable', [1; 2; 3]);
 %return;
 
 % Fancy stuff
-max_demonstrations_count = 5;
+max_demonstrations_count = 3;
 num_repeats = 10;
 data = [];
 for x = 1:max_demonstrations_count

### 2019-03-24 mpaper/transform_csv.py

@@ -1,8 +1,11 @@
 import pandas
+import numpy
 
 def normalise(data):
-    max_value = data.transform(lambda x: abs(x)).max().max()
-    return data.transform(lambda x: 500 * x / max_value + 500)
+    transformed = numpy.log(data)
+    max_value = transformed.transform(lambda x: abs(x)).max().max()
+    min_value = transformed.transform(lambda x: abs(x)).min().min()
+    return transformed.transform(lambda x: 500 * (x - min_value) / (max_value - min_value) + 500)
 
 df = pandas.read_csv('covariance_and_policy.csv', header=None)
 indices = df.iloc[:,:2]

### 2019-03-24 mpaper/paper.tex
A bit of evaluation

@@
 % 1. Use limits more often.
-% 2. Perturbation lemma can be replaced by continuity of A |-> A^-1.
-% 3. epsilon is treated as an arbitrary constant.
+% 2. Avoid saying that epsilon is arbitrary and using phrases like 'sufficiently small'.
+% TODO: Q should be lowercase
 
@@
     \centering
-    \includegraphics[width=0.8\columnwidth]{sage1}
-    \includegraphics[width=0.8\columnwidth]{sage2}
-  \end{minipage}%
-  \begin{minipage}{.5\textwidth}
-    \centering
-    \includegraphics[width=0.8\columnwidth]{elbo_over_B_short}
-  \end{minipage}
-  \caption{How a well-defined problem with a correct solutions becomes a
-    well-defined problem with an incorrect solution. In each plot, the function
-    we are trying to optimise is in blue, and its derivative is in red. The
-    plots on the left are for $q(\mathbf{u})$ and its derivatives with respect
-    to $B_{1,1}$ (at the top) and $B_{2,1}$ (at the bottom), whereas the plots
-    on the right are for $\mathbb{E}[v]$ and its corresponding derivatives.}
-  \label{fig:B_gradient}
-\end{figure*}
-
-\begin{figure*}
-  \centering
-  \begin{minipage}{.5\textwidth}
-    \centering
     \includegraphics[width=0.8\columnwidth]{convergence1}
@@
 
-Unfortunately, we are not able to use $\frac{\partial \mathcal{L}}{\partial
-  \mathbf{B}}$ in order to optimise $\mathbf{B}$. We illustrate the problem in
-Figure \ref{fig:B_gradient}. On the left side of the figure, we plot how
-$q(\mathbf{u})$ behaves as a function of a diagonal and a non-diagonal element
-of $\mathbf{B}$. Both functions have maximum values that can be attained by
-following the corresponding derivatives. However, when these derivatives are
-used to estimate $\frac{\partial \mathbb{E}[v]}{\partial \mathbf{B}}$, the
-resulting derivatives no longer match their corresponding functions, although
-the functions themselves still have optimal values: at or below $1$ for
-$B_{1,1}$ and at $0$ for $B_{2,1}$. This leads us to consider two restrictions
-of the model in order to investigate its convergence behaviour:
-\begin{itemize}
-\item In \textbf{Scenario 1}, we remove $\frac{\partial \mathbb{E}[v]}{\partial
-    \mathbf{B}}$ from $\frac{\partial \mathcal{L}}{\partial \mathbf{B}}$. This
-  essentially optimises $\mathbf{B}$ to match $\Kuu$, because then
-  $\frac{\partial \mathcal{L}}{\partial \mathbf{B}}$ becomes
-  $-\frac{\partial}{\partial \mathbf{B}} \DKL{q(\mathbf{u})}{p(\mathbf{u})}$,
-  i.e., we are optimising $\mathbf{B}$ to minimise the difference between the
-  prior and the posterior of $\mathbf{u}$.
-\item In \textbf{Scenario 2}, we set $\mathbf{B} = \mathbf{I}_m$, and do not
-  optimise it at all.
-\end{itemize}
-
 We plot how $\mathcal{L}$ as well as policies $\pi(a_1 \mid s_1)$, $\pi(a_2 \mid
@@
 the set of demonstrations $\mathcal{D}$, so we would expect to see their
-probabilities converge to values above $0.5$. Note that due to the stochastic
-nature of our MDP model, we do not expect to see any probability reach exactly
-$1$. The third policy, however, has no relevant data in $\mathcal{D}$, so the
-maximum causal entropy framework would put the probability at around $0.5$.
-
+probabilities converge to values near $1$. The third policy, however, has no
+data in $\mathcal{D}$ that would guide its value, so the maximum causal entropy
+framework would put the probability at around $0.5$.
+
+% TODO: rewrite this
 As we ran the algorithm with two stopping conditions---$300$ iterations and the
@@
 
+% TODO: rewrite
 Figure \ref{fig:convergence2} shows how the parameters of the model converge in
@@
 
-This leaves us with two models: one that converges to reasonable-but-suboptimal
-values, and one that diverges to infinite variance but also provides correct
-policies.
-
-\begin{figure*}
-  \begin{tikzpicture}[
-    /pgfplots/colormap/viridis,
-    ellipse2/.style={
-      draw,
-      ellipse,
-      /utils/exec={
-        \pgfplotscolormapdefinemappedcolor{#1}},
-      fill=mapped color},
-    edge2/.style={
-      ultra thick,
-      /utils/exec={
-        \pgfplotscolormapdefinemappedcolor{#1}},
-      draw=mapped color}]
-    \csvreader[
-    no head,
-    column count=11
-    ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
-      \node[ellipse2=\Soneone] (s1) at (3*\x, 2*\y) {$s_1$};
-      \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
-      \node[ellipse2=\Sthreethree] (s3) at (3*\x+2, 2*\y) {$s_3$};
-      \draw[edge2=\Stwoone] (s1) -- (s2);
-      \draw[edge2=\Sthreeone] (s1) to (s3);
-      \draw[edge2=\Sthreetwo] (s2) to (s3);
-    }
-    \draw [<->] (2.5, 7.5) node (yaxis) [above] {$\zeta_2$}
-    |- (11.5, 1.5) node (xaxis) [right] {$\zeta_1$};
-    \foreach \x in {1, 2, 3} {
-      \node (x\x) at (1+3*\x, 1.3) {$\x$};
-    }
-    \foreach \y in {1, 2, 3} {
-      \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
-    }
-  \end{tikzpicture}
+  % \begin{tikzpicture}[
+  %   /pgfplots/colormap/viridis,
+  %   ellipse2/.style={
+  %     draw,
+  %     ellipse,
+  %     /utils/exec={
+  %       \pgfplotscolormapdefinemappedcolor{#1}},
+  %     fill=mapped color},
+  %   edge2/.style={
+  %     ultra thick,
+  %     /utils/exec={
+  %       \pgfplotscolormapdefinemappedcolor{#1}},
+  %     draw=mapped color}]
+  %   \csvreader[
+  %   no head,
+  %   column count=11
+  %   ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
+  %     \node[ellipse2=\Soneone] (s1) at (3*\x, 2*\y) {$s_1$};
+  %     \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
+  %     \node[ellipse2=\Sthreethree] (s3) at (3*\x+2, 2*\y) {$s_3$};
+  %     \draw[edge2=\Stwoone] (s1) -- (s2);
+  %     \draw[edge2=\Sthreeone] (s1) to (s3);
+  %     \draw[edge2=\Sthreetwo] (s2) to (s3);
+  %   }
+  %   \draw [<->] (2.5, 7.5) node (yaxis) [above] {$\zeta_2$}
+  %   |- (11.5, 1.5) node (xaxis) [right] {$\zeta_1$};
+  %   \foreach \x in {1, 2, 3} {
+  %     \node (x\x) at (1+3*\x, 1.3) {$\x$};
+  %   }
+  %   \foreach \y in {1, 2, 3} {
+  %     \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
+  %   }
+  % \end{tikzpicture}
+
+\begin{figure}
+  \centering
   \begin{tikzpicture}
@@
         bar shift=0pt,
-        %legend pos=outer south,
         legend style={
@@
   \end{tikzpicture}
-\end{figure*}
-
+  \caption{How policy probabilities change after adding more demonstrations. The
+    $x$-axis ($y$-axis) shows how many copies of $\zeta_1$ ($\zeta_2$) were in
+    $\mathcal{D}$. Each plot shows the values of three policies after 300
+    iterations (averaged out over 10 runs). The bottom (top) of each bar plot
+    corresponds to probability 0 (1), and dashed lines mark probability 0.5.}
+  \label{fig:policies}
+\end{figure}
+
+We would expect stochastic probabilities to converge closer to extreme
+values (i.e., 0 and 1) as we add more data to the model. We test this hypothesis
+with a small-scale experiment in Figure \ref{fig:policies}. The policy $\pitwo$,
+in particular, confirms our prediction. As this policy denotes the probability
+of going from $s_2$ to $s_1$, we expect it to stay around $0.5$ as long as we
+have equal amounts of demonstrations $\zeta_1$ and $\zeta_2$ (this corresponds
+to the plots at coordinates $(1, 1)$, $(2, 2)$, and $(3, 3)$). But as we
+increase the number of times state $s_1$ appears in our demonstrations (going
+to the right across any row of plots), the model recognises the increasing value
+of state $s_1$, and the probability increases. In contrast, going up across the
+plots reduces the probability, as state $s_3$ becomes more valuable, and the
+model prefers it over state $s_1$. Interestingly, the other two policy
+probabilities seem to increase with any added data. This is likely due to the
+fact that more demonstrations result in higher gradient values, and thus lead to
+better-converged policies.
+
+%The experiments were run with a constant step size of 0.1, discarding iterations where either $\bm\Sigma$ or $\bm\Gamma$ becomes near-singular.
+
+%Diagonal and non-diagonal entries of the covariance matrix are normalised to
+%$[0, 1]$ separately, so as to emphasize the variation in each. There were no
+%negative entries of the covariance matrix after convergence.
+
+%More data creates bigger gradients---need smaller step size. Dynamic step sizes
+%don't work very well because they become zero after a big spike (and spikes do
+%happen).
 \section{Related Work} % 0.5-1 page
@@
 \subsection{Variational Inference}
-% none of these approaches really work in our situation because they tend to
-% depend on properties of how GPs are used in regression that are no longer true
-% in IRL
 
@@
   limiting assumptions opens up many new research directions.
-\item My excuse for the buggy implementation...
 \end{itemize}

### 2019-03-24 code/VIGPIRL/vigpirlrun.m

@@ -37,11 +37,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   % wrapper: vector of parameters -> scalar value * gradient vector
   function [elbo, grad] = wrapper(parameter_vector)
     gp = vigpirlunpackparam(gp, parameter_vector);
-
-    %disp(gp);
-    %disp(gp.B * gp.B');
-    %disp(gp.mu);
- 
     matrices = vigpirlkernel(gp);
     zz = mvnrnd(gp.mu', gp.B * gp.B', algorithm_params.samples_count);
     [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, zz, matrices);
@@ -303,11 +298,6 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
     %grad(d+2:d+m+1) = 0; % B diagonal
     %grad(d+2*m+2:end) = 0; % rest of B
 
-    %fprintf('Hyperparameters:\n');
-    %disp(old_hyperparameters);
-    %fprintf('Gradient:\n');
-    %disp(grad);
-
     hyperparameter_history = horzcat(hyperparameter_history, old_hyperparameters);
 
     % Make the derivative of B weaker
@@ -350,10 +340,10 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   fprintf('\n');
   time = toc;
 
-  %figure('Units', 'centimeters', 'Position', [0 0 15 20], 'PaperPositionMode', 'auto');
-  %subplot(2, 1, 1);
-  %stem(elbo_list);
-  %xlabel('number of iterations');
+  % figure('Units', 'centimeters', 'Position', [0 0 15 20], 'PaperPositionMode', 'auto');
+  % subplot(2, 1, 1);
+  % stem(elbo_list);
+  % xlabel('number of iterations');
   % ylabel('$\mathcal{L}$', 'Interpreter', 'latex');
 
   % subplot(2, 1, 2);
@@ -383,12 +373,11 @@ function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
   %plot(hyperparameter_history(7,:), 'b--');
   %plot(hyperparameter_history(8,:), 'b-.');
   %legend('$\lambda_0$', '$\log\lambda_1$', '$\mu_1$', '$\mu_2$', '$\mu_3$', 'Interpreter', 'latex', 'Location', 'westoutside');
[…]

### 2019-03-24 mpaper/transform_csv.py

@@ -5,7 +5,7 @@ def normalise(data):
     transformed = numpy.log(data)
     max_value = transformed.transform(lambda x: abs(x)).max().max()
     min_value = transformed.transform(lambda x: abs(x)).min().min()
-    return transformed.transform(lambda x: 500 * (x - min_value) / (max_value - min_value) + 500)
+    return transformed.transform(lambda x: 1000 * (x - min_value) / (max_value - min_value))
 
 df = pandas.read_csv('covariance_and_policy.csv', header=None)
 indices = df.iloc[:,:2]

### 2019-03-25 mpaper/paper.tex
New data + reorganising files

@@
     \centering
-    \includegraphics[width=0.8\columnwidth]{convergence1}
+    \includegraphics[width=0.8\columnwidth]{figures/convergence1}
   \end{minipage}%
@@
     \centering
-    \includegraphics[width=0.8\columnwidth]{convergence2}
+    \includegraphics[width=0.8\columnwidth]{figures/convergence2}
   \end{minipage}
@@
   \centering
-  \includegraphics[width=\textwidth]{parameter_convergence1}
-  \includegraphics[width=\textwidth]{parameter_convergence2}
+  \includegraphics[width=\textwidth]{figures/parameter_convergence1}
   \caption{Convergence of all optimised parameters for Scenario~1 at the top and
@@
 
-  % \begin{tikzpicture}[
-  %   /pgfplots/colormap/viridis,
-  %   ellipse2/.style={
-  %     draw,
-  %     ellipse,
-  %     /utils/exec={
-  %       \pgfplotscolormapdefinemappedcolor{#1}},
-  %     fill=mapped color},
-  %   edge2/.style={
-  %     ultra thick,
-  %     /utils/exec={
-  %       \pgfplotscolormapdefinemappedcolor{#1}},
-  %     draw=mapped color}]
-  %   \csvreader[
-  %   no head,
-  %   column count=11
-  %   ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
-  %     \node[ellipse2=\Soneone] (s1) at (3*\x, 2*\y) {$s_1$};
-  %     \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
-  %     \node[ellipse2=\Sthreethree] (s3) at (3*\x+2, 2*\y) {$s_3$};
-  %     \draw[edge2=\Stwoone] (s1) -- (s2);
-  %     \draw[edge2=\Sthreeone] (s1) to (s3);
-  %     \draw[edge2=\Sthreetwo] (s2) to (s3);
-  %   }
-  %   \draw [<->] (2.5, 7.5) node (yaxis) [above] {$\zeta_2$}
-  %   |- (11.5, 1.5) node (xaxis) [right] {$\zeta_1$};
-  %   \foreach \x in {1, 2, 3} {
-  %     \node (x\x) at (1+3*\x, 1.3) {$\x$};
-  %   }
-  %   \foreach \y in {1, 2, 3} {
-  %     \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
-  %   }
-  % \end{tikzpicture}
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[
+    /pgfplots/colormap/violet,
+    ellipse2/.style={
+      draw,
+      ellipse,
+      /utils/exec={
+        \pgfplotscolormapdefinemappedcolor{#1}},
+      fill=mapped color},
+    edge2/.style={
+      ultra thick,
+      /utils/exec={
+        \pgfplotscolormapdefinemappedcolor{#1}},
+      draw=mapped color}]
+    \csvreader[
+    no head,
+    column count=11
+    ]{data/median_covariance_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo}{%
+      \node[ellipse2=\Soneone] (s1) at (3*\x, 2*\y) {$s_1$};
+      \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
+      \node[ellipse2=\Sthreethree] (s3) at (3*\x+2, 2*\y) {$s_3$};
+      \draw[edge2=\Stwoone] (s1) -- (s2);
+      \draw[edge2=\Sthreeone] (s1) to (s3);
+      \draw[edge2=\Sthreetwo] (s2) to (s3);
+    }
+    \draw [<->] (2.5, 7.5) node (yaxis) [above] {$\zeta_2$}
+    |- (11.5, 1.5) node (xaxis) [right] {$\zeta_1$};
+    \foreach \x in {1, 2, 3} {
+      \node (x\x) at (1+3*\x, 1.3) {$\x$};
+    }
+    \foreach \y in {1, 2, 3} {
+      \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
+    }
+  \end{tikzpicture}
+  \end{figure}
 
@@
     column count=11
-    ]{covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
+    ]{data/covariance_and_policy_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo, 9=\ponetwo, 10=\ptwoone, 11=\pthreetwo}{%
       \begin{axis}[

### 2019-03-25 code/singletest.m

@@ -36,8 +36,8 @@ mdp_data = struct('discount', 0.9, 'states', 3, 'actions', 2);
 mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
 mdp_data.sa_p(1:3, 1:2, 1) = 1;
 feature_data = struct('splittable', [1; 2; 3]);
-%example_samples = {[1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2], [1, 1], [3, 2]};
-%wrapper(mdp_data, feature_data, example_samples);
+example_samples = {[1, 1], [3, 2]};
+wrapper(mdp_data, feature_data, example_samples);
 %vigpirlrun(struct(), mdp_data, 'linearmdp', feature_data, example_samples);
 %return;
 

### 2019-03-25 mpaper/transform_csv.py

@@ -1,17 +1,0 @@
-import pandas
-import numpy
-
-def normalise(data):
-    transformed = numpy.log(data)
-    max_value = transformed.transform(lambda x: abs(x)).max().max()
-    min_value = transformed.transform(lambda x: abs(x)).min().min()
-    return transformed.transform(lambda x: 1000 * (x - min_value) / (max_value - min_value))
-
-df = pandas.read_csv('covariance_and_policy.csv', header=None)
-indices = df.iloc[:,:2]
-diagonal = df.iloc[:,[2,6,10]]
-nondiagonal = df.iloc[:,[3,4,7]]
-policy = df.iloc[:,[11,12,16]]
-
-together = pandas.concat([indices, normalise(diagonal), normalise(nondiagonal), policy], axis=1)
-together.to_csv('covariance_and_policy_transformed.csv', header=None, index=False)

### 2019-03-27 mpaper/paper.tex
Some evaluation and a lot of slides

@@
 % 2. Avoid saying that epsilon is arbitrary and using phrases like 'sufficiently small'.
-% TODO: Q should be lowercase
 
@@
   \bm\Gamma &= \Krr - \mathbf{S}\Kru, \\
-  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru, \\
-  Q &= (\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} - \bm\mu).
+  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru.
 \end{align*}
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/convergence_new}
+  \caption{Convergence of $\mathcal{L}$ (at the top) as well as several example
+    policies (at the bottom).}
+  \label{fig:convergence1}
+\end{figure}
+
 \begin{figure*}
   \centering
-  \begin{minipage}{.5\textwidth}
-    \centering
-    \includegraphics[width=0.8\columnwidth]{figures/convergence1}
-  \end{minipage}%
-  \begin{minipage}{.5\textwidth}
-    \centering
-    \includegraphics[width=0.8\columnwidth]{figures/convergence2}
-  \end{minipage}
-  \caption{The convergence of $\mathcal{L}$ (at the top) and several example
-    policies (at the bottom) over a number of iterations for Scenario~1 on the
-    left and Scenario~2 on the right.}
-  \label{fig:convergence1}
-\end{figure*}
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\textwidth]{figures/parameter_convergence1}
-  \caption{Convergence of all optimised parameters for Scenario~1 at the top and
-  Scenario~2 at the bottom. In order to represent different variables on the
-  same scale, some variables have been $\log$-transformed. Colours denote which
-  vector or matrix each scalar comes from: black for $\bm\lambda$,
-  \textcolor{blue}{blue} for $\bm\mu$, \textcolor{red}{red} for diagonal
-  elements of $\mathbf{B}$, and \textcolor{green}{green} for its non-diagonal
-  elements.}
+  \includegraphics[width=\textwidth]{figures/parameter_convergence_new}
+  \caption{Convergence of variational parameters. In order to represent
+    different variables on the same scale, some variables have been
+    $\log$-transformed. Colours denote which vector or matrix each scalar comes
+    from: black for $\bm\lambda$, \textcolor{blue}{blue} for $\bm\mu$,
+    \textcolor{red}{red} for diagonal elements of $\mathbf{B}$, and
+    \textcolor{green}{green} for its non-diagonal elements.}
   \label{fig:convergence2}
@@
 probabilities converge to values near $1$. The third policy, however, has no
-data in $\mathcal{D}$ that would guide its value, so the maximum causal entropy
-framework would put the probability at around $0.5$.
-
-% TODO: rewrite this
-As we ran the algorithm with two stopping conditions---$300$ iterations and the
-$\ell_1$-norm of the change in parameter values being below $0.01$---note that
-the algorithm terminated early in Scenario~2 but not in Scenario~1, although
-there is clear convergence in both cases. An important difference is that
-$\pi(a_1 \mid s_1)$ and $\pi(a_2 \mid s_3)$ converge closer and closer to $1$
-with the slightly-restricted model of Scenario~1, but stabilise at $0.6$ in
-Scenario~2 due to an important part of the model being completely fixed. In both
-cases, $\pi(a_1 \mid s_2)$ converges to $0.5$ as expected.
-
-% TODO: rewrite
-Figure \ref{fig:convergence2} shows how the parameters of the model converge in
-both scenarios. The algorithm seems to converge fine in Scenario~2, but most of
-the parameters fail to stabilise in Scenario~1. Most importantly, the diagonal
-values of the $\mathbf{B}$ matrix diverge to positive infinity, leading to higher
-variance in ELBO estimates seen in Figure \ref{fig:convergence1}. The
-non-diagonal values, however, converge just fine. Both $\lambda_0$ and
-$\lambda_1$ also diverge to positive infinity, while the elements of $\bm\mu$
-diverge, but in `the right' direction: as $\mu_2$ increases while $\mu_1$ and
-$\mu_3$ decrease, the policies in Figure \ref{fig:convergence1} converge to
-their optimal values.
-
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[
-    /pgfplots/colormap/violet,
-    ellipse2/.style={
-      draw,
-      ellipse,
-      /utils/exec={
-        \pgfplotscolormapdefinemappedcolor{#1}},
-      fill=mapped color},
-    edge2/.style={
-      ultra thick,
-      /utils/exec={
-        \pgfplotscolormapdefinemappedcolor{#1}},
-      draw=mapped color}]
-    \csvreader[
-    no head,
-    column count=11
-    ]{data/median_covariance_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo}{%
-      \node[ellipse2=\Soneone] (s1) at (3*\x, 2*\y) {$s_1$};
-      \node[ellipse2=\Stwotwo] (s2) at (3*\x+1, 2*\y+1) {$s_2$};
-      \node[ellipse2=\Sthreethree] (s3) at (3*\x+2, 2*\y) {$s_3$};
-      \draw[edge2=\Stwoone] (s1) -- (s2);
-      \draw[edge2=\Sthreeone] (s1) to (s3);
-      \draw[edge2=\Sthreetwo] (s2) to (s3);
-    }
-    \draw [<->] (2.5, 7.5) node (yaxis) [above] {$\zeta_2$}
-    |- (11.5, 1.5) node (xaxis) [right] {$\zeta_1$};
-    \foreach \x in {1, 2, 3} {
-      \node (x\x) at (1+3*\x, 1.3) {$\x$};
-    }
-    \foreach \y in {1, 2, 3} {
-      \node (y\y) at (2.3, 0.5+2*\y) {$\y$};
-    }
-  \end{tikzpicture}
-  \end{figure}
+data in $\mathcal{D}$ to guide its value, so the maximum causal entropy
+framework would put the probability at around $0.5$. The first thing to note is
+that while $\mathcal{L}$ converges in just a few iterations, policies continue
+to improve for much longer. In addition, all policies behave as expected:
+$\pitwo$ stays around $0.5$, while both $\pione$ and $\pithree$ keep increasing.
+Although they stay at around $0.75$ (and not 1), this may be optimal for the
+maximum causal entropy model (and this question will be addressed in more detail
+with a different experiment).
+
+We also plot how the parameters of the model converge in Figure
+\ref{fig:convergence2}. While some parameters could benefit from a higher number of
+iterations\footnote{The limit of 300 iterations was chosen so that the initial
+  increase in $\mathcal{L}$ would not be overshadowed by a long straight line
+  afterwards.}, none look alarmingly wrong. Note that $\mu_2$, a variable
+closely related to $r(s_2)$, stabilises at a positive value whereas both $\mu_1$
+and $\mu_2$ become negative.
 
@@
   \end{tikzpicture}
-  \caption{How policy probabilities change after adding more demonstrations. The
-    $x$-axis ($y$-axis) shows how many copies of $\zeta_1$ ($\zeta_2$) were in
+  \caption{Changes in policies after adding more demonstrations. The $x$-axis
+    ($y$-axis) shows how many copies of $\zeta_1$ ($\zeta_2$) are in
     $\mathcal{D}$. Each plot shows the values of three policies after 300
     iterations (averaged out over 10 runs). The bottom (top) of each bar plot
-    corresponds to probability 0 (1), and dashed lines mark probability 0.5.}
+    corresponds to probability 0 (1), and dashed lines mark probability
+    0.5.}
   \label{fig:policies}
@@
 
-We would expect stochastic probabilities to converge closer to extreme
-values (i.e., 0 and 1) as we add more data to the model. We test this hypothesis
-with a small-scale experiment in Figure \ref{fig:policies}. The policy $\pitwo$,
-in particular, confirms our prediction. As this policy denotes the probability
-of going from $s_2$ to $s_1$, we expect it to stay around $0.5$ as long as we
-have equal amounts of demonstrations $\zeta_1$ and $\zeta_2$ (this corresponds
-to the plots at coordinates $(1, 1)$, $(2, 2)$, and $(3, 3)$). But as we
-increase the number of times state $s_1$ appears in our demonstrations (going
-to the right across any row of plots), the model recognises the increasing value
-of state $s_1$, and the probability increases. In contrast, going up across the
-plots reduces the probability, as state $s_3$ becomes more valuable, and the
-model prefers it over state $s_1$. Interestingly, the other two policy
-probabilities seem to increase with any added data. This is likely due to the
-fact that more demonstrations result in higher gradient values, and thus lead to
+We would expect policies to converge closer to extreme values (i.e., 0 and 1) as
+we add more data to the model. We test this hypothesis with a small-scale
+experiment in Figure \ref{fig:policies}. The policy $\pitwo$, in particular,
+confirms our prediction. As this policy denotes the probability of going from
+$s_2$ to $s_1$, we expect it to stay around $0.5$ as long as we have equal
+amounts of $\zeta_1$ and $\zeta_2$ (this corresponds to the plots at coordinates
+$(1, 1)$, $(2, 2)$, and $(3, 3)$). But as we increase the number of times state
+$s_1$ appears in our demonstrations (going to the right across any row of
+plots), the model recognises the increasing value of state $s_1$, and the
+probability increases. In contrast, going up across the plots reduces the
+probability, as state $s_3$ becomes more valuable, and the model prefers it over
+state $s_1$. Interestingly, the other two policy probabilities seem to increase
+regardless of what data is added. This is likely due to the fact that a higher
+number of demonstrations results in higher gradient values, and thus leads to
 better-converged policies.
 
-%The experiments were run with a constant step size of 0.1, discarding iterations where either $\bm\Sigma$ or $\bm\Gamma$ becomes near-singular.
-
+\begin{figure}[!ht]
+  \centering
+  \begin{tikzpicture}[
+    /pgfplots/colormap/violet,
+    ellipse2/.style={
+      draw,
+      ellipse,
+      /utils/exec={
+        \pgfplotscolormapdefinemappedcolor{#1}},
+      fill=mapped color},
+    edge2/.style={
+      ultra thick,
+      /utils/exec={
+        \pgfplotscolormapdefinemappedcolor{#1}},
+      draw=mapped color}]
+    \csvreader[
+    no head,
+    column count=11
+    ]{data/median_covariance_transformed.csv}{1=\x, 2=\y, 3=\Soneone, 4=\Stwotwo, 5=\Sthreethree, 6=\Stwoone, 7=\Sthreeone, 8=\Sthreetwo}{%
+      \node[ellipse2=\Soneone] (s1) at (2*\x, 2*\y) {$s_1$};
+      \node[ellipse2=\Stwotwo] (s2) at (2*\x+0.5, 2*\y+1) {$s_2$};
+      \node[ellipse2=\Sthreethree] (s3) at (2*\x+1, 2*\y) {$s_3$};
+      \draw[edge2=\Stwoone] (s1) -- (s2);
+      \draw[edge2=\Sthreeone] (s1) to (s3);
+      \draw[edge2=\Sthreetwo] (s2) to (s3);
+    }
+    \draw [<->] (1.5, 7.5) node (yaxis) [above] {$\zeta_2$}
+    |- (8, 1.5) node (xaxis) [right] {$\zeta_1$};
+    \foreach \x in {1, 2, 3} {
+      \node (x\x) at (0.5+2*\x, 1.3) {$\x$};
+    }
+    \foreach \y in {1, 2, 3} {
+      \node (y\y) at (1.3, 0.5+2*\y) {$\y$};
+    }
+    \begin{axis}[
+      colormap/violet,
+      scale only axis,
+      width=0pt,
+      height=0pt,
+      hide axis,
+      colorbar horizontal,
+      point meta min=-1,
+      point meta max=1,
+      xshift=40,
+      yshift=30,
+      colorbar style={
+        width=6.5cm,
+        xtick={-1,0,1},
+        xlabel=normalised covariance,
+      }]
+      \addplot [draw=none] coordinates {(0,0)};
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Changes in reward covariances after adding more demonstrations. The
+    meaning of external axes is the same as in Figure \ref{fig:policies}.
+    Colours denote reward covariance values, e.g., the colour of node $s_1$
+    denotes the variance of $r(s_1)$, and the colour of the edge between nodes
+    $s_1$ and $s_2$ denotes the covariance between $r(s_1)$ and $r(s_2)$. The
+    colours represent median covariances across 10 runs, normalised to the
+    interval $[-1, 1]$ while preserving their positivity/negativity.}
+  \label{fig:covariance}
+\end{figure}
+
+Finally, we would like to see similar changes in reward covariances. Firstly, we
+will show how to derive the covariance matrix for $\mathbf{r}$ from the
+posterior distributions of $\mathbf{u}$ and $\mathbf{r} \mid \mathbf{u}$.
+Remember that $\mathbf{u} \sim \mathcal{N}(\bm\mu, \bm\Sigma)$ and $\mathbf{r}
+\mid \mathbf{u} \sim \mathcal{N}(\mathbf{Su}, \bm\Gamma)$, and let $\mathbf{r}
+\sim \mathcal{N}(\bm\mu', \bm\Sigma')$, where our goal is to find $\bm\mu'$ and
+$\bm\Sigma'$. Then
+\[
+  \bm\mu' = \mathbb{E}[\mathbf{r}] = \mathbb{E}[\mathbf{Su}] = \mathbf{S}\bm\mu.
+\]
+Furthermore,
+\begin{align*}
+  \mathbb{E}_{\mathbf{r}}[\mathbf{rr}^\intercal] &= \bm\Gamma + \mathbb{E}_{\mathbf{u}}[\mathbf{Suu}^\intercal\mathbf{S}^\intercal] = \bm\Gamma + \mathbf{S}\mathbb{E}_{\mathbf{u}}[\mathbf{uu}^\intercal]\mathbf{S}^\intercal \\
+                                                 &= \bm\Gamma + \mathbf{S}(\bm\Sigma + \bm\mu\bm\mu^\intercal)\mathbf{S}^\intercal
+\end{align*}
+by the identity for $\mathbb{E}[\mathbf{xx}^\intercal]$ from the book by
+Petersen and Pedersen \cite{petersen2008matrix}. Now we can find the covariance
+matrix as follows:
+\begin{align*}
+  \bm\Sigma' &= \mathbb{E}[(\mathbf{r} - \mathbf{S}\bm\mu)(\mathbf{r} -
+               \mathbf{S}\bm\mu)^\intercal] \\
+             &= \mathbb{E}[\mathbf{rr}^\intercal - \mathbf{r}\bm\mu^\intercal\mathbf{S}^\intercal - \mathbf{S}\bm\mu\mathbf{r}^\intercal + \mathbf{S}\bm\mu\bm\mu^\intercal\mathbf{S}^\intercal] \\
+             &= \bm\Gamma + \mathbf{S}\bm\Sigma\mathbf{S}^\intercal.
+\end{align*}
+Therefore, $\mathbf{r} \sim \mathcal{N}(\mathbf{S}\bm\mu, \bm\Gamma +
+\mathbf{S}\bm\Sigma\mathbf{S}^\intercal)$.
+
+Secondly, we normalise the elements of this new covariance matrix into the
+interval $[-1, 1]$ for visualisation purposes. Diagonal (state) and non-diagonal
+(edge) covariances are normalised separately in order to represent the full
+range of values more clearly. All covariances are normalised according to this
+rule:
+\[
+  x \mapsto
+  \begin{cases*}
+    x/M & if $x \ge 0$ \\
+    -x/m & otherwise,
+  \end{cases*}
+\]
+where $M$ is the maximum (diagonal or non-diagonal) covariance across all data,
+and $m$ is the minimum. This way, all covariances are scaled to $[-1, 1]$, with
+both extreme points guaranteed to be reached unless all values are on one side
+of zero on the real number axis.
+
+Unfortunately, the resulting plot in Figure \ref{fig:covariance} shows no clear
+pattern. The only worthwhile observation is that there are only two instances of
+negative covariance, and both of them are between states $s_2$ and $s_3$.
+Perhaps there is a good (currently unknown) reason for why adding more data does
+not reduce variance, or perhaps this plot could benefit from a higher number of
+iterations and more repeated experiments to ensure the accuracy of the results.
+% or a different MDP, like a star
+
+% EXTRA INFO:
+% Discarding iterations where either $\bm\Sigma$ or $\bm\Gamma$ becomes near-singular.
 %Diagonal and non-diagonal entries of the covariance matrix are normalised to
@@
 %negative entries of the covariance matrix after convergence.
-
 %More data creates bigger gradients---need smaller step size. Dynamic step sizes
@@
 \item Why my theoretical contributions are important...
-\item Reasonable results with convergence: policies do converge to optimal
-  values, even at the extremes (i.e., $0$ and $1$).
+\item Reasonable results with policy convergence.
 \item Shown how to eliminate the deterministic training conditional
@@
   limiting assumptions opens up many new research directions.
+\item Determining how reward covariance depends on data would be the next step.
 \end{itemize}
@@
   \item
-    \[
-      \begin{split}
-        \frac{\partial q(\mathbf{u})}{\partial m} &=
-        q(\mathbf{u})\dm\left[-\frac{Q}{2}\right]
-        \\
-        &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
-        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} -
-        \bm\mu] \\
-        &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} +
-        \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
-      \end{split}
-    \]
+    \begin{align*}
+      \frac{\partial q(\mathbf{u})}{\partial m} &= q(\mathbf{u})\dm\left[-\frac{(\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} - \bm\mu)}{2}\right] \\
+                                                &= -\frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} + \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu)\dm[\mathbf{u} - \bm\mu] \\
+                                                &= \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} + \bm\Sigma^{-\intercal})(\mathbf{u} - \bm\mu).
+    \end{align*}
   \item An online tool by Laue et

### 2019-03-27 code/VIGPIRL/vigpirldefaultparams.m

@@ -12,8 +12,9 @@ default_params = struct(...
     'inducing_pts_count',16,...
     'samples_count',100,...
     'learning_rate',0.1,...
+    'lambda1_learning_rate',1,...
     'B_learning_rate',0.1,...
-    'num_iterations',300,...
+    'num_iterations',500,...
     'required_precision', 0.01);
 
 % Set parameters.

### 2019-03-27 code/singletest.m

@@ -39,7 +39,7 @@ feature_data = struct('splittable', [1; 2; 3]);
 example_samples = {[1, 1], [3, 2]};
 wrapper(mdp_data, feature_data, example_samples);
 %vigpirlrun(struct(), mdp_data, 'linearmdp', feature_data, example_samples);
-%return;
+return;
 
 % Fancy stuff 1
 %max_demonstrations_count = 3;

### 2019-03-30 code/VIGPIRL/full_gradient.m
Covariance experiment with cliques!

@@ -40,7 +40,6 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
     unique_lambda_part = arrayfun(@lambda_derivative, 1:size(matrices.Kuu_grad, 3));
     unique_mu_part = (u' - gp.mu)' * (Sigma_inv + Sigma_inv');
     unique_B_part = 2 * (Sigma_inv * U * Sigma_inv - Sigma_inv) * gp.B;
-    %unique_B_part = zeros(length(gp.mu));
 
     unique_part = vertcat(2, unique_lambda_part', diag(unique_B_part),...
       unique_mu_part', get_lower_triangle(unique_B_part));
@@ -97,11 +96,11 @@ function [elbo, grad] = full_gradient(mdp_data, demonstrations, counts, gp, z, m
 
   % TESTING
   %fprintf('Not estimated: %f, estimated: %f, total: %f\n', not_estimated(2), estimated_grad(2), changes(2));
-  p_derivative = -0.5 * estimated_grad(2) + counts' * (matrices.Kru_grad(:, :, 1)' -...
-    matrices.Kru' * Kuu_inv * matrices.Kuu_grad(:, :, 1)) * Kuu_inv * gp.mu;
-  kl_derivative = 0.5 * (trace(Kuu_inv * matrices.Kuu_grad(:, :, 1) * Kuu_inv * Sigma) +...
-    gp.mu' * Kuu_inv * matrices.Kuu_grad(:, :, 1) * Kuu_inv * gp.mu -...
-    trace(Kuu_inv * matrices.Kuu_grad(:, :, 1)));
+  % p_derivative = -0.5 * estimated_grad(2) + counts' * (matrices.Kru_grad(:, :, 1)' -...
+  %   matrices.Kru' * Kuu_inv * matrices.Kuu_grad(:, :, 1)) * Kuu_inv * gp.mu;
+  % kl_derivative = 0.5 * (trace(Kuu_inv * matrices.Kuu_grad(:, :, 1) * Kuu_inv * Sigma) +...
+  %   gp.mu' * Kuu_inv * matrices.Kuu_grad(:, :, 1) * Kuu_inv * gp.mu -...
+  %   trace(Kuu_inv * matrices.Kuu_grad(:, :, 1)));
   %fprintf('d/d_lambda0 of E[log p(D | r)]: %f\n', p_derivative);
   %fprintf('average v: %f\n', 0.5 * estimated_grad(1));
[…]

### 2019-03-30 code/VIGPIRL/vigpirldefaultparams.m

@@ -5,16 +5,16 @@ function algorithm_params = vigpirldefaultparams(algorithm_params)
 default_params = struct(...
     ...% These are initial values.
     'ard_init',1,...
-    'noise_init',0.01,...
+    'noise_init',0.1,...
     'rbf_init',5,...
     ...% These parameters control how the inducing points are selected.
     'inducing_pts','examplesplus',...
     'inducing_pts_count',16,...
     'samples_count',100,...
-    'learning_rate',0.1,...
-    'lambda1_learning_rate',1,...
-    'B_learning_rate',0.1,...
-    'num_iterations',500,...
+    'learning_rate',0.001,...
+    'lambda1_learning_rate',0.001,...
+    'B_learning_rate',0.001,...
+    'num_iterations',100,...
     'required_precision', 0.01);
 
 % Set parameters.

### 2019-03-30 mpaper/paper.tex
Covariance experiment with cliques!
Describing the new plots and a few additions to the presentation

@@
 
-\section{Evaluation} \label{sec:evaluation}
+\section{Experiments} \label{sec:evaluation}
 % showing deep understanding
 
-In order to fully understand the model's behaviour, we focus on a three-state
+In order to fully understand the model's behaviour, we start with a three-state
 MDP where the agent can deterministically move from any state to any other
@@
 rewards in order to reflect this.
+
+\paragraph{Convergence}
 
@@
 s_3)$, and $\pi(a_1 \mid s_2)$ converge over a number of iterations in Figure
-\ref{fig:convergence1}. The first two policies correspond to actions taken in
-the set of demonstrations $\mathcal{D}$, so we would expect to see their
-probabilities converge to values near $1$. The third policy, however, has no
-data in $\mathcal{D}$ to guide its value, so the maximum causal entropy
-framework would put the probability at around $0.5$. The first thing to note is
-that while $\mathcal{L}$ converges in just a few iterations, policies continue
-to improve for much longer. In addition, all policies behave as expected:
-$\pitwo$ stays around $0.5$, while both $\pione$ and $\pithree$ keep increasing.
-Although they stay at around $0.75$ (and not 1), this may be optimal for the
-maximum causal entropy model (and this question will be addressed in more detail
-with a different experiment).
+\ref{fig:convergence1}\footnote{In all experiments, we discard runs that
+  result in a near-singular $\bm\Sigma$ or $\bm\Gamma$, repeatedly testing each
+  combination of parameters as many times as necessary.}. The first two policies
+correspond to actions taken in the set of demonstrations $\mathcal{D}$, so we
+would expect to see their probabilities converge to values near $1$. The third
+policy, however, has no data in $\mathcal{D}$ to guide its value, so the maximum
+causal entropy framework would put the probability at around $0.5$. The first
+thing to note is that while $\mathcal{L}$ converges in just a few iterations,
+policies continue to improve for much longer. In addition, all policies behave
+as expected: $\pitwo$ stays around $0.5$, while both $\pione$ and $\pithree$
+keep increasing. Although they stay at around $0.75$ (and not 1), this may be
+optimal for the maximum causal entropy model (and this question will be
+addressed in more detail with a different experiment).
 
@@
 and $\mu_2$ become negative.
+
+\paragraph{Adding More Data}
 
@@
     $\mathcal{D}$. Each plot shows the values of three policies after 300
-    iterations (averaged out over 10 runs). The bottom (top) of each bar plot
+    iterations (averaged out over ten runs). The bottom (top) of each bar plot
     corresponds to probability 0 (1), and dashed lines mark probability
@@
       draw,
-      ellipse,
+      circle,
       /utils/exec={
@@
     $s_1$ and $s_2$ denotes the covariance between $r(s_1)$ and $r(s_2)$. The
-    colours represent median covariances across 10 runs, normalised to the
+    colours represent median covariances across ten runs, normalised to the
     interval $[-1, 1]$ while preserving their positivity/negativity.}
@@
 Unfortunately, the resulting plot in Figure \ref{fig:covariance} shows no clear
-pattern. The only worthwhile observation is that there are only two instances of
-negative covariance, and both of them are between states $s_2$ and $s_3$.
-Perhaps there is a good (currently unknown) reason for why adding more data does
-not reduce variance, or perhaps this plot could benefit from a higher number of
-iterations and more repeated experiments to ensure the accuracy of the results.
-% or a different MDP, like a star
-
-% EXTRA INFO:
-% Discarding iterations where either $\bm\Sigma$ or $\bm\Gamma$ becomes near-singular.
-%Diagonal and non-diagonal entries of the covariance matrix are normalised to
-%$[0, 1]$ separately, so as to emphasize the variation in each. There were no
-%negative entries of the covariance matrix after convergence.
-%More data creates bigger gradients---need smaller step size. Dynamic step sizes
-%don't work very well because they become zero after a big spike (and spikes do
-%happen).
+pattern, suggesting that adding more of the same demonstrations does not result
+in lower covariances. The only other observation from this plot is that there
+are only two instances of negative covariance, and both of them are between
+states $s_2$ and $s_3$.
+
+\paragraph{Structure versus Randomness}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \foreach \i in {0,...,9}
+    \node[draw,circle] (v\i) at (360/10 * \i:3cm) {$s_{\i}$};
+    \foreach \i in {0,...,9}
+    \foreach \j in {0,...,\i}
+    \draw[gray] (v\i) -- (v\j);
+
+    \draw[->,thick,blue] (v9) -- (v2);
+    \draw[->,thick,blue] (v7) -- (v8);
+    \draw[->,thick,blue] (v9) -- (v4);
+    \draw[->,thick,blue] (v3) -- (v0);
+    \draw[->,thick,blue] (v2) -- (v4);
+    \draw[->,thick,blue] (v4) -- (v6);
+    \draw[->,thick,blue] (v6) -- (v5);
+
+    \begin{scope}[yshift=-7cm]
+    \foreach \i in {0,...,9}
+    \node[draw,circle] (v\i) at (360/10 * \i:3cm) {$s_{\i}$};
+    \foreach \i in {0,...,9}
+    \foreach \j in {0,...,\i}
+    \draw[gray] (v\i) -- (v\j);
+
+    \draw[->,thick,blue] (v3) -- (v0);
+    \draw[->,thick,blue] (v4) -- (v0);
+    \draw[->,thick,blue] (v6) -- (v0);
+    \draw[->,thick,blue] (v2) -- (v0);
+    \draw[->,thick,blue] (v5) -- (v0);
+    \draw[->,thick,blue] (v8) -- (v0);
+    \draw[->,thick,blue] (v9) -- (v0);
+    \end{scope}
+  \end{tikzpicture}
+  \caption{The clique MDP with two sets of demonstrations denoted by blue arrows.}
+  \label{fig:cliques}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figures/boxplots}
+  \caption{Box plots of the difference between (absolute values of) reward
+    covariances in random and semi-structured scenarios from Figure
+    \ref{fig:cliques}, averaged out over 100 runs and grouped into four key
+    categories. `First state' refers to the variance of $r(s_0)$, the reward of
+    the state targeted by all demonstrations in the semi-structured case. `Other
+    states' are variances of other states of the MDP, i.e., diagonal entries of
+    the covariance matrix. `Incident edges' refers to covariances between
+    $r(s_0)$ and all other states, i.e., the first row or column of the
+    covariance matrix. Finally, `other edges' is the category for the remaining
+    covariances between pairs of rewards.}
+  \label{fig:covariance_boxplots}
+\end{figure}
+
+If having more copies of the same demonstrations does not affect covariances, we
+can test if covariances tend to be lower in situations that exhibit more
+structure, i.e., demonstrations provide a clearer picture of which states are
+more valuable. For this, we consider a completely-deterministic MDP with ten
+states and nine actions that allow the agent to move from any state to any other
+state---the \emph{clique MDP}. We denote the states by $\mathcal{S} = \{ s_i
+\}_{i=0}^9$, and, similarly to the previous example, set up a single feature $f$
+such that $f(s_i) = i$ for $i = 0, \dots, 9$. Finally, $\gamma = 0.9$ as before.
+We consider two scenarios, with 100 randomly generated demonstrations in each.
+The scenarios are visualised in Figure \ref{fig:cliques}, although we only show
+a small subset of demonstrations. In the first scenario, we draw both the
+starting state and the action uniformly at random. In the second scenario, we
+draw the starting state uniformly from $\mathcal{S} \setminus \{ s_0 \}$, and
+the action always points to $s_0$. We would expect at least variances of state
+rewards to be lower in the second scenario in order to reflect the more
+structured, certain behavioural pattern expressed by the demonstrations.
+
+Indeed, Figure \ref{fig:covariance_boxplots} shows exactly that. The first two
+box plots show that in roughly $3/4$ of the runs reward variances were higher in
+the random scenario\footnote{Note that outliers were removed from the plot.
+  However, all outliers were positive.}. Interestingly, the same applies to
+covariances.
+
+% TODO: conclude the experiments
+
 \section{Related Work} % 0.5-1 page
@@
 \subsection{Further Work}
+%TODO: how optimal policies depend on the structure of the mdp and available
+%data (i.e., why don't our policies converge to 1)
 

### 2019-03-30 code/singletest.m

@@ -158,7 +158,6 @@ function [values, groups] = meaningful_classes(matrix_of_differences)
     a = b + 1;
     b = b + length(non_neighbours);
     [groups{a:b}] = deal('other edges');
-    %groups = horzcat(zeros(1, length(state)), ones(1, length(other_diag)), 2 * ones(1, length(neighbours)), 3 * ones(1, length(non_neighbours)));
 end
 
 function covariance = covariance_matrix(result)

### 2019-04-08 mpaper/paper.tex
An example of notation and lots of editing

@@
 \usepackage{pgfplots}
+\usepackage{blkarray}
 
@@
 \newtheorem{problem}[theorem]{Open Problem}
+\newtheorem{example}[theorem]{Example}
 
@@
 \newcommand{\V}{V_{\mathbf{r}}}
-\newcommand{\dx}{\,d\mathbf{r}\,d\mathbf{u}}
+\newcommand{\dx}{\,\mathrm{d}\mathbf{r}\,\mathrm{d}\mathbf{u}}
 \newcommand{\vbound}{\frac{\rinf + \log|\mathcal{A}|}{1 - \gamma}}
@@
 state-action pairs.
+% TODO (must): along with inverse optimal control, mention imitation learning
+% and apprenticeship learning
 
@@
 \cite{DBLP:conf/icml/PieterN04,DBLP:conf/icml/NgR00,ziebart2008maximum}.
-This assumption makes the models unable to represent many reward structures. Out
-of the non-linear models proposed to date, none can answer the questions posed
-in the first paragraph. Quite often, the models assume that rewards have no
-variance \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}. In this paper,
-we show how that assumption can be lifted by switching from maximum likelihood
-estimation to \emph{variational inference} (VI), i.e., we approximate the
-posterior distribution of the model by optimising the parameters of a simpler
-distribution to make it similar to the posterior. This approach can prove useful
-in four major ways:
+This assumption severely restricts what reward structures can be modelled (i.e.,
+the properties of the function that represents rewards across states). Out of
+the non-linear models proposed to date, none can answer the questions posed in
+the first paragraph. Quite often, the models assume that rewards have no
+variance \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}.
+
+In this paper, we show how that assumption can be lifted by switching from
+maximum likelihood estimation to \emph{variational inference} (VI), i.e., we
+approximate the posterior distribution of the model by optimising the parameters
+of a simpler distribution to make it similar to the posterior. This approach can
+prove useful in four major ways:
 \begin{enumerate}
@@
 
-Our main contribution is a lengthy proof in Section \ref{sec:proof} that shows
+Our main contribution is a proof in Section~\ref{sec:proof} that shows
 how VI can be applied to the maximum-entropy IRL model with Gaussian processes
 (GPs) proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. We describe
-how we adapt the model and set up the VI problem in Section \ref{sec:model}.
-Finally, in Section \ref{sec:evaluation} we examine the convergence properties
+how we adapt the model and set up the VI problem in Section~\ref{sec:model}.
+Finally, in Section~\ref{sec:evaluation} we examine the convergence properties
 of our model and its ability to deduce optimal policies in practice.
+% TODO: expand on this more
 
@@
 In this section, we introduce definitions and mathematical details relevant to
-the problem. We begin by formally defining what the problem is.
+the problem. We begin by formally defining the problem.
 
@@
   factor; and $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ is the reward
-  vector\footnote{Depending on the situation, we will sometimes represent
-    rewards as a function $r : \mathcal{S} \to \mathbb{R}$.}.
+  vector.
 \end{definition}
+
+While it is more common to represent rewards as a function $r : \mathcal{S} \to
+\mathbb{R}$, the vector notation is generally more appropriate in our situation.
+However, we switch between the two notations as needed.
 
@@
   $|\mathcal{S}| \times d$ feature matrix $\mathbf{X}$ (where $d$ is the number
-  of features), and a set of expert demonstrations $\mathcal{D} = \{\zeta_i
+  of features), and a set of demonstrations $\mathcal{D} = \{\zeta_i
   \}_{i=1}^N$, where each demonstration $\zeta_i = \{ (s_{i,t}, a_{i,t})
-  \}_{t=1}^T$ is a multiset of state-action pairs representing optimal actions
-  executed by an expert, find the reward function that maximises the probability
+  \}_{t=1}^T$ is a multiset of state-action pairs that represents optimal (or
+  near-optimal) actions, find the reward function that maximises the probability
   of observing the demonstrations, i.e.,
@@
 
-The optimal (deterministic) \emph{policy} $\pi : \mathcal{S} \to \mathcal{A}$
-(i.e., a choice of actions for each state that maximises reward over time) is
-usually constructed by defining a \emph{value (utility) function} $\V :
-\mathcal{S} \to \mathbb{R}$ that measures how good a state is based on the
-reward $\mathbf{r}$ as well as the structure of the MDP. One can then find $\V$
-by applying the Bellman backup operator until convergence to every $s \in
-\mathcal{S}$ (the technique is known as \emph{value iteration})
-\cite{DBLP:books/daglib/0023820}:
+\begin{example}
+  Figure \ref{fig:example} shows a possible setup for IRL. In this case,
+  \begin{align*}
+    \mathcal{S} &= \{ s_1, s_2, s_3, s_4 \}, \\
+    \mathcal{A} &= \{ l, r, u, d \}, \\
+    \mathcal{T}(s, a, s') &= \begin{cases}
+      1 & \text{if } (s, a, s') = (s_1, r, s_2) \\
+      1 & \text{if } (s, a, s') = (s_1, d, s_4) \\
+      1 & \text{if } (s, a, s') = (s_2, l, s_1) \\
+      1 & \text{if } (s, a, s') = (s_2, d, s_3) \\
+      1 & \text{if } (s, a, s') = (s_3, l, s_4) \\
+      1 & \text{if } (s, a, s') = (s_3, u, s_2) \\
+      1 & \text{if } (s, a, s') = (s_4, r, s_3) \\
+      1 & \text{if } (s, a, s') = (s_4, u, s_1) \\
+      0 & \text{otherwise,}
+    \end{cases} \\
+      \mathbf{X} &= \begin{blockarray}{ccc}
+        & f_1 & f_2 \\
+        \begin{block}{c[cc]}
+          s_1 & 1 & 2 \\
+          s_2 & 2 & 2 \\
+          s_3 & 2 & 1 \\
+          s_4 & 1 & 1 \\
+        \end{block}
+      \end{blockarray},
+  \end{align*}
+  and $\gamma \in [0, 1)$ can be assigned any value. The distribution of example
+  state-action pairs into separate demonstrations is irrelevant to our model
+  (and possible all other models), so we can keep each pair in its own set as
+  follows:
+  \[
+    \mathcal{D} = \{ \{(s_1, d)\}, \{(s_2, l)\}, \{(s_4, u)\}, \{(s_4, u)\} \}.
+  \]
+
+  A solution to this problem is then an assignment of rewards to states, e.g.,
+  \[
+    \mathbf{r} = \begin{bmatrix}
+      2 \\
+      0 \\
+      -1 \\
+      1
+    \end{bmatrix} \quad \text{or} \quad r(s) = \begin{cases}
+      2 & \text{if } s = s_1 \\
+      0 & \text{if } s = s_2 \\
+      -1 & \text{if } s = s_3 \\
+      1 & \text{if } s = s_4.
+    \end{cases}
+  \]
+\end{example}
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}[state/.style={circle,draw,fill=lightgray}]
+    \node [state] (s1) {$s_1$};
+    \node [state,right=2 of s1] (s2) {$s_2$};
+    \node [state,below=2 of s2] (s3) {$s_3$};
+    \node [state,left=2 of s3] (s4) {$s_4$};
+
+    \draw[->,bend left,dashed] (s1) edge node[above] {$r$} (s2);
+    \draw[->,bend left] (s2) edge node[below] {$l(1)$} (s1);
+    \draw[->,bend left] (s1) edge node[right] {$d(1)$} (s4);
+    \draw[->,bend left] (s4) edge node[left] {$u(2)$} (s1);
+    \draw[->,bend left,dashed] (s4) edge node[above] {$r$} (s3);
+    \draw[->,bend left,dashed] (s3) edge node[below] {$l$} (s4);
+    \draw[->,bend left,dashed] (s2) edge node[right] {$d$} (s3);
+    \draw[->,bend left,dashed] (s3) edge node[left] {$u$} (s2);
+
+    \draw [<->] (-1.3, 0.8) node (yaxis) [above] {$f_2$}
+    |- (3.5, -3.6) node (xaxis) [right] {$f_1$};
+    \draw[shift={(0,-3.6)}] (0pt,2pt) -- (0pt,-2pt) node[below] {$1$};
+    \draw[shift={(2.65,-3.6)}] (0pt,2pt) -- (0pt,-2pt) node[below] {$2$};
+    \draw[shift={(-1.3,-2.6)}] (2pt,0pt) -- (-2pt,0pt) node[left] {$1$};
+    \draw[shift={(-1.3,0)}] (2pt,0pt) -- (-2pt,0pt) node[left] {$2$};
+  \end{tikzpicture}
+  \caption{An example MDP with deterministic actions. States are represented by
+    grey circles, actions by directed labelled edges between them. The numbers
+    in parentheses next to some actions' names denote how many times that
+    state-action pair appears in $\mathcal{D}$. If that number is zero, the edge
+    is dashed. Finally, the axes assign two features to each state.}
+  \label{fig:example}
+\end{figure}
+
+The set of demonstrations $\mathcal{D}$ often comes from observed human
+behaviour and constitutes the most significant part of the data. The optimal
+(deterministic) \emph{policy} $\pi : \mathcal{S} \to \mathcal{A}$ (i.e., a
+choice of actions for each state that maximises reward over time) is usually
+constructed by defining a \emph{value (utility) function} $\V : \mathcal{S} \to
+\mathbb{R}$ that measures how good a state is based on the reward $\mathbf{r}$
+as well as the structure of the MDP. One can then find $\V$ by applying the
+Bellman backup operator until convergence to every $s \in \mathcal{S}$ (the
+technique is known as \emph{value iteration}) \cite{DBLP:books/daglib/0023820}:
 \[
@@
 
-% TODO: rewrite the 'we write' sentence and the sentence before that
-As we want learned rewards to generalise to previously unseen states and
-for states to have a notion of similarity, the IRL definition also includes $d$
-features associated with each state. In this paper, we will focus on modelling
-rewards as a function from feature space to $\mathbb{R}$ using a \emph{Gaussian
-  process}. A GP is defined as a collection of random variables, any finite
-number of which has a joint Gaussian distribution
-\cite{DBLP:books/lib/RasmussenW06}. We write $r \sim \mathcal{GP}(0, k)$ to say
-that $r$ is a GP with mean $0$ and covariance function $k$. \emph{Covariance
-  functions} (also known as \emph{kernels}) take two state feature vectors as
-input and quantify how similar the two states are, in a sense that we would
-expect high covariance scores to be associated with similar rewards.
+A significant part of learning is about being able to generalise to previously
+unseen input states (otherwise the problem becomes knowledge representation
+instead of learning). For this reason, we represent each state by a
+$d$-dimensional feature vector. This way, similar states are located in close
+proximity to one another, and the ML model can predict rewards for new states
+based on their similarity to training data.
+
+In this paper, we model rewards as a function from feature space to
+$\mathbb{R}$ using a \emph{Gaussian process}. A GP represents an infinite
+collection of random variables, any finite number of which has a joint Gaussian
+distribution \cite{DBLP:books/lib/RasmussenW06}. A GP is defined by its mean
+function (which is always $\mathbf{x} \mapsto 0$ in our case) and covariance
+function, which we denote by $k$. \emph{Covariance functions} (also known as
+\emph{kernels}) take two state feature vectors as input and quantify how similar
+the two states are, in a sense that we would expect high covariance scores to be
+associated with similar rewards.
 
@@
 
-Our goal is then to use VI in order to approximate $\posterior$.
-Let $\approximation$ denote our approximation. VI aims to optimise this
-approximation by minimising the \emph{Kullback-Leibler} (KL) divergence between
-the original probability distribution and our approximation. KL divergence
-(asymmetrically) measures the difference between two probability distributions
-and can be defined as \cite{blei2017variational}
+Our goal is then to use VI in order to approximate
+$p(\mathbf{u},~\mathbf{r}~\mid~\mathcal{D})$. Let $\approximation$ denote our
+approximation. VI aims to optimise this approximation by minimising the
+\emph{Kullback-Leibler} (KL) divergence between the original probability
+distribution and our approximation. KL divergence (asymmetrically) measures the
+difference between two probability distributions and is defined as
+\cite{blei2017variational}
 \begin{align*}
@@
     \mathcal{L} &= \mathbb{E}_{\approximation}\left[ \log \frac{\pfull}{\approximation} \right] \\
-    &= \iiint \log \frac{\pfull}{\approximation} \approximation\dx.
+    &= \iint \log \frac{\pfull}{\approximation} \approximation\dx.
   \end{split}
@@
 Thus, instead of minimising KL divergence, we focus on maximising $\mathcal{L}$
-by optimising the values of parameters to be defined in Section \ref{sec:model}.
+by optimising the values of parameters to be defined in Section~\ref{sec:model}.
 Note that hereafter we drop the subscript notation, as all expected values will
@@
 convergence theorem which plays a major role in our theoretical results in
-Section \ref{sec:proof}.
+Section~\ref{sec:proof}.
 
@@
   \[
-    |f_n| \le g \text{ a.e. on $X$ for all $n$.}
+    |f_n| \le g \text{ almost everywhere on $X$ for all $n$.}
   \]
@@
   \[
-    \lim_{n \to \infty} \int_X f_n\,d\mu = \int_X f\,d\mu.
+    \lim_{n \to \infty} \int_X f_n\,\mathrm{d}\mu = \int_X f\,\mathrm{d}\mu.
   \]
@@
 define the GP, consider what parameters should be used to optimise $\mathcal{L}$
-and how they should be initialised, and in Section \ref{sec:elbo}, we derive our
+and how they should be initialised, and in Section~\ref{sec:elbo}, we derive our
 final expression for $\mathcal{L}$.
@@
 $\mathbf{u}$ involves $\Kuu^{-1}$, and $\mathcal{L}$ has an expected value that
-cannot be eliminated (see Section \ref{sec:elbo}), we are unable to show that
+cannot be eliminated (see Section~\ref{sec:elbo}), we are unable to show that
 such an $\mathcal{L}$ is well-defined. More generally, we pose the following
@@
 $\mathbf{D} = \mathbf{O}_m$ in order to retain full covariance structure.
+
+The resulting model is summarised in Figure~\ref{fig:graphical_model}. We rely
+on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and our model.
+Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
+and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
+parameters of the kernel, $\bm\lambda$. The two remaining dependencies stem from
+the fact that the approximating distribution for $\mathbf{u}$ is
+$\mathcal{N}(\bm\mu, \mathbf{BB}^\intercal)$.
 
@@
 
-The resulting model is summarised in Figure \ref{fig:graphical_model}. We rely
-on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and our model.
-Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
-and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
-parameters of the kernel, $\bm\lambda$. The two remaining dependencies stem from
-the fact that the approximating distribution for $\mathbf{u}$ is
-$\mathcal{N}(\bm\mu, \mathbf{BB}^\intercal)$.
-
 As we want to restrict some parameters (namely, $\bm\lambda$ and the diagonal of
@@
 
-Firstly, let us return to \eqref{eq:elbo} and write
+Firstly, let us return to Equation~\eqref{eq:elbo} and write
 \[
@@
 \]
-By substituting in \eqref{eq:full} and \eqref{eq:approximation}, we get
+By substituting in Equations~\eqref{eq:full} and \eqref{eq:approximation}, we get
 \begin{align*}
@@
 computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
-complicates the situation. The goal of this section is to show how Theorem
-\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
-anyway\footnote{This technique is inspired by black box VI
+complicates the situation. The goal of this section is to show how
+Theorem~\ref{thm:lebesgue} can be applied to our model in order to derive the
+gradient anyway\footnote{This technique is inspired by black box VI
   \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
@@
 \mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.
+
+Our main goal is Theorem~\ref{thm:main}, which allows us to move differentiation
+inside the integral. In order to prove it, we use a number of intermediate
+results. We start by stating a few derivatives of PDFs and covariance matrices,
+and bound their values with some easy-to-deal-with polynomials. We then provide
+a sketch proof of the measurability of MDP value functions as well as new upper
+and lower bounds on their values. After another quick lemma, we tackle the main
+proof of this paper. See Figure~\ref{fig:theorems} for an overview of how these
+results fit together.
 
@@
     \node [rectangle, draw] (boundedness) at (3, -2) {\textsc{Boundedness}};
-    \node [rectangle, draw] (lemma1) at (-2, -1) {\textsc{Lemma \ref{lemma:bound1}}};
-    \node [rectangle, draw] (lemma2) at (0, -1) {\textsc{Lemma \ref{lemma:bound2}}};
-    \node [rectangle, draw] (lemma3) at (2, -1) {\textsc{Lemma \ref{lemma:bound3}}};
-    \node [rectangle, draw] (lemma4) at (3, -3) {\textsc{Lemma \ref{lemma:integral_of_r}}};
-    \node [rectangle, draw] (theorem) at (0, -2) {\textsc{Theorem \ref{thm:main}}};
+    \node [rectangle, draw] (lemma1) at (-2, -1) {\textsc{Lemma~\ref{lemma:bound1}}};
+    \node [rectangle, draw] (lemma2) at (0, -1) {\textsc{Lemma~\ref{lemma:bound2}}};
+    \node [rectangle, draw] (lemma3) at (2, -1) {\textsc{Lemma~\ref{lemma:bound3}}};
+    \node [rectangle, draw] (lemma4) at (3, -3) {\textsc{Lemma~\ref{lemma:integral_of_r}}};
+    \node [rectangle, draw] (theorem) at (0, -2) {\textsc{Theorem~\ref{thm:main}}};
 
@@
 \end{figure}
-
-Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
-inside the integral. In order to prove it, we use a number of intermediate
-results. We start by stating a few derivatives of PDFs and covariance matrices,
-and bound their values with some easy-to-deal-with polynomials. We then provide
-a sketch proof of the measurability of MDP value functions as well as new upper
-and lower bounds on their values. After another quick lemma, we tackle the main
-proof of this paper. See Figure \ref{fig:theorems} for an overview of how these
-results fit together.
 
@@
   \begin{equation} \label{eq:convergence1}
-    \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \mathbf{K}
+    \lim_{\epsilon \to 0} \mathbf{K}|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \mathbf{K},
   \end{equation}
@@
   \begin{equation} \label{eq:convergence2}
-    \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} \to \frac{\partial \mathbf{K}}{\partial \lambda_i}
+    \lim_{\epsilon \to 0} \left. \frac{\partial \mathbf{K}}{\partial \lambda_i} \right|_{\lambda_i = c(\mathbf{r}, \mathbf{u})} = \frac{\partial \mathbf{K}}{\partial \lambda_i}.
   \end{equation}
-  as $\epsilon \to 0$. For \eqref{eq:convergence1}, the result is obvious if $i
-  = 0$. Otherwise, it follows from the continuity of the exponential function.
-  For \eqref{eq:convergence2}, observe that if $i=0$, then
+  For Equation~\eqref{eq:convergence1}, the result is obvious if $i = 0$.
+  Otherwise, it follows from the continuity of the exponential function. For
+  Equation~\eqref{eq:convergence2}, observe that if $i=0$, then
   \[
@@
   \]
-  so the same reasoning applies as in case of \eqref{eq:convergence1}.
+  so the same reasoning applies as in case of Equation~\eqref{eq:convergence1}.
 
@@
   \end{align*}
-  as $\epsilon \to 0$. For $i > 0$, by \eqref{eq:convergence1} and continuity of
-  $\mathbf{A} \mapsto \mathbf{A}^{-1}$ we immediate get \eqref{eq:limit3}.
+  as $\epsilon \to 0$. For $i > 0$, by Equation~\eqref{eq:convergence1} and
+  continuity of $\mathbf{A} \mapsto \mathbf{A}^{-1}$ we immediate get
+  Equation~\eqref{eq:limit3}.
 
@@
 
-  From \eqref{eq:gamma_convergence} and continuity of $\mathbf{A} \mapsto
+  From Equation~\eqref{eq:gamma_convergence} and continuity of $\mathbf{A} \mapsto
   \mathbf{A}^{-1}$ we can immediately deduce that
@@
   \end{equation}
-  Similarly, from \eqref{eq:det_gamma} and \eqref{eq:gamma_inverse} we get that
+  Similarly, from Equations~\eqref{eq:det_gamma} and \eqref{eq:gamma_inverse} we
+  get that
   \[
@@
 \begin{proof}
-  Using Lemma \ref{lemma:derivatives},
+  Using Lemma~\ref{lemma:derivatives},
   \[
@@
   we use the maximum norm $\lVert \cdot \rVert_\infty$ on both vectors and
-  matrices. We can apply Lemma \ref{lemma:perturbation} to $\bm\Sigma$ and
+  matrices. We can apply Lemma~\ref{lemma:perturbation} to $\bm\Sigma$ and
   $\mathbf{E}(\mathbf{r}, \mathbf{u})$ since
@@
       \;&\V(s) \in \mathbb{R} \text{ for all } s \in \mathcal{S} \setminus \{ s_0 \}; \\
-      \;&\text{\eqref{eq:linear_mdp} is satisfied by all } s \in
+      \;&\text{Equation~\eqref{eq:linear_mdp} is satisfied by all } s \in
       \mathcal{S}
@@
   bound, then the bound remains satisfied throughout value iteration:
-  \begin{equation} \label{eq:bound}
+  \[
     |\V(s)| \le \vbound.
-  \end{equation}
+  \]
 \end{proposition}
 \begin{proof}
-  We begin by considering \eqref{eq:bound} without taking the absolute value of
-  $\V(s)$, i.e.,
+  We begin by considering the desired inequality without taking the absolute
+  value of $\V(s)$, i.e.,
   \begin{equation} \label{eq:positive_bound}
@@
   and assuming that the initial values of $\{ \V(s) \mid s \in \mathcal{S} \}$
-  already satisfy \eqref{eq:positive_bound}. Recall that for each $s \in
-  \mathcal{S}$, the value of $\V(s)$ is updated by applying
-  \eqref{eq:update_rule}. Note that both the natural logarithm and the
+  already satisfy Inequality~\eqref{eq:positive_bound}. Recall that for each $s
+  \in \mathcal{S}$, the value of $\V(s)$ is updated by applying
+  Equation~\eqref{eq:update_rule}. Note that both the natural logarithm and the
   exponential function are increasing, $\gamma > 0$, and the $\mathcal{T}$
@@
   where we use the fact that $r(s) \ge -\rinf - 2\log|\mathcal{A}|$. Combining
-  \eqref{eq:positive_bound} and \eqref{eq:negative_bound} gives
-  \eqref{eq:bound}.
+  Inequalities~\eqref{eq:positive_bound} and \eqref{eq:negative_bound} gives
+  the desired result.
 \end{proof}
@@
   \[
-    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le a +
-    \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1,
+    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid
+    \mathbf{u})\,\mathrm{d}\mathbf{r} \le a + \lVert \Kru^\intercal \Kuu^{-1}
+    \mathbf{u} \rVert_1,
   \]
@@
   \[
-    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le \int
-    \lVert \mathbf{r} \rVert_1 q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} =
+    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid
+    \mathbf{u})\,\mathrm{d}\mathbf{r} \le \int \lVert \mathbf{r} \rVert_1
+    q(\mathbf{r} \mid \mathbf{u})\,\mathrm{d}\mathbf{r} =
     \sum_{i=1}^{|\mathcal{S}|} \mathbb{E}[|r_i|].
@@
   \]
-  \eqref{eq:to_prove} follows from Theorem \ref{thm:lebesgue} as soon as we show
-  that both $f$ and $f_n$ are measurable and find a non-negative integrable
-  function $g$ such that for all $n$, $\mathbf{r}$, $\mathbf{u}$,
+  Equation~\eqref{eq:to_prove} follows from Theorem~\ref{thm:lebesgue} as soon
+  as we show that both $f$ and $f_n$ are measurable and find a non-negative
+  integrable function $g$ such that for all $n$, $\mathbf{r}$, $\mathbf{u}$,
   \[
@@
   \]
-  The MDP value function is measurable by Proposition \ref{thm:measurability}.
+  The MDP value function is measurable by Proposition~\ref{thm:measurability}.
   The result of multiplying or adding measurable functions (e.g., PDFs) to a
@@
   since PDFs are non-negative. An upper bound for
-  $|\V(s)|$ is given by Proposition \ref{thm:bound}, while
+  $|\V(s)|$ is given by Proposition~\ref{thm:bound}, while
   \[
@@
   is also integrable. Depending on what $t$ represents, we can use one of the
-  Lemmas \ref{lemma:bound1}, \ref{lemma:bound2}, and \ref{lemma:bound3}, which
+  Lemmas~\ref{lemma:bound1}, \ref{lemma:bound2}, and \ref{lemma:bound3}, which
   gives us two polynomials $p_1(\mathbf{u}), p_2(\mathbf{u}) \in
@@
   \]
-  We can now apply Lemma \ref{lemma:integral_of_r}, which allows us to integrate
+  We can now apply Lemma~\ref{lemma:integral_of_r}, which allows us to integrate
   out $\mathbf{r}$, and we are left with showing the existence of
   \begin{equation} \label{eq:remaining_integral}
-    \int \left( a + \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1 \right) \max \{|p_1(\mathbf{u})|, |p_2(\mathbf{u})| \} q(\mathbf{u})\,d\mathbf{u},
+    \int \left( a + \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1 \right) \max \{|p_1(\mathbf{u})|, |p_2(\mathbf{u})| \} q(\mathbf{u})\,\mathrm{d}\mathbf{u},
   \end{equation}
@@
       \end{aligned}
-    \right\} q(\mathbf{u})\,d\mathbf{u} = \int \max \left\{
+    \right\} q(\mathbf{u})\,\mathrm{d}\mathbf{u} = \int \max \left\{
       \begin{aligned}
@@
       \end{aligned}
-    \right\}\,d\mathbf{u}
+    \right\}\,\mathrm{d}\mathbf{u}
   \]
@@
   \mathbb{R}_1[\mathbf{u}]$, a similar argument can be applied to the rest of
-  \eqref{eq:remaining_integral} as well.
+  Integral~\eqref{eq:remaining_integral} as well.
 \end{proof}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figures/convergence_new}
-  \caption{Convergence of $\mathcal{L}$ (at the top) as well as several example
-    policies (at the bottom).}
-  \label{fig:convergence1}
-\end{figure}
-
-\begin{figure*}
-  \centering
-  \includegraphics[width=\textwidth]{figures/parameter_convergence_new}
-  \caption{Convergence of variational parameters. In order to represent
-    different variables on the same scale, some variables have been
-    $\log$-transformed. Colours denote which vector or matrix each scalar comes
-    from: black for $\bm\lambda$, \textcolor{blue}{blue} for $\bm\mu$,
-    \textcolor{red}{red} for diagonal elements of $\mathbf{B}$, and
-    \textcolor{green}{green} for its non-diagonal elements.}
-  \label{fig:convergence2}
-\end{figure*}
-
 We plot how $\mathcal{L}$ as well as policies $\pi(a_1 \mid s_1)$, $\pi(a_2 \mid
-s_3)$, and $\pi(a_1 \mid s_2)$ converge over a number of iterations in Figure
-\ref{fig:convergence1}\footnote{In all experiments, we discard runs that
+s_3)$, and $\pi(a_1 \mid s_2)$ converge over a number of iterations in
+Figure~\ref{fig:convergence1}\footnote{In all experiments, we discard runs that
   result in a near-singular $\bm\Sigma$ or $\bm\Gamma$, repeatedly testing each
@@
 
-We also plot how the parameters of the model converge in Figure
-\ref{fig:convergence2}. While some parameters could benefit from a higher number of
-iterations\footnote{The limit of 300 iterations was chosen so that the initial
-  increase in $\mathcal{L}$ would not be overshadowed by a long straight line
-  afterwards.}, none look alarmingly wrong. Note that $\mu_2$, a variable
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/convergence_new}
+  \caption{Convergence of $\mathcal{L}$ (at the top) as well as several example
+    policies (at the bottom)}
+  \label{fig:convergence1}
+\end{figure}
+
+We also plot how the parameters of the model converge in
+Figure~\ref{fig:convergence2}. While some parameters could benefit from a higher
+number of iterations\footnote{The limit of 300 iterations was chosen so that the
+  initial increase in $\mathcal{L}$ would not be overshadowed by a long straight
+  line afterwards.}, none look alarmingly wrong. Note that $\mu_2$, a variable
 closely related to $r(s_2)$, stabilises at a positive value whereas both $\mu_1$
@@
 
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{figures/parameter_convergence_new}
+  \caption{Convergence of variational parameters. In order to represent
+    different variables on the same scale, some variables have been
+    $\log$-transformed. Colours denote which vector or matrix each scalar comes
+    from: black for $\bm\lambda$, \textcolor{blue}{blue} for $\bm\mu$,
+    \textcolor{red}{red} for diagonal elements of $\mathbf{B}$, and
+    \textcolor{green}{green} for its non-diagonal elements.}
+  \label{fig:convergence2}
+\end{figure*}
+
 \paragraph{Adding More Data}
+
+We would expect policies to converge closer to extreme values (i.e., 0 and 1) as
+we add more data to the model. We test this hypothesis with a small-scale
+experiment in Figure~\ref{fig:policies}. The policy $\pitwo$, in particular,
+confirms our prediction. As this policy denotes the probability of going from
+$s_2$ to $s_1$, we expect it to stay around $0.5$ as long as we have equal
+amounts of $\zeta_1$ and $\zeta_2$ (this corresponds to the plots at coordinates
+$(1, 1)$, $(2, 2)$, and $(3, 3)$). But as we increase the number of times state
+$s_1$ appears in our demonstrations (going to the right across any row of
+plots), the model recognises the increasing value of state $s_1$, and the
+probability increases. In contrast, going up across the plots reduces the
+probability, as state $s_3$ becomes more valuable, and the model prefers it over
+state $s_1$. Interestingly, the other two policy probabilities seem to increase
+regardless of what data is added. This is likely due to the fact that a higher
+number of demonstrations results in higher gradient values, and thus leads to
+better-converged policies.
 
@@
 
-We would expect policies to converge closer to extreme values (i.e., 0 and 1) as
-we add more data to the model. We test this hypothesis with a small-scale
-experiment in Figure \ref{fig:policies}. The policy $\pitwo$, in particular,
-confirms our prediction. As this policy denotes the probability of going from
-$s_2$ to $s_1$, we expect it to stay around $0.5$ as long as we have equal
-amounts of $\zeta_1$ and $\zeta_2$ (this corresponds to the plots at coordinates
-$(1, 1)$, $(2, 2)$, and $(3, 3)$). But as we increase the number of times state
-$s_1$ appears in our demonstrations (going to the right across any row of
-plots), the model recognises the increasing value of state $s_1$, and the
-probability increases. In contrast, going up across the plots reduces the
-probability, as state $s_3$ becomes more valuable, and the model prefers it over
-state $s_1$. Interestingly, the other two policy probabilities seem to increase
-regardless of what data is added. This is likely due to the fact that a higher
-number of demonstrations results in higher gradient values, and thus leads to
-better-converged policies.
+Finally, we would like to see similar changes in reward covariances. Firstly, we
+will show how to derive the covariance matrix for $\mathbf{r}$ from the
+posterior distributions of $\mathbf{u}$ and $\mathbf{r} \mid \mathbf{u}$.
+Remember that $\mathbf{u} \sim \mathcal{N}(\bm\mu, \bm\Sigma)$ and $\mathbf{r}
+\mid \mathbf{u} \sim \mathcal{N}(\mathbf{Su}, \bm\Gamma)$, and let $\mathbf{r}
+\sim \mathcal{N}(\bm\mu', \bm\Sigma')$, where our goal is to find $\bm\mu'$ and
+$\bm\Sigma'$. Then
+\[
+  \bm\mu' = \mathbb{E}[\mathbf{r}] = \mathbb{E}[\mathbf{Su}] = \mathbf{S}\bm\mu.
+\]
+Furthermore,
+\begin{align*}
+  \mathbb{E}_{\mathbf{r}}[\mathbf{rr}^\intercal] &= \bm\Gamma + \mathbb{E}_{\mathbf{u}}[\mathbf{Suu}^\intercal\mathbf{S}^\intercal] = \bm\Gamma + \mathbf{S}\mathbb{E}_{\mathbf{u}}[\mathbf{uu}^\intercal]\mathbf{S}^\intercal \\
+                                                 &= \bm\Gamma + \mathbf{S}(\bm\Sigma + \bm\mu\bm\mu^\intercal)\mathbf{S}^\intercal
+\end{align*}
+by an identity for $\mathbb{E}[\mathbf{xx}^\intercal]$ by Petersen and Pedersen
+\cite{petersen2008matrix}. Now we can find the covariance matrix as follows:
+\begin{align*}
+  \bm\Sigma' &= \mathbb{E}[(\mathbf{r} - \mathbf{S}\bm\mu)(\mathbf{r} -
+               \mathbf{S}\bm\mu)^\intercal] \\
+             &= \mathbb{E}[\mathbf{rr}^\intercal - \mathbf{r}\bm\mu^\intercal\mathbf{S}^\intercal - \mathbf{S}\bm\mu\mathbf{r}^\intercal + \mathbf{S}\bm\mu\bm\mu^\intercal\mathbf{S}^\intercal] \\
+             &= \bm\Gamma + \mathbf{S}\bm\Sigma\mathbf{S}^\intercal.
+\end{align*}
+Therefore, $\mathbf{r} \sim \mathcal{N}(\mathbf{S}\bm\mu, \bm\Gamma +
+\mathbf{S}\bm\Sigma\mathbf{S}^\intercal)$.
+
+Secondly, we normalise the elements of this new covariance matrix into the
+interval $[-1, 1]$ for visualisation purposes. Diagonal (state) and non-diagonal
+(edge) covariances are normalised separately in order to represent the full
+range of values more clearly. All covariances are normalised according to this
+rule:
+\[
+  x \mapsto
+  \begin{cases*}
+    x/M & if $x \ge 0$ \\
+    -x/m & otherwise,
+  \end{cases*}
+\]
+where $M$ is the maximum (diagonal or non-diagonal) covariance across all data,
+and $m$ is the minimum. This way, all covariances are scaled to $[-1, 1]$, with
+both extreme points guaranteed to be reached unless all values are on one side
+of zero on the real number axis.
+
+Unfortunately, the resulting plot in Figure~\ref{fig:covariance} shows no clear
+pattern, suggesting that adding more of the same demonstrations does not result
+in lower covariances. The only other observation from this plot is that there
+are only two instances of negative covariance, and both of them are between
+states $s_2$ and $s_3$.
 
@@
     \foreach \x in {1, 2, 3} {
-      \node (x\x) at (0.5+2*\x, 1.3) {$\x$};
+      \draw[shift={(0.5+2*\x, 1.5)}] (0pt,2pt) -- (0pt,-2pt) node[below] {$\x$};
     }
     \foreach \y in {1, 2, 3} {
-      \node (y\y) at (1.3, 0.5+2*\y) {$\y$};
+      \draw[shift={(1.5, 0.5+2*\y)}] (2pt,0pt) -- (-2pt,0pt) node[left] {$\y$};
     }
@@
   \caption{Changes in reward covariances after adding more demonstrations. The
-    meaning of external axes is the same as in Figure \ref{fig:policies}.
+    meaning of external axes is the same as in Figure~\ref{fig:policies}.
     Colours denote reward covariance values, e.g., the colour of node $s_1$
@@
 
-Finally, we would like to see similar changes in reward covariances. Firstly, we
-will show how to derive the covariance matrix for $\mathbf{r}$ from the
-posterior distributions of $\mathbf{u}$ and $\mathbf{r} \mid \mathbf{u}$.
-Remember that $\mathbf{u} \sim \mathcal{N}(\bm\mu, \bm\Sigma)$ and $\mathbf{r}
-\mid \mathbf{u} \sim \mathcal{N}(\mathbf{Su}, \bm\Gamma)$, and let $\mathbf{r}
-\sim \mathcal{N}(\bm\mu', \bm\Sigma')$, where our goal is to find $\bm\mu'$ and
-$\bm\Sigma'$. Then
-\[
-  \bm\mu' = \mathbb{E}[\mathbf{r}] = \mathbb{E}[\mathbf{Su}] = \mathbf{S}\bm\mu.
-\]
-Furthermore,
-\begin{align*}
-  \mathbb{E}_{\mathbf{r}}[\mathbf{rr}^\intercal] &= \bm\Gamma + \mathbb{E}_{\mathbf{u}}[\mathbf{Suu}^\intercal\mathbf{S}^\intercal] = \bm\Gamma + \mathbf{S}\mathbb{E}_{\mathbf{u}}[\mathbf{uu}^\intercal]\mathbf{S}^\intercal \\
-                                                 &= \bm\Gamma + \mathbf{S}(\bm\Sigma + \bm\mu\bm\mu^\intercal)\mathbf{S}^\intercal
-\end{align*}
-by the identity for $\mathbb{E}[\mathbf{xx}^\intercal]$ from the book by
-Petersen and Pedersen \cite{petersen2008matrix}. Now we can find the covariance
-matrix as follows:
-\begin{align*}
-  \bm\Sigma' &= \mathbb{E}[(\mathbf{r} - \mathbf{S}\bm\mu)(\mathbf{r} -
-               \mathbf{S}\bm\mu)^\intercal] \\
-             &= \mathbb{E}[\mathbf{rr}^\intercal - \mathbf{r}\bm\mu^\intercal\mathbf{S}^\intercal - \mathbf{S}\bm\mu\mathbf{r}^\intercal + \mathbf{S}\bm\mu\bm\mu^\intercal\mathbf{S}^\intercal] \\
-             &= \bm\Gamma + \mathbf{S}\bm\Sigma\mathbf{S}^\intercal.
-\end{align*}
-Therefore, $\mathbf{r} \sim \mathcal{N}(\mathbf{S}\bm\mu, \bm\Gamma +
-\mathbf{S}\bm\Sigma\mathbf{S}^\intercal)$.
-
-Secondly, we normalise the elements of this new covariance matrix into the
-interval $[-1, 1]$ for visualisation purposes. Diagonal (state) and non-diagonal
-(edge) covariances are normalised separately in order to represent the full
-range of values more clearly. All covariances are normalised according to this
-rule:
-\[
-  x \mapsto
-  \begin{cases*}
-    x/M & if $x \ge 0$ \\
-    -x/m & otherwise,
-  \end{cases*}
-\]
-where $M$ is the maximum (diagonal or non-diagonal) covariance across all data,
-and $m$ is the minimum. This way, all covariances are scaled to $[-1, 1]$, with
-both extreme points guaranteed to be reached unless all values are on one side
-of zero on the real number axis.
-
-Unfortunately, the resulting plot in Figure \ref{fig:covariance} shows no clear
-pattern, suggesting that adding more of the same demonstrations does not result
-in lower covariances. The only other observation from this plot is that there
-are only two instances of negative covariance, and both of them are between
-states $s_2$ and $s_3$.
-
 \paragraph{Structure versus Randomness}
 
-\begin{figure}
+If having more copies of the same demonstrations does not affect covariances, we
+can test if covariances tend to be lower in situations that exhibit more
+structure, i.e., demonstrations provide a clearer picture of which states are
+more valuable. For this, we consider a completely-deterministic MDP with ten
+states and nine actions that allow the agent to move from any state to any other
+state---the \emph{clique MDP}. We denote the states by $\mathcal{S} = \{ s_i
+\}_{i=0}^9$, and, similarly to the previous example, set up a single feature $f$
+such that $f(s_i) = i$ for $i = 0, \dots, 9$. Finally, $\gamma = 0.9$ as before.
+We consider two scenarios, with 100 randomly generated demonstrations in each.
+The scenarios are visualised in Figure~\ref{fig:cliques}, although we only show
+a small subset of demonstrations. In the first scenario, we draw both the
+starting state and the action uniformly at random. In the second scenario, we
+draw the starting state uniformly from $\mathcal{S} \setminus \{ s_0 \}$, and
+the action always points to $s_0$. We would expect at least variances of state
+rewards to be lower in the second scenario in order to reflect the more
+structured, certain behavioural pattern expressed by the demonstrations.
+
+\begin{figure}[!ht]
   \centering
@@
   \end{tikzpicture}
-  \caption{The clique MDP with two sets of demonstrations denoted by blue arrows.}
+  \caption{The clique MDP with two sets of demonstrations denoted by blue arrows}
   \label{fig:cliques}
 \end{figure}
+
+Indeed, Figure~\ref{fig:covariance_boxplots} shows exactly that. The first two
+box plots show that in roughly $3/4$ of the runs reward variances were higher in
+the random scenario\footnote{Note that outliers were removed from the plot.
+  However, all outliers were positive.}. Interestingly, the same applies to
+covariances.
 
@@
   \caption{Box plots of the difference between (absolute values of) reward
-    covariances in random and semi-structured scenarios from Figure
-    \ref{fig:cliques}, averaged out over 100 runs and grouped into four key
-    categories. `First state' refers to the variance of $r(s_0)$, the reward of
-    the state targeted by all demonstrations in the semi-structured case. `Other
-    states' are variances of other states of the MDP, i.e., diagonal entries of
-    the covariance matrix. `Incident edges' refers to covariances between
-    $r(s_0)$ and all other states, i.e., the first row or column of the
+    covariances in random and semi-structured scenarios from
+    Figure~\ref{fig:cliques}, averaged out over 100 runs and grouped into four
+    key categories. `First state' refers to the variance of $r(s_0)$, the reward
+    of the state targeted by all demonstrations in the semi-structured case.
+    `Other states' are variances of other states of the MDP, i.e., diagonal
+    entries of the covariance matrix. `Incident edges' refers to covariances
+    between $r(s_0)$ and all other states, i.e., the first row or column of the
     covariance matrix. Finally, `other edges' is the category for the remaining
@@
 \end{figure}
-
-If having more copies of the same demonstrations does not affect covariances, we
-can test if covariances tend to be lower in situations that exhibit more
-structure, i.e., demonstrations provide a clearer picture of which states are
-more valuable. For this, we consider a completely-deterministic MDP with ten
-states and nine actions that allow the agent to move from any state to any other
-state---the \emph{clique MDP}. We denote the states by $\mathcal{S} = \{ s_i
-\}_{i=0}^9$, and, similarly to the previous example, set up a single feature $f$
-such that $f(s_i) = i$ for $i = 0, \dots, 9$. Finally, $\gamma = 0.9$ as before.
-We consider two scenarios, with 100 randomly generated demonstrations in each.
-The scenarios are visualised in Figure \ref{fig:cliques}, although we only show
-a small subset of demonstrations. In the first scenario, we draw both the
-starting state and the action uniformly at random. In the second scenario, we
-draw the starting state uniformly from $\mathcal{S} \setminus \{ s_0 \}$, and
-the action always points to $s_0$. We would expect at least variances of state
-rewards to be lower in the second scenario in order to reflect the more
-structured, certain behavioural pattern expressed by the demonstrations.
-
-Indeed, Figure \ref{fig:covariance_boxplots} shows exactly that. The first two
-box plots show that in roughly $3/4$ of the runs reward variances were higher in
-the random scenario\footnote{Note that outliers were removed from the plot.
-  However, all outliers were positive.}. Interestingly, the same applies to
-covariances.
 
@@
     we have
-    \[
+    \begin{align*}
       \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} =
-      -\frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) \dl[(\mathbf{r} -
-      \mathbf{Su})^\intercal\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su}) +
-      \log|\bm\Gamma|].
-    \]
+      -\frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) \dl[&(\mathbf{r} -
+      \mathbf{Su})^\intercal\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su}) \\
+      &+ \log|\bm\Gamma|].
+    \end{align*}
     The same online tool can be used to show that
@@
 \]
-by Theorem \ref{thm:main} and Lemma \ref{lemma:derivatives}. Hence,
+by Theorem~\ref{thm:main} and Lemma~\ref{lemma:derivatives}. Hence,
 \[
@@
 \]
-By Theorem \ref{thm:main},
+By Theorem~\ref{thm:main},
 \[
@@
 \]
-and Lemma \ref{lemma:derivatives} gives
+and Lemma~\ref{lemma:derivatives} gives
 \begin{gather*}
@@
 \]
-by Theorem \ref{thm:main} and Lemma \ref{lemma:derivatives}. Thus,
+by Theorem~\ref{thm:main} and Lemma~\ref{lemma:derivatives}. Thus,
 \begin{align*}
@@
 \end{align*}
-where the remaining derivatives can be found in Lemma \ref{lemma:derivatives}.
+where the remaining derivatives can be found in Lemma~\ref{lemma:derivatives}.
 

### 2019-04-08 code/singletest.m

@@ -18,13 +18,14 @@ addpaths;
 %   'training_samples',64,'verbosity',1));
 
 % VIGPIRL
-%test_result = runtest('vigpirl',struct(),'linearmdp',...
-%    'gridworld',struct('n',8,'determinism',1,'seed',1,'continuous',0),...
-%    struct('training_sample_lengths',32,'training_samples',16,'verbosity',2));
+test_result = runtest('vigpirl',struct(),'linearmdp',...
+    'gridworld',struct('n',8,'determinism',1,'seed',1,'continuous',0),...
+    struct('training_sample_lengths',32,'training_samples',16,'verbosity',2));
 
 % Visualize solution.
-%printresult(test_result);
-%visualize(test_result);
+printresult(test_result);
+visualize(test_result);
+return;
 
 % Two states
 %mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0, 'states', 2, 'actions', 1);

### 2019-04-11 mpaper/paper.tex
Editing (mostly introduction and background)

@@
   function of a Markov decision process that explains observed behaviour. Many
-  approaches are only able to construct reward functions as linear combinations
+  approaches are only able to express reward functions as linear combinations
   of state features. Out of those that can handle nonlinearity, none can provide
@@
 many others have benefited from an approach called \emph{inverse reinforcement
-  learning} (IRL) (also known as inverse optimal control). IRL proposes a way to
-learn behaviour from \emph{demonstrations} that typically come from human
-actions. More formally, the IRL problem asks us to find a reward function for a
-Markov decision process (MDP), where demonstrations are encoded as sets of
-state-action pairs.
-% TODO (must): along with inverse optimal control, mention imitation learning
-% and apprenticeship learning
+  learning} (IRL), also known as apprenticeship learning and inverse optimal
+control. IRL proposes a way to learn behaviour from demonstrations that
+typically come from human actions. More formally, the IRL problem asks us to
+find a reward function for a Markov decision process (MDP), where demonstrations
+are encoded as sets of state-action pairs.
 
@@
 
-% TODO: need to rewrite the first sentence
-However, most IRL models in the literature make a convenient yet unjustified
+Most IRL models in the literature make a convenient yet unjustified
 assumption that the reward function can be expressed as a linear combination of
@@
 the non-linear models proposed to date, none can answer the questions posed in
-the first paragraph. Quite often, the models assume that rewards have no
-variance \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}.
+the first paragraph. More specifically, without knowing the ground truth for the
+learning problem, we have no reliable way to know whether the learned behaviour
+is optimal and which parts of the model could benefit from more data. Quite
+often, the models assume that rewards have no variance
+\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}.
 
@@
 of a simpler distribution to make it similar to the posterior. This approach can
-prove useful in four major ways:
+prove useful in three major ways:
 \begin{enumerate}
-\item By working with full distributions instead of point estimates, we can
-  expect more precise reward predictions.
 \item Variance estimates can be used to guide what data should be collected
@@
   adjustments or more data.
-\item By adopting a more Bayesian approach, we automatically incorporate Occam's
-  razor into the model that guards against overfitting
+\item By adopting a more Bayesian approach we can set prior distributions on
+  random variables. This can help prevent overfitting as well as encode
+  additional information about the reward function
   \cite{DBLP:conf/uai/JinDAS17}.
@@
 
-Our main contribution is a proof in Section~\ref{sec:proof} that shows
-how VI can be applied to the maximum-entropy IRL model with Gaussian processes
-(GPs) proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. We describe
-how we adapt the model and set up the VI problem in Section~\ref{sec:model}.
-Finally, in Section~\ref{sec:evaluation} we examine the convergence properties
-of our model and its ability to deduce optimal policies in practice.
-% TODO: expand on this more
+% TODO: could raise intractable posterior as the main reason, but would have to
+% rephrase the whole section
+No fully Bayesian nonlinear IRL model has been proposed yet. We will uncover one
+reason for this when we set up our IRL model for VI in Section~\ref{sec:model}.
+Applying VI to this model in a theoretically sound way (without introducing
+unjustified simplifying assumptions) requires us to use Lebesgue's dominated
+convergence theorem in order to derive the gradient of the expectation of the
+MDP value function---we provide a complete proof for this (and some other
+results) in Section~\ref{sec:proof}. Finally, in Section~\ref{sec:evaluation} we
+demonstrate the model's correctness and properties experimentally using several
+example scenarios.
 
@@
 
-\section{The Problem} \label{sec:problem}
+\section{Background} % 0.5-1 page
+% TODO: edit this: we can probably extend this (and the introduction)
+
+The IRL problem itself was originally proposed by Russell in 1998
+\cite{DBLP:conf/colt/Russell98}. Most of the early approaches had the
+aforementioned reward linearity assumption. One of the first papers on the
+subject by Ng and Russell \cite{DBLP:conf/icml/NgR00} introduced several linear
+programming algorithms and identified an important issue: there are typically
+many reward functions that can explain the data equally well. This problem was
+solved by Ziebart et al. \cite{ziebart2008maximum} with the introduction of IRL
+based on the principles of maximum causal entropy in a linearly-solvable MDP.
+
+Levine et al. \cite{DBLP:conf/nips/LevinePK11} were the first to lift the
+linearity assumption without imposing additional restrictions on the problem.
+They do, however, model rewards as having no variance---our work removes this
+restriction without any compromises.
+
+An alternative to GPs for modelling nonlinear functions is, of course, neural
+networks. Wulfmeier et al. \cite{wulfmeier2015maximum} have shown how they can
+be used in the IRL setting. While this approach benefits from constant-time
+inference and the ability to learn complex features from data, neural
+networks often need significantly more data for the weights across all layers to
+stabilise.
+
+Recently, Jin et al. \cite{DBLP:conf/uai/JinDAS17} have adapted the model
+proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11} to use deep GPs,
+harnessing the power of deep learning to make the model less dependent on what
+features are provided. Although they use VI, their approximating distribution
+for rewards at inducing points is simply the Dirac $\delta$ function, which is
+essentially equivalent to the assumption of no variance.
+
+We argue that developing an IRL model with a variational approximation of the
+full posterior distribution has eluded previous research in the area for two
+reasons. First, it is not directly relevant to the goal of producing accurate
+point estimates of rewards. However, we described several key benefits in the
+previous section, and we postulate how this work could affect the wider fields
+of reinforcement learning and artificial intelligence in
+Section~\ref{sec:conclusions}. Second, as it will become apparent in
+Sections~\ref{sec:model} and \ref{sec:proof}, optimising such a VI model
+involves estimates of gradients, establishing correctness of which requires a
+significant amount of work.
+
+\subsection{Variational Inference}
+
+VI has come a long way from the initial \emph{mean field} approach where each
+variable is approximated by a univariate Gaussian, independent of all other
+variables. Recent years have brought improvements in both computational
+complexity and flexibility of the variational approximation
+\cite{blei2017variational}. In this section, we will mention some of the most
+important and relevant developments in the field, while referring the interested
+reader to recent survey articles
+\cite{blei2017variational,DBLP:journals/corr/abs-1711-05597} for a broader
+overview of the area.
+
+A major development in approximating complex distributions came in the form of
+\emph{normalizing flows}, i.e., a collection of invertible
+functions---parametrised by additional variational parameters---that are applied
+to latent variables \cite{DBLP:conf/icml/RezendeM15}. A sequence of such
+functions can transform a simple initial distribution (e.g., uniform or
+Gaussian) into an arbitrarily complex one.
+%A major challenge in applying this
+%work to IRL is related to variational parameters being used to compute the MDP
+%value function, i.e., how can we take the derivative of the value function with
+%respect to such a variational parameter? Alternatively, perhaps one can
+%construct a different model that would make this question moot.
+
+Another approach to flexibility in modelling could come from considering
+different GP kernels. For instance, Wilson and Adams \cite{pmlr-v28-wilson13}
+show how all \emph{stationary} (i.e., invariant to translations) kernels can be
+generated (or at least approximated) from a mixture of Gaussians in their
+spectral representation using Bochner's theorem
+\cite{bochner1959lectures,DBLP:journals/technometrics/Woodard00}. It looks
+promising to combine these kernels with the variational Fourier features
+approach by Hensman et al. \cite{DBLP:journals/jmlr/HensmanDS17} that leverages
+the same spectral representations for efficient VI.
+
+While these and many other advancements in VI promise great improvements, there
+are hidden challenges when it comes to applying them to VI. Many suggested
+improvements implicitly impose restrictions on the model that are unlikely to be
+satisfied in the context of IRL, e.g., the evidence of the model being a
+Gaussian. Nevertheless, the work outlined here seems adaptable to the IRL
+setting and can make variational approximations both more accurate and faster.
+
+\section{Problem Description} \label{sec:problem}
 
@@
   $|\mathcal{S}| \times d$ feature matrix $\mathbf{X}$ (where $d$ is the number
-  of features), and a set of demonstrations $\mathcal{D} = \{\zeta_i
+  of features), and a set of \emph{demonstrations} $\mathcal{D} = \{\zeta_i
   \}_{i=1}^N$, where each demonstration $\zeta_i = \{ (s_{i,t}, a_{i,t})
@@
 \begin{example}
-  Figure \ref{fig:example} shows a possible setup for IRL. In this case,
-  \begin{align*}
-    \mathcal{S} &= \{ s_1, s_2, s_3, s_4 \}, \\
-    \mathcal{A} &= \{ l, r, u, d \}, \\
+  Figure~\ref{fig:example} shows a possible setup for IRL. In this case,
+  $\mathcal{S} = \{ s_1, s_2, s_3, s_4 \}$, $\mathcal{A} = \{ l, r, u, d \}$,
+  \begin{align*}
     \mathcal{T}(s, a, s') &= \begin{cases}
@@
 \end{equation}
-where
-\begin{align*}
-  p(\mathbf{u}) &= \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu) \\
-                &= \frac{1}{(2\pi)^{m/2}|\Kuu|^{1/2}}\exp \left( -\frac{1}{2} \mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} \right) \\
-                &= \exp\left(-\frac{1}{2}\mathbf{u}^\intercal\Kuu^{-1}\mathbf{u} - \frac{1}{2}\log|\Kuu| - \frac{m}{2}\log 2\pi\right)
-\end{align*}
-is the GP prior \cite{DBLP:books/lib/RasmussenW06}. The GP posterior is then a
+where $p(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{0}, \Kuu)$ is the GP
+prior \cite{DBLP:books/lib/RasmussenW06}. The GP posterior is then a
 multivariate Gaussian \cite{DBLP:conf/nips/LevinePK11} defined as
@@
 
-Our goal is then to use VI in order to approximate
-$p(\mathbf{u},~\mathbf{r}~\mid~\mathcal{D})$. Let $\approximation$ denote our
-approximation. VI aims to optimise this approximation by minimising the
+We can then use VI in order to approximate
+$p(\mathbf{u},~\mathbf{r}~\mid~\mathcal{D})$. This is done by minimising the
 \emph{Kullback-Leibler} (KL) divergence between the original probability
 distribution and our approximation. KL divergence (asymmetrically) measures the
-difference between two probability distributions and is defined as
+difference between two probability distributions. Let $\approximation$ denote
+our approximation. The KL divergence is then defined as
 \cite{blei2017variational}
@@
 The last term is both hard to compute and constant with respect to
-$\approximation$ \cite{blei2017variational}, so we can remove it from our
-optimisation objective. The negation of what remains is known as the
-\emph{evidence lower bound} (ELBO) and is defined as
-\cite{DBLP:books/lib/Bishop07,blei2017variational}
+$\approximation$, so we can remove it from our optimisation objective. The
+negation of what remains is known as the \emph{evidence lower bound} (ELBO) and
+is defined as \cite{DBLP:books/lib/Bishop07,blei2017variational}
 \begin{equation} \label{eq:elbo}
@@
 be with respect to $\approximation$.
-
-\section{Background}
-
-We introduce a few definitions and results from linear algebra, numerical
-analysis, and measure theory that will be used later in the paper. Namely, we
-will use several different vector and matrix norms, consider how an inverse of a
-matrix changes with a small perturbation, and introduce Lebesgue's dominated
-convergence theorem which plays a major role in our theoretical results in
-Section~\ref{sec:proof}.
-
-\begin{definition}[Norms]
-  For any finite-dimensional vector $\mathbf{x} = (x_1, \dots, x_n)^\intercal$,
-  its \emph{maximum norm} (\emph{$\ell_\infty$-norm}) is
-  \[
-    \lVert \mathbf{x} \rVert_\infty = \max_i |x_i|
-  \]
-  whereas its \emph{$\ell_1$-norm} is
-  \[
-    \lVert \mathbf{x} \rVert_1 = \sum_{i = 1}^n |x_i|.
-  \]
-  Let $\mathbf{A}$ be a matrix. For any vector norm $\lVert
-  \cdot \rVert_p$, we can also define its \emph{induced norm} for matrices as
-  \[
-    \lVert \mathbf{A} \rVert_p = \sup_{\mathbf{x} \ne \mathbf{0}} \frac{\lVert
-      \mathbf{Ax} \rVert_p}{\lVert \mathbf{x} \rVert_p}.
-  \]
-  In particular, for $p = \infty$, we have
-  \[
-    \lVert \mathbf{A} \rVert_\infty = \max_i \sum_{j} |A_{i,j}|.
-  \]
-\end{definition}
-
-\begin{lemma}[Perturbation Lemma
-  \cite{layton2014numerical}] \label{lemma:perturbation}
-  Let $\lVert \cdot \rVert$ be any matrix norm, and let $\mathbf{A}$ and
-  $\mathbf{E}$ be matrices such that $\mathbf{A}$ is invertible and $\lVert
-  \mathbf{A}^{-1} \rVert \lVert \mathbf{E} \rVert < 1$, then $\mathbf{A} +
-  \mathbf{E}$ is invertible, and
-  \[
-    \lVert (\mathbf{A} + \mathbf{E})^{-1} \rVert \le \frac{\lVert
-      \mathbf{A}^{-1} \rVert}{1 - \lVert \mathbf{A}^{-1} \rVert \lVert
-      \mathbf{E} \rVert}.
-  \]
-\end{lemma}
-
-\begin{theorem}[Dominated Convergence Theorem
-  \cite{royden2010real}] \label{thm:lebesgue}
-  Let $(X, \mathcal{M}, \mu)$ be a measure space and $\{ f_n \}$ a sequence of
-  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise a.e. on $X$
-  and the function $f$ is measurable. Assume there is a non-negative function
-  $g$ that is integrable over $X$ and dominates the sequence $\{ f_n \}$ on $X$
-  in the sense that
-  \[
-    |f_n| \le g \text{ almost everywhere on $X$ for all $n$.}
-  \]
-  Then $f$ is integrable over $X$ and
-  \[
-    \lim_{n \to \infty} \int_X f_n\,\mathrm{d}\mu = \int_X f\,\mathrm{d}\mu.
-  \]
-\end{theorem}
 
@@
     Bayesian network. The only observed variable (representing the
-    demonstrations) is in a gray circle, modelled latent variables are in white
+    demonstrations) is in a grey circle, modelled latent variables are in white
     circles, and the variational parameters are at the bottom.}
@@
 \]
+We will need to optimise this expression in order to train our model.
 
 \section{Theoretical Justification} \label{sec:proof}
-% TODO: Possible improvements:
-% 1. Use limits more often.
-% 2. Avoid saying that epsilon is arbitrary and using phrases like 'sufficiently small'.
 
@@
 computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
-complicates the situation. The goal of this section is to show how
-Theorem~\ref{thm:lebesgue} can be applied to our model in order to derive the
-gradient anyway\footnote{This technique is inspired by black box VI
-  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
-  problem and requires significantly more work to prove correctness.}. After
-showing that the theorem applies to our situation, we can estimate
+complicates the situation. Indeed, in order to take the derivative of an
+expected value, certain conditions must be satisfied. Without proving the
+validity of our algorithm, we risk building it on faulty assumptions and having
+it misbehave---perhaps all the time, perhaps only in certain situations. The
+goal of this section is to show how Lebesgue's dominated convergence theorem can
+be applied to our model in order to derive $\nabla\mathbb{E}[v]$, i.e., the
+derivative of $\mathbb{E}[v]$ with respect to our variational parameters
+$\bm\mu$, $\mathbf{B}$, and $\bm\lambda$. This technique is inspired by black
+box VI \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at
+the problem and requires significantly more work to prove correctness. After
+showing that the theorem applies to our situation, we can then estimate
 $\nabla\mathbb{E}[v]$ with
@@
 inside the integral. In order to prove it, we use a number of intermediate
-results. We start by stating a few derivatives of PDFs and covariance matrices,
-and bound their values with some easy-to-deal-with polynomials. We then provide
-a sketch proof of the measurability of MDP value functions as well as new upper
-and lower bounds on their values. After another quick lemma, we tackle the main
-proof of this paper. See Figure~\ref{fig:theorems} for an overview of how these
-results fit together.
+results. We begin by introducing a few important tools and techniques in
+Section~\ref{sec:math_preliminaries}. In Section~\ref{sec:derivatives}, we state
+a few derivatives of PDFs and covariance matrices and bound their values with
+some easy-to-work-with polynomials. In Section~\ref{sec:propositions}, we
+provide a sketch proof of the measurability of the MDP value function as well as
+new upper and lower bounds on their values. After another quick lemma, the main
+proof of the paper is in Section~\ref{sec:the_proof}. See
+Figure~\ref{fig:theorems} for an overview of how these results fit together.
 
@@
     results. An arrow from $A$ to $B$ means that $A$ was used to prove $B$.
-    Results from the literature are in gray.}
+    Results from the literature are in grey.}
   \label{fig:theorems}
@@
 
-Before that, however, we define a few extra variables in order to simplify
-expressions of derivatives:
+\subsection{Mathematical Preliminaries} \label{sec:math_preliminaries}
+
+We introduce a few definitions and results from linear algebra, numerical
+analysis, and measure theory that will be used later in the paper. Namely, we
+will use several different vector and matrix norms, consider how an inverse of a
+matrix changes with a small perturbation, and introduce Lebesgue's dominated
+convergence theorem which plays a major role in Section~\ref{sec:the_proof}.
+
+\begin{definition}[Norms]
+  For any finite-dimensional vector $\mathbf{x} = (x_1, \dots, x_n)^\intercal$,
+  its \emph{maximum norm} (\emph{$\ell_\infty$-norm}) is
+  \[
+    \lVert \mathbf{x} \rVert_\infty = \max_i |x_i|
+  \]
+  whereas its \emph{$\ell_1$-norm} is
+  \[
+    \lVert \mathbf{x} \rVert_1 = \sum_{i = 1}^n |x_i|.
+  \]
+  Let $\mathbf{A}$ be a matrix. For any vector norm $\lVert
+  \cdot \rVert_p$, we can also define its \emph{induced norm} for matrices as
+  \[
+    \lVert \mathbf{A} \rVert_p = \sup_{\mathbf{x} \ne \mathbf{0}} \frac{\lVert
+      \mathbf{Ax} \rVert_p}{\lVert \mathbf{x} \rVert_p}.
+  \]
+  In particular, for $p = \infty$, we have
+  \[
+    \lVert \mathbf{A} \rVert_\infty = \max_i \sum_{j} |A_{i,j}|.
+  \]
+\end{definition}
+
+\begin{lemma}[Perturbation Lemma
+  \cite{layton2014numerical}] \label{lemma:perturbation}
+  Let $\lVert \cdot \rVert$ be any matrix norm, and let $\mathbf{A}$ and
+  $\mathbf{E}$ be matrices such that $\mathbf{A}$ is invertible and $\lVert
+  \mathbf{A}^{-1} \rVert \lVert \mathbf{E} \rVert < 1$, then $\mathbf{A} +
+  \mathbf{E}$ is invertible, and
+  \[
+    \lVert (\mathbf{A} + \mathbf{E})^{-1} \rVert \le \frac{\lVert
+      \mathbf{A}^{-1} \rVert}{1 - \lVert \mathbf{A}^{-1} \rVert \lVert
+      \mathbf{E} \rVert}.
+  \]
+\end{lemma}
+
+\begin{theorem}[Dominated Convergence Theorem
+  \cite{royden2010real}] \label{thm:lebesgue}
+  Let $(X, \mathcal{M}, \mu)$ be a measure space and $\{ f_n \}$ a sequence of
+  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise a.e. on $X$
+  and the function $f$ is measurable. Assume there is a non-negative function
+  $g$ that is integrable over $X$ and dominates the sequence $\{ f_n \}$ on $X$
+  in the sense that
+  \[
+    |f_n| \le g \text{ almost everywhere on $X$ for all $n$.}
+  \]
+  Then $f$ is integrable over $X$ and
+  \[
+    \lim_{n \to \infty} \int_X f_n\,\mathrm{d}\mu = \int_X f\,\mathrm{d}\mu.
+  \]
+\end{theorem}
+
+\subsection{Derivatives of PDFs and Their Bounds} \label{sec:derivatives}
+
+The goal of this section is to find derivatives of $q(\mathbf{u})$ and
+$q(\mathbf{r} \mid \mathbf{u})$ with respect to variational parameters and
+bound their values. Note that throughout this section the word `constant' means
+`constant with respect to $\mathbf{u}$ and $\mathbf{r}$'. We begin by
+introducing a few extra variables in order to simplify expressions of
+derivatives:
 \begin{align*}
@@
 \end{align*}
-Also note that throughout this section the word `constant' means `constant with
-respect to $\mathbf{u}$ and $\mathbf{r}$'.
 
@@
 
+\subsection{The MDP Value Function} \label{sec:propositions}
+
+In this section, we present two results that are necessary for the main proof in
+Section \ref{sec:the_proof} but also interesting in their own right. Firstly, we
+sketch a proof for the measurability of the MDP value function. Afterwards, we
+establish upper and lower bounds on said function.
+
 \begin{remark}
-  MDP values are characterised by both a state and a reward function/vector. For
-  the rest of this section, we will think of the value function as $V :
-  \mathcal{S} \to \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a
-  state $s \in \mathcal{S}$ and returns a function $V(s) :
-  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ that takes a reward vector
-  $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state
-  $s$, $\V(s) \in \mathbb{R}$. Given a reward vector, the function $V(s)$
-  computes the values of all states and returns the value of state $s$.
+  MDP values are characterised by both a state and a reward function/vector. In
+  this section, we will think of the value function as $V : \mathcal{S} \to
+  \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a state $s \in
+  \mathcal{S}$ and returns a function $V(s) : \mathbb{R}^{|\mathcal{S}|} \to
+  \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
+  \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$, $\V(s) \in
+  \mathbb{R}$. Given a reward vector, the function $V(s)$ computes the values of
+  all states and returns the value of state $s$.
 \end{remark}
@@
 
+\subsection{Differentiability} \label{sec:the_proof}
+
 Our main theorem is a specialised version of an integral differentiation result
@@
 
+Having proven that differentiation can be moved inside the integral allows us to
+use $\nabla\mathcal{L}$ for optimisation, knowing that our approach is sound.
+This allows us to implement a simple optimisation algorithm with a constant step
+size. The algorithm was implemented in MATLAB and is compatible with the
+toolkit\footnote{\url{https://graphics.stanford.edu/projects/gpirl/}} for
+evaluating IRL algorithms developed by Levine et al.
+\cite{DBLP:conf/nips/LevinePK11}. Note that due to the stochastic nature of the
+gradient, dynamic step size algorithms such as AdaGrad
+\cite{DBLP:journals/jmlr/DuchiHS11} and AdaDelta
+\cite{DBLP:journals/corr/abs-1212-5701} proved ineffective, as spikes in the
+gradient estimates would drive the step size down to zero. Similarly, although
+we check the $\ell_1$-norm of the difference between old and new parameter
+values in order to detect convergence, the algorithm usually only stops when the
+maximum number of iterations is exhausted.
+% TODO (must-y): An algorithm box / pseudo code outlining the algorithms (with
+% links to the many equations) so the reader can easily get an impression of the
+% (training) algorithm and the complexity.
+
 \section{Experiments} \label{sec:evaluation}
@@
 
-In order to fully understand the model's behaviour, we start with a three-state
-MDP where the agent can deterministically move from any state to any other
-state. More formally, we set $\mathcal{S} = \{ s_1, s_2, s_3 \}$, $\mathcal{A} =
-\{ a_1, a_2 \}$,
+The main goal of this section is use experimental evidence in order to show that
+the implementation agrees with theory and to investigate its properties such as
+convergence and how the learned model changes with varying amounts and types of
+data. We begin with a three-state MDP where the agent can deterministically move
+from any state to any other state. More formally, we set $\mathcal{S} = \{ s_1,
+s_2, s_3 \}$, $\mathcal{A} = \{ a_1, a_2 \}$,
 \begin{align*}
@@
 and create two demonstrations $\zeta_1 = \{ (s_1, a_1) \}$ and $\zeta_2 = \{
-(s_3, a_2) \}$ that correspond to moving from $s_1$ and $s_3$ to $s_2$.
+(s_3, a_2) \}$ that correspond to moving from $s_1$ and from $s_3$ to $s_2$.
 Therefore, we would expect the reward of $s_2$ to be higher than the other two
@@
 
-Unfortunately, the resulting plot in Figure~\ref{fig:covariance} shows no clear
+However, the resulting plot in Figure~\ref{fig:covariance} shows no clear
 pattern, suggesting that adding more of the same demonstrations does not result
@@
 
-% TODO: conclude the experiments
-
-\section{Related Work} % 0.5-1 page
-
-The IRL problem itself was originally proposed by Russell in 1998
-\cite{DBLP:conf/colt/Russell98}. Most of the early approaches had the
-aforementioned reward linearity assumption. One of the first papers on the
-subject by Ng and Russell \cite{DBLP:conf/icml/NgR00} introduced several linear
-programming algorithms and identified an important issue: there are typically
-many reward functions that can explain the data equally well. This problem was
-solved by Ziebart et al. \cite{ziebart2008maximum} with the introduction of IRL
-based on the principles of maximum causal entropy in a linearly-solvable MDP.
-
-Levine et al. \cite{DBLP:conf/nips/LevinePK11} were the first to lift the
-linearity assumption without imposing additional restrictions on the problem.
-They do, however, model rewards as having no variance---our work removes this
-restriction without any compromises.
-
-Recently, Jin et al. \cite{DBLP:conf/uai/JinDAS17} have adapted the model
-proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11} to use deep GPs,
-harnessing the power of deep learning to make the model less dependent on what
-features are provided. Although they use VI, their approximating distribution
-for rewards at inducing points is simply the Dirac $\delta$ function, which is
-essentially equivalent to the assumption of no variance.
-
-An alternative to GPs for modelling nonlinear functions is, of course, neural
-networks. Wulfmeier et al. \cite{wulfmeier2015maximum} have shown how they can
-be used in the IRL setting. While this approach benefits from constant-time
-inference and the ability to learn complex features from data, neural
-networks often need significantly more data for the weights across all layers to
-stabilise.
-
-\subsection{Variational Inference}
-
-Since the focus of our work was on proving feasibility rather than ensuring
-performance, we simply used a combination of Gaussians for our variational
-approximation. However, while VI was initially focused on approximating
-distributions using simplistic models where all variables are independent
-\cite{blei2017variational}, the last few years have brought many advances in
-approximating more complex distributions and greatly reducing the computational
-complexity of the task. Adapting some of them to IRL is a nontrivial but highly
-valuable undertaking.
-
-For approximating complex distributions, Rezende and Mohamed
-\cite{DBLP:conf/icml/RezendeM15} suggest using \emph{normalising flows}, i.e., a
-collection of invertible functions---parametrised by additional variational
-parameters---that are applied to latent variables. A major challenge in applying
-their work to IRL is related to variational parameters being used to compute the
-MDP value function, i.e., how can we take the derivative of the value function
-with respect to such a variational parameter? Alternatively, perhaps one can
-construct a different model that would make this question moot.
-
-Another approach to flexibility in modelling could come from considering
-different GP kernels. For instance, Wilson and Adams \cite{pmlr-v28-wilson13}
-show how all \emph{stationary} (i.e., invariant to translations) kernels can be
-generated (or at least approximated) from a mixture of Gaussians in their
-spectral representation using Bochner's theorem
-\cite{bochner1959lectures,DBLP:journals/technometrics/Woodard00}. It looks
-promising to combine these kernels with the variational Fourier features
-approach by Hensman et al. \cite{DBLP:journals/jmlr/HensmanDS17} that leverages
-the same spectral representations for efficient VI.
-
-\section{Conclusion}
+\section{Conclusions and Further Work} \label{sec:conclusions}
 % TODO: write
@@
 
-\subsection{Further Work}
-%TODO: how optimal policies depend on the structure of the mdp and available
-%data (i.e., why don't our policies converge to 1)
+% FURTHER WORK
 
@@
 
-It is also worth noting the approach presented in this paper requires solving
-$S$ MDPs for every iteration of optimising the parameters (where $S$ is the
-number of samples drawn from $\approximation$). There are at least two ways to
-reduce or eliminate this performance bottleneck:
-\begin{itemize}
-\item The MDP value function could be approximated, allowing for some minor
-  mistakes in the resulting policy.
-\item Perhaps there is a good way to use information about previously computed
-  value functions for similar rewards to hasten the current computation. One
-  simple way to do this would be by initialising the current values to the
-  optimal values of the previous MDP value function computation.
-\end{itemize}
+It is also worth noting that the approach presented in this paper requires solving
+$S$ MDPs for every iteration of optimisation (where $S$ is the
+number of samples drawn from $\approximation$). The running time could be
+significantly improved by reusing the solutions of previous MDPs to initialise
+the new (unsolved) MDP in order to hasten the convergence of value iteration.
 

### 2019-04-12 mpaper/paper.tex
Short conclusions and pseudocode!

@@
 \usepackage{blkarray}
+\usepackage[boxed]{algorithm2e}
 
@@
 e^{\lambda'_i}$ and optimise $\lambda'_i$ using the chain rule:
-\[
+\begin{equation} \label{eq:transformation}
   \frac{\partial \mathcal{L}}{\partial \lambda'_i} = e^{\lambda'_i}
   \frac{\partial \mathcal{L}}{\partial \lambda_i}.
-\]
+\end{equation}
 This way, we restrict $\lambda_i$ to positive values while allowing $\lambda'_i$
@@
   values of $\mathbf{B}$ can often result in a near-singular $\bm\Sigma$.}:
-\begin{alignat*}{3}
-  \mu_i &\sim \mathcal{U}(0, 1) \quad &&\text{for } i = 1, \dots, m, \\
+%\begin{equation} \label{eq:initialisation}
+\begin{alignat}{3}
+  \mu_i &\sim \mathcal{U}(0, 1) \quad &&\text{for } i = 1, \dots, m, \label{eq:first_initialisation} \\
   \lambda_0 &\sim \chi^2_5, && \\
@@
   \diag(\mathbf{B}) &\sim \chi^2_4, && \\
-  \text{the rest of } \mathbf{B} &\sim \mathcal{N}(0, 1). &&
-\end{alignat*}
+  \text{the rest of } \mathbf{B} &\sim \mathcal{N}(0, 1). \label{eq:last_initialisation} &&
+\end{alignat}
+%\end{equation}
 The initialisation of $\bm\mu$ mirrors the initialisation of $\mathbf{r}$ in
@@
 use $\nabla\mathcal{L}$ for optimisation, knowing that our approach is sound.
-This allows us to implement a simple optimisation algorithm with a constant step
-size. The algorithm was implemented in MATLAB and is compatible with the
+This allows us to implement a simple optimisation algorithm, the description of
+which is in the next section.
+
+\section{Experiments} \label{sec:evaluation}
+% showing deep understanding
+
+The main goal of this section is use experimental evidence to show that the
+implementation agrees with theory and to investigate its properties such as
+convergence and how the learned model changes with varying amounts and types of
+data. We begin with an outline of the implementation in
+Algorithm~\ref{alg:optimisation}. The algorithm was implemented in MATLAB and is
+compatible with the
 toolkit\footnote{\url{https://graphics.stanford.edu/projects/gpirl/}} for
 evaluating IRL algorithms developed by Levine et al.
-\cite{DBLP:conf/nips/LevinePK11}. Note that due to the stochastic nature of the
-gradient, dynamic step size algorithms such as AdaGrad
-\cite{DBLP:journals/jmlr/DuchiHS11} and AdaDelta
-\cite{DBLP:journals/corr/abs-1212-5701} proved ineffective, as spikes in the
-gradient estimates would drive the step size down to zero. Similarly, although
-we check the $\ell_1$-norm of the difference between old and new parameter
-values in order to detect convergence, the algorithm usually only stops when the
-maximum number of iterations is exhausted.
-% TODO (must-y): An algorithm box / pseudo code outlining the algorithms (with
-% links to the many equations) so the reader can easily get an impression of the
-% (training) algorithm and the complexity.
-
-\section{Experiments} \label{sec:evaluation}
-% showing deep understanding
-
-The main goal of this section is use experimental evidence in order to show that
-the implementation agrees with theory and to investigate its properties such as
-convergence and how the learned model changes with varying amounts and types of
-data. We begin with a three-state MDP where the agent can deterministically move
-from any state to any other state. More formally, we set $\mathcal{S} = \{ s_1,
-s_2, s_3 \}$, $\mathcal{A} = \{ a_1, a_2 \}$,
+\cite{DBLP:conf/nips/LevinePK11}. We decided to use a constant step size because
+the stochastic nature of the gradient makes dynamic step size algorithms such as
+AdaGrad \cite{DBLP:journals/jmlr/DuchiHS11} and AdaDelta
+\cite{DBLP:journals/corr/abs-1212-5701} ineffective, as spikes in the gradient
+estimates drive the step size down to zero. Similarly, although we check for
+convergence using the $\ell_\infty$ norm, the algorithm usually only stops when
+the maximum number of iterations is exhausted.
+
+\begin{algorithm}
+  \SetKwData{parameters}{parameters}
+  \SetKwData{stepSize}{stepSize}
+  \SetKwData{ttolerance}{tolerance}
+  \SetKwData{numIterations}{numIterations}
+  \SetKwFunction{estimatedPart}{estimatedPart}
+  \SetKwFunction{exactPart}{exactPart}
+  \SetKwFunction{pack}{pack}
+  \SetKwFunction{unpack}{unpack}
+  \SetKwFunction{valueIteration}{valueIteration}
+  \SetKwFunction{transformGradient}{transformGradient}
+  \SetKwProg{Function}{Function}{}{end}
+  \KwData{number of samples $S$, constant step size \stepSize, convergence
+    criteria \ttolerance and \numIterations}
+  \KwResult{optimised \parameters}
+  initialise \parameters as in Equations~\eqref{eq:first_initialisation}--\eqref{eq:last_initialisation}\;
+  \For{$i \gets 1$ \KwTo \numIterations}{
+    $\{ \mathbf{u}_s \}_{s=1}^S \sim \mathcal{N}(\parameters.\bm\mu, \parameters.\bm\Sigma)$\;
+    $\nabla\mathcal{L} \gets \exactPart{\parameters} + \frac{1}{S} \sum_{s=1}^S\estimatedPart{\parameters, $\mathbf{u}_s$}$\;
+    $\mathbf{v} \gets$ \pack{\parameters}\;
+    $\mathbf{v}' \gets \mathbf{v} + \stepSize \times \transformGradient{$\nabla\mathcal{L}$}$\;
+    $\parameters \gets \unpack{$\mathbf{v}'$}$\;
+    \lIf{$\lVert \mathbf{v}' - \mathbf{v} \rVert_\infty < \ttolerance$}{break}
+  }
+  \Function{\estimatedPart{\parameters, $\mathbf{u}$}}{
+    $\mathbf{r} \sim \mathcal{N}(\parameters.\mathbf{Su}, \parameters.\bm\Gamma)$\;
+    $\V \gets \valueIteration{$\mathbf{r}$}$\;
+    \Return{the estimated part of $\nabla\mathcal{L}$ according to $\mathbf{r}$,
+      $\V$, and \parameters\;}
+  }
+  \caption{The optimisation procedure. All parameters are kept in a single variable
+    \textsf{parameters}. Converting it to/from vector form and using $\log$,
+    $\exp$, and Equation~\eqref{eq:transformation} to handle the optimisation of
+    parameters restricted to positive values is handled by functions
+    \texttt{pack}, \texttt{unpack}, and \texttt{transformGradient}. Each
+    derivative from Appendix~\ref{appendix:derivatives} is split into parts
+    outside and inside $\mathbb{E}[\cdot]$ which are handled differently by
+    functions \texttt{exactPart} and \texttt{estimatedPart}. The function
+    \texttt{valueIteration} simply applies Equation~\eqref{eq:update_rule} until
+    convergence.}
+  \label{alg:optimisation}
+\end{algorithm}
+
+We will first focus on a simple three-state MDP where the agent can
+deterministically move from any state to any other state. More formally, we set
+$\mathcal{S} = \{s_1, s_2, s_3 \}$, $\mathcal{A} = \{ a_1, a_2 \}$,
 \begin{align*}
@@
 and create two demonstrations $\zeta_1 = \{ (s_1, a_1) \}$ and $\zeta_2 = \{
-(s_3, a_2) \}$ that correspond to moving from $s_1$ and from $s_3$ to $s_2$.
-Therefore, we would expect the reward of $s_2$ to be higher than the other two
-rewards in order to reflect this.
+(s_3, a_2) \}$ that correspond to moving from $s_1$ and $s_3$ to $s_2$. With
+this setup, we would expect the reward of $s_2$ to be higher than the other two
+rewards.
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}
   \centering
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!t]
   \centering
@@
 
-\begin{figure}[!ht]
+% TODO: must shorten something by ~1 line
+\begin{figure}[!t]
   \centering
@@
 \end{figure}
+% TODO: probably need to conclude the experiments
 
 \section{Conclusions and Further Work} \label{sec:conclusions}
-% TODO: write
 % Show how we achieved our claimed contributions
 
-\begin{itemize}
-\item Why my theoretical contributions are important...
-\item Reasonable results with policy convergence.
-\item Shown how to eliminate the deterministic training conditional
-  assumption---a common weakness of previous approaches.
-\item Being able to model rewards using full probability distributions without
-  limiting assumptions opens up many new research directions.
-\item Determining how reward covariance depends on data would be the next step.
-\end{itemize}
+In this paper, we showed how VI can be used to solve the IRL problem. We
+proved the validity of our approach, establishing subsidiary results such as new
+bounds for the linearly solvable MDP model along the way. The implementation of
+the model was well-behaved and showed that covariance can indeed be used to
+recognise a well-learned model from a model which is less sure about what
+policies are optimal.
 
 % FURTHER WORK
-
-An interesting extension to our work would be to consider IRL in the context
-of a reinforcement learning (RL) agent. Suppose we have an agent whose purpose
-is to learn optimal behaviour from observing other agents using IRL. It could
-then take reward variance estimates into account when choosing what states to
-visit next. It would have to handle the balance between exploration and
-exploitation similarly to many RL agents, but the information about rewards
-would come from observing (presumably near-optimal) behaviour exhibited by other
-agents rather than directly from the environment.
+Building on the that, an interesting question for MDP (or, perhaps, dynamical
+systems) research would be: how much does a reward have to change in order to
+affect the deterministic policy? A simple answer to this question would allow us
+to use variance estimates in order to quantify the model's confidence regarding
+optimal behaviour.
 
@@
 
-Finally, building on the idea that variance estimates can be used to judge
-whether the model has learned optimal policy, an interesting question for
-MDP (or, perhaps, dynamical systems) research would be: how much does a reward
-have to change in order to affect the deterministic policy? A simple answer to
-this question would allow us to use variance estimates in order to quantify the
-model's confidence regarding optimal behaviour.
+Finally, an interesting extension to our work would be to consider IRL in the
+context of a reinforcement learning (RL) agent. Suppose we have an agent whose
+purpose is to learn optimal behaviour from observing other agents using IRL. It
+could then take reward variance estimates into account when choosing what states
+to visit next. It would have to handle the balance between exploration and
+exploitation similarly to many RL agents, but the information about rewards
+would come from observing (presumably near-optimal) behaviour exhibited by other
+agents rather than directly from the environment.
 
@@
 
-\section{Derivatives of the ELBO}
+\section{Derivatives of the ELBO} \label{appendix:derivatives}
 

### 2019-04-13 mpaper/paper.tex
Major refactoring

@@
 around people so that it learns to predict where people are going and adjust its
-path accordingly. The ML algorithm would use data about various possible
-situations. But do we have enough data to ensure reasonably optimal behaviour?
-Perhaps the robot behaves well in most situations, but fails in less common
-scenarios. Can the ML model itself describe its weaknesses so that we could
-ensure it is exposed to sufficiently many uncommon or difficult situations?
-
-This learning problem
-\cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} as well as
+path accordingly. The ML algorithm would learn from data about proper actions in
+a range of different situations. But do we have enough data to ensure reasonably
+optimal behaviour? Perhaps the robot behaves well in most situations, but fails
+in less common scenarios. Can the ML model itself describe its weaknesses so
+that we could ensure it is exposed to sufficiently many uncommon or difficult
+situations?
+
+This learning problem \cite{DBLP:journals/ijrr/KretzschmarSSB16} as well as
 many others have benefited from an approach called \emph{inverse reinforcement
@@
 find a reward function for a Markov decision process (MDP), where demonstrations
-are encoded as sets of state-action pairs.
+are encoded as sets of state-action pairs, showing observed actions in a number
+of states.
 
@@
 learning the reward function rather than the policy itself makes the model more
-transferable to new environments---a minor change in the environment can
+transferable to new environments: a minor change in the environment can
 reorganise the whole policy but only have a local effect in the reward structure
@@
 \cite{DBLP:conf/icml/PieterN04,DBLP:conf/icml/NgR00,ziebart2008maximum}.
-This assumption severely restricts what reward structures can be modelled (i.e.,
-the properties of the function that represents rewards across states). Out of
-the non-linear models proposed to date, none can answer the questions posed in
-the first paragraph. More specifically, without knowing the ground truth for the
-learning problem, we have no reliable way to know whether the learned behaviour
-is optimal and which parts of the model could benefit from more data. Quite
-often, the models assume that rewards have no variance
+This assumption severely restricts what reward structures can be modelled. Out
+of the non-linear models proposed to date, none can answer the questions posed
+in the first paragraph. More specifically, without knowing the ground truth for
+the learning problem, we have no reliable way to know whether the learned
+behaviour is optimal and which parts of the model could benefit from more data.
+Quite often, the models assume that rewards have no variance
 \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}.
@@
 
-% TODO: could raise intractable posterior as the main reason, but would have to
-% rephrase the whole section
 No fully Bayesian nonlinear IRL model has been proposed yet. We will uncover one
 reason for this when we set up our IRL model for VI in Section~\ref{sec:model}.
-Applying VI to this model in a theoretically sound way (without introducing
+Applying VI to this model in a theoretically sound way (and without introducing
 unjustified simplifying assumptions) requires us to use Lebesgue's dominated
@@
 
-For any matrix $\mathbf{A}$, we use either $A_{i,j}$ or
+For any matrix $\mathbf{A}$, we write either $A_{i,j}$ or
 $[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
@@
 \section{Background} % 0.5-1 page
-% TODO: edit this: we can probably extend this (and the introduction)
 
@@
 solved by Ziebart et al. \cite{ziebart2008maximum} with the introduction of IRL
-based on the principles of maximum causal entropy in a linearly-solvable MDP.
+based on the principle of maximum causal entropy in a linearly-solvable MDP.
 
 Levine et al. \cite{DBLP:conf/nips/LevinePK11} were the first to lift the
-linearity assumption without imposing additional restrictions on the problem.
-They do, however, model rewards as having no variance---our work removes this
-restriction without any compromises.
+linearity assumption without imposing additional restrictions on the problem by
+modelling rewards with a Gaussian process (GP). They do, however, assume that
+rewards have no variance---our work removes this restriction without any
+compromises.
 
@@
 reasons. First, it is not directly relevant to the goal of producing accurate
-point estimates of rewards. However, we described several key benefits in the
-previous section, and we postulate how this work could affect the wider fields
-of reinforcement learning and artificial intelligence in
-Section~\ref{sec:conclusions}. Second, as it will become apparent in
-Sections~\ref{sec:model} and \ref{sec:proof}, optimising such a VI model
-involves estimates of gradients, establishing correctness of which requires a
-significant amount of work.
+point estimates of rewards. However, we described several key benefits of
+variance prediction in the previous section, and we postulate how this work
+could affect the wider fields of reinforcement learning and artificial
+intelligence in Section~\ref{sec:conclusions}. Second, as it will become
+apparent in Sections~\ref{sec:model} and \ref{sec:proof}, optimising such a VI
+model involves estimating gradients, and establishing correctness of these
+estimates requires a significant amount of theoretical groundwork.
 
@@
 Gaussian) into an arbitrarily complex one.
-%A major challenge in applying this
-%work to IRL is related to variational parameters being used to compute the MDP
-%value function, i.e., how can we take the derivative of the value function with
-%respect to such a variational parameter? Alternatively, perhaps one can
-%construct a different model that would make this question moot.
-
-Another approach to flexibility in modelling could come from considering
-different GP kernels. For instance, Wilson and Adams \cite{pmlr-v28-wilson13}
-show how all \emph{stationary} (i.e., invariant to translations) kernels can be
-generated (or at least approximated) from a mixture of Gaussians in their
-spectral representation using Bochner's theorem
-\cite{bochner1959lectures,DBLP:journals/technometrics/Woodard00}. It looks
+
+Another approach to flexibility in modelling can be achieved by considering
+different GP kernels, i.e., functions that model similarity between pairs of
+points. For instance, Wilson and Adams \cite{pmlr-v28-wilson13} show how all
+\emph{stationary} (i.e., invariant to translations) kernels can be generated (or
+at least approximated) from a mixture of Gaussians in their spectral
+representation using Bochner's theorem \cite{bochner1959lectures}. It looks
 promising to combine these kernels with the variational Fourier features
 approach by Hensman et al. \cite{DBLP:journals/jmlr/HensmanDS17} that leverages
-the same spectral representations for efficient VI.
+the same spectral representations for efficient VI. This algorithm, however, is
+currently restricted to Mat\'ern kernels, which could be restrictive,
+considering how IRL models tend to use kernels that can adjust the importance of
+each feature \cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
 
 While these and many other advancements in VI promise great improvements, there
-are hidden challenges when it comes to applying them to VI. Many suggested
+are hidden challenges when it comes to applying them to IRL. Many suggested
 improvements implicitly impose restrictions on the model that are unlikely to be
 satisfied in the context of IRL, e.g., the evidence of the model being a
-Gaussian. Nevertheless, the work outlined here seems adaptable to the IRL
+Gaussian. Nevertheless, the work mentioned here seems adaptable to the IRL
 setting and can make variational approximations both more accurate and faster.
@@
 
-In this section, we introduce definitions and mathematical details relevant to
-the problem. We begin by formally defining the problem.
+In this section, we describe the problem in more detail, introducing the
+notation and definitions that will be used throughout the paper.
 
@@
   \mathcal{A}, \mathcal{T}, \gamma, \mathbf{r} \}$, where $\mathcal{S}$ and
-  $\mathcal{A}$ are sets of states and actions, respectively; $\mathcal{T} :
-  \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is a function
-  defined so that $\mathcal{T}(s, a, s')$ is the probability of moving to state
-  $s'$ after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the discount
-  factor; and $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ is the reward
+  $\mathcal{A}$ are sets of states and actions, respectively; $\mathcal{T}
+  \colon \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is a
+  function defined so that $\mathcal{T}(s, a, s')$ is the probability of moving
+  to state $s'$ after taking action $a$ in state $s$; $\gamma \in [0, 1)$ is the
+  discount factor; and $\mathbf{r} \in \mathbb{R}^{|\mathcal{S}|}$ is the reward
   vector.
@@
 
-While it is more common to represent rewards as a function $r : \mathcal{S} \to
-\mathbb{R}$, the vector notation is generally more appropriate in our situation.
-However, we switch between the two notations as needed.
+While it is more common to represent rewards as a function $r \colon \mathcal{S}
+\to \mathbb{R}$, the vector notation is generally more appropriate in our
+situation. However, we switch between the two notations as needed.
 
@@
   \}_{t=1}^T$ is a multiset of state-action pairs that represents optimal (or
-  near-optimal) actions, find the reward function that maximises the probability
+  near-optimal) actions, find a reward function that maximises the probability
   of observing the demonstrations, i.e.,
@@
   state-action pairs into separate demonstrations is irrelevant to our model
-  (and possible all other models), so we can keep each pair in its own set as
-  follows:
+  (and presumably all other models) and is only a relic of the practical
+  motivation behind the problem's initial definition. Hence, we can keep each
+  pair in its own set as follows:
   \[
@@
       -1 & \text{if } s = s_3 \\
-      1 & \text{if } s = s_4.
-    \end{cases}
+      1 & \text{if } s = s_4
+    \end{cases}.
   \]
@@
     state-action pair appears in $\mathcal{D}$. If that number is zero, the edge
-    is dashed. Finally, the axes assign two features to each state.}
+    is dashed. Finally, the axes are used to assign two features to each state.}
   \label{fig:example}
@@
 
-The set of demonstrations $\mathcal{D}$ often comes from observed human
-behaviour and constitutes the most significant part of the data. The optimal
-(deterministic) \emph{policy} $\pi : \mathcal{S} \to \mathcal{A}$ (i.e., a
-choice of actions for each state that maximises reward over time) is usually
-constructed by defining a \emph{value (utility) function} $\V : \mathcal{S} \to
-\mathbb{R}$ that measures how good a state is based on the reward $\mathbf{r}$
-as well as the structure of the MDP. One can then find $\V$ by applying the
-Bellman backup operator until convergence to every $s \in \mathcal{S}$ (the
-technique is known as \emph{value iteration}) \cite{DBLP:books/daglib/0023820}:
+The optimal (deterministic) \emph{policy} $\pi \colon \mathcal{S} \to
+\mathcal{A}$ (i.e., a choice of actions for each state that maximises reward
+over time) is usually constructed by defining a \emph{value (utility) function}
+$\V \colon \mathcal{S} \to \mathbb{R}$ that measures how good a state is based
+on the reward $\mathbf{r}$ as well as the structure of the MDP. One can then
+find $\V$ by applying the Bellman backup operator until convergence to all
+states $s \in \mathcal{S}$ (this algorithm is known as \emph{value iteration})
+\cite{DBLP:books/daglib/0023820}:
 \[
@@
 
-However, we follow previous work on GP IRL
-\cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}, and use a
+However, we follow previous work on IRL with GPs
+\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11} and use a
 \emph{linearly solvable} (or \emph{maximum causal entropy}) MDP with a
-stochastic policy that defines probability distributions over actions (instead
-of suggesting a single action for each state) \cite{ziebart2008maximum}. This
+stochastic policy that defines probability distributions over actions instead
+of suggesting a single action for each state \cite{ziebart2008maximum}. This
 type of MDP can be solved by applying the `soft' version of the operator
@@
 instead of learning). For this reason, we represent each state by a
-$d$-dimensional feature vector. This way, similar states are located in close
-proximity to one another, and the ML model can predict rewards for new states
-based on their similarity to training data.
+$d$-dimensional feature vector. This way, similar states are represented by
+similar features, and the ML model can predict rewards for new states based on
+their similarity to training data.
 
@@
 $\mathbb{R}$ using a \emph{Gaussian process}. A GP represents an infinite
-collection of random variables, any finite number of which has a joint Gaussian
-distribution \cite{DBLP:books/lib/RasmussenW06}. A GP is defined by its mean
-function (which is always $\mathbf{x} \mapsto 0$ in our case) and covariance
-function, which we denote by $k$. \emph{Covariance functions} (also known as
-\emph{kernels}) take two state feature vectors as input and quantify how similar
-the two states are, in a sense that we would expect high covariance scores to be
-associated with similar rewards.
+collection of random variables from a joint (infinitely-dimensional) probability
+distribution, any finite number of which has a Gaussian distribution
+\cite{DBLP:books/lib/RasmussenW06}. A GP is defined by its mean function (which
+is always $\mathbf{x} \mapsto 0$ in our case) and covariance function, which we
+denote by $k$. \emph{Covariance functions} (also known as \emph{kernels}) take
+two state feature vectors as input and quantify how similar the two states are,
+in a sense that we would expect high covariance scores to be associated with
+similar rewards.
 
@@
 Thus, instead of minimising KL divergence, we focus on maximising $\mathcal{L}$
-by optimising the values of parameters to be defined in Section~\ref{sec:model}.
-Note that hereafter we drop the subscript notation, as all expected values will
-be with respect to $\approximation$.
+by optimising the values of parameters that we define in
+Section~\ref{sec:model}. Note that hereafter we drop the subscript notation, as
+all expected values will be with respect to $\approximation$.
 
@@
 define the GP, consider what parameters should be used to optimise $\mathcal{L}$
-and how they should be initialised, and in Section~\ref{sec:elbo}, we derive our
+and how they should be initialised, and in Section~\ref{sec:elbo} we derive our
 final expression for $\mathcal{L}$.
@@
 `always'. A proof of such a result would allow us to model $\bm\lambda$ instead
-of treating it as a variational parameter, and would help to guard against
-overfitting. For now, $\bm\lambda$ will have to be treated as a variational
+of treating it as a variational parameter, which would make the model fully
+Bayesian. For now, $\bm\lambda$ will have to be treated as a variational
 parameter.
 
-It remains to decide on the approximation distribution for $\mathbf{u}$ and
+It remains to decide on the approximating distribution for $\mathbf{u}$ and
 $\mathbf{r}$. As is commonly done when applying VI to GPs
@@
 The initialisation of $\bm\mu$ mirrors the initialisation of $\mathbf{r}$ in
-previous work by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. In contrast,
-while they have constant initial values for $\bm\lambda$, we sample from
-$\chi^2$ distributions centred around those values ($5$ for $\lambda_0$ and $1$
-for any other $\lambda_i$). The distributions for initial values of $\mathbf{B}$
-are simply set to provide a reasonable spread of positive values for the
-diagonal, and both positive and negative values for all other entries in the
-matrix.
+previous work \cite{DBLP:conf/nips/LevinePK11}. In contrast, while they have
+constant initial values for $\bm\lambda$, we sample from $\chi^2$ distributions
+centred around those values ($5$ for $\lambda_0$ and $1$ for any other
+$\lambda_i$). The distributions for initial values of $\mathbf{B}$ are simply
+set to provide a reasonable spread of positive values for the diagonal, and both
+positive and negative values for all other entries in the matrix.
 
@@
 provide a sketch proof of the measurability of the MDP value function as well as
-new upper and lower bounds on their values. After another quick lemma, the main
+new upper and lower bounds on its values. After another quick lemma, the main
 proof of the paper is in Section~\ref{sec:the_proof}. See
@@
   $\mathbf{E}$ be matrices such that $\mathbf{A}$ is invertible and $\lVert
-  \mathbf{A}^{-1} \rVert \lVert \mathbf{E} \rVert < 1$, then $\mathbf{A} +
+  \mathbf{A}^{-1} \rVert \lVert \mathbf{E} \rVert < 1$. Then $\mathbf{A} +
   \mathbf{E}$ is invertible, and
@@
   Let $(X, \mathcal{M}, \mu)$ be a measure space and $\{ f_n \}$ a sequence of
-  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise a.e. on $X$
-  and the function $f$ is measurable. Assume there is a non-negative function
-  $g$ that is integrable over $X$ and dominates the sequence $\{ f_n \}$ on $X$
-  in the sense that
+  measurable functions on $X$ for which $\{ f_n \} \to f$ pointwise almost
+  everywhere on $X$, and the function $f$ is measurable. Assume there is a
+  non-negative function $g$ that is integrable over $X$ and dominates the
+  sequence $\{ f_n \}$ on $X$ in the sense that
   \[
@@
 \end{align*}
+Next, we state a few key derivatives, derivations of which can be found in
+Appendix~\ref{appendix:proofs}.
 
@@
 \begin{lemma} \label{lemma:bound1}
-  Let $i \in \{ 0, \dots, d \}$ be arbitrary, and let $c :
+  Let $i \in \{ 0, \dots, d \}$ be arbitrary, and let $c \colon
   \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\lambda_i - \epsilon,
@@
 \begin{lemma} \label{lemma:bound2}
-  Let $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
+  Let $c \colon \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (a, b) \subset
   \mathbb{R}$ be an arbitrary bounded function. Then, for $i = 1, \dots, m$,
@@
   \[
-    c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\Sigma_{i,j} - \epsilon,
-    \Sigma_{i,j} + \epsilon) \subset \mathbb{R}
+    c \colon \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to (\Sigma_{i,j} -
+    \epsilon, \Sigma_{i,j} + \epsilon) \subset \mathbb{R}
   \]
@@
   MDP values are characterised by both a state and a reward function/vector. In
-  this section, we will think of the value function as $V : \mathcal{S} \to
+  this section, we will think of the value function as $V \colon \mathcal{S} \to
   \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$, i.e., $V$ takes a state $s \in
-  \mathcal{S}$ and returns a function $V(s) : \mathbb{R}^{|\mathcal{S}|} \to
-  \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
+  \mathcal{S}$ and returns a function $V(s) \colon \mathbb{R}^{|\mathcal{S}|}
+  \to \mathbb{R}$ that takes a reward vector $\mathbf{r} \in
   \mathbb{R}^{|\mathcal{S}|}$ and returns a value of the state $s$, $\V(s) \in
@@
 \begin{proposition}[Measurability] \label{thm:measurability}
-  MDP value functions $V(s) : \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$ (for $s
-  \in \mathcal{S}$) are Lebesgue measurable.
+  MDP value functions $V(s) \colon \mathbb{R}^{|\mathcal{S}|} \to \mathbb{R}$
+  (for $s \in \mathcal{S}$) are Lebesgue measurable.
 \end{proposition}
@@
   and assuming that the initial values of $\{ \V(s) \mid s \in \mathcal{S} \}$
-  already satisfy Inequality~\eqref{eq:positive_bound}. Recall that for each $s
-  \in \mathcal{S}$, the value of $\V(s)$ is updated by applying
+  already satisfy Inequality~\eqref{eq:positive_bound}. We will show that it
+  remains satisfied using an inductive argument. Recall that for each $s \in
+  \mathcal{S}$, the value of $\V(s)$ is updated by applying
   Equation~\eqref{eq:update_rule}. Note that both the natural logarithm and the
@@
     \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid
-    \mathbf{u})\,\mathrm{d}\mathbf{r} \le a + \lVert \Kru^\intercal \Kuu^{-1}
-    \mathbf{u} \rVert_1,
+    \mathbf{u})\,\mathrm{d}\mathbf{r} \le a + \lVert \mathbf{Su} \rVert_1,
   \]
@@
   \]
-  where $\xi_i = \left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right]_i$, $\sigma_i =
-  \sqrt{[\Krr - \Kru^\intercal\Kuu^{-1}\Kru]_{i,i}}$\footnote{The expression
-    under the square root sign is non-negative because $\Krr -
-    \Kru^\intercal\Kuu^{-1}\Kru$ is a covariance matrix of a Gaussian
-    distribution, hence also positive semi-definite, which means that its
-    diagonal entries are non-negative.}, and $\Phi$ is the cumulative
-  distribution function of the standard Gaussian. Furthermore,
+  where $\xi_i = \left[\mathbf{Su}\right]_i$, $\sigma_i =
+  \sqrt{\bm\Gamma_{i,i}}$\footnote{The expression
+    under the square root sign is non-negative because $\bm\Gamma$ is a
+    covariance matrix of a Gaussian distribution, hence also positive
+    semi-definite, which means that its diagonal entries are non-negative.}, and
+  $\Phi$ is the cumulative distribution function of the standard Gaussian.
+  Furthermore,
   \[
@@
   as $\sigma_i$ is non-negative, and $\Phi(x) \in [0, 1]$ for all $x$. Since
-  \[ \sum_{i=1}^{|\mathcal{S}|} |\xi_i| = \lVert \Kru^\intercal \Kuu^{-1}
-    \mathbf{u} \rVert_1, \]
+  \[
+    \sum_{i=1}^{|\mathcal{S}|} |\xi_i| = \lVert \mathbf{Su} \rVert_1,
+  \]
   we can set
@@
   \[
-    \frac{q(\mathbf{r} \mid \mathbf{u})|_{t = t_n} - q(\mathbf{r} \mid \mathbf{u})}{t_n - t} = \left.
-      \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial t} \right|_{t = c(\mathbf{r},
-      \mathbf{u})}
-  \]
-  for some function $c : \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to
+    \frac{q(\mathbf{r} \mid \mathbf{u})|_{t = t_n} - q(\mathbf{r} \mid
+      \mathbf{u})}{t_n - t} = \left. \frac{\partial q(\mathbf{r} \mid
+        \mathbf{u})}{\partial t} \right|_{t = c(\mathbf{r}, \mathbf{u})}
+  \]
+  for some function $c \colon \mathbb{R}^{|\mathcal{S}|} \times \mathbb{R}^m \to
   (\min\{t, t_n\}, \max\{t, t_n\})$ due to the mean value theorem (since $q$ is
@@
   \begin{equation} \label{eq:remaining_integral}
-    \int \left( a + \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1 \right) \max \{|p_1(\mathbf{u})|, |p_2(\mathbf{u})| \} q(\mathbf{u})\,\mathrm{d}\mathbf{u},
+    \int \left( a + \lVert \mathbf{Su} \rVert_1 \right) \max \{|p_1(\mathbf{u})|, |p_2(\mathbf{u})| \} q(\mathbf{u})\,\mathrm{d}\mathbf{u},
   \end{equation}
@@
   values are integrable, and the maximum of two integrable functions is also
-  integrable. Since $\lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1 \in
-  \mathbb{R}_1[\mathbf{u}]$, a similar argument can be applied to the rest of
+  integrable. Since $\lVert \mathbf{Su} \rVert_1 \in \mathbb{R}_1[\mathbf{u}]$,
+  a similar argument can be applied to the rest of
   Integral~\eqref{eq:remaining_integral} as well.
@@
 
-The main goal of this section is use experimental evidence to show that the
-implementation agrees with theory and to investigate its properties such as
+The main goal of this section is to use experimental evidence to show that the
+implementation agrees with the theory and to investigate its properties such as
 convergence and how the learned model changes with varying amounts and types of
 data. We begin with an outline of the implementation in
-Algorithm~\ref{alg:optimisation}. The algorithm was implemented in MATLAB and is
+Algorithm~\ref{alg:optimisation}. The algorithm was implemented in MATLAB to be
 compatible with the
@@
 estimates drive the step size down to zero. Similarly, although we check for
-convergence using the $\ell_\infty$ norm, the algorithm usually only stops when
-the maximum number of iterations is exhausted.
+convergence using the $\ell_\infty$-norm, the algorithm usually only stops when
+the maximum number of iterations is exhausted. Throughout the experiments, we
+discard runs that result in a near-singular $\bm\Sigma$ or $\bm\Gamma$,
+repeatedly testing each combination of parameters as many times as necessary
+until a non-failing run is achieved.
 
@@
     $\V \gets \valueIteration{$\mathbf{r}$}$\;
-    \Return{the estimated part of $\nabla\mathcal{L}$ according to $\mathbf{r}$,
-      $\V$, and \parameters\;}
+    \Return{the estimated part of $\nabla\mathcal{L}$ according to $\mathbf{u}$,
+      $\mathbf{r}$, $\V$, and \parameters\;}
   }
   \caption{The optimisation procedure. All parameters are kept in a single variable
-    \textsf{parameters}. Converting it to/from vector form and using $\log$,
-    $\exp$, and Equation~\eqref{eq:transformation} to handle the optimisation of
-    parameters restricted to positive values is handled by functions
-    \texttt{pack}, \texttt{unpack}, and \texttt{transformGradient}. Each
-    derivative from Appendix~\ref{appendix:derivatives} is split into parts
-    outside and inside $\mathbb{E}[\cdot]$ which are handled differently by
-    functions \texttt{exactPart} and \texttt{estimatedPart}. The function
+    \textsf{parameters}. Converting it to/from vector form and taking logarithms,
+    exponentiating, and using Equation~\eqref{eq:transformation} to handle the
+    optimisation of parameters restricted to positive values is handled by
+    functions \texttt{pack}, \texttt{unpack}, and \texttt{transformGradient}.
+    Each derivative from Appendix~\ref{appendix:derivatives} is split into parts
+    outside and inside $\mathbb{E}[\cdot]$ that are handled by functions
+    \texttt{exactPart} and \texttt{estimatedPart}, respectively. The function
     \texttt{valueIteration} simply applies Equation~\eqref{eq:update_rule} until
@@
 also set the inducing points to be equal to the three states in $\mathcal{S}$,
-add a single feature $f : \mathcal{S} \to \mathbb{R}$ such that
+add a single feature $f \colon \mathcal{S} \to \mathbb{R}$ such that
 \[
@@
 s_3)$, and $\pi(a_1 \mid s_2)$ converge over a number of iterations in
-Figure~\ref{fig:convergence1}\footnote{In all experiments, we discard runs that
-  result in a near-singular $\bm\Sigma$ or $\bm\Gamma$, repeatedly testing each
-  combination of parameters as many times as necessary.}. The first two policies
+Figure~\ref{fig:convergence1}. The first two policies
 correspond to actions taken in the set of demonstrations $\mathcal{D}$, so we
@@
 closely related to $r(s_2)$, stabilises at a positive value whereas both $\mu_1$
-and $\mu_2$ become negative.
+and $\mu_2$ move to negative values as expected.
 
@@
 $s_2$ to $s_1$, we expect it to stay around $0.5$ as long as we have equal
-amounts of $\zeta_1$ and $\zeta_2$ (this corresponds to the plots at coordinates
-$(1, 1)$, $(2, 2)$, and $(3, 3)$). But as we increase the number of times state
-$s_1$ appears in our demonstrations (going to the right across any row of
+amounts of $\zeta_1$ and $\zeta_2$---this corresponds to plots at coordinates
+$(1, 1)$, $(2, 2)$, and $(3, 3)$. Moreover, as we increase the number of times
+state $s_1$ appears in our demonstrations (going to the right across any row of
 plots), the model recognises the increasing value of state $s_1$, and the
@@
 probability, as state $s_3$ becomes more valuable, and the model prefers it over
-state $s_1$. Interestingly, the other two policy probabilities seem to increase
-regardless of what data is added. This is likely due to the fact that a higher
-number of demonstrations results in higher gradient values, and thus leads to
+state $s_1$. Interestingly, the other two policies seem to increase regardless
+of what data is added. This is likely due to the fact that a higher number of
+demonstrations results in higher gradient values, and thus leads to
 better-converged policies.
@@
 
-Finally, we would like to see similar changes in reward covariances. Firstly, we
+We would also like to see similar changes in reward covariances. Firstly, we
 will show how to derive the covariance matrix for $\mathbf{r}$ from the
@@
 \sim \mathcal{N}(\bm\mu', \bm\Sigma')$, where our goal is to find $\bm\mu'$ and
-$\bm\Sigma'$. Then
+$\bm\Sigma'$. Then,
 \[
@@
 
-If having more copies of the same demonstrations does not affect covariances, we
-can test if covariances tend to be lower in situations that exhibit more
-structure, i.e., demonstrations provide a clearer picture of which states are
-more valuable. For this, we consider a completely-deterministic MDP with ten
+If having more copies of the same demonstrations has no effect on covariances,
+we can test if covariances tend to be lower in situations that exhibit more
+structure, i.e., where demonstrations provide a clearer picture of which states
+are more valuable. For this, we consider a completely deterministic MDP with ten
 states and nine actions that allow the agent to move from any state to any other
@@
 rewards to be lower in the second scenario in order to reflect the more
-structured, certain behavioural pattern expressed by the demonstrations.
-
-% TODO: must shorten something by ~1 line
+structured, certain behavioural pattern expressed by the demonstrations. Indeed,
+Figure~\ref{fig:covariance_boxplots} shows exactly that. The first two box plots
+show that in roughly $3/4$ of the runs reward variances were higher in the
+random scenario\footnote{Note that outliers were removed from the plot. However,
+  all outliers were positive.}. Interestingly, the same applies to covariances.
+
 \begin{figure}[!t]
@@
 \end{figure}
-
-Indeed, Figure~\ref{fig:covariance_boxplots} shows exactly that. The first two
-box plots show that in roughly $3/4$ of the runs reward variances were higher in
-the random scenario\footnote{Note that outliers were removed from the plot.
-  However, all outliers were positive.}. Interestingly, the same applies to
-covariances.
 
@@
 \end{figure}
-% TODO: probably need to conclude the experiments
 
@@
 % FURTHER WORK
-Building on the that, an interesting question for MDP (or, perhaps, dynamical
+Building on that, an interesting question for MDP (or, perhaps, dynamical
 systems) research would be: how much does a reward have to change in order to
@@
 $S$ MDPs for every iteration of optimisation (where $S$ is the
-number of samples drawn from $\approximation$). The running time could be
-significantly improved by reusing the solutions of previous MDPs to initialise
-the new (unsolved) MDP in order to hasten the convergence of value iteration.
+number of samples drawn from $\approximation$). Thus, any improvements in MDP
+solving speed are highly desirable. Two simple improvements can come from
+requiring less precision when running value iteration and from reusing the
+solutions of previous MDPs to initialise the new (unsolved) MDP.
 

### 2019-04-13 code/VIGPIRL/deactivate_some_derivatives.m

@@ -1,0 +1,17 @@
+function gradient = deactivate_some_derivatives(gradient, algorithm_params, d, m)
+  if algorithm_params.disable_optimising_lambda0
+    gradient(1) = 0;
+  end
+  if algorithm_params.disable_optimising_lambda
+    gradient(2:d+1) = 0;
+  end
+  if algorithm_params.disable_optimising_mu
+    grad(d+m+2:d+2*m+1) = 0;
+  end
+  if algorithm_params.disable_optimising_B_diagonal
+    grad(d+2:d+m+1) = 0;
+  end
+  if algorithm_params.disable_optimising_rest_of_B
+    grad(d+2*m+2:end) = 0;
+  end
+end

### 2019-04-13 code/VIGPIRL/vigpirldefaultparams.m

@@ -1,21 +1,27 @@
-% Fill in default parameters for the GPIRL algorithm.
+% Fill in default parameters for the VIGPIRL algorithm.
 function algorithm_params = vigpirldefaultparams(algorithm_params)
 
 % Create default parameters.
 default_params = struct(...
-    ...% These are initial values.
-    'ard_init',1,...
-    'noise_init',0.01,...
-    'rbf_init',5,...
-    ...% These parameters control how the inducing points are selected.
-    'inducing_pts','examplesplus',...
-    'inducing_pts_count',16,...
-    'samples_count',100,...
-    'learning_rate',0.00001,...
-    'lambda1_learning_rate',0.00001,...
-    'B_learning_rate',0.00001,...
-    'num_iterations',100,...
-    'required_precision', 0.01);
+    'ard_init', 1,... % Mean initial value for lambda1
+    'noise_init', 0.01,... % 2 * sigma^2, the noise factor
+    'rbf_init', 5,... % Mean initial value for lambda0
+    'inducing_pts', 'examplesplus',...
+    'inducing_pts_count', 16,... % Number of inducing points
+    'samples_count', 100,... % S, the number of samples
+    'learning_rate', 0.1,... % Step sizes
+    'lambda1_learning_rate', 0.1,...
+    'B_learning_rate', 0.1,...
+    'num_iterations', 10,... % Max number of iterations for optimisation
+    'required_precision', 0.01,... % Alternatively, tolerance for convergence
+    'random_initial_B', false,...
+    'disable_optimising_lambda0', false,... % Mostly for debugging
+    'disable_optimising_lambda', false,...
+    'disable_optimising_mu', false,...
+    'disable_optimising_B_diagonal', false,...
+    'disable_optimising_rest_of_B', false,...
+    'step_size_algorithm', 'constant'... % Other values are 'AdaGrad' and 'AdaDelta'
+);
 
 % Set parameters.
-algorithm_params = filldefaultparams(algorithm_params,default_params);
+algorithm_params = filldefaultparams(algorithm_params, default_params);

### 2019-04-13 code/VIGPIRL/vigpirlrun.m

@@ -1,380 +1,24 @@
-                                % GP-based non-linear IRL algorithm.
-function irl_result = vigpirlrun(algorithm_params,mdp_data,mdp_model,...
-                                 feature_data,demonstrations,~,~)
+% GP-based non-linear IRL algorithm with variational inference
+function irl_result = vigpirlrun(algorithm_params, mdp_data, mdp_model,...
+                                 feature_data, demonstrations, ~, ~)
+  % algorithm_params - parameters of the GP IRL algorithm.
+  % mdp_data - definition of the MDP to be solved.
+  % demonstrations - cell array containing examples.
+  % irl_result - result of IRL algorithm (see bottom of file).
 
-                    % algorithm_params - parameters of the GP IRL algorithm.
-                    % mdp_data - definition of the MDP to be solved.
-                    % demonstrations - cell array containing examples.
-                    % irl_result - result of IRL algorithm (see bottom of file).
+  model = Vigpirl(algorithm_params, mdp_data, mdp_model, feature_data, demonstrations);
+  [model, time] = model.run();
+  model.convergence_plot();
+  model.policy_convergence_plot();
+  model.parameter_convergence_plot();
 
-                                % Fill in default parameters.
-  algorithm_params = vigpirldefaultparams(algorithm_params);
-
-                      % Get state-action counts and initial state distributions.
-  [mu_sa, counts] = vigpirlgetstatistics(demonstrations, mdp_data);
-
-                                % Create initial GP.
-  gp = vigpirlinit(algorithm_params, feature_data);
-                                % Choose inducing points.
-  %gp = vigpirlgetinducingpoints(gp,mu_sa,algorithm_params);
-  gp.X_u = gp.X;
-  m = size(gp.X_u, 1);
-  gp.mu = rand(1, m)';
-  %gp.mu = [-5; 0; 4]; % TEMP
-
[…]

### 2019-04-13 code/singletest.m

@@ -18,14 +18,14 @@ addpaths;
 %   'training_samples',64,'verbosity',1));
 
 % VIGPIRL
-test_result = runtest('vigpirl',struct(),'linearmdp',...
-    'gridworld',struct('n',8,'determinism',1,'seed',1,'continuous',0),...
-    struct('training_sample_lengths',32,'training_samples',16,'verbosity',2));
+%test_result = runtest('vigpirl',struct(),'linearmdp',...
+%    'gridworld',struct('n',8,'determinism',1,'seed',1,'continuous',0),...
+%    struct('training_sample_lengths',32,'training_samples',16,'verbosity',2));
 
 % Visualize solution.
-printresult(test_result);
-visualize(test_result);
-return;
+%printresult(test_result);
+%visualize(test_result);
+%return;
 
 % Two states
 %mdp_data = struct('sa_s', [2; 1], 'sa_p', [1; 1], 'discount', 0, 'states', 2, 'actions', 1);
@@ -33,14 +33,16 @@ return;
 %example_samples = {[1, 1]};
 
 % Three states
-%mdp_data = struct('discount', 0.9, 'states', 3, 'actions', 2);
-%mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
-%mdp_data.sa_p(1:3, 1:2, 1) = 1;
-%feature_data = struct('splittable', [1; 2; 3]);
-%example_samples = {[1, 1], [3, 2]};
+mdp_data = struct('discount', 0.9, 'states', 3, 'actions', 2);
+mdp_data.sa_s(:, :, 1) = [2, 3; 1, 3; 1, 2];
+mdp_data.sa_p(1:3, 1:2, 1) = 1;
+feature_data = struct('splittable', [1; 2; 3]);
+example_samples = {[1, 1], [3, 2]};
 %wrapper(mdp_data, feature_data, example_samples);
-%vigpirlrun(struct(), mdp_data, 'linearmdp', feature_data, example_samples);
-%return;
+vigpirlrun(struct(), mdp_data, 'linearmdp', feature_data, example_samples);
+%model = Vigpirl(struct(), mdp_data, 'linearmdp', feature_data, example_samples);
+%model.elbo_and_derivative_plot();
+return;
 
 % Fancy stuff 1
 %max_demonstrations_count = 3;

### 2019-04-16 mpaper/paper.tex
A bit more refactoring

@@
   where $\xi_i = \left[\mathbf{Su}\right]_i$, $\sigma_i =
-  \sqrt{\bm\Gamma_{i,i}}$\footnote{The expression
-    under the square root sign is non-negative because $\bm\Gamma$ is a
-    covariance matrix of a Gaussian distribution, hence also positive
-    semi-definite, which means that its diagonal entries are non-negative.}, and
-  $\Phi$ is the cumulative distribution function of the standard Gaussian.
-  Furthermore,
+  \sqrt{\Gamma_{i,i}}$\footnote{The diagonal of $\bm\Gamma$ is non-negative
+    because $\bm\Gamma$ is a covariance matrix of a Gaussian distribution, hence
+    also positive semi-definite.}, and $\Phi$ is the cumulative distribution
+  function of the standard Gaussian. Furthermore,
   \[

### 2019-04-16 code/VIGPIRL/Vigpirl.m

@@ -1,3 +1,6 @@
+% This class contains methods for running the algorithm as well as many
+% experiments and plots. Some plotting methods run their own experiments, some
+% should be called after run().
 classdef Vigpirl
     properties
         algorithm_params
@@ -8,12 +11,14 @@ classdef Vigpirl
         m                   % Number of inducing points
         mdp_data
 
-        elbo_list = [];
-        hyperparameter_history = [];
+        % Properties used to track how things change with each iteration
+        elbo_history = [];
+        parameter_history = [];
         grad_history = [];
         policy_history = [];
     end
     methods
+        % mdp should belong to class Mdp, algorithm_params can be an empty struct
         function obj = Vigpirl(mdp, algorithm_params)
             % Fill in default parameters
             obj.algorithm_params = vigpirldefaultparams(algorithm_params);
@@ -42,14 +47,16 @@ classdef Vigpirl
             end
         end
 
+        % Run the optimisation algorithm. Updates the object with new values of
+        % gp, extends all the history properties, and returns running time
         function [obj, time] = run(obj)
             % Set up variables for dynamic step size algorithms
-            num_hyperparameters = obj.m + obj.d + 1 + obj.m*(obj.m+1)/2;
+            num_parameters = obj.m + obj.d + 1 + obj.m * (obj.m + 1) / 2;
             if strcmp(obj.algorithm_params.step_size_algorithm, 'AdaGrad')
-                G = zeros(num_hyperparameters, 1);
+                G = zeros(num_parameters, 1);
             elseif strcmp(obj.algorithm_params.step_size_algorithm, 'AdaDelta')
-                E_g = zeros(num_hyperparameters, 1);
-                E_x = zeros(num_hyperparameters, 1);
+                E_g = zeros(num_parameters, 1);
+                E_x = zeros(num_parameters, 1);
                 epsilon = 1e-6;
                 rho = 0.6;
             end
@@ -59,86 +66,103 @@ classdef Vigpirl
                 fprintf('.');
                 % Compute the gradient
                 matrices = vigpirlkernel(obj.gp);
-                z = mvnrnd(obj.gp.mu', obj.gp.B * obj.gp.B', obj.algorithm_params.samples_count);
-                [elbo, grad] = full_gradient(obj.mdp_data, obj.demonstrations, obj.counts, obj.gp, z, matrices);
[…]

### 2019-04-16 code/VIGPIRL/vigpirldefaultparams.m

@@ -9,9 +9,9 @@ default_params = struct(...
     'lambda_init', 1,... % Mean initial value for lambda_i, i > 0
     'random_initial_B', false,... % Should initial B be random or constant?
     ... % Learning rates (step sizes)
-    'learning_rate', 0.001,...
-    'lambda1_learning_rate', 0.001,...
-    'B_learning_rate', 0.001,...
+    'learning_rate', 0.01,...
+    'lambda1_learning_rate', 0.01,...
+    'B_learning_rate', 0.01,...
     ... % Disabling optimisation of parameters (mostly for debugging)
     'disable_optimising_lambda0', false,...
     'disable_optimising_lambda', false,...

### 2019-04-16 code/singletest.m

@@ -28,11 +28,23 @@ addpaths;
 %return;
 
 mdp = ThreeStateMdp();
+
+% One run, many plots
+%model = Vigpirl(mdp, struct());
+%model = model.run();
+%model.convergence_plots();
+%model.elbo_and_policy_convergence_plots();
+%model.parameter_convergence_plot();
+
+% Plots that run their own experiments
+%model = Vigpirl(mdp, struct());
+%model.elbo_contour_plot();
+%model.policy_contour_plot();
+%model.elbo_vs_gamma_plot();
+%model.elbo_and_derivative_plot();
+
+% Fancy experiments specific to the three-state MDP
 %mdp.covariance_and_policy_with_more_data();
 %mdp.covariances_with_more_data();
-
-model = Vigpirl(mdp, struct());
-model.run();
-%model.elbo_and_derivative_plot();
 
 %clique_experiment();

### 2019-04-16 code/tests.m

@@ -3,7 +3,7 @@ function tests = tests
   tests = functiontests(localfunctions);
 end
 
-function test_full_gradient(testCase)
+function test_gradient(testCase)
   function c = k(x, y)
     c = rbf * exp(-0.5 * 2 * (x - y)^2);
   end
@@ -31,10 +31,10 @@ function test_full_gradient(testCase)
   mdp_solution = linearmdpsolve(mdp_data, r);
   v = counts' * mdp_solution.v;
 
-  [~, full_gradient_answer] = full_gradient(mdp_data, demonstrations, counts, gp, u, matrices, true);
+  [~, compute_gradient_answer] = compute_gradient(mdp_data, demonstrations, counts, gp, u, matrices, true);
 
   lambda0_derivative = 116 + 0.5 * (417 * v - 25);
-  verifyEqual(testCase, full_gradient_answer(1), lambda0_derivative);
+  verifyEqual(testCase, compute_gradient_answer(1), lambda0_derivative);
 
   B11_derivative = -8 * v;
   B22_derivative = -24 * v;
@@ -43,13 +43,13 @@ function test_full_gradient(testCase)
   B31_derivative = 0;
   B32_derivative = 0;
 
-  verifyEqual(testCase, full_gradient_answer(3), B11_derivative);
-  verifyEqual(testCase, full_gradient_answer(4), B22_derivative);
-  verifyEqual(testCase, full_gradient_answer(5), B33_derivative);
+  verifyEqual(testCase, compute_gradient_answer(3), B11_derivative);
+  verifyEqual(testCase, compute_gradient_answer(4), B22_derivative);
+  verifyEqual(testCase, compute_gradient_answer(5), B33_derivative);
 
-  verifyEqual(testCase, full_gradient_answer(9), B21_derivative);
-  verifyEqual(testCase, full_gradient_answer(10), B31_derivative);
-  verifyEqual(testCase, full_gradient_answer(11), B32_derivative);
+  verifyEqual(testCase, compute_gradient_answer(9), B21_derivative);
+  verifyEqual(testCase, compute_gradient_answer(10), B31_derivative);
+  verifyEqual(testCase, compute_gradient_answer(11), B32_derivative);
 end
 
 function test_hyperparameter_packing(testCase)
